### 2015-04-21 MLCV/ex1/report/template.tex
parallel prog and svm for mlfcv

\documentclass[journal, a4paper]{IEEEtran}

\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath}

\begin{document}

	\title{SVM for MNIST Handwritten Digits Classification}
	\author{Adam Kosiorek}
	\thanks{Advisor: M.Sc. Caner Hazibras & M.Sc. Phillip , Chair for Computer Vision, TUM, SS 2015.}
	\markboth{Machine Learning for Applications in Computer Vision}{}
	\maketitle

\begin{abstract}
  This report describes classification of handwritten digits from the MNIST dataset by an SVM classifier.
\end{abstract}

\section{Introduction}
    \PARstart{T}{he} first assignment was to explore the SVM classifier for the MNIST handwritten digits recognition.

\section{Dataset}
    The MNIST dataset is comprised of 60000 training and 10000 testing samples \cite{MNIST}. Each sample is a 28x28 pixel black and white image of a single centered digit. It is widly used as a benchmark for comparing machine learning algorithms.

\section{Experiments}
    All experiments were carried out on a notebook with Intel i7-2670QM quad core CPU and 8Gb RAM with Python SciPy package.
    \subsection{Dataset Preparation}
	It is known that processing input data prior to classification can improve results. In case of MNIST and SVM, however, state of the arts results have been achieved without preprocessing, that is, with raw pixels \cite{VSVM}. I compare the influence of mean intensity subtraction and ZCA-Whitening \cite{UFLDL} on classification accuracy.

    \subsection{Classifier Tuning}
	The SVM classifier can be adapted to a particular task by choosing a kernel type and adjusting several parameters. Since the parameters might not be independent it is often necessary to perform an exhaustive grid search. Python's sklearn.grid\_search.GridSearchCV is well suited for this task. I explored linear kernel, rbf kernel with gamma between 1e-6 and 1 and a polynomial kernel of degree between 2 and 9. For all kernels the C parameter was taken from the range between 1e-2 and 1e4. Tuning was performed on a 10\% subset of the data - 6000 training and 1000 testing samples. Only the best performing classifiers were evaluated on the whole dataset.

\section {Results and Discussion}

	\begin{table}[!hbt]
		\begin{center}
		\label{tab:results}
		\begin{tabular}{|c|c|}
			% To create a horizontal line, type \hline
			\hline
			% To end a column type &
			% For a linebreak type \\
			Information message length & $k=16000$ bit \\
			\hline
			Radio segment size & $b=160$ bit \\
			\hline
			Rate of component codes & $R_{cc}=1/3$\\
			\hline
			Polynomial of component encoders & $[1 , 33/37 , 25/37]_8$\\
			\hline
		\end{tabular}
		\end{center}
	\end{table}

% Now we need a bibliography:
\begin{thebibliography}{5}

	\bibitem{MNIST}
	Y.~LeCun, L.~Bottou, Y.~Bengio, and P.~Haffner. ``Gradient-based learning applied to document recognition.'' Proceedings of the IEEE, 86(11):2278-2324, November 1998.

	\bibitem{MJH06} % Conference paper
	T.~Mayer, H.~Jenkac, and J.~Hagenauer. Turbo base-station cooperation for intercell interference cancellation. {\em IEEE Int. Conf. Commun. (ICC)}, Istanbul, Turkey, pp.~356--361, June 2006.

	\bibitem{VSVM}
	D.~Decoste and B.~Schölkopf. ``Training invariant support vector machines.'' Machine learning 46, no. 1-3 (2002): 161-190.

	\bibitem{Proakis} % Book
	J.~G.~Proakis. {\em Digital Communications}. McGraw-Hill Book Co.,
	New York, USA, 3rd edition, 1995.

	\bibitem{MNIST_Website} % Web document
	Y.~LeCun, C.~Cortes, C.J.C.~Burges. The MNIST database of handwritten digits.
	\url{http://yann.lecun.com/exdb/mnist/}.

	\bibitem{UFLDL} % Web document
	A.~Ng, J,~Ngiam, C.Y.~Foo, Y.~Mai and C~Suen. ``Unsupervised Feature Learning and Deep Learning Tutorial.''
	\url{http://ufldl.stanford.edu/wiki/index.php/UFLDL_Tutorial}.

	\bibitem{5}
	IEEE Transactions \LaTeX and Microsoft Word Style Files.
	\url{http://www.ieee.org/web/publications/authors/transjnl/index.html}

\end{thebibliography}

% Your document ends here!
\end{document}

### 2015-04-21 MLCV/ex1/classify.py

@@ -1,0 +1,105 @@
+__author__ = 'Adam Kosiorek'
+
+import mnist
+from sklearn import svm
+from sklearn.grid_search import GridSearchCV
+import numpy as np
+import cv2
+
+MNIST_PATH = '../data/mnist'
+
+def reshapeDataset(dataset):
+    shape = dataset.shape
+    numOfFeatures = shape[1] * shape[2]
+    return dataset.reshape((shape[0], numOfFeatures))
+
+def deskew(img):
+    m = cv2.moments(img)
+    if abs(m['mu02']) < 1e-2:
+        return img.copy()
+    skew = m['mu11']/m['mu02']
+    M = np.float32([[1, skew, -0.5 * img.shape[0] * skew], [0, 1, 0]])
+    img = cv2.warpAffine(img, M, img.shape, flags=cv2.WARP_INVERSE_MAP | cv2.INTER_LINEAR)
+    return img
+
+def deskewSet(dataset):
+    data = [deskew(img) for img in dataset]
+    return np.asarray(data)
+
+def accuracy(predictions, labels):
+    matches = np.count_nonzero(predictions == labels)
+    return float(matches) / labels.shape[0]
+
+def whiten(dataset, filter=None):
+    if filter:
+        filter, avg = filter
+        data = dataset.T
+        data = data - np.tile(avg, [data.shape[1], 1]).T
+        data = filter.dot(data)
+        return data.T
+
+    else:
+        eps = 1e-4
+
+        data = dataset.T
+        avg = np.mean(data, axis=1)
+        data = data - np.tile(avg, [data.shape[1], 1]).T
+        sigma = data.dot(data.T) / data.shape[1]
+        U, S, V = np.linalg.svd(sigma)
+
+        filter = U.dot(np.diag(1 / np.sqrt(S + eps))).dot(U.T)
+
+        data = filter.dot(data)
+        return data.T, (filter, avg)
+
+def subtractMean(dataset, mean=None):
+    returnMean = False
+    if mean is None:
+        mean = np.mean(dataset, axis=0)
+        returnMean = True
+    data = dataset - np.tile(mean, [dataset.shape[0], 1])
+    if returnMean:
+        data = (data, mean)
+    return data
+
+trainSet, trainLabels = mnist.load_mnist('training', path=MNIST_PATH, selection=slice(0, 60000))
+testSet, testLabels = mnist.load_mnist('testing', path=MNIST_PATH, selection=slice(0, 10000))
+
[…]

### 2015-04-21 MLCV/mnist.py

@@ -1,0 +1,127 @@
+import os, struct
+from array import array as pyarray 
+from numpy import append, array, int8, uint8, zeros
+
+def load_mnist(dataset="training", digits=None, path=None, asbytes=False, selection=None, return_labels=True, return_indices=False):
+    """
+    Loads MNIST files into a 3D numpy array.
+
+    You have to download the data separately from [MNIST]_. It is recommended
+    to set the environment variable ``MNIST`` to point to the folder where you
+    put the data, so that you don't have to select path. On a Linux+bash setup,
+    this is done by adding the following to your ``.bashrc``::
+
+        export MNIST=/path/to/mnist
+
+    Parameters
+    ----------
+    dataset : str 
+        Either "training" or "testing", depending on which dataset you want to
+        load. 
+    digits : list 
+        Integer list of digits to load. The entire database is loaded if set to
+        ``None``. Default is ``None``.
+    path : str 
+        Path to your MNIST datafiles. The default is ``None``, which will try
+        to take the path from your environment variable ``MNIST``. The data can
+        be downloaded from http://yann.lecun.com/exdb/mnist/.
+    asbytes : bool
+        If True, returns data as ``numpy.uint8`` in [0, 255] as opposed to
+        ``numpy.float64`` in [0.0, 1.0].
+    selection : slice
+        Using a `slice` object, specify what subset of the dataset to load. An
+        example is ``slice(0, 20, 2)``, which would load every other digit
+        until--but not including--the twentieth.
+    return_labels : bool
+        Specify whether or not labels should be returned. This is also a speed
+        performance if digits are not specified, since then the labels file
+        does not need to be read at all.
+    return_indicies : bool
+        Specify whether or not to return the MNIST indices that were fetched.
+        This is valuable only if digits is specified, because in that case it
+        can be valuable to know how far
+        in the database it reached.
+
+    Returns
+    -------
+    images : ndarray
+        Image data of shape ``(N, rows, cols)``, where ``N`` is the number of images. If neither labels nor inices are returned, then this is returned directly, and not inside a 1-sized tuple.
+    labels : ndarray
+        Array of size ``N`` describing the labels. Returned only if ``return_labels`` is `True`, which is default.
+    indices : ndarray
[…]

### 2015-04-21 parallel_programming/histogram/main.c

@@ -1,0 +1,6 @@
+#include <stdlib.h>
+
+int main(int argc, char** argv) {
+
+	return EXIT_SUCCESS;
+}

### 2015-04-21 parallel_programming/histogram/histogram_seq.c
parallel prog histogram initial implementation

@@ -1,0 +1,19 @@
+#include "histogram.h"
+
+void get_histogram(unsigned int nBlocks,
+		   						 block_t *blocks,
+		   						 unsigned int* histogram,
+		   						 unsigned int num_threads) {
+
+	unsigned int i, j;
+
+	// build histogram
+	for (i=0; i<nBlocks; i++) {
+		for (j=0; j<BLOCKSIZE; j++) {
+			if (blocks[i][j] >= 'a' && blocks[i][j] <= 'z')
+				histogram[blocks[i][j]-'a']++;
+			else if(blocks[i][j] >= 'A' && blocks[i][j] <= 'Z')
+				histogram[blocks[i][j]-'A']++;
+		}
+	}
+}

### 2015-04-21 parallel_programming/histogram/main.c

@@ -1,6 +1,73 @@
+#include <stdio.h>
 #include <stdlib.h>
+#include <sys/stat.h>
+#include <time.h>
+#include <string.h>
 
-int main(int argc, char** argv) {
+#include "histogram.h"
 
-	return EXIT_SUCCESS;
+int main(int argc, char* argv[]) {
+
+	// declarations
+	struct stat st;
+	unsigned int nBlocks, i;
+	block_t *blocks;
+	struct timespec begin, end;
+	histogram_t histogram = {0};
+	char *filename = "war_and_peace.txt";
+	long num_threads = 1, repetitions=0;
+ 
+	// argument handling
+	if (argc < 2 || argc > 4) {
+		printf("usage: %s filename [#threads] [#repetition]\n", argv[0]);
+		return 1;
+	}
+	if (argc >= 2)
+		filename = argv[1];
+	if (argc >= 3)
+		if ((num_threads = strtol(argv[2], NULL, 0)) == 0 || num_threads < 0) {
+    			fprintf(stderr, "#threads not valid!\n");
+    			return 1;
+  		}
+	if (argc >= 4)
+		if ((repetitions = strtol(argv[3], NULL, 0)) == 0 || repetitions < 0) {
+    	fprintf(stderr, "#repetition not valid!\n");
+    	return 1;
+  	}
+
+ 	printf("\nProcess %s by %ld thread(s) with %ld repetition(s)\n\n", 
+		filename, num_threads, repetitions);
+
+	// get file size and allocate array
+	stat(filename, &st);
+	nBlocks = st.st_size / BLOCKSIZE;
+	nBlocks += (st.st_size % BLOCKSIZE) == 0 ? 0 : 1;
+	blocks = (block_t*)calloc(nBlocks, BLOCKSIZE * (repetitions+1));
+
+	// open file and read all blocks
+	FILE *fp = fopen(filename, "r");
+	if (fp == NULL) {
+		fprintf(stderr, "Could not open file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+	if (fread(blocks, BLOCKSIZE, nBlocks, fp) == 0) {
+		fprintf(stderr, "Could not read from file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+
+	// [optional] do the repetition
+	for (i = 1; i < repetitions; i++)
[…]

### 2015-04-21 parallel_programming/histogram/student/histogram_par.c

@@ -1,0 +1,71 @@
+#include <pthread.h>
+#include <stdlib.h>
+
+#include "histogram.h"
+
+#define SIZE 'z' - 'a' + 1
+
+struct Data {
+	int histogram[SIZE];
+	int num_blocks;
+	block_t* blocks;
+};
+
+void* build_histogram(void* arg) {
+
+	struct Data* data = (struct Data*)arg;
+
+	for (int i = 0; i < data->num_blocks; i++) {
+		for (int j = 0; j < BLOCKSIZE; j++) {
+			if (data->blocks[i][j] >= 'a' && data->blocks[i][j] <= 'z') {
+
+				data->histogram[data->blocks[i][j] - 'a']++;
+
+			} else if(data->blocks[i][j] >= 'A' && data->blocks[i][j] <= 'Z') {
+
+				data->histogram[data->blocks[i][j]-'A']++;
+			}	
+		}
+	}
+	return NULL;
+}
+
+void get_histogram(unsigned int nBlocks,
+			block_t *blocks,
+			histogram_t histogram,
+			unsigned int num_threads) {
+
+	pthread_t* thread = malloc(num_threads * sizeof(*thread));
+	struct Data* data = malloc(num_threads * sizeof(*data));
+	int blocksPerThread = nBlocks / num_threads;
+
+	// Spawn N-1 threads
+	--num_threads;
+	for(int i = 0; i < num_threads; i++) {
+		data[i].num_blocks = blocksPerThread;
+		data[i].blocks = blocks + i * blocksPerThread;
+		pthread_create(&thread[i], NULL, build_histogram, &data[i]);
+	}
+
+	// Process the last bit in the main thread
+	data[num_threads].num_blocks = blocksPerThread + nBlocks % num_threads;
+	data[num_threads].blocks = blocks + (num_threads - 2) * blocksPerThread;
+	build_histogram(&data[num_threads]);
+
+
+	// Wait for the workers to complete
+	for(int i = 0; i < num_threads; i++) {
+		pthread_join(thread[i], NULL);
+	}
+
+	// Accumulate the results
+	++num_threads;
+	for(int i = 0; i < num_threads; i++) {
+		for(int j = 0; j < SIZE; j++) {
+			histogram[j] += data[i].histogram[j];
+		}
+	}
+
+	free(thread);
+	free(data);
+}

### 2015-04-21 parallel_programming/histogram/student/histogram_par.c
histogram done, working on binary fan in for histogramming

@@ -1,71 +1,92 @@
 #include <pthread.h>
 #include <stdlib.h>
+#include <stdio.h>
 
 #include "histogram.h"
 
 #define SIZE 'z' - 'a' + 1
 
 struct Data {
-	int histogram[SIZE];
-	int num_blocks;
-	block_t* blocks;
+  int histogram[SIZE];
+  int num_blocks;
+  block_t* blocks;
+  int num_threads;
+  int thread_id;
+  pthread_barrier_t* barrier;
 };
 
 void* build_histogram(void* arg) {
 
-	struct Data* data = (struct Data*)arg;
+  struct Data* data = (struct Data*)arg;
 
-	for (int i = 0; i < data->num_blocks; i++) {
-		for (int j = 0; j < BLOCKSIZE; j++) {
-			if (data->blocks[i][j] >= 'a' && data->blocks[i][j] <= 'z') {
+  for (int i = 0; i < data->num_blocks; i++) {
+    for (int j = 0; j < BLOCKSIZE; j++) {
+      if (data->blocks[i][j] >= 'a' && data->blocks[i][j] <= 'z') {
 
-				data->histogram[data->blocks[i][j] - 'a']++;
+        data->histogram[data->blocks[i][j] - 'a']++;
 
-			} else if(data->blocks[i][j] >= 'A' && data->blocks[i][j] <= 'Z') {
+      } else if(data->blocks[i][j] >= 'A' && data->blocks[i][j] <= 'Z') {
 
-				data->histogram[data->blocks[i][j]-'A']++;
-			}	
-		}
-	}
-	return NULL;
+        data->histogram[data->blocks[i][j]-'A']++;
+      } 
+    }
+  }
+
+  pthread_barrier_wait(data->barrier);
+  int step = 2;
+  int id = data->thread_id;
+  while(thread_id > 0 && id % step == 0) {
+ 
+    step = step << 1;
+  }
+  return NULL;
 }
 
 void get_histogram(unsigned int nBlocks,
-			block_t *blocks,
-			histogram_t histogram,
-			unsigned int num_threads) {
+      block_t *blocks,
+      histogram_t histogram,
+      unsigned int num_threads) {
 
-	pthread_t* thread = malloc(num_threads * sizeof(*thread));
-	struct Data* data = malloc(num_threads * sizeof(*data));
-	int blocksPerThread = nBlocks / num_threads;
+ printf("#threads = %d\n", num_threads);
 
[…]

### 2015-04-23 parallel_programming/histogram/histogram_seq.c
working histogram

@@ -1,4 +1,5 @@
 #include "histogram.h"
+#include <stdio.h>
 
 void get_histogram(unsigned int nBlocks,
 		   						 block_t *blocks,
@@ -16,4 +17,8 @@ void get_histogram(unsigned int nBlocks,
 				histogram[blocks[i][j]-'A']++;
 		}
 	}
+
+//  for(i = 0; i < 'z' - 'a' + 1; i++) {
+//    printf("%c:\t%u\n", (char)('a'+i), histogram[i]);
+//  }
 }

### 2015-04-23 parallel_programming/histogram/student/histogram_par.c

@@ -7,62 +7,49 @@
 #define SIZE 'z' - 'a' + 1
 
 struct Data {
-  int histogram[SIZE];
-  int num_blocks;
+  unsigned int histogram[SIZE];
+  unsigned int num_blocks;
   block_t* blocks;
-  int num_threads;
-  int thread_id;
-  pthread_barrier_t* barrier;
 };
+
 
 void* build_histogram(void* arg) {
 
   struct Data* data = (struct Data*)arg;
 
+  for(int i = 0; i < SIZE; ++i) {
+    data->histogram[i] = 0;
+  }
+
   for (int i = 0; i < data->num_blocks; i++) {
     for (int j = 0; j < BLOCKSIZE; j++) {
       if (data->blocks[i][j] >= 'a' && data->blocks[i][j] <= 'z') {
 
-        data->histogram[data->blocks[i][j] - 'a']++;
+        data->histogram[data->blocks[i][j] - 'a'] += 1;
 
       } else if(data->blocks[i][j] >= 'A' && data->blocks[i][j] <= 'Z') {
 
-        data->histogram[data->blocks[i][j]-'A']++;
+        data->histogram[data->blocks[i][j]-'A'] += 1;
       } 
     }
   }
-
-  pthread_barrier_wait(data->barrier);
-  int step = 2;
-  int id = data->thread_id;
-  while(thread_id > 0 && id % step == 0) {
- 
-    step = step << 1;
-  }
   return NULL;
 }
+
 
 void get_histogram(unsigned int nBlocks,
       block_t *blocks,
       histogram_t histogram,
       unsigned int num_threads) {
 
- printf("#threads = %d\n", num_threads);
-
-  pthread_barrier_t barrier;
   pthread_t* thread = malloc(num_threads * sizeof(*thread));
   struct Data* data = malloc(num_threads * sizeof(*data));
-  int blocksPerThread = nBlocks / num_threads;
-  int remainingBlocks = nBlocks % num_threads;
-
-  pthread_barrier_init(&barrier, num_threads);
+  unsigned int blocksPerThread = nBlocks / num_threads;
+  unsigned int remainingBlocks = nBlocks % num_threads;
 
   // Spawn N threads to process the data
   for(int i = 0; i < num_threads; i++) {
     data[i].num_blocks = blocksPerThread;
-    data[i].num_threads = num_threads;
-    data[i].thread_it = i;
[…]

### 2015-04-23 MLCV/ex1/__init__.py
finish report for MLCV ex1

@@ -1,0 +1,1 @@
+__author__ = 'Adam Kosiorek'

### 2015-04-23 MLCV/ex1/classify.py

@@ -1,105 +1,0 @@
-__author__ = 'Adam Kosiorek'
-
-import mnist
-from sklearn import svm
-from sklearn.grid_search import GridSearchCV
-import numpy as np
-import cv2
-
-MNIST_PATH = '../data/mnist'
-
-def reshapeDataset(dataset):
-    shape = dataset.shape
-    numOfFeatures = shape[1] * shape[2]
-    return dataset.reshape((shape[0], numOfFeatures))
-
-def deskew(img):
-    m = cv2.moments(img)
-    if abs(m['mu02']) < 1e-2:
-        return img.copy()
-    skew = m['mu11']/m['mu02']
-    M = np.float32([[1, skew, -0.5 * img.shape[0] * skew], [0, 1, 0]])
-    img = cv2.warpAffine(img, M, img.shape, flags=cv2.WARP_INVERSE_MAP | cv2.INTER_LINEAR)
-    return img
-
-def deskewSet(dataset):
-    data = [deskew(img) for img in dataset]
-    return np.asarray(data)
-
-def accuracy(predictions, labels):
-    matches = np.count_nonzero(predictions == labels)
-    return float(matches) / labels.shape[0]
-
-def whiten(dataset, filter=None):
-    if filter:
-        filter, avg = filter
-        data = dataset.T
-        data = data - np.tile(avg, [data.shape[1], 1]).T
-        data = filter.dot(data)
-        return data.T
-
-    else:
-        eps = 1e-4
-
-        data = dataset.T
-        avg = np.mean(data, axis=1)
-        data = data - np.tile(avg, [data.shape[1], 1]).T
-        sigma = data.dot(data.T) / data.shape[1]
-        U, S, V = np.linalg.svd(sigma)
-
-        filter = U.dot(np.diag(1 / np.sqrt(S + eps))).dot(U.T)
-
-        data = filter.dot(data)
-        return data.T, (filter, avg)
-
-def subtractMean(dataset, mean=None):
-    returnMean = False
-    if mean is None:
-        mean = np.mean(dataset, axis=0)
-        returnMean = True
-    data = dataset - np.tile(mean, [dataset.shape[0], 1])
-    if returnMean:
-        data = (data, mean)
-    return data
-
-trainSet, trainLabels = mnist.load_mnist('training', path=MNIST_PATH, selection=slice(0, 60000))
-testSet, testLabels = mnist.load_mnist('testing', path=MNIST_PATH, selection=slice(0, 10000))
-
[…]

### 2015-04-23 MLCV/ex1/data.py

@@ -1,0 +1,105 @@
+__author__ = 'Adam Kosiorek'
+
+import mnist
+import cv2
+import numpy as np
+import pydot
+import matplotlib.pyplot as plt
+from StringIO import StringIO
+from sklearn import tree
+
+# MNIST dataset must be unpacked in this folder !
+MNIST_PATH = '../data/mnist'
+
+def reshapeDataset(dataset):
+    shape = dataset.shape
+    numOfFeatures = shape[1] * shape[2]
+    return dataset.reshape((shape[0], numOfFeatures))
+
+# Deskew function taken from:
+# http://opencv-python-tutroals.readthedocs.org/en/latest/py_tutorials/py_ml/py_svm/py_svm_opencv/py_svm_opencv.html
+def deskew(img):
+    m = cv2.moments(img)
+    if abs(m['mu02']) < 1e-2:
+        return img.copy()
+    skew = m['mu11']/m['mu02']
+    M = np.float32([[1, skew, -0.5 * img.shape[0] * skew], [0, 1, 0]])
+    img = cv2.warpAffine(img, M, img.shape, flags=cv2.WARP_INVERSE_MAP | cv2.INTER_LINEAR)
+    return img
+
+def deskewSet(dataset):
+    data = [deskew(img) for img in dataset]
+    return np.asarray(data)
+
+# ZCA-Whitening according to:
+# http://ufldl.stanford.edu/wiki/index.php/UFLDL_Tutorial
+def doWhiten(dataset, filter=None):
+    if filter:
+        filter, avg = filter
+        data = dataset.T
+        data = data - np.tile(avg, [data.shape[1], 1]).T
+        data = filter.dot(data)
+        return data.T
+
+    else:
+        eps = 1e-4
+        data = dataset.T
+        avg = np.mean(data, axis=1)
+        data = data - np.tile(avg, [data.shape[1], 1]).T
+        sigma = data.dot(data.T) / data.shape[1]
+        U, S, V = np.linalg.svd(sigma)
+
+        filter = U.dot(np.diag(1 / np.sqrt(S + eps))).dot(U.T)
+
+        data = filter.dot(data)
+        return data.T, (filter, avg)
+
+def doSubtractMean(dataset, mean=None):
+    returnMean = False
+    if mean is None:
+        mean = np.mean(dataset, axis=0)
+        returnMean = True
+    data = dataset - np.tile(mean, [dataset.shape[0], 1])
+    if returnMean:
+        data = (data, mean)
+    return data
+
[…]

### 2015-04-23 MLCV/ex1/random_forests.py

@@ -1,0 +1,34 @@
+from sklearn.ensemble import RandomForestClassifier
+from sklearn.grid_search import GridSearchCV
+from sklearn.metrics import accuracy_score
+from ex1.data import getData, plot_feature_importance, plot_tree
+
+
+trainData, testData = getData(fraction=1, deskew=True, subtractMean=True)
+trainSet, trainLabels = trainData
+testSet, testLabels = testData
+
+
+# # Parameter Grid Search with Cross-Validation
+# tuned_parameters = {
+#                      'n_estimators' : [2, 5, 10, 20, 30, 50, 70, 100],
+#                      'criterion' : ['gini', 'entropy'],
+#                      'max_depth' : [3, 5, 10, 30, 50, 70, 100],
+#                      'max_features' : [3, 5, 50, 70, 100, 350, 500, 650, 784]
+#                     }
+#
+# classifier = GridSearchCV(RandomForestClassifier(), tuned_parameters, cv=5, n_jobs=-1)
+# print classifier.fit(trainSet, trainLabels)
+# print classifier.best_params_
+# predictions = classifier.predict(testSet)
+# print accuracy_score(predictions, testLabels)
+
+# #Best Parameters from CrossValidation
+best_params = {'max_features': 50, 'n_estimators': 100, 'criterion': 'entropy', 'max_depth': 10}
+classifier = RandomForestClassifier(**best_params)
+print classifier.fit(trainSet, trainLabels)
+
+predictions = classifier.predict(testSet)
+print accuracy_score(predictions, testLabels)
+# plot_tree(classifier, 'random')
+plot_feature_importance(classifier, 'random')

### 2015-04-23 MLCV/ex1/svm.py

@@ -1,0 +1,35 @@
+__author__ = 'Adam Kosiorek'
+
+from sklearn import svm
+from sklearn.grid_search import GridSearchCV
+from sklearn.metrics import accuracy_score
+from ex1.data import getData
+from sklearn import tree
+
+trainData, testData = getData(deskew=True, subtractMean=True)
+trainSet, trainLabels = trainData
+testSet, testLabels = testData
+
+print 'fitting SVM'
+## Parameter Grid Search with Cross-Validation
+# c_range= [1e-2, 0.1, 1, 10, 100, 1000]
+# tuned_parameters = [{'kernel': ['rbf'],
+#                         'gamma': [1e-3, 1e-4],
+#                         'C': c_range},
+#                     {'kernel': ['linear'],
+#                         'C': c_range},
+#                     {'kernel':['poly'],
+#                         'C' : c_range,
+#                         'degree' : [2, 3, 5, 7, 9]}
+#                     ]
+#
+# classifier = GridSearchCV(svm.SVC(cache_size=2000), tuned_parameters, cv=5, n_jobs=-1)
+# print classifier.fit(trainSet, trainLabels)
+# print classifier.best_params_
+
+classifier = svm.SVC(kernel='poly', degree=2, C=1000)
+print classifier.fit(trainSet, trainLabels)
+
+
+predictions = classifier.predict(testSet)
+print accuracy_score(predictions, testLabels)

### 2015-04-23 MLCV/ex1/trees.py

@@ -1,0 +1,37 @@
+__author__ = 'Adam Kosiorek'
+
+
+
+from sklearn import tree
+from sklearn.grid_search import GridSearchCV
+from sklearn.metrics import accuracy_score
+from ex1.data import getData, plot_feature_importance, plot_tree
+
+
+trainData, testData = getData(fraction=1, deskew=True, subtractMean=True)
+trainSet, trainLabels = trainData
+testSet, testLabels = testData
+
+
+# Parameter Grid Search with Cross-Validation
+# tuned_parameters = {
+#                      'criterion' : ['gini', 'entropy'],
+#                      'max_depth' : [3, 5, 10, 30, 50, 70, 100],
+#                      'max_features' : [3, 5, 50, 70, 100, 350, 500, 650, 784]
+#                     }
+#
+# classifier = GridSearchCV(tree.DecisionTreeClassifier(), tuned_parameters, cv=5, n_jobs=-1)
+# print classifier.fit(trainSet, trainLabels)
+# print classifier.best_params_
+# predictions = classifier.predict(testSet)
+# print accuracy_score(predictions, testLabels)
+
+# #Best Parameters from CrossValidation
+best_params = {'max_features': 500, 'criterion': 'entropy', 'max_depth': 100}
+classifier = tree.DecisionTreeClassifier(**best_params)
+print classifier.fit(trainSet, trainLabels)
+
+predictions = classifier.predict(testSet)
+print accuracy_score(predictions, testLabels)
+plot_tree(classifier, 'tree')
+plot_feature_importance(classifier, 'tree')

### 2015-04-30 parallel_programming/ex1/histogram.c
add parallel prog ex2

@@ -1,0 +1,45 @@
+#include <stdio.h>
+
+#include "histogram.h"
+
+#define MAX(X,Y) ((X) > (Y) ? (X) : (Y))
+#define NLINES 20
+
+// get the maximum number of characters per letter
+int max_chars(histogram_t histogram) {
+	int i, max = 0;
+
+	for (i=0; i<NALPHABET; ++i)
+		max = MAX(histogram[i], max);
+
+	return max;
+}
+
+void print_histogram(histogram_t histogram) {
+
+	int max = max_chars(histogram);
+	int i, j, y;
+
+	// print line by line
+	for (i=NLINES; i>=0; i--) {
+
+		// calculate and print y labels
+		y = (i*max) / NLINES;
+		printf("%10d:", y);
+
+		// print histogram
+		for (j=0; j<NALPHABET; j++) {
+			if (histogram[j] >= y)
+				printf("%2c|", 0);
+			else
+				printf("%3c", 0);
+		}
+		printf("\n");
+	}
+
+	// print x labels
+	printf("%12c", 0);
+	for (j=0; j<NALPHABET; j++) {
+		printf("%2c%c", 0, 97+j);
+	}
+}

### 2015-04-30 parallel_programming/ex1/histogram.h

@@ -1,0 +1,20 @@
+#ifndef HISTOGRAM_H_
+#define HISTOGRAM_H_
+
+#define BLOCKSIZE 8192 	// block size in byte
+#define BYTESIZE 8 		// bits per byte
+#define NALPHABET 26	// number of letters in the alphabet
+
+typedef char block_t[BLOCKSIZE];
+typedef unsigned int histogram_t[NALPHABET];
+
+// produce histogram from blocks to histogram
+void get_histogram(unsigned int nBlocks,
+				   				 block_t *blocks,
+				   				 unsigned int* histogram,
+				   				 unsigned int num_threads);
+
+// dump histogram to stdout
+void print_histogram(histogram_t histogram);
+
+#endif /* HISTOGRAM_H_ */

### 2015-04-30 parallel_programming/ex1/histogram_seq.c

@@ -1,0 +1,22 @@
+#include "histogram.h"
+#include <stdio.h>
+
+void get_histogram(unsigned int nBlocks,
+		   						 block_t *blocks,
+		   						 unsigned int* histogram,
+		   						 unsigned int num_threads) {
+
+	unsigned int i, j;
+
+	// build histogram
+	for (i=0; i<nBlocks; i++) {
+		for (j=0; j<BLOCKSIZE; j++) {
+			if (blocks[i][j] >= 'a' && blocks[i][j] <= 'z')
+				histogram[blocks[i][j]-'a']++;
+			else if(blocks[i][j] >= 'A' && blocks[i][j] <= 'Z')
+				histogram[blocks[i][j]-'A']++;
+		}
+	}
+
+  printf("Number of blocks: %u\n", nBlocks);
+}

### 2015-04-30 parallel_programming/ex1/main.c

@@ -1,0 +1,73 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <sys/stat.h>
+#include <time.h>
+#include <string.h>
+
+#include "histogram.h"
+
+int main(int argc, char* argv[]) {
+
+	// declarations
+	struct stat st;
+	unsigned int nBlocks, i;
+	block_t *blocks;
+	struct timespec begin, end;
+	histogram_t histogram = {0};
+	char *filename = "war_and_peace.txt";
+	long num_threads = 1, repetitions=0;
+ 
+	// argument handling
+	if (argc < 2 || argc > 4) {
+		printf("usage: %s filename [#threads] [#repetition]\n", argv[0]);
+		return 1;
+	}
+	if (argc >= 2)
+		filename = argv[1];
+	if (argc >= 3)
+		if ((num_threads = strtol(argv[2], NULL, 0)) == 0 || num_threads < 0) {
+    			fprintf(stderr, "#threads not valid!\n");
+    			return 1;
+  		}
+	if (argc >= 4)
+		if ((repetitions = strtol(argv[3], NULL, 0)) == 0 || repetitions < 0) {
+    	fprintf(stderr, "#repetition not valid!\n");
+    	return 1;
+  	}
+
+ 	printf("\nProcess %s by %ld thread(s) with %ld repetition(s)\n\n", 
+		filename, num_threads, repetitions);
+
+	// get file size and allocate array
+	stat(filename, &st);
+	nBlocks = st.st_size / BLOCKSIZE;
+	nBlocks += (st.st_size % BLOCKSIZE) == 0 ? 0 : 1;
+	blocks = (block_t*)calloc(nBlocks, BLOCKSIZE * (repetitions+1));
+
+	// open file and read all blocks
+	FILE *fp = fopen(filename, "r");
+	if (fp == NULL) {
+		fprintf(stderr, "Could not open file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+	if (fread(blocks, BLOCKSIZE, nBlocks, fp) == 0) {
+		fprintf(stderr, "Could not read from file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+
+	// [optional] do the repetition
+	for (i = 1; i < repetitions; i++)
+		memcpy(blocks + i*nBlocks, blocks, nBlocks*BLOCKSIZE);
+
+	// build the histogram	
[…]

### 2015-04-30 parallel_programming/ex1/student/histogram_par.c

@@ -1,0 +1,119 @@
+#include <pthread.h>
+#include <stdlib.h>
+#include <stdio.h>
+
+#include "histogram.h"
+
+#define SIZE 'z' - 'a' + 1
+
+struct Data {
+  unsigned int histogram[SIZE];
+  unsigned int num_blocks;
+  block_t* blocks;
+  int thread_id;
+  int num_threads;
+  struct Data* data;
+  pthread_barrier_t* barrier;
+};
+
+
+void* build_histogram(void* arg) {
+
+  struct Data* data = (struct Data*)arg;
+
+  for(int i = 0; i < SIZE; ++i) {
+    data->histogram[i] = 0;
+  }
+
+  for (int i = 0; i < data->num_blocks; i++) {
+    for (int j = 0; j < BLOCKSIZE; j++) {
+      if (data->blocks[i][j] >= 'a' && data->blocks[i][j] <= 'z') {
+
+        data->histogram[data->blocks[i][j] - 'a'] += 1;
+
+      } else if(data->blocks[i][j] >= 'A' && data->blocks[i][j] <= 'Z') {
+
+        data->histogram[data->blocks[i][j]-'A'] += 1;
+      } 
+    }
+  }
+
+
+  if(data->num_threads > 1) {
+    unsigned int step = 2;
+
+    while(data->thread_id % step == 0 && step <= data->num_threads) {
+      printf("Id #%u at step = %u\n", data->thread_id, step);
+      if(pthread_barrier_wait(data->barrier) == PTHREAD_BARRIER_SERIAL_THREAD) {
+        printf("Id %u init at step = %u\n", data->thread_id, step);
+        pthread_barrier_init(data->barrier, NULL, data->num_threads / step / 2);
+      }
+
+      unsigned int index = data->thread_id + step / 2;
+      if(index < data->num_threads) {
+        for(int i = 0; i < SIZE; ++i) {
+          data->histogram[i] += data->data[index].histogram[i];
+        }
+      }
+
+      step *= 2;
+    }
+  }
+  return NULL;
+}
+
+
+void get_histogram(unsigned int nBlocks,
+      block_t *blocks,
+      histogram_t histogram,
+      unsigned int num_threads) {
+
+  pthread_t* thread = malloc(num_threads * sizeof(*thread));
+  struct Data* data = calloc(num_threads, sizeof(*data));
+  unsigned int blocksPerThread = nBlocks / num_threads;
[…]

### 2015-04-30 parallel_programming/ex2/histogram.h

@@ -1,0 +1,20 @@
+#ifndef HISTOGRAM_H_
+#define HISTOGRAM_H_
+
+#define BYTESIZE 8 		// bits per byte
+#define NALPHABET 26	// number of letters in the alphabet
+#define CHUNKSIZE 8192	// number of chunks for a task
+#define TERMINATOR (char)255
+
+typedef unsigned int histogram_t[NALPHABET];
+
+// produce histogram from blocks to histogram
+void get_histogram(char *input,
+				   	 unsigned int* histogram,
+				   	 unsigned int num_threads,
+						 unsigned int chunk_size);
+
+// dump histogram to stdout
+void print_histogram(histogram_t histogram);
+
+#endif /* HISTOGRAM_H_ */

### 2015-04-30 parallel_programming/ex2/histogram_ref.c

@@ -1,0 +1,13 @@
+#include "histogram.h"
+
+void get_histogram_ref(char* buffer,
+				   		  unsigned int *histogram) {
+
+   // build histogram
+   for (int i=0; buffer[i]!=TERMINATOR; i++) {
+      if (buffer[i] >= 'a' && buffer[i] <= 'z')
+         histogram[buffer[i]-'a']++;
+      else if(buffer[i] >= 'A' && buffer[i] <= 'Z')
+         histogram[buffer[i]-'A']++;
+   }
+}

### 2015-04-30 parallel_programming/ex2/histogram_ref.h

@@ -1,0 +1,7 @@
+#ifndef HISTOGRAM_REF_H_
+#define HISTOGRAM_REF_H_
+
+void get_histogram_ref(char* buffer,
+				   		  unsigned int *histogram);
+
+#endif /* HISTOGRAM_REF_H_ */

### 2015-04-30 parallel_programming/ex2/histogram_seq.c

@@ -1,0 +1,16 @@
+#include "histogram.h"
+
+void get_histogram(char *buffer,
+		   			 unsigned int* histogram,
+		   			 unsigned int num_threads,
+						 unsigned int chunk_size) {
+
+
+	// build histogram
+	for (int i=0; buffer[i]!=TERMINATOR; i++) {
+		if (buffer[i] >= 'a' && buffer[i] <= 'z')
+			histogram[buffer[i]-'a']++;
+		else if(buffer[i] >= 'A' && buffer[i] <= 'Z')
+			histogram[buffer[i]-'A']++;
+	}
+}

### 2015-04-30 parallel_programming/ex2/main.c

@@ -1,0 +1,73 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <sys/stat.h>
+#include <time.h>
+#include <string.h>
+
+#include "histogram.h"
+
+int main(int argc, char* argv[]) {
+
+	// declarations
+	struct stat st;
+	unsigned int i, buffer_size;
+	char *buffer;
+	struct timespec begin, end;
+	histogram_t histogram = {0};
+	char *filename = "war_and_peace.txt";
+	long num_threads = 1, repetitions=0;
+ 
+	// argument handling
+	if (argc < 2 || argc > 4) {
+		printf("usage: %s filename [#threads] [#repetition]\n", argv[0]);
+		return 1;
+	}
+	if (argc >= 2)
+		filename = argv[1];
+	if (argc >= 3)
+		if ((num_threads = strtol(argv[2], NULL, 0)) == 0 || num_threads < 0) {
+    			fprintf(stderr, "#threads not valid!\n");
+    			return 1;
+  		}
+	if (argc >= 4)
+		if ((repetitions = strtol(argv[3], NULL, 0)) == 0 || repetitions < 0) {
+    	fprintf(stderr, "#repetition not valid!\n");
+    	return 1;
+  	}
+
+ 	printf("\nProcess %s by %ld thread(s) with %ld repetition(s)\n\n", 
+		filename, num_threads, repetitions);
+
+	// get file size and allocate array
+	stat(filename, &st);
+	buffer_size = (unsigned int)st.st_size * (repetitions+1) + CHUNKSIZE;
+	buffer = (char*)malloc(buffer_size);
+	memset(buffer, TERMINATOR, buffer_size);
+
+	// open file and read all characters
+	FILE *fp = fopen(filename, "r");
+	if (fp == NULL) {
+		fprintf(stderr, "Could not open file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+	if (fread(buffer, 1, st.st_size, fp) == 0) {
+		fprintf(stderr, "Could not read from file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+
+	// [optional] do the repetition
+	for (i = 1; i < repetitions; i++)
+		memcpy(buffer + i*st.st_size, buffer, st.st_size);
+
+	// build the histogram	
+	clock_gettime(CLOCK_MONOTONIC, &begin);
[…]

### 2015-04-30 parallel_programming/ex2/student/histogram_par.c

@@ -1,0 +1,100 @@
+#include <pthread.h>
+#include <stdlib.h>
+#include <stdio.h>
+
+#include "histogram.h"
+
+
+struct Data {
+
+  pthread_mutex_t* mutex;
+  char** globalBuffer;
+  unsigned int chunkSize;
+  unsigned int* histogram;
+};
+
+
+void buildHistogram(const char *buffer,
+             int size,
+		   			 unsigned int* histogram) {
+
+	// build histogram
+	for (int i=0; i < size; i++) {
+		if (buffer[i] >= 'a' && buffer[i] <= 'z')
+			histogram[buffer[i]-'a']++;
+		else if(buffer[i] >= 'A' && buffer[i] <= 'Z')
+			histogram[buffer[i]-'A']++;
+    else if(buffer[i] == TERMINATOR)
+      break;
+  } 
+}
+
+char* getDataChunk(struct Data* data) {
+
+  pthread_mutex_lock(data->mutex);
+
+  char* dataChunkPtr = *data->globalBuffer;
+  if(*dataChunkPtr == TERMINATOR) {
+    dataChunkPtr = NULL;
+  } else {
+    *data->globalBuffer += data->chunkSize;
+  }
+  pthread_mutex_unlock(data->mutex);
+
+  //printf("%p\n", dataChunkPtr);
+  return dataChunkPtr;
+}
+
+void* histogramThreadFun(void* ptr) {
+
+  struct Data* data = (struct Data*)ptr;
+  unsigned int size = data->chunkSize;
+  unsigned int* histogram = calloc(NALPHABET, sizeof(*histogram));
+  const char* dataChunk = getDataChunk(data);
+  while(dataChunk) {
+    buildHistogram(dataChunk, size, histogram);
+    dataChunk = getDataChunk(data);
+  }
+
+  data->histogram = histogram;
+  return NULL;
+}
+
+
+
+void get_histogram(char *buffer,
+		   			 unsigned int* histogram,
+		   			 unsigned int num_threads,
+						 unsigned int chunk_size) {
+
+
+  pthread_mutex_t mutex = PTHREAD_MUTEX_INITIALIZER;
+  pthread_t* thread = malloc(num_threads * sizeof(*thread));
+  struct Data* data = malloc(num_threads * sizeof(*data));
+
+
+  for(int i = 0; i < num_threads; ++i) {
+    data[i].mutex = &mutex;
[…]

### 2015-04-30 parallel_programming/ex2/unit_test.c

@@ -1,0 +1,56 @@
+#include <stdio.h>
+#include <string.h>
+#include <stdlib.h>
+#include <sys/stat.h>
+
+#include "histogram.h"
+#include "histogram_ref.h"
+
+int main(int argc, char **argv)
+{
+	int failed = 0;
+	histogram_t histogram_ref = {0};
+	struct stat st;
+	char *buffer;
+	unsigned int buffer_size;
+	char *filename = "war_and_peace.txt";
+
+   // get file size and allocate array
+   stat(filename, &st);
+   buffer_size = (unsigned int)st.st_size + CHUNKSIZE;
+   buffer = (char*)malloc(buffer_size);
+   memset(buffer, (char)255, buffer_size);
+
+	// open file and read all blocks
+	FILE *fp = fopen(filename, "r");
+	if (fp == NULL) {
+		fprintf(stderr, "Could not open file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+	if (fread(buffer, 1, st.st_size, fp) == 0) {
+		fprintf(stderr, "Could not read from file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+
+	// build the histograms	
+	get_histogram_ref(buffer, histogram_ref);
+
+	for (int nThreads = 1; nThreads <= 8; nThreads++) {
+		histogram_t histogram = {0};		
+		get_histogram(buffer, histogram, nThreads, CHUNKSIZE);
+		for (int i = 0; i < NALPHABET; i++) {
+			if (histogram_ref[i] != histogram[i]) {
+				fprintf(stderr, "Computation with %d threads failed:\n", nThreads);
+				fprintf(stderr, "Wrong result for %c: %d (correct: %d)\n",
+						i+97, histogram[i], histogram_ref[i]);
+				failed = 1;
+				break;
+			}
+		}
+		if (failed)
+			break;
+	}
+	/* end test */
+
+	return failed;
+}

### 2015-05-01 parallel_programming/ex2/student/histogram_par.c
clean parallel_programming

@@ -4,36 +4,35 @@
 
 #include "histogram.h"
 
+#pragma GCC optimize (3)
 
 struct Data {
 
-  pthread_mutex_t* mutex;
+  pthread_mutex_t* restrict mutex;
   char** globalBuffer;
   unsigned int chunkSize;
-  unsigned int* histogram;
+  unsigned int* restrict histogram;
+  char pad[64 - sizeof(int) - 3 * sizeof(int*)];
 };
 
 
-void buildHistogram(const char *buffer,
+void buildHistogram(const char* restrict buffer,
              int size,
-		   			 unsigned int* histogram) {
+		   			 unsigned int* restrict histogram) {
 
 	// build histogram
 	for (int i=0; i < size; i++) {
-		if (buffer[i] >= 'a' && buffer[i] <= 'z')
-			histogram[buffer[i]-'a']++;
-		else if(buffer[i] >= 'A' && buffer[i] <= 'Z')
-			histogram[buffer[i]-'A']++;
-    else if(buffer[i] == TERMINATOR)
-      break;
-  } 
+    if ((buffer[i]|32) >= 'a' && (buffer[i]|32) <= 'z')
+      histogram[(buffer[i]|32)-'a']++;
+    }
 }
 
-char* getDataChunk(struct Data* data) {
+
+char* getDataChunk(struct Data* restrict data) {
 
   pthread_mutex_lock(data->mutex);
 
-  char* dataChunkPtr = *data->globalBuffer;
+  char* restrict dataChunkPtr = *data->globalBuffer;
   if(*dataChunkPtr == TERMINATOR) {
     dataChunkPtr = NULL;
   } else {
@@ -41,16 +40,15 @@ char* getDataChunk(struct Data* data) {
   }
   pthread_mutex_unlock(data->mutex);
 
-  //printf("%p\n", dataChunkPtr);
   return dataChunkPtr;
 }
 
-void* histogramThreadFun(void* ptr) {
+void* histogramThreadFun(void* restrict ptr) {
 
-  struct Data* data = (struct Data*)ptr;
+  struct Data* restrict data = (struct Data*)ptr;
   unsigned int size = data->chunkSize;
   unsigned int* histogram = calloc(NALPHABET, sizeof(*histogram));
-  const char* dataChunk = getDataChunk(data);
+  const char* restrict dataChunk = getDataChunk(data);
   while(dataChunk) {
     buildHistogram(dataChunk, size, histogram);
[…]

### 2015-05-01 parallel_programming/histogram/histogram.c

@@ -1,45 +1,0 @@
-#include <stdio.h>
-
-#include "histogram.h"
-
-#define MAX(X,Y) ((X) > (Y) ? (X) : (Y))
-#define NLINES 20
-
-// get the maximum number of characters per letter
-int max_chars(histogram_t histogram) {
-	int i, max = 0;
-
-	for (i=0; i<NALPHABET; ++i)
-		max = MAX(histogram[i], max);
-
-	return max;
-}
-
-void print_histogram(histogram_t histogram) {
-
-	int max = max_chars(histogram);
-	int i, j, y;
-
-	// print line by line
-	for (i=NLINES; i>=0; i--) {
-
-		// calculate and print y labels
-		y = (i*max) / NLINES;
-		printf("%10d:", y);
-
-		// print histogram
-		for (j=0; j<NALPHABET; j++) {
-			if (histogram[j] >= y)
-				printf("%2c|", 0);
-			else
-				printf("%3c", 0);
-		}
-		printf("\n");
-	}
-
-	// print x labels
-	printf("%12c", 0);
-	for (j=0; j<NALPHABET; j++) {
-		printf("%2c%c", 0, 97+j);
-	}
-}

### 2015-05-01 parallel_programming/histogram/histogram.h

@@ -1,20 +1,0 @@
-#ifndef HISTOGRAM_H_
-#define HISTOGRAM_H_
-
-#define BLOCKSIZE 8192 	// block size in byte
-#define BYTESIZE 8 		// bits per byte
-#define NALPHABET 26	// number of letters in the alphabet
-
-typedef char block_t[BLOCKSIZE];
-typedef unsigned int histogram_t[NALPHABET];
-
-// produce histogram from blocks to histogram
-void get_histogram(unsigned int nBlocks,
-				   				 block_t *blocks,
-				   				 unsigned int* histogram,
-				   				 unsigned int num_threads);
-
-// dump histogram to stdout
-void print_histogram(histogram_t histogram);
-
-#endif /* HISTOGRAM_H_ */

### 2015-05-01 parallel_programming/histogram/histogram_seq.c

@@ -1,24 +1,0 @@
-#include "histogram.h"
-#include <stdio.h>
-
-void get_histogram(unsigned int nBlocks,
-		   						 block_t *blocks,
-		   						 unsigned int* histogram,
-		   						 unsigned int num_threads) {
-
-	unsigned int i, j;
-
-	// build histogram
-	for (i=0; i<nBlocks; i++) {
-		for (j=0; j<BLOCKSIZE; j++) {
-			if (blocks[i][j] >= 'a' && blocks[i][j] <= 'z')
-				histogram[blocks[i][j]-'a']++;
-			else if(blocks[i][j] >= 'A' && blocks[i][j] <= 'Z')
-				histogram[blocks[i][j]-'A']++;
-		}
-	}
-
-//  for(i = 0; i < 'z' - 'a' + 1; i++) {
-//    printf("%c:\t%u\n", (char)('a'+i), histogram[i]);
-//  }
-}

### 2015-05-01 parallel_programming/histogram/main.c

@@ -1,73 +1,0 @@
-#include <stdio.h>
-#include <stdlib.h>
-#include <sys/stat.h>
-#include <time.h>
-#include <string.h>
-
-#include "histogram.h"
-
-int main(int argc, char* argv[]) {
-
-	// declarations
-	struct stat st;
-	unsigned int nBlocks, i;
-	block_t *blocks;
-	struct timespec begin, end;
-	histogram_t histogram = {0};
-	char *filename = "war_and_peace.txt";
-	long num_threads = 1, repetitions=0;
- 
-	// argument handling
-	if (argc < 2 || argc > 4) {
-		printf("usage: %s filename [#threads] [#repetition]\n", argv[0]);
-		return 1;
-	}
-	if (argc >= 2)
-		filename = argv[1];
-	if (argc >= 3)
-		if ((num_threads = strtol(argv[2], NULL, 0)) == 0 || num_threads < 0) {
-    			fprintf(stderr, "#threads not valid!\n");
-    			return 1;
-  		}
-	if (argc >= 4)
-		if ((repetitions = strtol(argv[3], NULL, 0)) == 0 || repetitions < 0) {
-    	fprintf(stderr, "#repetition not valid!\n");
-    	return 1;
-  	}
-
- 	printf("\nProcess %s by %ld thread(s) with %ld repetition(s)\n\n", 
-		filename, num_threads, repetitions);
-
-	// get file size and allocate array
-	stat(filename, &st);
-	nBlocks = st.st_size / BLOCKSIZE;
-	nBlocks += (st.st_size % BLOCKSIZE) == 0 ? 0 : 1;
-	blocks = (block_t*)calloc(nBlocks, BLOCKSIZE * (repetitions+1));
-
-	// open file and read all blocks
-	FILE *fp = fopen(filename, "r");
-	if (fp == NULL) {
-		fprintf(stderr, "Could not open file %s", filename);
-		exit(EXIT_FAILURE);
-	}
-	if (fread(blocks, BLOCKSIZE, nBlocks, fp) == 0) {
-		fprintf(stderr, "Could not read from file %s", filename);
-		exit(EXIT_FAILURE);
-	}
-
-	// [optional] do the repetition
-	for (i = 1; i < repetitions; i++)
-		memcpy(blocks + i*nBlocks, blocks, nBlocks*BLOCKSIZE);
-
-	// build the histogram	
[…]

### 2015-05-01 parallel_programming/histogram/student/histogram_par.c

@@ -1,89 +1,0 @@
-#include <pthread.h>
-#include <stdlib.h>
-#include <stdio.h>
-
-#include "histogram.h"
-
-#define SIZE 'z' - 'a' + 1
-
-struct Data {
-  unsigned int histogram[SIZE];
-  unsigned int num_blocks;
-  block_t* blocks;
-};
-
-
-void* build_histogram(void* arg) {
-
-  struct Data* data = (struct Data*)arg;
-
-  for(int i = 0; i < SIZE; ++i) {
-    data->histogram[i] = 0;
-  }
-
-  for (int i = 0; i < data->num_blocks; i++) {
-    for (int j = 0; j < BLOCKSIZE; j++) {
-      if (data->blocks[i][j] >= 'a' && data->blocks[i][j] <= 'z') {
-
-        data->histogram[data->blocks[i][j] - 'a'] += 1;
-
-      } else if(data->blocks[i][j] >= 'A' && data->blocks[i][j] <= 'Z') {
-
-        data->histogram[data->blocks[i][j]-'A'] += 1;
-      } 
-    }
-  }
-  return NULL;
-}
-
-
-void get_histogram(unsigned int nBlocks,
-      block_t *blocks,
-      histogram_t histogram,
-      unsigned int num_threads) {
-
-  pthread_t* thread = malloc(num_threads * sizeof(*thread));
-  struct Data* data = malloc(num_threads * sizeof(*data));
-  unsigned int blocksPerThread = nBlocks / num_threads;
-  unsigned int remainingBlocks = nBlocks % num_threads;
-
-  // Spawn N threads to process the data
-  for(int i = 0; i < num_threads; i++) {
-    data[i].num_blocks = blocksPerThread;
-
-    if(remainingBlocks > 0) {
-        data[i].num_blocks += 1;
-        --remainingBlocks;
-    }
-
-    data[i].blocks = blocks;
-    blocks += data[i].num_blocks;
-    pthread_create(&thread[i], NULL, build_histogram, &data[i]);
-  }
-
-  // Wait for the workers to complete
-  for(int i = 0; i < num_threads; i++) {
-    pthread_join(thread[i], NULL);
-  }
-
-  for(int i = 0; i < SIZE; ++i) {
-    histogram[i] = 0;
-  }
-
-  // Accumulate the results
-  for(int i = 0; i < num_threads; i++) {
[…]

### 2015-05-01 HiWi/multiply_report/multiply_report.tex
add HiWi report

\documentclass[journal, a4paper]{IEEEtran}

\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath} 
\usepackage{booktabs}
\usepackage{listings}

\begin{document}

	\title{Report on Implementation of Fast Multiplication Routines in Fortran}
	\author{B.Sc. Adam Kosiorek, M.Sc. Nils Zander}	
	\markboth{Institute for Computation in Engineering, Technische Universit\"at M\"unchen}{}
	\maketitle

\begin{abstract}
 This report describes the derivation of performance models and development of fast matrix multiplication routines in Fortran for Intel Xeon E5-2689 Sandy Bridge CPU for the ADHOC++ project.
\end{abstract}

\section{Introduction}
    \PARstart{T}{he} majority of numerically intensive software projects can be reduced on the lowest evel to linear algebra problems. It is hardly surprising that the main suite of tools for them - Basic Linear Algebra Subroutines (BLAS) - is amongst the most heavily optimized software packages out there. Using advanced data distribution and load balancing techniques, it leverages data locality and various levels of parallelism to boost performance. Many BLAS implementations are, however, ill-suited for small sized problems. As shown later, Intel MKL~-- arguably the fastest implementation for x86 CPUs~\cite{MKL}~-- is easily outperformed by the most na\"ive general matrix-matrix multiplication code when inputs are not big enough. In many cases it is therefore possible and necessary to devise custom matrix multiplication routines optimized for smaller problems. We develop performance model and optimized code for the following cases of matrix-matrix multiplication:
    \begin{equation}
     C = Ax + C
    \end{equation}
    where $C \in \mathcal{R}^{N \times K}$, $A \in \mathcal{R}^{N \times M}$, $x \in \mathcal{R}^{M \times K}$ and $N \in \mathcal{N}^+$, $K \in \{1, 2, 3\}$

\section{Performance Model}
    Theoretical performance models, while not perfectly accurate, can hugely benefit development of optimized numerical programming routines by providing maximum performance estimates as a goal to strive for. Maximum theoretical performace can be computed by: 
    \begin{equation}
     P_{peak} = IPC * FPI * Frequency
    \end{equation}
    where $IPC$ - instructions per cycle, $FPI$ - floating point operations per instruction, $Frequency$ - CPU's frequency. Intel Xeon E5-2689 operates with the frequency of $2.9$ GHz and is capable of executing 1, 2 or 4 FPI, depending on the mode (Scalar, SSE, AVX) [need ref]. IPC depends on the numerical algorithm or the average number of floating point instructions executed per cycle to compute a single element of the output matrix. To derive IPC and $P_{peak}$, we refer to table \ref{tab:low_level}, which lists instructions that can be executed in one cycle by the Sandy Bridge CPU and to table \ref{tab:req}, which depicts number of each low level instructons needed to be executed to compute one entry of $C$ matrix. Finally, table \ref{tab:perf} summarizes results.

      \begin{table}[!hbt]
	\begin{center}
	  \caption{Low level instructions per cycle.}
	  \label{tab:low_level}
	  \begin{tabular}{c|c|c}
		  Instruction	& Scalar & AVX \\
		  \hline
		  load		& 1 or 2 & 1 \\
		  \hline
		  store		& 1 or 0 & 0.5 \\
		  \hline
		  multiply	& 1 & 1 \\
		  \hline
		  add 		& 1 & 1 \\
	  \end{tabular}
	\end{center}
      \end{table}

      \begin{equation}      
	  K=1:\;\;C(i, 1) = C(i, 1) + A(i, 1) * x(1, 1) 
      \end{equation}

      \begin{equation}      
	\begin{split}
	  K=2:\;\;C(i, 1) = C(i, 1)  + A(i, 1) * x(1, 1) \\+ A(i, 2) * x(2, 1)
	\end{split}
      \end{equation}

      \begin{equation}      
	\begin{split}
	  K=3:\;\;C(i, 1) = C(i, 1) + A(i, 1) * x(1, 1) \\+ A(i, 2) * x(2, 1) \\+ A(i, 3) * x(3, 1)
	\end{split}
      \end{equation}

  \begin{table}[!hbt]
    \begin{center}
      \caption{Low level instructions required to compute one entry of $C$ matrix.}
      \label{tab:req}
      \begin{tabular}{c|c|c|c|c}
	K	& load & store & multiply & add \\
	\hline
	1	& 2 & 1 & 1 & 1 \\
	\hline
	2	& 3 & 1 & 2 & 2 \\
	\hline
	3	& 4 & 1 & 3 & 3
      \end{tabular}
    \end{center}
  \end{table}

  \begin{table}[!hbt]
    \begin{center}
      \caption{Floating point instructions per cycle and peak theoretical performance for Scalar and AVX modes.}
      \label{tab:perf}
      \begin{tabular}{c|c|c|c|c}
		& \multicolumn{2}{c}{IPC} & \multicolumn{2}{c}{$P_{peak}$ [GFlops]} \\
	K	& Scalar & AVX	& Scalar & AVX \\
	\hline
	1	& 1 & 1 & 2.9 & 11.60 \\
	\hline
	2	& 2 & 1.33 & 5.8 & 15.47 \\
	\hline
	3	& 2 & 1.5 & 5.8 & 17.40
      \end{tabular}
    \end{center}
  \end{table}

\section{Implementation and Results}

  Fortran is the programming language traditionally used to implement high performance numerical algorithms. One reason is that it makes weaker assumptions about memory layout of variables than e.g. C, thus making vectorization easier.  We decided to use Intel's Fortran compiler ifort, which often produces faster code than GNU's gfortran \cite{intel_comp}, but also provides easier to read and more complete optimization feedback. Performance maximization requires code vectorization or execution in AVX mode. The following subsections describe constraints and counter-measures that enabled successful vectorization.

  \subsection{Memory Layout}
    Fortran assumes column-major memory layout, that is, a single column of a matrix is comprised of elements adjacent in memory. In our case the routines are executed from C/C++, which assumes row-major memory layout. Normally, that would require declaration of transposed matrices in Fortran: e.g. double A[sizeA1][sizeA2] in C/C++ would become DOUBLE PRECISION A(sizeA2, sizeA1) in Fortran. Efficient multiplication, however, requires iteration along rows of the original matrices, which transfers to impossible to vectorize iteration along the innermost dimension in Fortran. Thus, we resort to computing indices manually and declaring matrices as one dimensional vectors i.e. DOUBLE~PRECISION~A(sizeA1~$*$~sizeA2).

    Another limiting factor is memory alignment. High performance AVX instructions can be used only when data are aligned at appropriate memory boundaries. Even if that is the case, the compiler cannot be sure if every possible input will be memory-aligned. One way to ensure it is to use inline compiler directives, in this case: !DEC\$ VECTOR ALIGNED just before a loop referencing vectors.

  \subsection{Data Dependency - Read after Write}
    Listing \ref{lst:raw} contains the Read after Write dependency. It prevents vectorization, since the same element in memory is referenced multiple times. One possible solution is to introduce a local variable that would accumulate the results and store it in C(i) only once as in listing \ref{lst:noraw}. Compilers than unrolls the outer loop containing code from listing \ref{lst:noraw} and vectorizes the result.
    \begin{lstlisting}[label=lst:raw, caption=Read after Write dependency.]
      DO i = 1, 3
	C(i) = C(i) + A(index + i) * x(i)
      END DO
    \end{lstlisting}

    \begin{lstlisting}[label=lst:noraw, caption=Read after Write dependency resolved.]
      sum = 0
      DO i = 1, 3
	sum = sum + A(index + i) * x(i)
      END DO
      C(i) = C(i) + sum
    \end{lstlisting}

   \subsection{Results}

  Achieved results, summarized in table \ref{tab:results} are not perfect and are true only for matrices small enough to fit in L2 cache. Divergence of theoretical and achieved results might stem from the fact that we did not take memory bandwitch or cache effects into consideration.

  \begin{table}[!hbt]
    \begin{center}
      \caption{Achieved results and comparision with theoretical peak performance}
      \label{tab:results}
      \begin{tabular}{c|c|c}
	K	& $P_{achieved}$ [Gflops] & $P_{achieved} / P_{peak}$\\
	\hline
	1	& 9.17 & 0.79 \\
	\hline
	2	& 11.89 & 0.77 \\
	\hline
	3	& 13.96 & 0.8
      \end{tabular}
    \end{center}
  \end{table}

\begin{thebibliography}{5}

	\bibitem{MKL}
	Intel Math Kernel Library: \url{https://software.intel.com/en-us/intel-mkl}

	\bibitem{intel_comp}
	Intel® Parallel Studio XE 2015 Product Brief: \url{https://goo.gl/vc5eYo}

\end{thebibliography}

% Your document ends here!
\end{document}

### 2015-05-01 HiWi/multiply_report/multiply_report.tex
HiWi report spell check

@@
 \section{Introduction}
-    \PARstart{T}{he} majority of numerically intensive software projects can be reduced on the lowest evel to linear algebra problems.
+    \PARstart{T}{he} majority of numerically intensive software projects can be reduced on the lowest level to linear algebra problems.
 It is hardly surprising that the main suite of tools for them - Basic Linear Algebra Subroutines (BLAS) - is amongst the most heavily optimized software packages out there.
@@
     Theoretical performance models, while not perfectly accurate, can hugely benefit development of optimized numerical programming routines by providing maximum performance estimates as a goal to strive for.
-Maximum theoretical performace can be computed by:
+Maximum theoretical performance can be computed by:
     \begin{equation}
@@
 IPC depends on the numerical algorithm or the average number of floating point instructions executed per cycle to compute a single element of the output matrix.
-To derive IPC and $P_{peak}$, we refer to table \ref{tab:low_level}, which lists instructions that can be executed in one cycle by the Sandy Bridge CPU and to table \ref{tab:req}, which depicts number of each low level instructons needed to be executed to compute one entry of $C$ matrix.
+To derive IPC and $P_{peak}$, we refer to table \ref{tab:low_level}, which lists instructions that can be executed in one cycle by the Sandy Bridge CPU and to table \ref{tab:req}, which depicts number of each low level instructions needed to be executed to compute one entry of $C$ matrix.
 Finally, table \ref{tab:perf} summarizes results.
@@
    \subsection{Results}  
-
-  Achieved results, summarized in table \ref{tab:results} are not perfect and are true only for matrices small enough to fit in L2 cache.
-Divergence of theoretical and achieved results might stem from the fact that we did not take memory bandwitch or cache effects into consideration.
+    Achieved results, summarized in table \ref{tab:results}, are not perfect and are true only for matrices small enough to fit in L2 cache.
+Divergence of theoretical and achieved results might stem from the fact that we did not consider memory bandwidth or cache effects.
 
@@
     \begin{center}
-      \caption{Achieved results and comparision with theoretical peak performance}
+      \caption{Achieved results and comparison with theoretical peak performance}
       \label{tab:results}

### 2015-05-05 MLCV/ex2/classify.py
add MLCV ex2

@@ -1,0 +1,110 @@
+#!/usr/bin/env python
+
+import sys
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy
+
+# import caffe
+CAFFE_ROOT = '/home/adam/workspace/caffe/caffe/'
+CAFFE_PYTHON_ROOT = os.path.join(CAFFE_ROOT, 'python')
+sys.path.append(CAFFE_PYTHON_ROOT)
+import caffe
+
+
+MODEL_FILE = 'models/bvlc_reference_caffenet/deploy.prototxt'
+MODEL_FILE = os.path.join(CAFFE_ROOT, MODEL_FILE)
+PRETRAINED = 'models/bvlc_reference_caffenet/bvlc_reference_caffenet.caffemodel'
+PRETRAINED = os.path.join(CAFFE_ROOT, PRETRAINED)
+MEAN_FILE = os.path.join(CAFFE_ROOT, 'python/caffe/imagenet/ilsvrc_2012_mean.npy')
+SYNSET_FILE = os.path.join(CAFFE_ROOT, 'data/ilsvrc12/synset_words.txt')
+DATA_FOLDER = 'data'
+IMG_SIZE = (256, 256)
+
+PROB_LAYER = 'prob'
+
+def read_synset():
+    synset = []
+    with open(SYNSET_FILE) as f:
+        for line in f:
+            splitted = line.strip().split()
+            synset.append((splitted[0], ' '.join(splitted[1:])))
+
+    synset.sort()
+    synset = [entry[1] for entry in synset]
+    return synset
+
+def resizeToLonger(img, longer):
+    imgWidth = img.shape[1]
+    imgHeight = img.shape[0]
+
+    ratio = 0;
+    if imgWidth > imgHeight:
+        imgHeight = int(imgHeight / float(imgWidth) * longer)
+        imgWidth = longer 
+    else:
+        imgWidth = int(imgWidth / float(imgHeight) * longer)
+        imgHeight = longer
+
+    print imgHeight, imgWidth
+    return caffe.io.resize_image(img, (imgHeight, imgWidth)) 
+
+def padTo(img, width, height):
+    wPad = (width - img.shape[1]) / 2
+    wPadOdd = (width - img.shape[1]) % 2
+    hPad = (height - img.shape[0]) / 2
+    hPadOdd = (height - img.shape[0]) % 2
[…]

### 2015-05-05 MLCV/ex1/report/template.tex
report for ex2 MLCV

@@
 
-	\title{SVM for MNIST Handwritten Digits Classification}
+	\title{Assignment I\\MNIST Handwritten Digits Classification}
 	\author{Adam Kosiorek}	
-	\thanks{Advisor: M.Sc. Caner Hazibras & M.Sc. Phillip , Chair for Computer Vision, TUM, SS 2015.}
 	\markboth{Machine Learning for Applications in Computer Vision}{}
@@
 
-\begin{abstract}
-  This report describes classification of handwritten digits from the MNIST dataset by an SVM classifier.
-\end{abstract}
+%\begin{abstract}
+%  This report describes classification of handwritten digits from the MNIST dataset by an SVM classifier.
+%\end{abstract}
 
-\section{Introduction}
-    \PARstart{T}{he} first assignment was to explore the SVM classifier for the MNIST handwritten digits recognition.
+\section{Objective}
+    \PARstart{T}{he} first assignment was to explore the SVM classifier, Decision Trees and Random Forests for the MNIST handwritten digits recognition.
 
@@
 Each sample is a 28x28 pixel black and white image of a single centered digit.
-It is widly used as a benchmark for comparing machine learning algorithms.
+It is widely used as a benchmark for comparing machine learning algorithms.
 
@@
 In case of MNIST and SVM, however, state of the arts results have been achieved without preprocessing, that is, with raw pixels \cite{VSVM}.
-I compare the influence of mean intensity subtraction and ZCA-Whitening \cite{UFLDL} on classification accuracy.
+I examine the influence of mean intensity subtraction and momentum-based deskewing on classification accuracy.
 
     \subsection{Classifier Tuning}
-	The SVM classifier can be adapted to a particular task by choosing a kernel type and adjusting several parameters.
-Since the parameters might not be independent it is often necessary to perform an exhaustive grid search.
-Python's sklearn.grid\_search.GridSearchCV is well suited for this task.
-I explored linear kernel, rbf kernel with gamma between 1e-6 and 1 and a polynomial kernel of degree between 2 and 9.
-For all kernels the C parameter was taken from the range between 1e-2 and 1e4.
-Tuning was performed on a 10\% subset of the data - 6000 training and 1000 testing samples.
-Only the best performing classifiers were evaluated on the whole dataset.
-
-\section {Results and Discussion}
+	Most classifiers can be fine-tuned to a particular task by adjusting their hyper-parameters.
+They are rarely independent of each other and it is necessary to perform an exhaustive grid search in the hyper-parameter space.
+I chose to do so with 10\% of the training data and 5-fold cross-validation.
+The parameters optimized for each classifier are shown in table \ref{tab:parameters}.
 
@@
 		\begin{center}
-		\label{tab:results}
-		\begin{tabular}{|c|c|}
-			% To create a horizontal line, type \hline
+		\caption{Parameters optimized for each classifier}
+		\label{tab:parameters}
+		\begin{tabular}{c|c}
+			Classifier & Parameters \\
 			\hline
-			% To end a column type &
-			% For a linebreak type \\
-			Information message length & $k=16000$ bit \\
+			SVM & kernel type, C, gamma (for RBF), degree (for polynomial) \\
 			\hline
-			Radio segment size & $b=160$ bit \\
+			Decision Tree & cost function, max tree depth, number of features \\
 			\hline
-			Rate of component codes & $R_{cc}=1/3$\\
-			\hline
-			Polynomial of component encoders & $[1 , 33/37 , 25/37]_8$\\
-			\hline
+			Random Forests & number of estimators, cost function, \\ 
+			& max estimator depth, number of features\\
 		\end{tabular}
@@
 
-% Now we need a bibliography:
+	Classifiers with the best performing parameter combinations were trained on the full dataset.
+
+\section {Results and Discussion}
+    Table \ref{tab:results} summarizes results.
+Mean subtraction and deskewing applied together delivered the highest accuracy for SVM.
+Therefore, this composition was used in the rest of the experiments.
+Parameter tuning resulted in the following parameters for each classifier:
+    \begin{itemize}
+     \item SVM: \hfill polynomial kernel of degree 2, $C = 1000$
+     \item Decision Tree: 500 max features, 100 max depth, entropy criterion
+     \item Random Forest: 50 max features, 10 max depth, entropy criterion, 100 estimators
+    \end{itemize}
+
+    \begin{table}[!hbt]
+	    \begin{center}
+	    \caption{Results}
+	    \label{tab:results}
+	    \begin{tabular}{c|c|c|c}
+		    Classifier & Preprocessing  & Parameters  & Accuracy \\
+		    \hline
+			      & none             &            & 98.02\% \\
+		    SVM       & mean             & fine-tuned & 98.28\% \\
+			      & deskew           &            & 98.70\% \\
+			      & mean \& deskew   &            & 98.84\% \\
+		    \hline
+		    Decision & mean \& deskew    & default    & 90.44\% \\
+		    Tree     & mean \& deskew    & fine-tuned & 91.08\% \\
+		    \hline
+		    Random   & mean \& deskew    & default    & 96.01\% \\
+		    Forest   & mean \& deskew    & fine-tuned & 96.28\% \\
+	    \end{tabular}
+	    \end{center}
+    \end{table}
+
+    In this setting SVM outperforms Decision Tree and Random Forest classifiers.
+Random Forest is far better than the basic Decision Tree, even though it uses the latter as its building block.
+Each tree of Random Forest's ensemble uses only up to 10\% of the features and levels used by Decision Tree classifier.
+
+    Achieved performance is far from the state-of-the-art level, however.
+The highest classification performance up to date is by a convolutional neural network \cite{MNIST_Website}.
+The best performing SVM classifier leverages robustness to affine-transformations by a special way of training \cite{VSVM}.
+
+    \begin{figure}[!hbt]
+      \includegraphics[width=0.4\textwidth]{tree_importance}
+      \caption{Pixel importance for fine-tuned Decision Tree classifier}
+      \label{fig:tree_importance}
+    \end{figure}
+    \begin{figure}[!hbt]
+      \includegraphics[width=0.4\textwidth]{random_importance}
+      \caption{Pixel importance for fine-tuned Random Forest classifier}
+      \label{fig:random_importance}
+    \end{figure}
+
+    Figures \ref{fig:tree_importance} and \ref{fig:random_importance} depict pixel importance as seen by fine-tuned Decision Tree and Random Forest classifiers respectively.
+It seems that due to the limited depth and number of features the individual trees of Random Forest better cover the feature space of the MNSIT dataset, allowing for improved generalization.
+
 \begin{thebibliography}{5}
@@
 
-	\bibitem{MJH06} % Conference paper
-	T.~Mayer, H.~Jenkac, and J.~Hagenauer. Turbo base-station cooperation for intercell interference cancellation. {\em IEEE Int. Conf. Commun. (ICC)}, Istanbul, Turkey, pp.~356--361, June 2006.
-
 	\bibitem{VSVM}
 	D.~Decoste and B.~Schölkopf. ``Training invariant support vector machines.'' Machine learning 46, no. 1-3 (2002): 161-190.
-
-	\bibitem{Proakis} % Book
-	J.~G.~Proakis. {\em Digital Communications}. McGraw-Hill Book Co.,
-	New York, USA, 3rd edition, 1995.
 
@@
 
-	\bibitem{5}
-	IEEE Transactions \LaTeX and Microsoft Word Style Files.
-	\url{http://www.ieee.org/web/publications/authors/transjnl/index.html}
-
 \end{thebibliography}

### 2015-05-05 MLCV/ex2/report/report.tex
report for ex2 MLCV

\documentclass[journal, a4paper]{IEEEtran}

\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath} 
\usepackage{subcaption}

\begin{document}

	\title{Assignment II\\Neural Networks and Deep Learning}
	\author{Adam Kosiorek}	
	\markboth{Machine Learning for Applications in Computer Vision}{}
	\maketitle

\section{Objective}
    \PARstart{T}{he} aim of the second assignment was to develop thorough understanding of deep neural networks. Firstly, through literature research and derivation of backpropagation algorithm for arbitrary activation functions we gained insight into the inner workings of feed-forward neural networks. Secondly, we learned how to use a state-of-the-art open-source neural network package \textit{caffe} \cite{caffe} by classifing given images.

\section{Neural Networks and Deep Learning}
    Neural networks are nothing more than a composition of series of non-linear functions and matrix-matrix products. Yet, due to clever learning algorithms, they can be trained to approximate any function. Before we can teach a network, however, we need a dataset of inputs and ground truth pairs as well as a cost function. The latter quantifies differences between network's output and the ground truth for a given input and is being minimized by the training procedure on the whole dataset. The term ``deep learning'' describes neural networks with hidden layers - the layers between input and output layers. It is believed that deep networks learn feature hierarchies, thus alleviating the need for feature engineering.

\section{Deriving Backpropagation}
    In his book, M.~Nielsen derived the Backpropgation algorithm for a sigmoid activation function $\sigma$ \cite{MLDL}. Here, following the notaion, I derive it for an arbitrary function $f$.

    TODO

\section {Classification}
  \textit{caffe} is easy to use. Written in C++ with CUDA acceleration, it is high performance, yet, due to Python and Matlab bindings and configuration-oriented architecture it is user-friendly. After completing \textit{ImageNet classification} Notebook I decided to write a custom script for performing this assignment \cite{code}.

  \subsection{Configuration}
    Every neural network in \textit{caffe} needs to be configured before it becomes operational. It needs a deployment configuration file - Google Protocol Buffer file defining the network with free (e.g. non-database) inputs, and a trained model. Since we wanted the output class name after classification we also need a synset file with class names and identifiers from ILSVRC2012 \cite{ImageNet}

  \subsection{Data Preparation}
    Provided images are 3-channel with various sizes and aspect ratios, while the reference \textit{caffe} model operates on $256\times 256\times 3$ data. Convolutional neural networks are scale invariant but are also sensitive to changes in aspect ratio. I have therefore resized each input image to 256 px on the longer size and then padded it with zeros to form a square matrix. Padding with zeros should not affect classification, since zeros do not contribute to convolution results. Moreover, the training data for the ImageNet model had intensity values between 0 and 255 and was preprocessed by subtracting mean value computed on the whole dataset and I preprocessed input images accordingly.

  \subsection{Results and Discussion}
    Figure \ref{fig:imgs} depicts images given for classification and table \ref{tab:results} shows results. Only images (f) and (g) were classified correctly, as might be suggested by high classification probability and low entropy. Images (e) and (h) can be seen as similar to their predicted categories with lower classification probability and higher entropy. All other images have very low probability and high entropy, suggesting that the classifier is wrong. It seems that classification probability and entropy can be good measures of prediction credibility. According to expectations, the lower the probability (the wider it is spread), the higher the entropy.

  \begin{figure}[h]
    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_1}
      \caption{}
      \label{fig:a}
    \end{subfigure}%
    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_2}
      \caption{}
      \label{fig:b}
    \end{subfigure}

    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_3}
      \caption{}
      \label{fig:c}
    \end{subfigure}%
    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_4}
      \caption{}
      \label{fig:d}
    \end{subfigure}

    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_5}
      \caption{}
      \label{fig:e}
    \end{subfigure}%
    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_6}
      \caption{}
      \label{fig:f}
    \end{subfigure}

    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_7}
      \caption{}
      \label{fig:g}
    \end{subfigure}%
    \begin{subfigure}{.25\textwidth}
      \centering
      \includegraphics[width=.8\linewidth]{../data/IMG_8}
      \caption{}
      \label{fig:h}
    \end{subfigure}

    \caption{Images for Classification.}
    \label{fig:imgs}
  \end{figure}

  \begin{table}[h]
    \begin{center}
    \caption{Classification Results}
    \label{tab:results}
    \begin{tabular}{c|c|c|c|c}
      \# & \#class  & class name  & probability & entropy \\ \hline
      a & 428 & barrow, garden cart, lawn cart, \ldots & 0.07 & 5.22\\ \hline
      b & 311 & grasshopper, hopper & 0.14 & 3.72\\ \hline
      c & 403 & aircraft carrier, carrier, flattop, \ldots & 0.13 & 4.67\\ \hline
      d & 620 & laptop, laptop computer & 0.12 & 4.63\\ \hline
      e & 673 & mouse, computer mouse & 0.41 & 2.75\\ \hline
      f & 947 & mushroom & 0.76 & 1.17\\ \hline
      g & 834 & suit, suit of clothes & 0.62 & 1.09\\ \hline
      h & 980 & volcano & 0.36 & 3.32
    \end{tabular}
    \end{center}
  \end{table}

\begin{thebibliography}{5}

	\bibitem{code}
	Assignment Code: \url{https://github.com/akosiorek/CSE/tree/master/MLCV/ex2/}

	\bibitem{MLDL} % Web document
	M.~A.~Nielsen, ``Neural Networks and Deep Learning'', Determination Press, 2015
	\url{http://neuralnetworksanddeeplearning.com}.

	\bibitem{caffe}
	Y.~Jia, E.~Shelhamer, J.~Donahue, S.~Karayev, J.~Long, R.~Girshick, S.~Guadarrama and T.~Darrell, ``Caffe: Convolutional Architecture for Fast Feature Embedding.'' arXiv preprint arXiv:1408.5093, 2014.

	\bibitem{ImageNet}
	\url{http://image-net.org/challenges/LSVRC/2012/index}

\end{thebibliography}

% Your document ends here!
\end{document}

### 2015-05-05 MLCV/ex2/report/report.tex
derived backpropagation for ex2 MLCV, check spelling, improve format

@@
     \PARstart{T}{he} aim of the second assignment was to develop thorough understanding of deep neural networks.
-Firstly, through literature research and derivation of backpropagation algorithm for arbitrary activation functions we gained insight into the inner workings of feed-forward neural networks.
-Secondly, we learned how to use a state-of-the-art open-source neural network package \textit{caffe} \cite{caffe} by classifing given images.
+Firstly, through literature research and derivation of Backpropagation algorithm for arbitrary activation functions we gained insight into the inner workings of feed-forward neural networks.
+Secondly, we learned how to use a state-of-the-art open-source neural network package \textit{caffe} \cite{caffe} by classifying given images.
 
@@
 \section{Deriving Backpropagation}
-    In his book, M.~Nielsen derived the Backpropgation algorithm for a sigmoid activation function $\sigma$ \cite{MLDL}. Here, following the notaion, I derive it for an arbitrary function $f$.
+    In his book, M.~Nielsen derived the Backpropgation algorithm for a sigmoid activation function $\sigma$ \cite{MLDL}. Here, following the notation, I derive it for an arbitrary function $f$.
 
-    TODO
+    Let $\delta_j^L$ be the error of $j^{th}$ output neuron, then:
+    \begin{equation}
+      \delta_j^L = \frac{\partial C(a^L)}{\partial z_j^L} = \frac{\partial C(a^L)}{\partial a_j^L} \frac{\partial a_j^L}{\partial z_j^L}
+    \end{equation}
+
+    where:
+
+    \begin{equation}
+      \frac{\partial a_j^L}{\partial z_j^L} = \frac{\partial f(z^L)}{\partial z_j^L} = f'(z_j^L)
+    \end{equation}
+
+    and thus:
+
+    \begin{equation}
+      \delta^L = \nabla C(a^L) \odot f'(z^L)
+    \end{equation}
+
+    For the previous layer, denoted $l$:
+
+    \begin{equation}
+    \begin{gathered}
+      \delta^l = \frac{\partial C(a^L)}{\partial z^l} = \frac{\partial a^L}{\partial z^l} \nabla C(a^L) = \\ = \frac{\partial z^L}{\partial z^l} (f'(z^L) \odot \nabla C(a^L) ) = \frac{\partial z^L}{\partial z^l} \delta^L = \\ = diag(f'(z^l))(w^L)^T \delta^L = f'(z^L) \odot ((w^L)^T \delta^L)
+    \end{gathered}
+    \end{equation}
+
+    Since relations between any two adjacent layers are the same it holds that:
+
+    \begin{equation}
+      \delta^l  = f'(z^{l+1}) \odot ((w^{l+1})^T \delta^{l+1})
+    \end{equation}
+
+    The final equations for gradients w.r.t. weights and biases do not depend on the activation function and has the same form as in \cite{MLDL}.
 
@@
 Written in C++ with CUDA acceleration, it is high performance, yet, due to Python and Matlab bindings and configuration-oriented architecture it is user-friendly.
-After completing \textit{ImageNet classification} Notebook I decided to write a custom script for performing this assignment \cite{code}.
+After completing \textit{ImageNet classification} Notebook I have written a custom script for the assignment~\cite{code}.
 
   \subsection{Configuration}
-    Every neural network in \textit{caffe} needs to be configured before it becomes operational.
+    Every neural network in \textit{caffe} needs to be configured before being operational.
 It needs a deployment configuration file - Google Protocol Buffer file defining the network with free (e.g.
-non-database) inputs, and a trained model.
+non-database) inputs and a trained model.
 Since we wanted the output class name after classification we also need a synset file with class names and identifiers from ILSVRC2012 \cite{ImageNet}
@@
     Provided images are 3-channel with various sizes and aspect ratios, while the reference \textit{caffe} model operates on $256\times 256\times 3$ data.
-Convolutional neural networks are scale invariant but are also sensitive to changes in aspect ratio.
+Convolutional neural networks are scale invariant but are sensitive to changes in aspect ratio.
 I have therefore resized each input image to 256 px on the longer size and then padded it with zeros to form a square matrix.

### 2015-05-06 MLCV/ex2/report/report.tex
stylistic correction

@@
 	\author{Adam Kosiorek}	
-	\markboth{Machine Learning for Applications in Computer Vision}{}
+	\markboth{Machine Learning for Applications in Computer Vision, Institute of Computer Vision, Technische Universit\'at M\"unchen}{}
 	\maketitle
@@
 \section{Deriving Backpropagation}
-    In his book, M.~Nielsen derived the Backpropgation algorithm for a sigmoid activation function $\sigma$ \cite{MLDL}. Here, following the notation, I derive it for an arbitrary function $f$.
+    In his book, M.~Nielsen derived the Backpropgation algorithm for the sigmoid activation function $\sigma$ \cite{MLDL}. Here, following the notation, I derive it for an arbitrary function $f$.
 
@@
     \begin{gathered}
-      \delta^l = \frac{\partial C(a^L)}{\partial z^l} = \frac{\partial a^L}{\partial z^l} \nabla C(a^L) = \\ = \frac{\partial z^L}{\partial z^l} (f'(z^L) \odot \nabla C(a^L) ) = \frac{\partial z^L}{\partial z^l} \delta^L = \\ = diag(f'(z^l))(w^L)^T \delta^L = f'(z^L) \odot ((w^L)^T \delta^L)
+      \delta^l = \frac{\partial C(a^L)}{\partial z^l} = \frac{\partial a^L}{\partial z^l} \nabla C(a^L) = \\ = \frac{\partial z^L}{\partial z^l} (f'(z^L) \odot \nabla C(a^L) ) = \frac{\partial z^L}{\partial z^l} \delta^L = \\ = diag(f'(z^l))(w^L)^T \delta^L = f'(z^l) \odot ((w^L)^T \delta^L)
     \end{gathered}
@@
     \begin{equation}
-      \delta^l  = f'(z^{l+1}) \odot ((w^{l+1})^T \delta^{l+1})
+      \delta^l  = f'(z^l) \odot ((w^{l+1})^T \delta^{l+1})
     \end{equation}
 
-    The final equations for gradients w.r.t. weights and biases do not depend on the activation function and has the same form as in \cite{MLDL}.
+    The final equations for gradients w.r.t. weights and biases do not depend on the activation function and have the same form as in \cite{MLDL}.
 

### 2015-05-07 HiWi/multiply_report/multiply_report.tex
add plots and transposed cases to fortran multiplication report

@@
 \usepackage{booktabs}
+
+%% Code Listing setup
 \usepackage{listings}
+\usepackage{color}
+
+\definecolor{mygreen}{rgb}{0,0.6,0}
+\definecolor{mygray}{rgb}{0.5,0.5,0.5}
+\definecolor{mymauve}{rgb}{0.58,0,0.82}
+
+\lstset{ %
+  backgroundcolor=\color{white},   % choose the background color
+  basicstyle=\footnotesize,        % size of fonts used for the code
+  breaklines=true,                 % automatic line breaking only at whitespace
+  captionpos=t,                    % sets the caption-position to bottom
+  commentstyle=\color{mygreen},    % comment style
+  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
+  keywordstyle=\color{blue},       % keyword style
+  stringstyle=\color{mymauve},     % string literal style
+}
+
+\def\arraystretch{1.25}
 
@@
 \begin{abstract}
- This report describes the derivation of performance models and development of fast matrix multiplication routines in Fortran for Intel Xeon E5-2689 Sandy Bridge CPU for the ADHOC++ project.
+ This report describes the derivation of performance models and development of fast matrix multiplication routines in Fortran for Intel Xeon E5-2690 Sandy Bridge CPU for the ADHOC++ project.
 \end{abstract}
@@
 \section{Performance Model}
-    Theoretical performance models, while not perfectly accurate, can hugely benefit development of optimized numerical programming routines by providing maximum performance estimates as a goal to strive for.
+    Theoretical performance models, while not perfectly accurate, can hugely benefit development of optimized numerical programming routines by providing a maximum performance estimate as a goal to strive for.
 Maximum theoretical performance can be computed by:
     \begin{equation}
-     P_{peak} = IPC * FPI * Frequency
+     P_{peak} = IPC \cdot FPI \cdot Frequency
     \end{equation}
     where $IPC$ - instructions per cycle, $FPI$ - floating point operations per instruction, $Frequency$ - CPU's frequency.
-Intel Xeon E5-2689 operates with the frequency of $2.9$ GHz and is capable of executing 1, 2 or 4 FPI, depending on the mode (Scalar, SSE, AVX) [need ref].
-IPC depends on the numerical algorithm or the average number of floating point instructions executed per cycle to compute a single element of the output matrix.
-To derive IPC and $P_{peak}$, we refer to table \ref{tab:low_level}, which lists instructions that can be executed in one cycle by the Sandy Bridge CPU and to table \ref{tab:req}, which depicts number of each low level instructions needed to be executed to compute one entry of $C$ matrix.
+Intel Xeon E5-2690 operates with the frequency of $2.9$ GHz and is capable of executing 1, 2 or 4 $FPI$, depending on the operation mode (Scalar, SSE, AVX).
+$IPC$ depends on the numerical algorithm or the average number of floating point instructions executed per cycle to compute a single element of the output matrix.
+To derive $IPC$ and $P_{peak}$, we refer to table \ref{tab:low_level}, which lists instructions that can be executed in one cycle by the Sandy Bridge CPU and to table \ref{tab:req}, which depicts number of each low level instructions needed to be executed to compute one entry of $C$ matrix.
 Finally, table \ref{tab:perf} summarizes results.
+It is obvious that our case can hugely benefit from AVX operation mode.
 
@@
       \begin{equation}      
-	  K=1:\;\;C(i, 1) = C(i, 1) + A(i, 1) * x(1, 1) 
+	  K=1:\;\;C(i, 1) = C(i, 1) + A(i, 1) \cdot x(1, 1) 
       \end{equation}
@@
 	\begin{split}
-	  K=2:\;\;C(i, 1) = C(i, 1)  + A(i, 1) * x(1, 1) \\+ A(i, 2) * x(2, 1)
+	  K=2:\;\;C(i, 1) = C(i, 1)  + A(i, 1) \cdot x(1, 1) \\+ A(i, 2) \cdot x(2, 1)
 	\end{split}
@@
 	\begin{split}
-	  K=3:\;\;C(i, 1) = C(i, 1) + A(i, 1) * x(1, 1) \\+ A(i, 2) * x(2, 1) \\+ A(i, 3) * x(3, 1)
+	  K=3:\;\;C(i, 1) = C(i, 1) + A(i, 1) \cdot x(1, 1) \\+ A(i, 2) \cdot x(2, 1) \\+ A(i, 3) \cdot x(3, 1)
 	\end{split}
@@
   Fortran is the programming language traditionally used to implement high performance numerical algorithms.
-One reason is that it makes weaker assumptions about memory layout of variables than e.g.
-C, thus making vectorization easier.
+One reason is that it makes stronger assumptions about memory layout of variables than \textit{e.g.} the C language, thus making vectorization easier.
 We decided to use Intel's Fortran compiler ifort, which often produces faster code than GNU's gfortran \cite{intel_comp}, but also provides easier to read and more complete optimization feedback.
-Performance maximization requires code vectorization or execution in AVX mode.
-The following subsections describe constraints and counter-measures that enabled successful vectorization.
+To use AVX instructions, we can rely on one of the following: (1) use compiler intrinsics or (2) write standard Fortran code in a way that enables automatic vectorization.
+The following sections describe constraints and counter-measures that enabled successful vectorization.
 
-  \subsection{Memory Layout}
+  \subsection{Memory Layout and Iteration}
     Fortran assumes column-major memory layout, that is, a single column of a matrix is comprised of elements adjacent in memory.
-In our case the routines are executed from C/C++, which assumes row-major memory layout.
-Normally, that would require declaration of transposed matrices in Fortran: e.g.
-double A[sizeA1][sizeA2] in C/C++ would become DOUBLE PRECISION A(sizeA2, sizeA1) in Fortran.
-Efficient multiplication, however, requires iteration along rows of the original matrices, which transfers to impossible to vectorize iteration along the innermost dimension in Fortran.
-Thus, we resort to computing indices manually and declaring matrices as one dimensional vectors i.e.
-DOUBLE~PRECISION~A(sizeA1~$*$~sizeA2).
+In our case, the routines are executed from C/C++, which assumes row-major memory layout.
+Normally, that would require declaration of transposed matrices in Fortran: \textit{e.g.} \texttt{double A[sizeA1][sizeA2]} in C/C++ would become \texttt{DOUBLE PRECISION A(sizeA2, sizeA1)} in Fortran.
+Efficient multiplication, however, requires iteration along rows of the original matrices, which transfers to impossible to vectorize (\textit{i.e.} Scalar mode) iteration along the innermost dimension in Fortran.
+Thus, we resort to computing indices manually and declaring matrices as one dimensional vectors \textit{i.e.} \texttt{DOUBLE~PRECISION~A(sizeA1~$\cdot$~sizeA2)}.
 
+    This approach improves performance, but is still slow when compared to $P_{peak}$.
+A closer inspection of Assembly code reveals that AVX instructions are used, but inefficiently for $M > 1$.
+To understand it, we have to explicitly look at the data layout in memory.
+Equation \ref{eq:mat} represents the case of $M = 3$ as seen in Fortran (vector $C$ omitted for brevity).
+
+  \begin{equation}
+    x^TA^T = \begin{pmatrix}
+	  x_1 & x_2 & x_3
+	  \end{pmatrix}
+	  \begin{pmatrix}
+	    a_{1, 1} & a_{1, 2} & a_{1, 3} & \cdots \\
+	    a_{2, 1} & a_{2, 2} & a_{2, 3} & \cdots \\
+	    a_{3, 1} & a_{3, 2} & a_{3, 3} & \cdots \\
+	  \end{pmatrix}
+	  \label{eq:mat}
+  \end{equation}
+
+  Here, elements $a_{1, i}$, $a_{2, i}$, $a_{3, i}$ are contiguous in memory.
+Since AVX operates on 4 elements at once and only 3 elements are available, one element would have to be discarded.
+Compiler, therefore, chooses another way by unrolling the multiplication loop four times and vectorizing along the rows of the matrix.
+Now, four elements of C can be computed in three steps (using $a_{1, i:i+4}$ first, $a_{2, i:i+4}$ next and finally $a_{3, i:i+4}$).
+An inevitable performance loss occurs due to the elements $a_{j, i:i+4}$ being non-contiguous, so they have to be loaded one by one into the AVX register.
+
+\begin{equation}
+  Ax = 
+        \begin{pmatrix}
+          a_{1, 1} & a_{1, 2} & a_{1, 3} \\
+          a_{2, 1} & a_{2, 2} & a_{2, 3} \\
+          a_{3, 1} & a_{3, 2} & a_{3, 3} \\
+          \vdots & \vdots & \vdots \\
+        \end{pmatrix}
+        \begin{pmatrix}
+         x_1 \\ x_2 \\ x_3
+        \end{pmatrix}
+        \label{eq:mat_transposed}
+\end{equation}
+
+  The opposite situation occurs with the transposed version of the problem, as shown in equation \ref{eq:mat_transposed}.
+Again, computation of one element of the $C$ vector would be inefficient, because one entry would have to be discarded from the AVX register.
+When unrolled 4 times, however, vectorization can proceed along memory-contiguous columns.
+Packed AVX load instructions lead to higher performance.
+
+  \subsection{Data Alignment}
     Another limiting factor is memory alignment.
@@
 Even if that is the case, the compiler cannot be sure if every possible input will be memory-aligned.
-One way to ensure it is to use inline compiler directives, in this case: !DEC\$ VECTOR ALIGNED just before a loop referencing vectors.
+One way to ensure it is to use inline compiler directives, in this case: \texttt{!DEC\$ VECTOR ALIGNED} just before vector references.
 
@@
 It prevents vectorization, since the same element in memory is referenced multiple times.
-One possible solution is to introduce a local variable that would accumulate the results and store it in C(i) only once as in listing \ref{lst:noraw}.
-Compilers than unrolls the outer loop containing code from listing \ref{lst:noraw} and vectorizes the result.
-    \begin{lstlisting}[label=lst:raw, caption=Read after Write dependency.]
+One possible solution is to introduce a local variable that would accumulate the results and store it in \texttt{C(i)} only once as in listing \ref{lst:noraw}.
+The compiler than unrolls the outer loop with the code from listing \ref{lst:noraw} and vectorizes the result.
+    \begin{lstlisting}[language=Fortran,label=lst:raw, caption=Read after Write dependency.]
       DO i = 1, 3
@@
 
-    \begin{lstlisting}[label=lst:noraw, caption=Read after Write dependency resolved.]
+    \begin{lstlisting}[language=Fortran,label=lst:noraw, caption=Read after Write dependency resolved.]
       sum = 0
@@
 
+    This optimization is relevant only to the normal (\textit{i.e}. non-transposed) setting.
+
    \subsection{Results}  
@@
 Divergence of theoretical and achieved results might stem from the fact that we did not consider memory bandwidth or cache effects.
+They also suggest that using transposed matrix $A$ might be extremely beneficial in terms of speed.
+Even performing transposition before multiplication could be justified when a number of different $x$ vectors is to be multiplied.
+Figures \ref{fig:1x1}, \ref{fig:2x1} and \ref{fig:3x1} compare performance of multiplication routines of the normal and transposed cases with Intel's MKL.
+MKL's functions \texttt{cblas\_daxpy} was used for \texttt{[1x1]} case and \texttt{cblas\_dgemv} for cases \texttt{[2x1]} and \texttt{[3x1]}.
 
@@
 	\hline
-	1	& 9.17 & 0.79 \\
+	$1$	& 9.17 & 0.79 \\
 	\hline
-	2	& 11.89 & 0.77 \\
+	$2$	& 9.64 & 0.62 \\
 	\hline
-	3	& 13.96 & 0.8
+	$3$	& 7.15 & 0.41 \\
+	\hline
+	$2^T$	& 11.89 & 0.77 \\
+	\hline
+	$3^T$	& 13.96 & 0.8 \\
       \end{tabular}
@@
   \end{table}
+
+  \begin{figure}
+   \includegraphics[width=0.5\textwidth]{1x1}
+   \caption{Performance of $C = Ax + C$, where $x \in \mathcal{R}$}
+   \label{fig:1x1}
+  \end{figure}
+
+  \begin{figure}
+   \includegraphics[width=0.5\textwidth]{2x1}
+   \caption{Performance of $C = Ax + C$, where $x \in \mathcal{R}^{2\times 1}$}
+   \label{fig:2x1}
+  \end{figure}
+
+  \begin{figure}
+   \includegraphics[width=0.5\textwidth]{3x1}
+   \caption{Performance of $C = Ax + C$, where $x \in \mathcal{R}^{3\times 1}$}
+   \label{fig:3x1}
+  \end{figure}
 

### 2015-05-07 HiWi/plot_from_csv.m

@@ -1,0 +1,54 @@
+function plot_from_csv()
+    close all, clear all, clc
+    
+    dim = '2x1';
+    p_peak = 15.47;
+    total_cols = 5;
+    
+    cache_size = 32 * 2^10;
+    cache_boundary = cache_size / 8 / total_cols;
+    normalData = sprintf('normal/%s.txt', dim);
+    transposedData = sprintf('transposed/%s.txt', dim);
+    blasData = sprintf('blas/%s.txt', dim);
+    
+    if ismember('transposedData', who)
+        data_files = {normalData, transposedData, blasData};
+    else
+        data_files = {normalData blasData};
+    end
+    colors = 'rkb';
+    
+    max_num = 0;
+    max_perf = 0;
+    
+    figure(1)
+    hold on
+    grid on
+    for i = 1:numel(data_files)
+        data = csvread(data_files{i});
+        data = data(:, 1:21);
+        plot(data(1, :), data(2, :), strcat(colors(i), '*-'), 'LineWidth', 2)
+        
+        max_num = max(max_num, max(data(1, :)));
+        max_perf = max(max_perf, max(data(2, :)));
+    end
+    
+    x = linspace(0, max(cache_boundary, max_num), 10);
+    y = p_peak * ones(size(x));
+    plot(x, y, 'm')
+    
+    y = linspace(0, max(max_perf, p_peak), 2);
+    x = cache_boundary * ones(size(y));
+    plot(x, y, 'g')
+    
+    xlabel('Number of rows of A')
+    ylabel('Performance [GFlops]')
+    %title(sprintf('Performance of C = Ax + C with x = [%s]', dim))
+    if ismember('transposedData', who)        
+        legend('Normal', 'Transposed', 'MKL', 'P_{peak}', 'Cache limit')
+    else
+        legend('Normal', 'MKL', 'P_{peak}', 'Cache limit')
+    end
+
+    print('-depsc', dim)
+end

### 2015-05-08 MLCV/ex2/report/report.tex
add case for only 1 neuron modified

@@
       \delta_j^L = \frac{\partial C(a^L)}{\partial z_j^L} = \frac{\partial C(a^L)}{\partial a_j^L} \frac{\partial a_j^L}{\partial z_j^L}
+      \label{eq:error_L}
     \end{equation}
@@
       \frac{\partial a_j^L}{\partial z_j^L} = \frac{\partial f(z^L)}{\partial z_j^L} = f'(z_j^L)
+      \label{eq:activ_L}
     \end{equation}
@@
     The final equations for gradients w.r.t. weights and biases do not depend on the activation function and have the same form as in \cite{MLDL}.
+
+    If, on the other hand, only a single neuron in one of the layers has a different activation function, the errors for that layer have to be computed for one neuron at a time.
+Suppose that all neurons but the neuron $j$ in layer $l$ use the sigmoid function $\sigma$ and this one neuron uses the function $f$.
+If $l = L$ equation \ref{eq:error_L} holds.
+Equation \ref{eq:activ_L} changes, however, for all but the $j^{th}$ neuron and it reads:
+
+    \begin{equation}
+     \frac{\partial a_j^L}{\partial z_j^L} = \frac{\partial \sigma(z^L)}{\partial z_j^L} = \sigma'(z_j^L)
+    \end{equation}
+
+    If $l < L$, then the error for the odd neuron reads:
+
+    \begin{equation}
+      \delta_j^l = f'(z_j^L) \sum_i w_{i, j}^{l+1} \delta_i^{l+1}
+    \end{equation}
+
+    and for all the other neurons of that layer:
+
+    \begin{equation}
+      k \neq j: \;\; \delta_k^l = \sigma'(z_k^L) \sum_i w_{i, k}^{l+1} \delta_i^{l+1}
+    \end{equation}
+
+    The simplicity and symmetry of matrix-matrix multiplication is therefore lost and errors have to be computed element-wise. Again, the gradient equations do not change.
 

### 2015-05-08 parallel_programming/ex3/histogram.c
ex 3

@@ -1,0 +1,69 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+#include "histogram.h"
+
+#define MAX(X,Y) ((X) > (Y) ? (X) : (Y))
+#define NLINES 20
+
+static char *buffer;
+static unsigned int buffer_size = 0;
+
+// get the maximum number of characters per letter
+int max_chars(histogram_t histogram) {
+	int i, max = 0;
+
+	for (i=0; i<NALPHABET; ++i)
+		max = MAX(histogram[i], max);
+
+	return max;
+}
+
+void print_histogram(histogram_t histogram) {
+
+	int max = max_chars(histogram);
+	int i, j, y;
+
+	// print line by line
+	for (i=NLINES; i>=0; i--) {
+
+		// calculate and print y labels
+		y = (i*max) / NLINES;
+		printf("%10d:", y);
+
+		// print histogram
+		for (j=0; j<NALPHABET; j++) {
+			if (histogram[j] >= y)
+				printf("%2c|", 0);
+			else
+				printf("%3c", 0);
+		}
+		printf("\n");
+	}
+
+	// print x labels
+	printf("%12c", 0);
+	for (j=0; j<NALPHABET; j++) {
+		printf("%2c%c", 0, 97+j);
+	}
+}
+
+void set_buffer( char *b,
+					  unsigned int size ) {
+	
+	buffer = b;
+	buffer_size = size;
+}
+
+int get_chunk(char *chunk) {
+
+	int size = (buffer_size > CHUNKSIZE) ? CHUNKSIZE
+													 : buffer_size;
+
+	memcpy(chunk, buffer, size);
+	buffer += size;
+	buffer_size -= size;
+	
+	return size;
+}

### 2015-05-08 parallel_programming/ex3/histogram.h

@@ -1,0 +1,24 @@
+#ifndef HISTOGRAM_H_
+#define HISTOGRAM_H_
+
+#define NALPHABET 26	// number of letters in the alphabet
+#define CHUNKSIZE 8192	// number of chunks for a task
+
+typedef unsigned int histogram_t[NALPHABET];
+
+// produce histogram from blocks to histogram
+void get_histogram(unsigned int* histogram,
+				   	 unsigned int num_threads);
+
+
+// copies next chunk into *chunk and returns the size of the chunk
+int get_chunk(char *chunk);
+
+// internal: sets the global buffer and its size
+void set_buffer(char *buffer,
+					 unsigned int buffer_size);
+
+// dump histogram to stdout
+void print_histogram(histogram_t histogram);
+
+#endif /* HISTOGRAM_H_ */

### 2015-05-08 parallel_programming/ex3/histogram_seq.c

@@ -1,0 +1,23 @@
+#include <stdlib.h>
+
+#include "histogram.h"
+
+void get_histogram(unsigned int* histogram,
+		   			 unsigned int num_threads) {
+
+	int size = 0;
+	char *buffer = malloc(CHUNKSIZE);
+
+	while ((size = get_chunk(buffer)) > 0) {
+
+		// build histogram
+		for (int i=0; i<size; i++) {
+			if (buffer[i] >= 'a' && buffer[i] <= 'z')
+				histogram[buffer[i]-'a']++;
+			else if(buffer[i] >= 'A' && buffer[i] <= 'Z')
+				histogram[buffer[i]-'A']++;
+		}
+	}
+
+	free(buffer);
+}

### 2015-05-08 parallel_programming/ex3/main.c

@@ -1,0 +1,75 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <sys/stat.h>
+#include <time.h>
+#include <string.h>
+
+#include "histogram.h"
+
+int main(int argc, char* argv[]) {
+
+	// declarations
+	struct stat st;
+	unsigned int i, buffer_size;
+	struct timespec begin, end;
+	char *buffer;
+	histogram_t histogram = {0};
+	char *filename = "war_and_peace.txt";
+	long num_threads = 1, repetitions=0;
+ 
+	// argument handling
+	if (argc < 2 || argc > 4) {
+		printf("usage: %s filename [#threads] [#repetition]\n", argv[0]);
+		return 1;
+	}
+	if (argc >= 2)
+		filename = argv[1];
+	if (argc >= 3)
+		if ((num_threads = strtol(argv[2], NULL, 0)) == 0 || num_threads < 0) {
+    			fprintf(stderr, "#threads not valid!\n");
+    			return 1;
+  		}
+	if (argc >= 4)
+		if ((repetitions = strtol(argv[3], NULL, 0)) == 0 || repetitions < 0) {
+    	fprintf(stderr, "#repetition not valid!\n");
+    	return 1;
+  	}
+
+ 	printf("\nProcess %s by %ld thread(s) with %ld repetition(s)\n\n", 
+		filename, num_threads, repetitions);
+
+	// get file size and allocate array
+	stat(filename, &st);
+	buffer_size = (unsigned int)st.st_size * (repetitions+1);
+	buffer = (char*)malloc(buffer_size);
+
+	// open file and read all characters
+	FILE *fp = fopen(filename, "r");
+	if (fp == NULL) {
+		fprintf(stderr, "Could not open file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+	if (fread(buffer, 1, st.st_size, fp) == 0) {
+		fprintf(stderr, "Could not read from file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+
+	// [optional] do the repetition
+	for (i = 1; i <= repetitions; i++)
+		memcpy(buffer + i*st.st_size, buffer, st.st_size);
+
+	// set the buffer
+	set_buffer(buffer, buffer_size);
+
+	// build the histogram	
+	clock_gettime(CLOCK_MONOTONIC, &begin);
+	get_histogram(histogram, num_threads);
[…]

### 2015-05-08 parallel_programming/ex3/student/histogram_par.c

@@ -1,0 +1,190 @@
+#include <pthread.h>
+#include <stdlib.h>
+#include <string.h>
+
+#include "histogram.h"
+
+void build_histogram(unsigned int* histogram,
+		   			 char* buffer, unsigned int size) {
+
+  // build histogram
+  for (int i=0; i<size; i++) {
+    if (buffer[i] >= 'a' && buffer[i] <= 'z')
+      histogram[buffer[i]-'a']++;
+    else if(buffer[i] >= 'A' && buffer[i] <= 'Z')
+      histogram[buffer[i]-'A']++;
+  }
+}
+
+int incrementCyclic(int i, int limit) {
+  return (i + 1) % limit;
+}
+
+struct Data {
+  char data[CHUNKSIZE];
+  int size;
+};
+
+struct CyclicBuffer {
+  struct Data** buffer;
+  int size;
+  int read;
+  int write;
+  int occupied;
+  pthread_mutex_t mutex;
+  pthread_cond_t condRead;
+  pthread_cond_t condWrite;
+};
+
+void buffer_init(struct CyclicBuffer* buffer, int size) {
+  
+  buffer->buffer = malloc(size * sizeof(*buffer));
+  buffer->size = size;
+  buffer->read = 0;
+  buffer->write = 0;
+  buffer->occupied = 0;
+  pthread_mutex_init(&buffer->mutex, NULL);
+  pthread_cond_init(&buffer->condRead, NULL);
+  pthread_cond_init(&buffer->condWrite, NULL);
+}
+
+void buffer_destroy(struct CyclicBuffer* buffer) {
+  for(int i = 0; i < buffer->size; ++i) {
+    free(buffer->buffer[i]);
+  }
+  free(buffer->buffer);
+  pthread_mutex_destroy(&buffer->mutex);
+  pthread_cond_destroy(&buffer->condRead);
+  pthread_cond_destroy(&buffer->condWrite);
+  free(buffer);
+}
+
+struct Data* buffer_read(struct CyclicBuffer* buffer) {
+  pthread_mutex_lock(&buffer->mutex);
+
+  //wait for data to be read
+  if(buffer->write == buffer->read) {
+    pthread_cond_wait(&buffer->condRead, &buffer->mutex);
+  }
+
+  // read data and free a slot
+  struct Data* data = buffer->buffer[buffer->read];
+  buffer->buffer[buffer->read] = NULL;
+  buffer->read = incrementCyclic(buffer->read, buffer->size);
+  buffer->occupied--;
+
[…]

### 2015-05-08 parallel_programming/ex3/unit_test.c

@@ -1,0 +1,58 @@
+#include <stdio.h>
+#include <string.h>
+#include <stdlib.h>
+#include <sys/stat.h>
+
+#include "histogram.h"
+#include "histogram_ref.h"
+
+#define NTHREADS 4
+
+int main(int argc, char **argv)
+{
+	int failed = 0;
+	histogram_t histogram_ref = {0};
+	struct stat st;
+	char *buffer;
+	unsigned int buffer_size;
+	char *filename = "war_and_peace.txt";
+
+   // get file size and allocate array
+   stat(filename, &st);
+   buffer_size = (unsigned int)st.st_size;
+   buffer = (char*)malloc(buffer_size);
+
+	// open file and read all blocks
+	FILE *fp = fopen(filename, "r");
+	if (fp == NULL) {
+		fprintf(stderr, "Could not open file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+	if (fread(buffer, 1, st.st_size, fp) == 0) {
+		fprintf(stderr, "Could not read from file %s", filename);
+		exit(EXIT_FAILURE);
+	}
+
+	// set the buffer
+	set_buffer(buffer, buffer_size);
+
+	// build the histogram (sequential)
+	get_histogram_ref(histogram_ref);
+
+	// build the histogram (parallel)
+	set_buffer(buffer, buffer_size);
+	histogram_t histogram = {0};
+	get_histogram(histogram, NTHREADS);
+	for (int i = 0; i < NALPHABET; i++) {
+		if (histogram_ref[i] != histogram[i]) {
+			fprintf(stderr, "Computation with %d threads failed:\n", NTHREADS);
+			fprintf(stderr, "Wrong result for %c: %d (correct: %d)\n",
+					i+97, histogram[i], histogram_ref[i]);
+			failed = 1;
+			break;
+		}
+	}
+
+	free(buffer);
+	return failed;
+}

### 2015-05-09 parallel_programming/ex3/histogram_seq.c
ex3 works, but does not pass due to hellgrid checks

@@ -1,6 +1,10 @@
 #include <stdlib.h>
+#include <stdio.h>
 
 #include "histogram.h"
+
+
+int chunkNum = 0;
 
 void get_histogram(unsigned int* histogram,
 		   			 unsigned int num_threads) {
@@ -9,7 +13,7 @@ void get_histogram(unsigned int* histogram,
 	char *buffer = malloc(CHUNKSIZE);
 
 	while ((size = get_chunk(buffer)) > 0) {
-
+  printf("%d\n", chunkNum++);
 		// build histogram
 		for (int i=0; i<size; i++) {
 			if (buffer[i] >= 'a' && buffer[i] <= 'z')

### 2015-05-09 parallel_programming/ex3/student/histogram_par.c

@@ -1,8 +1,15 @@
 #include <pthread.h>
 #include <stdlib.h>
 #include <string.h>
+#include <stdio.h>
+#include <unistd.h>
+#include <errno.h>
+#include <assert.h>
 
 #include "histogram.h"
+
+
+#define WAIT_TIME (int)9e8
 
 void build_histogram(unsigned int* histogram,
 		   			 char* buffer, unsigned int size) {
@@ -61,21 +68,35 @@ void buffer_destroy(struct CyclicBuffer* buffer) {
 
 struct Data* buffer_read(struct CyclicBuffer* buffer) {
   pthread_mutex_lock(&buffer->mutex);
+  int waitStatus = 0;
+  struct timespec timeSpec;
+  struct Data* data = NULL;
 
   //wait for data to be read
-  if(buffer->write == buffer->read) {
-    pthread_cond_wait(&buffer->condRead, &buffer->mutex);
+  while(buffer->occupied <= 0) {
+    timeSpec.tv_sec = time(0);
+    timeSpec.tv_nsec = WAIT_TIME;
+
+    waitStatus = pthread_cond_timedwait(&buffer->condRead, &buffer->mutex, &timeSpec);
+    if(waitStatus == ETIMEDOUT) {
+      break;
+    }
   }
 
-  // read data and free a slot
-  struct Data* data = buffer->buffer[buffer->read];
-  buffer->buffer[buffer->read] = NULL;
-  buffer->read = incrementCyclic(buffer->read, buffer->size);
-  buffer->occupied--;
 
-  // signal that a slot has just been freed
-  pthread_cond_signal(&buffer->condWrite);
+  if(waitStatus != ETIMEDOUT) {
+    assert(buffer->occupied > 0);
 
+    // read data and free a slot
+    data = buffer->buffer[buffer->read];
+
+    buffer->buffer[buffer->read] = NULL;
+    buffer->read = incrementCyclic(buffer->read, buffer->size);
+    buffer->occupied--;
+
+    // signal that a slot has just been freed
+    pthread_cond_signal(&buffer->condWrite);
+  }
   pthread_mutex_unlock(&buffer->mutex);
   return data;
 }
@@ -84,13 +105,15 @@ void buffer_write(struct CyclicBuffer* buffer, struct Data* data) {
   pthread_mutex_lock(&buffer->mutex);
 
   // wait for a slot to write
-  if(buffer->occupied == buffer->size) {
+  while(buffer->occupied >= buffer->size) {
[…]

### 2015-05-10 parallel_programming/ex3/student/histogram_par.c
working ex3

@@ -137,7 +137,7 @@ struct ProducerData {
 void* producer(void* arg) {
 
   struct ProducerData* data = (struct ProducerData*)arg;
-  *data->keepWorking = 1;
+  //*data->keepWorking = 1;
 
   struct Data* chunk = malloc(sizeof(*chunk)); 
   chunk->size = get_chunk(chunk->data);
@@ -149,13 +149,16 @@ void* producer(void* arg) {
     chunk->size = get_chunk(chunk->data);
   }
 
+  pthread_mutex_lock(&data->buffer->mutex);
   while(data->buffer->occupied > 0) {
+    pthread_cond_wait(&data->buffer->condWrite, &data->buffer->mutex);
     pthread_cond_signal(&data->buffer->condRead);
   };
 
-  free(chunk);
   *data->keepWorking = 0;
+  pthread_mutex_unlock(&data->buffer->mutex);
 
+  free(chunk);
   return NULL;
 }
 
@@ -164,14 +167,19 @@ void* consumer(void* arg) {
   struct ConsumerData* data = (struct ConsumerData*)arg;
   data->histogram = calloc(NALPHABET, sizeof(*data->histogram));
 
+  pthread_mutex_lock(&data->buffer->mutex);
+  //asm("" : "=m" (data->keepWorking));
   while(*data->keepWorking) {
-
+    pthread_mutex_unlock(&data->buffer->mutex);
     struct Data* chunk = buffer_read(data->buffer);
     if(chunk != NULL) {
       build_histogram(data->histogram, chunk->data, chunk->size);
       free(chunk);
     }
+    pthread_mutex_lock(&data->buffer->mutex);
+    //asm("" : "=m" (data->keepWorking));
   }
+  pthread_mutex_unlock(&data->buffer->mutex);
 
   return NULL;
 }

### 2015-05-15 MLCV/ex3/report/report.tex
ex3 boosting and gaussian processes

\documentclass[journal, a4paper]{IEEEtran}

\usepackage{graphicx}
\usepackage{url}
\usepackage{amsmath} 
\usepackage{subcaption}

\begin{document}

	\title{Assignment III\\Boosting and Gaussian Processes}
	\author{Adam Kosiorek}	
	\markboth{Machine Learning for Applications in Computer Vision, Institute of Computer Vision, Technische Universit\'at M\"unchen}{}
	\maketitle

\section{Objective}
    \PARstart{T}{he} aim of the third assignment was to get insight into workings of Boosting and Gaussian Process (GP) algorithms by comparing performance and estimating uncertainty of several variants of the algorithms.

\section{Dataset}
    Boosting algorithms were compared on the MNIST dataset is comprised of 60000 training and 10000 testing samples \cite{MNIST}. Each sample is a 28x28 pixel black and white image of a single centered digit. It is widely used as a benchmark for comparing machine learning algorithms.

    GPs were evaluated on an artificial 2-class dataset. It was randomly generated from a gaussian distribution with expected values and covariance matrices as denoted by $\bar{x}$ and $\Sigma$ in equations \ref{eq:mean} and \ref{eq:cov}.

    \begin{equation}
     \bar{x}_{positive} = \begin{pmatrix} 0.75 \\ 0 \end{pmatrix} \;\; 
     \bar{x}_{negative} = \begin{pmatrix} -0.75 \\ 0 \end{pmatrix}
     \label{eq:mean}
    \end{equation}
    \begin{equation}
     \Sigma_{positive} = \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix} \;\; 
     \Sigma_{negative} = \begin{pmatrix} 1 & 0.95 \\ 0.95 & 1 \end{pmatrix}
     \label{eq:cov}
    \end{equation}

    \begin{figure}[h]      
      \begin{subfigure}{.22\textwidth}
	\centering
	\includegraphics[width=.9\linewidth]{GP_trainset}
	\caption{}
	\label{fig:GP_a}
      \end{subfigure}
      \begin{subfigure}{.22\textwidth}
	\centering
	\includegraphics[width=.9\linewidth]{GP_testset}
	\caption{}
	\label{fig:GP_b}
      \end{subfigure}

      \caption{(a) trainset and (b) testset for GPs. \textbf{+} denotes positive samples and \textbf{o} denotes negative samples.}
      \label{fig:GP_data}      
    \end{figure}

\section{Methods}
    Boosting is a means of combining weak classifiers, that is, classifiers only slightly better than random classifiers, in order to create a strong classifier. It does so by sequentially minimizing an exponential loss function by adding parameterized basis functions to the growing ensemble of weak classifiers. A basis function is any weak classifier capable of handling weighted data, \textit{e.g.} decision trees or decision stumps.

    Gaussian Process is a collection of random variables, any finite number of which have a joint Gaussian distribution. Since the number of variables can be inifnite, a Gaussian Distribution is a distribution over functions. Therefore, if we have a mean function and covariance function, we can sample from the Gaussian Process. The inifite number of variables can be handled due to marginalization property, which allows operation on finite vectors only.

\section {Experiments and Results}
  All experiments were carried out on raw data (without any preprocessing) on a notebook with Intel i7-2670QM quad core CPU and 8Gb RAM in Matlab using Matlab Boosting Framework \cite{MATLAB_BOOST} and Gaussian Processes for Machine Learning (GPML) toolboxes \cite{GPML}. Assignment code is available on GitHub \cite{code}.

  \subsection{Boosting}
      I compare AdaBoost, RUSBoost, LPBoost and TotalBoost algorithms trained and tested on small part of the dataset (1000 training and testing samples) and full dataset with different ensemble sizes. The ensemble size is a hyperparameter for the first two algorithms and it is automatically chosen for the last two, which are self-terminating. LPBoost terminated with 128 and 8 classifiers, while TotalBoost terminated with 28 and 51 classifiers when trained on 1000 samples and on the whole dataset respectively. AdaBoost significantly outperforms all other algorithms. According to \cite{MATLAB_BOOST} TotalBoost and LPBoost are ill-suited for problems with large number of observations, which is the case here, while RUSBoost is especially effective for unbalanced data classification problems. Obviously, increased size of the training set positively impact classification accuracy.

    \begin{figure}[h]    
      \centering
      \includegraphics[width=\linewidth]{boost_plot_small}
      \caption{Classification accuracy of Boosting classifiers trained and tested on 1000 samples.}
      \label{fig:boost_accuracy_small}
    \end{figure}

    \begin{figure}[h]    
      \centering
      \includegraphics[width=\linewidth]{boost_plot_big}
      \caption{Classification accuracy of Boosting classifiers trained and tested on the whole dataset.}
      \label{fig:boost_accuracy_small}
    \end{figure}

  \subsection{Gaussian Processes}
      GPs output log probabilties instead of class labels. To compute accuracy and uncertainty I have taken the exponent of the log probability and thresholded it with a value of $0.5$, taking class label $l=1$ for probabilities greater than the threshold value and $l=-1$ for probabilities smaller than the threshold value. Table \ref{tab:GP_acc} shows accuracy of GPs with different covariance, likelihood and inference functions while figures \ref{fig:GP_true} and \ref{fig:GP_false} depict uncertainty histogram for one GP variants with the highest accuracy, namely with linear isotropic (LINiso) covariance function.

    \begin{table}[h]
      \begin{center}
	\caption{GP accuracy}
	\label{tab:GP_acc}
	\begin{tabular}{c|c|c|c}
	  covariance & likelihood  & inference  & accuracy \\ \hline
	  SEard & Erf & EP & 0.62\\ \hline
	  SEiso & Erf & EP & 0.63\\ \hline
	  LINard & Erf & EP & 0.68\\ \hline
	  LINard & Gaus & Exact & 0.55\\ \hline
	  LINard & Logistic & EP & 0.68\\ \hline
	  LINiso & Erf & EP & 0.68
	\end{tabular}
      \end{center}
    \end{table}

    \begin{figure}[h]
      \centering
      \includegraphics[width=.8\linewidth]{covLINiso_gaussian_true}
      \caption{GP classification uncertainty histogram for correctly classified test samples.}
      \label{fig:GP_true}
    \end{figure}

    \begin{figure}[h]
      \centering
      \includegraphics[width=.8\linewidth]{covLINiso_gaussian_false}
      \caption{GP classification uncertainty histogram for incorrectly classified test samples.}
      \label{fig:GP_false}
    \end{figure}

    I believe that uncertainty value should be a positive real number between 0 and 1. According to the formula given in the assignment, however, I obtained negative values. Also, since

  \subsection{Discussion}

%   \begin{figure}[h]    
%     \centering
%     \includegraphics[width=.8\linewidth]{}
%     \caption{Images for Classification.}
%     \label{fig:imgs}
%   \end{figure}
% 
% 
%   \begin{table}[h]
%     \begin{center}
%     \caption{Classification Results}
%     \label{tab:results}
%     \begin{tabular}{c|c|c|c|c}
%       \# & \#class  & class name  & probability & entropy \\ \hline
%       a & 428 & barrow, garden cart, lawn cart, \ldots & 0.07 & 5.22\\ \hline
%       b & 311 & grasshopper, hopper & 0.14 & 3.72\\ \hline
%       c & 403 & aircraft carrier, carrier, flattop, \ldots & 0.13 & 4.67\\ \hline
%       d & 620 & laptop, laptop computer & 0.12 & 4.63\\ \hline
%       e & 673 & mouse, computer mouse & 0.41 & 2.75\\ \hline
%       f & 947 & mushroom & 0.76 & 1.17\\ \hline
%       g & 834 & suit, suit of clothes & 0.62 & 1.09\\ \hline
%       h & 980 & volcano & 0.36 & 3.32
%     \end{tabular}
%     \end{center}
%   \end{table}

\begin{thebibliography}{5}

    \bibitem{code}
    Assignment Code: \url{https://github.com/akosiorek/CSE/tree/master/MLCV/ex3/}

    \bibitem{MNIST}
    Y.~LeCun, L.~Bottou, Y.~Bengio, and P.~Haffner. ``Gradient-based learning applied to document recognition.'' Proceedings of the IEEE, 86(11):2278-2324, November 1998.

    \bibitem{GPML}
    C.~E.~Rasmussen and H.~Nickisch, ``Gaussian processes for machine learning (GPML) toolbox'', The Journal of Machine Learning Research, 2010, 11, 3011-3015.

    \bibitem{MATLAB_BOOST}
    Matlab Boosting Framework: \url{http://www.mathworks.com/help/stats/ensemble-methods.html}

\end{thebibliography}

% Your document ends here!
\end{document}

### 2015-05-15 MLCV/ex3/boost.m

@@ -1,0 +1,65 @@
+function boost()
+    clc; clear all; close all;    
+    
+    fraction = 1;
+    [trainSamples, trainLabels] = loadMnistData('train', fraction);
+    [testSamples, testLabels] = loadMnistData('test', fraction);
+    
+    lpBoost = fitensemble(trainSamples, trainLabels, 'LPBoost', 500, 'Tree', 'nprint', 1) 
+    save('lp.mat', 'lpBoost')
+    totalBoost = fitensemble(trainSamples, trainLabels, 'TotalBoost', 500, 'Tree', 'nprint', 1)
+    save('total.mat', 'totalBoost')
+    
+    iters = 100;
+    adaBoost = fitensemble(trainSamples, trainLabels, 'AdaBoostM2', iters, 'Tree', 'nprint', 1)  
+    save('ada.mat', 'adaBoost')
+    rusBoost = fitensemble(trainSamples, trainLabels, 'RUSBoost', 100, 'Tree', 'nprint', 1)
+    save('rus.mat', 'rusBoost')
+    
+    accuracy = zeros(10, 3);    
+    accuracy(1, 1) = adaBoost.NTrained;
+    accuracy(1, 2) = checkModel(adaBoost, testSamples, testLabels)     
+    accuracy(1, 3) = checkModel(rusBoost, testSamples, testLabels)
+   
+    for i=2:10
+        adaBoost  = adaBoost.resume(iters, 'nprint', 1)
+        save('ada.mat', 'adaBoost')
+        accuracy(i, 1) = adaBoost.NTrained;
+        accuracy(i, 2) = checkModel(adaBoost, testSamples, testLabels)
+        
+        rusBoost = rusBoost.resume(iters, 'nprint', 1)
+        save('rus.mat', 'rusBoost')
+        accuracy(i, 3) = checkModel(rusBoost, testSamples, testLabels)
+    end
+    
+
+    save('accuracy.mat', 'accuracy');
+    
+    figure(1)
+    grid on
+    hold on
+    xlabel('NLearners');
+    ylabel('Accuracy');
+    
+    plot(accuracy(:, 1), accuracy(:, 2), 'b*')
+    plot(accuracy(:, 1), accuracy(:, 3), 'r*')    
+    plot(lpBoost.NTrained, checkModel(lpBoost, testSamples, testLabels), 'g*')
[…]

### 2015-05-15 MLCV/ex3/boost_uncertainty.m

@@ -1,0 +1,4 @@
+function boost_uncertainty()
+   
+    
+end

### 2015-05-15 MLCV/ex3/gaussian.m

@@ -1,0 +1,127 @@
+function gaussian()
+    
+    clear all, close all, clc;
+
+    n1 = 80;        % number of data points from each class
+    n2 = 40;        
+    S1 = eye(2);    % the two covariance matrices
+    S2 = [1 0.95; 0.95 1];      
+    m1 = [0.75; 0];     % the two means
+    m2 = [-0.75; 0];    
+      
+    % train data
+    seed = 0.2;
+    [x1, y1] = createData(n1, 1, m1, S1, seed);
+    [x2, y2] = createData(n2, -1, m2, S2, seed);
+    trainData = [x1 x2]'; 
+    trainLabels = [y1; y2];
+    
+    figure(1)
+    hold on
+    grid on
+    plot(x1(1, :), x1(2, :), 'r+', 'LineWidth', 2)
+    plot(x2(1, :), x2(2, :), 'bo', 'LineWidth', 2)
+    xlabel('x')
+    ylabel('y')
+    
+    % test data
+    seed = 0.4;
+    [xs1, ys1] = createData(n1, 1, m1, S1, seed);
+    [xs2, ys2] = createData(n2, -1, m2, S2, seed);
+    testData = [xs1 xs2]'; 
+    testLabels = [ys1; ys2];
+    
+    print -depsc 'GP_trainset'
+    
+    figure(2)
+    hold on
+    grid on
+    plot(xs1(1, :), xs1(2, :), 'r+', 'LineWidth', 2)
+    plot(xs2(1, :), xs2(2, :), 'bo', 'LineWidth', 2)
+    xlabel('x')
+    ylabel('y')
+    
+    print -depsc 'GP_testset'
+    
+ 
+    
+    name = 'covLINard_likLogistic_infEP'
+    meanfunc = @meanConst; 
+    hyp.mean = 0;
+    covfunc = @covLINard;   
+    hyp.cov = log([1 1]);
+%     covfunc = @covSEard;   
+%     hyp.cov = log([1 1 1]);
+    likfunc = @likLogistic;
+%     hyp.lik = log([1]);
+    infFunc = @infEP;
+
+    hyp = minimize(hyp, @gp, -40, @infEP, meanfunc, covfunc, likfunc, trainData, trainLabels);
+    [a b c d lp] = gp(hyp, infFunc, meanfunc, covfunc, likfunc, trainData, trainLabels, testData, testLabels);
+
+    probs = getProbs(lp);
+    labels = toLabels(probs);
+    accuracy = checkModel(labels, testLabels) 
+    
[…]

### 2015-05-15 MLCV/ex3/loadMnistData.m

@@ -1,0 +1,28 @@
+function [samples, labels] = loadMnistData(name, varargin)    
+    dataFolder = '../data/mnist/';
+    
+    if strcmp(name, 'train')
+        labels = loadMNISTLabels(strcat(dataFolder, 'train-labels-idx1-ubyte'));
+        samples = loadMNISTImages(strcat(dataFolder, 'train-images-idx3-ubyte'));
+    elseif strcmp(name, 'test')
+        labels = loadMNISTLabels(strcat(dataFolder, 't10k-labels-idx1-ubyte'));
+        samples = loadMNISTImages(strcat(dataFolder, 't10k-images-idx3-ubyte'));
+    else
+        assert(0, 'Name must be "train" or "test"')
+    end
+    samples = samples';
+    N = size(labels, 1);
+    
+    if nargin == 2
+        fraction = varargin{1};
+        if fraction <= 1
+            N = N * fraction;
+        else
+            N = fraction;
+        end
+        labels = labels(1:N, :);
+        samples = samples(1:N, :);
+    end
+    
+    fprintf('Loaded %d samples and labels from %s set\n', N, name);
+end

### 2015-05-15 parallel_programming/ex4/main.c
ex4 works

@@ -1,0 +1,147 @@
+#include <unistd.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <complex.h>
+#include <omp.h>
+#include <time.h>
+
+#include "mandelbrot_set.h"
+
+int main(int argc, char **argv)
+{
+	struct timespec begin, end;
+	int num_threads = 1;
+	int max_iter = 1000; // sizeof(colors);
+	int x_resolution = 480;
+	int y_resolution = 380;
+	int palette_shift = 0;
+	double view_x0 = -2.00;
+	double view_x1 = +0.50;
+	double view_y0 = -1.25;
+	double view_y1 = +1.25;
+	char file_name[256] = "mandelbrot.ppm";
+	int no_output = 0;
+
+	double x_stepsize;
+	double y_stepsize;
+
+	int c;
+	while ((c = getopt(argc, argv, "t:p:i:r:v:n:f:")) != -1)
+	{
+		switch (c)
+		{
+		case 't':
+			if (sscanf(optarg, "%d", &num_threads) != 1)
+				goto error;
+			break;
+
+		case 'p':
+			if (sscanf(optarg, "%d", &palette_shift) != 1)
+				goto error;
+			break;
+
+		case 'i':
+			if (sscanf(optarg, "%d", &max_iter) != 1)
+				goto error;
+			break;
+
+		case 'r':
+			if (sscanf(optarg, "%dx%d", &x_resolution, &y_resolution) != 2)
+				goto error;
+			break;
+
+		case 'v':
+			if (sscanf(optarg, "[%lf,%lf]x[%lf,%lf]", &view_x0, &view_x1,
+			           &view_y0, &view_y1) != 4)
+				goto error;
+			break;
+		case 'n':
+			if (sscanf(optarg, "%d", &no_output) != 1)
+				goto error;
+			break;
+		case 'f':
+			strncpy(file_name, optarg, sizeof(file_name));
+			break;
+
+		case '?':
+error: printf(
+			    "Usage:\n"
+			    "-t \t number of threads used in computation\n"
+			    "-i \t maximum number of iterations per pixel\n"
+			    "-r \t image resolution to be computed\n"
+			    "-v \t view frame that should be computed\n"
+			    "-p \t shift palette to change colors\n"
+			    "-f \t output file name\n"
[…]

### 2015-05-15 parallel_programming/ex4/student/mandelbrot_set_par.c

@@ -1,0 +1,59 @@
+#include <unistd.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <complex.h>
+#include <omp.h>
+
+#include "mandelbrot_set.h"
+
+void mandelbrot_draw(int x_resolution, int y_resolution, int max_iter,
+                     double view_x0, double view_x1, double view_y0, double view_y1,
+                     double x_stepsize, double y_stepsize,
+                     int palette_shift, unsigned char (*img)[x_resolution][3],
+							int num_threads) {
+	double y;
+	double x;
+
+	complex double Z;
+	complex double C;
+
+	int k;
+
+  max_iter += palette_shift;
+
+  #pragma omp parallel for   \
+    num_threads(num_threads) \
+    private(x, y, Z, C, k)   \
+    schedule(static, 3)
+	for (int i = 0; i < y_resolution; i++)
+	{
+		for (int j = 0; j < x_resolution; j++)
+		{
+			y = view_y1 - i * y_stepsize;
+			x = view_x0 + j * x_stepsize;
+
+			Z = 0 + 0 * I;
+			C = x + y * I;
+
+			k = palette_shift;
+
+			do
+			{
+				Z = Z * Z + C;
+				k++;
+			} while (cabs(Z) < 2 && k < max_iter);
+
+			if (k == max_iter)
+			{
+				memcpy(img[i][j], "\0\0\0", 3);
+			}
+			else
+			{
+				int index = (k + palette_shift)
+				            % (sizeof(colors) / sizeof(colors[0]));
+				memcpy(img[i][j], colors[index], 3);
+			}
+		}
+	}
+}

### 2015-05-20 MLCV/ex3/report/report.tex
finish ex3 boosting and GPs

@@
 \section{Dataset}
-    Boosting algorithms were compared on the MNIST dataset is comprised of 60000 training and 10000 testing samples \cite{MNIST}.
+    Boosting algorithms were compared on the MNIST dataset, which is comprised of 60000 training and 10000 testing samples \cite{MNIST}.
 Each sample is a 28x28 pixel black and white image of a single centered digit.
@@
     GPs were evaluated on an artificial 2-class dataset.
-It was randomly generated from a gaussian distribution with expected values and covariance matrices as denoted by $\bar{x}$ and $\Sigma$ in equations \ref{eq:mean} and \ref{eq:cov}.
+It was randomly generated from a Gaussian distribution with expected values and covariance matrices as denoted by $\bar{x}$ and $\Sigma$ in equations \ref{eq:mean} and \ref{eq:cov}.
 
@@
 
-    \begin{figure}[h]      
+    \begin{figure}[ht]      
       \begin{subfigure}{.22\textwidth}
@@
 It does so by sequentially minimizing an exponential loss function by adding parameterized basis functions to the growing ensemble of weak classifiers.
-A basis function is any weak classifier capable of handling weighted data, \textit{e.g.} decision trees or decision stumps.
+Here, a basis function means any weak classifier capable of handling weighted data, \textit{e.g.} decision trees or decision stumps.
 
     Gaussian Process is a collection of random variables, any finite number of which have a joint Gaussian distribution.
-Since the number of variables can be inifnite, a Gaussian Distribution is a distribution over functions.
+Since the number of variables can be infinite, a Gaussian Distribution is a distribution over functions.
 Therefore, if we have a mean function and covariance function, we can sample from the Gaussian Process.
-The inifite number of variables can be handled due to marginalization property, which allows operation on finite vectors only.
+The infinite number of variables can be handled due to marginalization property, which allows operation on finite vectors only.
 
 \section {Experiments and Results}
-  All experiments were carried out on raw data (without any preprocessing) on a notebook with Intel i7-2670QM quad core CPU and 8Gb RAM in Matlab using Matlab Boosting Framework \cite{MATLAB_BOOST} and Gaussian Processes for Machine Learning (GPML) toolboxes \cite{GPML}.
+  All experiments were carried out on raw data (without any preprocessing) on a notebook with Intel i7-2670QM quad core CPU and 8GB RAM in Matlab using Matlab Boosting Framework \cite{MATLAB_BOOST} and Gaussian Processes for Machine Learning (GPML) toolboxes \cite{GPML}.
 Assignment code is available on GitHub \cite{code}.
@@
       I compare AdaBoost, RUSBoost, LPBoost and TotalBoost algorithms trained and tested on small part of the dataset (1000 training and testing samples) and full dataset with different ensemble sizes.
-The ensemble size is a hyperparameter for the first two algorithms and it is automatically chosen for the last two, which are self-terminating.
-LPBoost terminated with 128 and 8 classifiers, while TotalBoost terminated with 28 and 51 classifiers when trained on 1000 samples and on the whole dataset respectively.
+The ensemble size is a hyper-parameter for the first two algorithms and it is automatically chosen for the last two due to the self-termination property.
+LPBoost terminated with 128 and 6 classifiers, while TotalBoost terminated with 28 and 51 classifiers when trained on 1000 samples and on the whole dataset respectively.
 AdaBoost significantly outperforms all other algorithms.
 According to \cite{MATLAB_BOOST} TotalBoost and LPBoost are ill-suited for problems with large number of observations, which is the case here, while RUSBoost is especially effective for unbalanced data classification problems.
-Obviously, increased size of the training set positively impact classification accuracy.
+Obviously, an increased size of the training set as well as an increased number of weak classifiers positively impacts the classification accuracy.
 
-    \begin{figure}[h]    
+    \begin{figure}[ht]    
       \centering
@@
 
-    \begin{figure}[h]    
+    \begin{figure}[ht]    
       \centering
@@
       \caption{Classification accuracy of Boosting classifiers trained and tested on the whole dataset.}
-      \label{fig:boost_accuracy_small}
+      \label{fig:boost_accuracy_big}
+    \end{figure}
+
+    Figures \ref{fig:boost_true} and \ref{fig:boost_false} show classification uncertainty histograms for AdaBoost with 1000 weak classifiers and 10000 test samples.
+Both histograms resemble a Gaussian distribution.
+Boosting classifiers ouput scores instead of probabilities in Matlab and they had to be normalized before entropy could be computed.
+Since it is a multi-class classification problem, I used the following formula for computing classification entropy:
+
+    \begin{equation}
+      h = -\Sigma_{c=1}^C p_c * log(p_c)
+    \end{equation}
+
+     \begin{figure}[ht]
+      \centering
+      \includegraphics[width=.8\linewidth]{ada_correct}
+      \caption{AdaBoost classification uncertainty histogram for correctly classified test samples.}
+      \label{fig:boost_true}
+    \end{figure}
+
+    \begin{figure}[ht]
+      \centering
+      \includegraphics[width=.8\linewidth]{ada_incorrect}
+      \caption{AdaBoost classification uncertainty histogram for incorrectly classified test samples.}
+      \label{fig:boost_false}
     \end{figure}
@@
   \subsection{Gaussian Processes}
-      GPs output log probabilties instead of class labels.
+      GPs output log probabilities instead of class labels.
 To compute accuracy and uncertainty I have taken the exponent of the log probability and thresholded it with a value of $0.5$, taking class label $l=1$ for probabilities greater than the threshold value and $l=-1$ for probabilities smaller than the threshold value.
-Table \ref{tab:GP_acc} shows accuracy of GPs with different covariance, likelihood and inference functions while figures \ref{fig:GP_true} and \ref{fig:GP_false} depict uncertainty histogram for one GP variants with the highest accuracy, namely with linear isotropic (LINiso) covariance function.
+Table \ref{tab:GP_acc} shows accuracy of GPs with different covariance, likelihood and inference functions, while figures \ref{fig:GP_true} and \ref{fig:GP_false} depict uncertainty histograms for one of the GP variants with the highest accuracy, namely with linear isotropic (LINiso) covariance function.
 
-    \begin{table}[h]
+    \begin{table}[ht]
       \begin{center}
@@
 
-    \begin{figure}[h]
+    \begin{figure}[ht]
       \centering
-      \includegraphics[width=.8\linewidth]{covLINiso_gaussian_true}
+      \includegraphics[width=.8\linewidth]{covLINiso_likErf_infEP_true}
       \caption{GP classification uncertainty histogram for correctly classified test samples.}
@@
 
-    \begin{figure}[h]
+    \begin{figure}[ht]
       \centering
-      \includegraphics[width=.8\linewidth]{covLINiso_gaussian_false}
+      \includegraphics[width=.8\linewidth]{covLINiso_likErf_infEP_false}
       \caption{GP classification uncertainty histogram for incorrectly classified test samples.}
@@
     \end{figure}    
-
-    I believe that uncertainty value should be a positive real number between 0 and 1. According to the formula given in the assignment, however, I obtained negative values. Also, since
-
   \subsection{Discussion}

### 2015-05-20 MLCV/ex3/boost_uncertainty.m

@@ -1,4 +1,33 @@
 function boost_uncertainty()
+    clc; clear all; close all;
    
+    load ada.mat
+%     load rus.mat
+%     load lp.mat
+%     load total.mat
+    [testSamples, testLabels] = loadMnistData('test', 0.1);
     
+    
+    [labels, scores] = predict(adaBoost, testSamples);
+    
+    scores = scores ./ repmat(sum(scores, 2), 1, size(scores, 2));
+    
+    correctScores = scores(labels == testLabels, :);
+    incorrectScores = scores(labels ~= testLabels, :);
+    
+    correctUncertainty = uncertainty(correctScores);
+    incorrectUncertainty = uncertainty(incorrectScores);
+    
+    
+    plotHistogram(correctUncertainty, 10, 'ada_correct');
+    plotHistogram(incorrectUncertainty, 10, 'ada_incorrect');
+    
+    
+    
+    
+end
+
+
+function u = uncertainty(scores)     
+    u = -sum(scores .* log(scores), 2)    
 end

### 2015-05-20 MLCV/ex3/gaussian.m

@@ -45,14 +45,14 @@ function gaussian()
     
  
     
-    name = 'covLINard_likLogistic_infEP'
+    name = 'covLINiso_likErf_infEP'
     meanfunc = @meanConst; 
     hyp.mean = 0;
     covfunc = @covLINard;   
     hyp.cov = log([1 1]);
 %     covfunc = @covSEard;   
 %     hyp.cov = log([1 1 1]);
-    likfunc = @likLogistic;
+    likfunc = @likErf;
 %     hyp.lik = log([1]);
     infFunc = @infEP;
 
@@ -70,9 +70,10 @@ function gaussian()
     trueUncertainty = uncertainty(trueProbs);
     falseUncertainty = uncertainty(falseProbs);
     
-    
-    plotHistogram(trueUncertainty, 10, strcat(name, '_gaussian_true'));
-    plotHistogram(falseUncertainty, 10, strcat(name, '_gaussian_false'));
+    figure(3)
+    plotHistogram(trueUncertainty, 10, strcat(name, '_true'));
+    figure(4)
+    plotHistogram(falseUncertainty, 10, strcat(name, '_false'));
 end
 
 
@@ -87,7 +88,7 @@ function acc = checkModel(labels, groundtruth)
 end
 
 function u = uncertainty(p)
-    u = -(p .* log(p) + (1 - p .* log(1 - p)));
+    u = -(p .* log(p) + (1 - p) .* log(1 - p));
 end
 
 function probs = getProbs(logProbs)
@@ -98,30 +99,4 @@ function labels = toLabels(probs)
     labels = probs;
     labels(labels < 0.5) = -1;
     labels(labels > 0.5) = 1;
-end
-
-function plotHistogram(data, bins, name)
-   
-    minValue = min(data);
-    maxValue = max(data);
-    
-    x = linspace(minValue, maxValue, bins + 1);
-    xx = zeros(bins, 1);
-    y = zeros(bins, 1);
-    
-    for i = 1:bins
-        y(i) = nnz(data >= x(i) & data < x(i + 1)); 
-        xx(i) = (x(i) + x(i+1)) / 2;
-    end
-    
-    x = round(100*x)/100;
-    figure(100)    
-    plot(xx, y, '*');  
-    xlabel('Uncertainty')
-    ylabel('Number of data points')
-    grid on
-    set(gca, 'XTick', x);
-    axis([minValue maxValue 0 max(y)+1])
-    
-    print('-depsc', name);
 end

### 2015-05-20 MLCV/ex3/plotHistogram.m

@@ -1,0 +1,23 @@
+function plotHistogram(data, bins, name)   
+    minValue = min(data);
+    maxValue = max(data);
+    
+    x = linspace(minValue, maxValue, bins + 1);
+    xx = zeros(bins, 1);
+    y = zeros(bins, 1);
+    
+    for i = 1:bins
+        y(i) = nnz(data >= x(i) & data < x(i + 1)); 
+        xx(i) = (x(i) + x(i+1)) / 2;
+    end
+    
+%     x = round(1000*x)/1000;
+    plot(xx, y, '*');  
+    xlabel('Uncertainty')
+    ylabel('Number of data points')
+    grid on
+    set(gca, 'XTick', x);
+    axis([minValue maxValue 0 max(y)+1])
+    
+    print('-depsc', name);
+end

### 2015-05-20 parallel_programming/ex5/main.c
add assignment 5 and 6

@@ -1,0 +1,67 @@
+#include <unistd.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <string.h>
+#include "quicksort.h"
+#include "helper.h"
+
+#define NTHREADS 4
+#define ELEMS 1000000
+
+int main(int argc, char **argv)
+{
+	int elements = ELEMS;
+	int num_threads = NTHREADS;
+	int dump_array = 0;
+	int c;
+	while ((c = getopt(argc, argv, "t:e:d:")) != -1)
+	{
+		switch (c)
+		{
+		case 't':
+			if (sscanf(optarg, "%d", &num_threads) != 1)
+				goto error;
+			break;
+
+		case 'e':
+			if (sscanf(optarg, "%d", &elements) != 1)
+				goto error;
+			break;
+
+		case 'd':
+			if (sscanf(optarg, "%d", &dump_array) != 1)
+				goto error;
+			break;
+
+		case '?':
+			error: printf(
+			    "Usage:\n"
+			    "-t \t number of threads used in computation\n"
+			    "-e \t number of elements to be sorted using quicksort\n"
+			    "-d \t dump array option (0 or 1)\n"
+			    "\n"
+			    "Example:\n"
+			    "%s -t 4 -e 10000000 -d 1\n", argv[0]);
+			exit(EXIT_FAILURE);
+			break;
+		}
+	}
+
+	struct timespec start, stop;
+	int *a = random_int_array(elements, elements/2, 13);
+
+	if (dump_array == 1) print_array(a, elements);
+
+	clock_gettime(CLOCK_MONOTONIC, &start);
+
+	quicksort(a, 0, elements-1, num_threads);
+
+	clock_gettime(CLOCK_MONOTONIC, &stop);
+
+	if (dump_array == 1) print_array(a, elements);
+
+	printf("\nTime: %lf seconds\n", time_diff(&start, &stop, NULL));
+
+	return 0;
+}

### 2015-05-20 parallel_programming/ex5/quicksort_seq.c

@@ -1,0 +1,36 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <string.h>
+#include "quicksort.h"
+#include "helper.h"
+
+void quicksort(int *a, int left, int right, int num_threads)
+{
+	if(left < right)
+	{
+		int x = left, y = (left+right)/2, z =right;
+		int pivotIdx = (a[x] <= a[y])
+		    ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
+		    : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
+
+	  int pivotVal = a[pivotIdx];
+	  swap(a + pivotIdx, a + right);
+
+	  int swapIdx = left;
+
+	  for(int i=left; i < right; i++)
+	  {
+		  if(a[i] <= pivotVal)
+		  {
+			  swap(a + swapIdx, a + i);
+			  swapIdx++;
+		  }
+	  }
+	  swap(a + swapIdx, a + right);
+
+	  quicksort(a, left, swapIdx - 1, num_threads);
+	  quicksort(a, swapIdx + 1, right, num_threads);
+
+	}
+}

### 2015-05-20 parallel_programming/ex5/student/quicksort_par.c

@@ -1,0 +1,10 @@
+#include <omp.h>
+
+#include "quicksort.h"
+#include "helper.h"
+
+void quicksort(int *a, int left, int right, int num_threads)
+{
+
+	// insert your code here
+}

### 2015-05-20 parallel_programming/ex5/unit_test.c

@@ -1,0 +1,79 @@
+#include <unistd.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <string.h>
+#include "quicksort.h"
+#include "helper.h"
+
+#define NTHREADS 4
+#define ELEMS 1000000
+
+int main(int argc, char **argv)
+{
+	int elements = ELEMS;
+	int num_threads = NTHREADS;
+	int dump_array = 0;
+	int c;
+	while ((c = getopt(argc, argv, "t:e:d:")) != -1)
+	{
+		switch (c)
+		{
+		case 't':
+			if (sscanf(optarg, "%d", &num_threads) != 1)
+				goto error;
+			break;
+
+		case 'e':
+			if (sscanf(optarg, "%d", &elements) != 1)
+				goto error;
+			break;
+
+		case 'd':
+			if (sscanf(optarg, "%d", &dump_array) != 1)
+				goto error;
+			break;
+
+		case '?':
+			error: printf(
+			    "Usage:\n"
+			    "-t \t number of threads used in computation\n"
+			    "-e \t number of elements to be sorted using quicksort\n"
+			    "-d \t dump array option (0 or 1)\n"
+			    "\n"
+			    "Example:\n"
+			    "%s -t 4 -e 10000000 -d 1\n", argv[0]);
+			exit(EXIT_FAILURE);
+			break;
+		}
+	}
+
+	int *a = random_int_array(elements, elements / 2, 13);
+	if (dump_array == 1) print_array(a, elements);
+	int *b = calloc(elements, sizeof(int));
+
+
+	for (int i = 0; i < elements; ++i)
+	{
+		b[i] = a[i];
+	}
+
+
+
+	quicksort(a, 0, elements - 1, num_threads);
+	quicksort_ref(b, 0, elements - 1, num_threads);
+
+	if (dump_array == 1) print_array(a, elements);
+
+
+	for (int i = 0; i < elements; ++i)
+	{
+		if (a[i] != b[i]) {
+			printf("Unit test using Sections has failed.\nYour paralellized function does not produce the same result as the reference sequential function.\n");
+			break;
+		}
+	}
+
+	printf("Unit test using Sections completed successfully.\n");
+	return 0;
+}

### 2015-05-20 parallel_programming/ex6/helper.c

@@ -1,0 +1,103 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <string.h>
+#include "quicksort.h"
+#include "helper.h"
+
+double time_diff(const struct timespec *first, const struct timespec *second, struct timespec *diff)
+{
+	struct timespec tmp;
+	const struct timespec *tmp_ptr;
+
+	if (first->tv_sec > second->tv_sec || (first->tv_sec == second->tv_sec && first->tv_nsec > second->tv_nsec))
+	{
+		tmp_ptr = first;
+		first = second;
+		second = tmp_ptr;
+	}
+
+	tmp.tv_sec = second->tv_sec - first->tv_sec;
+	tmp.tv_nsec = second->tv_nsec - first->tv_nsec;
+
+	if (tmp.tv_nsec < 0)
+	{
+		tmp.tv_sec -= 1;
+		tmp.tv_nsec += 1000000000;
+	}
+
+	if (diff != NULL )
+	{
+		diff->tv_sec = tmp.tv_sec;
+		diff->tv_nsec = tmp.tv_nsec;
+	}
+
+	return tmp.tv_sec + tmp.tv_nsec / 1000000000.0;
+}
+
+
+int * random_int_array(long size, int num_swaps, unsigned int seed)
+{
+	srand(seed);
+
+	int *a = malloc(size * (long)sizeof(*a));
+
+	for(int i = 0; i < size; i++)
+		a[i] = i;
+
+	for(long i = 0; i < num_swaps; i++)
+	{
+		long idx0 = rand() % size;
+		long idx1 = rand() % size;
+
+		int tmp = a[idx0];
+		a[idx0] = a[idx1];
+		a[idx1] = tmp;
+	}
+	return a;
+}
+
+
+void print_array(int *a, int elements)
+{
+	for(int i=0; i < elements; i++)
+		printf("%4d ", a[i]);
+	printf("\n");
+}
+
+void swap(int *a, int *b)
+{
+	int tmp = *a;
+	*a = *b;
+	*b = tmp;
+}
+
+void quicksort_ref(int *a, int left, int right, int num_threads)
+{
+	if(left < right)
+	{
+		int x = left, y = (left+right)/2, z =right;
+		int pivotIdx = (a[x] <= a[y])
+		    ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
[…]

### 2015-05-20 parallel_programming/ex6/helper.h

@@ -1,0 +1,19 @@
+#ifndef HELPER_H_
+#define HELPER_H_
+
+#include <time.h>
+
+#include "quicksort.h"
+
+
+double time_diff(const struct timespec *first, const struct timespec *second, struct timespec *diff);
+
+int * random_int_array(long size, int num_swaps, unsigned int seed);
+
+void print_array(int *a, int elements);
+
+void swap(int *a, int *b);
+
+void quicksort_ref(int *a, int left, int right, int num_threads);
+
+#endif /* HELPER_H_ */

### 2015-05-20 parallel_programming/ex6/quicksort.h

@@ -1,0 +1,6 @@
+#ifndef QUICKSORT_H_
+#define QUICKSORT_H_
+
+void quicksort(int *a, int left, int right, int num_threads);
+
+#endif /* QUICKSORT_H_ */

### 2015-05-20 parallel_programming/ex6/unit_test.c

@@ -1,0 +1,78 @@
+#include <unistd.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <string.h>
+#include "quicksort.h"
+#include "helper.h"
+
+#define NTHREADS 4
+#define ELEMS 1000000
+
+int main(int argc, char **argv)
+{
+	int elements = ELEMS;
+	int num_threads = NTHREADS;
+	int dump_array = 0;
+	int c;
+	while ((c = getopt(argc, argv, "t:e:d:")) != -1)
+	{
+		switch (c)
+		{
+		case 't':
+			if (sscanf(optarg, "%d", &num_threads) != 1)
+				goto error;
+			break;
+
+		case 'e':
+			if (sscanf(optarg, "%d", &elements) != 1)
+				goto error;
+			break;
+
+		case 'd':
+			if (sscanf(optarg, "%d", &dump_array) != 1)
+				goto error;
+			break;
+
+		case '?':
+			error: printf(
+			    "Usage:\n"
+			    "-t \t number of threads used in computation\n"
+			    "-e \t number of elements to be sorted using quicksort\n"
+			    "-d \t dump array option (0 or 1)\n"
+			    "\n"
+			    "Example:\n"
+			    "%s -t 4 -e 10000000 -d 1\n", argv[0]);
+			exit(EXIT_FAILURE);
+			break;
+		}
+	}
+
+	int *a = random_int_array(elements, elements / 2, 13);
+	if (dump_array == 1) print_array(a, elements);
+	int *b = calloc(elements, sizeof(int));
+	//print_array(b, elements);
+
+	for (int i = 0; i < elements; ++i)
+	{
+		b[i] = a[i];
+	}
+
+
+	quicksort(a, 0, elements-1, num_threads);
+	quicksort_ref(b, 0, elements-1, num_threads);
+
+	if (dump_array == 1) print_array(a, elements);
+
+	for (int i = 0; i < elements; ++i)
+	{
+		if (a[i] != b[i]){
+			printf("Unit test using Tasks has failed.\nYour paralellized function does not produce the same result as the reference sequential function.\n");
+			break;
+		} 
+	}
+
+	printf("Unit test using Tasks completed successfully.\n");
+
+	return 0;
+}

### 2015-05-21 parallel_programming/ex5_tasks/quicksort_par.c
initial task based implementation, too small a speedup

@@ -1,0 +1,106 @@
+#include <omp.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <string.h>
+#include "quicksort.h"
+#include "helper.h"
+
+#define SEQ_THRESH 100
+#define INSERT_THRESH 10
+
+void insertion_sort(int *a, int left, int right)
+{
+  for(int i = left; i <= right; ++i)
+  {
+    for(int j = i; j > left && a[j-1] >= a[j]; --j)
+    {
+      swap(&a[j], &a[j-1]);
+    }
+  }
+}
+
+void quicksort_seq(int *a, int left, int right)
+{
+  if(right - left <= INSERT_THRESH)
+  {
+    insertion_sort(a, left, right);
+    return;
+  }
+
+  int x = left, y = (left+right)/2, z =right;
+  int pivotIdx = (a[x] <= a[y])
+    ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
+    : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
+
+  int pivotVal = a[pivotIdx];
+  swap(a + pivotIdx, a + right);
+
+  int swapIdx = left;
+
+  for(int i=left; i < right; i++)
+  {
+    if(a[i] <= pivotVal)
+    {
+      swap(a + swapIdx, a + i);
+      swapIdx++;
+    }
+  }
+  swap(a + swapIdx, a + right);
+
+  quicksort_seq(a, left, swapIdx - 1);
+  quicksort_seq(a, swapIdx + 1, right);
+}
+
+void quicksort_par(int *a, int left, int right)
+{
+
+  if(right - left <= SEQ_THRESH)
+  {
+    quicksort_seq(a, left, right);
+    return;
+  }
+
+  int x = left, y = (left+right)/2, z =right;
+  int pivotIdx = (a[x] <= a[y])
+    ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
+    : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
+
+  int pivotVal = a[pivotIdx];
+  swap(a + pivotIdx, a + right);
+
+  int swapIdx = left;
+
+  #pragma omp for ordered
+  for(int i=left; i < right; i++)
+  {
+    if(a[i] <= pivotVal)
+    {
+      swap(a + swapIdx, a + i);
+      swapIdx++;
+    }
+  }
[…]

### 2015-05-21 parallel_programming/ex5_tasks/student/quicksort_par.c

@@ -1,10 +1,0 @@
-#include <omp.h>
-
-#include "quicksort.h"
-#include "helper.h"
-
-void quicksort(int *a, int left, int right, int num_threads)
-{
-
-	// insert your code here
-}

### 2015-05-28 parallel_programming/auto_vec/data/plot_from_csv.m
auto-vec presentation

@@ -1,0 +1,40 @@
+function plot_from_csv()
+    close all, clear all, clc
+    
+    colors = 'rkbm';
+    names = {'scalar', 'sse4', 'avx'};
+    prefix = '2x1_';
+    suffix = '.txt';
+    
+    p_peak = 15.47;
+    total_cols = 5;    
+    cache_size = 32 * 2^10;
+    cache_boundary = cache_size / 8 / total_cols;       
+    
+    max_num = 0;
+    max_perf = 0;
+    
+    figure(1)
+    hold on
+    grid on
+    for i = 1:numel(names)
+        filename = strcat(strcat(prefix, names{i}), suffix);
+        data = csvread(filename);
+        plot(data(1, :), data(2, :), strcat(colors(i), '*-'), 'LineWidth', 2)
+        
+        max_num = max(max_num, max(data(1, :)));
+        max_perf = max(max_perf, max(data(2, :)));
+    end
+    
+    x = linspace(0, max(cache_boundary, max_num), 10);
+    y = p_peak * ones(size(x));
+    plot(x, y, 'm')
+    
+    y = linspace(0, max(max_perf, p_peak), 2);
+    x = cache_boundary * ones(size(y));
+    plot(x, y, 'g')
+    
+    xlabel('Number of rows of A')
+    ylabel('Performance [GFlops]')
+    legend('scalar', 'sse', 'avx', 'P_{peak}', 'Cache limit')
+end

### 2015-05-28 HiWi/matlab/multiply_tests.m
extracted partition fun

@@ -1,0 +1,35 @@
+function multiply_tests()
+    clc, clear all
+    
+    sizeA1 = 4;
+    sizeA2 = 3;
+    sizeB1 = 5;
+    
+    A = createA(sizeA1, sizeA2)
+    B = createB(sizeB1, sizeA1)
+    C = createC(sizeB1, sizeA2)
+    
+    BtimesA = B*A
+    CplusBtimesA = C + BtimesA
+    
+    
+end
+
+
+function M = reshapeToRowMajor(A, sizeA1, sizeA2)
+    M = reshape(A, sizeA2, sizeA1)';
+end
+
+function A = createA(sizeA1, sizeA2)
+   lengthA = sizeA1 * sizeA2;
+   A = reshapeToRowMajor(0:sizeA1*sizeA2-1, sizeA1, sizeA2);
+end
+
+function B = createB(sizeB1, sizeB2)
+    lengthB = sizeB1 * sizeB2;
+    B = reshapeToRowMajor(lengthB - (0:lengthB-1), sizeB1, sizeB2);    
+end
+
+function C = createC(sizeC1, sizeC2)
+    C = createA(sizeC1, sizeC2) + 1;
+end

### 2015-05-28 parallel_programming/ex5_sections/quicksort_par.c

@@ -1,0 +1,93 @@
+#include <omp.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <time.h>
+#include <string.h>
+#include "quicksort.h"
+#include "helper.h"
+
+#define SEQ_THRESH 100
+
+
+static void quicksort_seq(int *a, int left, int right)
+{
+	if(left > right)
+	{
+    return;
+  }
+  int x = left, y = (left+right)/2, z =right;
+  int pivotIdx = (a[x] <= a[y])
+      ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
+      : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
+
+  int pivotVal = a[pivotIdx];
+  swap(a + pivotIdx, a + right);
+
+  int swapIdx = left;
+
+  for(int i=left; i < right; i++)
+  {
+    if(a[i] <= pivotVal)
+    {
+      swap(a + swapIdx, a + i);
+      swapIdx++;
+    }
+  }
+  swap(a + swapIdx, a + right);
+
+  quicksort_seq(a, left, swapIdx - 1);
+  quicksort_seq(a, swapIdx + 1, right);
+}
+
+static void quicksort_par(int *a, int left, int right, int num_threads)
+{
+	if((right - left) < SEQ_THRESH)
+	{
+    quicksort_seq(a, left, right);
+    return;
+  }
+
+  int swapIdx;
+  #pragma omp parallel shared(swapIdx)
+  {
+  #pragma omp sections
+  { 
+    #pragma omp section
+    { 
+    int x = left, y = (left+right)/2, z = right;
+    int pivotIdx = (a[x] <= a[y])
+        ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
+        : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
+
+    int pivotVal = a[pivotIdx];
+    swap(a + pivotIdx, a + right);
+
+    swapIdx = left;
+
+    for(int i=left; i < right; i++)
+    {
+      if(a[i] <= pivotVal)
+      {
+        swap(a + swapIdx, a + i);
+        swapIdx++;
+      }
+    }
+    swap(a + swapIdx, a + right);
+    }
+
+    #pragma omp section
+    quicksort_par(a, left, swapIdx - 1, num_threads);
+
+    #pragma omp section
[…]

### 2015-05-28 parallel_programming/ex5_tasks/quicksort_par.c

@@ -6,7 +6,7 @@
 #include "quicksort.h"
 #include "helper.h"
 
-#define SEQ_THRESH 100
+#define SEQ_THRESH 128
 #define INSERT_THRESH 10
 
 void insertion_sort(int *a, int left, int right)
@@ -20,76 +20,54 @@ void insertion_sort(int *a, int left, int right)
   }
 }
 
-void quicksort_seq(int *a, int left, int right)
+
+int partition(int* a, int left, int right) {
+
+  int x = left, y = (left+right)/2, z =right;
+  int pivotIdx = (a[x] <= a[y])
+      ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
+      : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
+
+  int pivotVal = a[pivotIdx];
+  swap(a + pivotIdx, a + right);
+
+  int swapIdx = left;
+
+  for(int i=left; i < right; i++)
+  {
+    if(a[i] <= pivotVal)
+    {
+      swap(a + swapIdx, a + i);
+      swapIdx++;
+                            
+    }
+              
+  }
+  swap(a + swapIdx, a + right);
+
+  return swapIdx;
+}
+
+void quicksort_par(int *a, int left, int right)
 {
+
   if(right - left <= INSERT_THRESH)
   {
     insertion_sort(a, left, right);
     return;
   }
 
-  int x = left, y = (left+right)/2, z =right;
-  int pivotIdx = (a[x] <= a[y])
-    ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
-    : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
 
-  int pivotVal = a[pivotIdx];
-  swap(a + pivotIdx, a + right);
+  int swapIdx = partition(a, left, right);
 
-  int swapIdx = left;
-
-  for(int i=left; i < right; i++)
-  {
-    if(a[i] <= pivotVal)
-    {
-      swap(a + swapIdx, a + i);
-      swapIdx++;
-    }
-  }
-  swap(a + swapIdx, a + right);
-
-  quicksort_seq(a, left, swapIdx - 1);
-  quicksort_seq(a, swapIdx + 1, right);
-}
-
-void quicksort_par(int *a, int left, int right)
-{
-
-  if(right - left <= SEQ_THRESH)
-  {
[…]

### 2015-05-28 parallel_programming/ex5_sections/quicksort_par.c
assignments 5 tasks and sections work

@@ -11,10 +11,11 @@
 
 static void quicksort_seq(int *a, int left, int right)
 {
-	if(left > right)
+	if(left >= right)
 	{
     return;
   }
+  //printf("sq: left = %d, right = %d\n", left, right);
   int x = left, y = (left+right)/2, z =right;
   int pivotIdx = (a[x] <= a[y])
       ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
@@ -47,47 +48,49 @@ static void quicksort_par(int *a, int left, int right, int num_threads)
     return;
   }
 
+  //printf("left = %d, right = %d\n", left, right);
   int swapIdx;
-  #pragma omp parallel shared(swapIdx)
+  int x = left, y = (left+right)/2, z = right;
+  int pivotIdx = (a[x] <= a[y])
+      ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
+      : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
+
+  int pivotVal = a[pivotIdx];
+  swap(a + pivotIdx, a + right);
+
+  swapIdx = left;
+
+  for(int i=left; i < right; i++)
   {
-  #pragma omp sections
-  { 
-    #pragma omp section
-    { 
-    int x = left, y = (left+right)/2, z = right;
-    int pivotIdx = (a[x] <= a[y])
-        ? ((a[y] <= a[z]) ? y : ((a[x] < a[z]) ? z : x))
-        : ((a[x] <= a[z]) ? x : ((a[y] < a[z]) ? z : y));
-
-    int pivotVal = a[pivotIdx];
-    swap(a + pivotIdx, a + right);
-
-    swapIdx = left;
-
-    for(int i=left; i < right; i++)
+    if(a[i] <= pivotVal)
     {
-      if(a[i] <= pivotVal)
-      {
-        swap(a + swapIdx, a + i);
-        swapIdx++;
-      }
+      swap(a + swapIdx, a + i);
+      swapIdx++;
     }
-    swap(a + swapIdx, a + right);
-    }
+  }
+  swap(a + swapIdx, a + right);
 
-    #pragma omp section
+  #pragma omp parallel sections
+  { 
+    //#pragma omp section 
+    //{
     quicksort_par(a, left, swapIdx - 1, num_threads);
-
+    //}
     #pragma omp section
+    {
     quicksort_par(a, swapIdx + 1, right, num_threads);
     }
-    }
+  }
[…]

### 2015-05-28 parallel_programming/ex5_tasks/quicksort_par.c

@@ -21,7 +21,7 @@ void insertion_sort(int *a, int left, int right)
 }
 
 
-int partition(int* a, int left, int right) {
+inline int partition(int* a, int left, int right) {
 
   int x = left, y = (left+right)/2, z =right;
   int pivotIdx = (a[x] <= a[y])
@@ -60,11 +60,12 @@ void quicksort_par(int *a, int left, int right)
 
   int swapIdx = partition(a, left, right);
 
-  #pragma omp task firstprivate(a, left, swapIdx) untied final(swapIdx - 1 - left < SEQ_THRESH)
+  //#pragma omp task firstprivate(a, left, swapIdx) untied final(swapIdx - 1 - left < SEQ_THRESH)
+  #pragma omp task final(swapIdx - left < SEQ_THRESH)
   quicksort_par(a, left, swapIdx - 1);
 
 
-  #pragma omp task firstprivate(a, swapIdx, right) untied final(right - 1 - swapIdx < SEQ_THRESH)
+  //#pragma omp task firstprivate(a, swapIdx, right) untied final(right - 1 - swapIdx < SEQ_THRESH)
   quicksort_par(a, swapIdx + 1, right);
 
   #pragma omp taskwait
@@ -75,7 +76,7 @@ void quicksort(int *a, int left, int right, int num_threads) {
 
   //printf("num threads = %d\n", num_threads);
   omp_set_num_threads(num_threads);
-  omp_set_nested(1);
+  omp_set_nested(100);
   #pragma omp parallel
   {
     #pragma omp single nowait

### 2015-11-06 phd/sop/sop.tex
first decent SoP draft

\documentstyle[12pt]{article}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
\pagestyle{empty}

\begin{document}

\begin{center}
{\Large Statement of Purpose} \\[.3in]
{\large Adam Kosiorek}
\end{center}

\vspace*{.5in}

I like challenges. After reading an article about genetic engineering I have chosen Biology as my major in the secondary school. With the biggest challange I faced being memorization of some facts, I decided to change my major to Maths \& Physics I always loved. I would have chosen Informatics if only I had any idea what that meant. Coming from a background where CS is associated with technicians repairing “broken” PCs, I had an idea of CS very far from the romantic vision of young startup founders set out to change the world. Luckily, I have been exposed to Matrix and IRobot, both filled with autonomous intelligent machines ruleing the world. Fascinated, I decied to find out if it is possible and to how make it happen. Not that I wanted to destroy the world, but the sheer concept of creating a machine that could improve itself, or technological singularity, name I didn't know at the time, was enticing.

My journey began at the best tech university in the country, where I enrolled for the BSc in Robotics. I learned a lot about physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, have I realized how a truly intelligent entinty could be controled. This, paired with my passion for vision and perception I've developed as a photojournalist steered me towards my first computer vision course, where I learned about image processing and machine learning. I immedietely knew that this was the area I wanted to focus on. It was not only challangeing, since there weren't any general purpose systems good enough for deployment in the real world, but it was also well aligned for my original passion for intelligent machines.

I also try to learn from the best. That's why I went fn or an internship to the Samsung R\&D Lab in Warsaw, where I've been working on object classification in the Computer Vision group. I've been exposed there to the Bag of Words approach for classification a-nd consequently, to the state-of-the-art in keypoint detection and feature description, which helped me to later understand why representations learnd automatically by convolutional neural networks are essential. I focused mainly on improving the quantization step of BoW approach – that is building visual vocabulary. I took it a step further and implemented a similar system for 3D point cloud classifications with the Microsoft Kinect as my bachelor thesis. I continued my work at Samsung where I was introduced to deep learning.

My observation was that deep learning required huge computational resources to be any good and therefore I decided to do CSE@TUM to be able to harness supercomputers in favour of my research. I've been working with some algorithms developed there and it made it look like a good idea. The prof' Cremers lab further increased attractivity. So did the DAAD scholarship, which meant I could focus on my studies.

This summer I did an internship at Bloomberg in London, where I've been working on fraud detection in financial transactions. The problem can be cast as unsupervised anomaly detection with further verification in a supervised settings. I learned how difficult it can be to introduce an innovative approach in a corporate setting. This, together with my earlier industrial experiences conviced me that I do want to pursue a PhD. I love solving scientific problems that do not have “the best” solution or reference specification and I vastly prefer it to chiseling icons in a UI.

Currently I'm working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof. Cremers. Our recent project is the introspective capacity of neural networks. Introspective capacity is understood as the ability of a classifier to assess uncertainty of its result given the data [ref]. It is an important problem, because in mobile robotics or medical settings a wrong classification might lead to the lose of life. While high classification accuracy is certainly desired, it is even more important to know how certain classification results are. Our intermediate results show that neural networks, augmented by additional layers and a novel cost function, can be jointly trained for classification and uncertainty estimation. The topic might expand into my master thesis, which I am intending to write during the summer semester.

In my PhD research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks, topics that I do not have much experience in. The first problem fascinates me in that it seems to be similar to how humans learn. It is far easier to specify a reward function or to designate an expert as in the inverse version of the problem than to provide supervised training data. This means that a bulk of reinforcement learning can be done in the unsupervised fashion, which is, I believe, THE way of bringing the cutting edge AI technology to the masses. RNNs, on the other hand, are well suited for processing sequential information. As such, this model might contribute to decreasing the bulk of computation required e.g. in object classification in videos. One way of doing that would be to introduce beliefs about what might be in the image in the form of priors computed from the preceeding elements of the sequence. Another interesting use case would be optical flow computation from a sequence of images. The only end-to-end neural network for optical flow computation, while efficient and accurate, requires that the consecutive images of a video sequence are input in pairs. I would like to investigate an RNN designs capale of computing optical flow while presented with a single image at a time.

Passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely well motivated to take part in the Traction Europe workshop. I am confident that I will be able to contribute significantly to and benefit immensely from the workshop. Thank you for considering my application.
\end{document}

### 2015-11-06 phd/sop/stanford.tex
first decent SoP draft

\documentstyle[12pt]{article}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
\pagestyle{empty}

\begin{document}

\begin{center}
{\Large Statement of Purpose} \\[.3in]
{\large Robert S. French}
\end{center}

\vspace*{.5in}

My primary interests in computer science are computer architecture and
compiler technology.  I have participated in several projects during
my undergraduate career at MIT that have provided me with experience
in these areas, and I have recently specialized in parallel processing
systems.

My interest in computer architecture was first kindled during the
design contest for the MIT course ``Computation Structures.''  The
contest consists of redesigning an 8-bit microcoded processor to
improve its performance.  Our two-person team constructed a 16-bit
design for the processor and achieved at speedup of over 750 times.
As part of the design, I implemented a microcode assembler, a
microcode simulator, various levels of logic simulators, and over 80
pages of microcode.  The contest taught me a great deal about
low-level system design and the tradeoffs involved in designing fast
systems.

My next architecture project was for the course ``Digital Systems
Laboratory,'' in which my partner and I implemented a hardware
simulator for a highly-parallel, fine-grained optical computer.  The
system was based on recent research by Alan Huang at Bell Laboratories
in ``computational origami,'' a simple method of simulating an
arbitrarily large number of processors using a small set of processing
elements with long delay lines.  As part of the project I designed a
compiler and a placement and routing system for allocating processors.
I presented a paper on the algorithm at the MIT-ACM Undergraduate
Computer Science Conference in April, 1989, and was invited to present
my findings at Bell Labs.

I designed and implemented part of {\em Zephyr\/}, a distributed
multicast message system, as an employee of MIT's Project Athena.
Project Athena is an eight year experiment in the use of computers in
education, and consists of approximately 1,000 workstations connected
by a campus-wide network.  This setup provides a unique environment in
which to develop scalable distributed software.  I investigated
various communications methods, and was the primary architect of the
protocol and the authentication system used by {\em Zephyr\/}.  Our
development team published a paper on {\em Zephyr\/} in the Winter
1988 Usenix Conference Proceedings.

My current research is for my undergraduate thesis, in which I am
investigating the use of fractals to measure communication locality in
distributed systems.  I hope to extend this research and develop new
algorithms for efficient processor allocation.

I plan to continue my study of architecture and compiler technology in
graduate school, and I feel that Stanford is the ideal place to study
these fields.  Stanford offers an excellent selection of courses and
supports a vigorous research program in my areas of interest.  I enjoy
developing theories and building hardware systems on which to test
them, and hope to be involved in applied research at Stanford.

After receiving a Ph.D., I plan to pursue an academic career in
computer science research.

\end{document}

### 2015-11-07 phd/sop/sop.tex
change font and add notes
rework research outlook and spellcheck

@@
-\documentstyle[12pt]{article}
+\documentclass[12pt]{article}
+\usepackage[utf8]{inputenc}
 \setlength{\oddsidemargin}{0in}
@@
 After reading an article about genetic engineering I have chosen Biology as my major in the secondary school.
-With the biggest challange I faced being memorization of some facts, I decided to change my major to Maths \& Physics I always loved.
+With the biggest challenge I faced being memorization of some facts, I decided to change my major to Maths \& Physics I always loved.
 I would have chosen Informatics if only I had any idea what that meant.
-Coming from a background where CS is associated with technicians repairing “broken” PCs, I had an idea of CS very far from the romantic vision of young startup founders set out to change the world.
-Luckily, I have been exposed to Matrix and IRobot, both filled with autonomous intelligent machines ruleing the world.
-Fascinated, I decied to find out if it is possible and to how make it happen.
-Not that I wanted to destroy the world, but the sheer concept of creating a machine that could improve itself, or technological singularity, name I didn't know at the time, was enticing.
+Coming from a background where CS is associated with technicians repairing “broken” PCs, my idea of CS was very far from the romantic vision of young startup founders set out to change the world.
+Luckily, I have been exposed to Matrix and IRobot, both filled with autonomous intelligent machines ruling the world.
+Fascinated, I decided to find out if it is possible and how make it happen.
+Not that I wanted to destroy the world, no, but the sheer concept of creating a machine that could improve itself, or the technological singularity, name I didn't know at the time, was enticing.
 
 My journey began at the best tech university in the country, where I enrolled for the BSc in Robotics.
-I learned a lot about physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, have I realized how a truly intelligent entinty could be controled.
-This, paired with my passion for vision and perception I've developed as a photojournalist steered me towards my first computer vision course, where I learned about image processing and machine learning.
-I immedietely knew that this was the area I wanted to focus on.
-It was not only challangeing, since there weren't any general purpose systems good enough for deployment in the real world, but it was also well aligned for my original passion for intelligent machines.
+I learned a lot about physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, have I realized how a truly intelligent entity could be controlled.
+This, paired with my passion for vision and perception I've developed as a photojournalist, steered me towards my first computer vision course, where I learned about image processing and machine learning.
+I immediately knew that this was the area I wanted to focus on.
+It was not only challenging, since there weren't any general purpose systems good enough for deployment in the real world, but it was also well aligned with my original passion for intelligent machines.
 
-I also try to learn from the best.
-That's why I went fn or an internship to the Samsung R\&D Lab in Warsaw, where I've been working on object classification in the Computer Vision group.
-I've been exposed there to the Bag of Words approach for classification a-nd consequently, to the state-of-the-art in keypoint detection and feature description, which helped me to later understand why representations learnd automatically by convolutional neural networks are essential.
-I focused mainly on improving the quantization step of BoW approach – that is building visual vocabulary.
-I took it a step further and implemented a similar system for 3D point cloud classifications with the Microsoft Kinect as my bachelor thesis.
-I continued my work at Samsung where I was introduced to deep learning.
+To learn from the best, I went for an internship to the Computer Vision Lab at Samsung R\&D in Warsaw.
+My project was to design and implement a Bag of Words-based pipeline for object classification.
+Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of automatically learned low level features in convolutional neural networks (CNNs).
+I focused my research mainly on the visual codebook generation within Bag of Words.
+The final step was to optimize it and port to the Android platform.
+In my bachelor thesis I investigated how spatial information affect classification performance by implementing a similar system for Kinect-gathered point cloud classification.
+I continued my work at Samsung, where I worked on deep learning techniques for object classification and duplicate detection.
 
-My observation was that deep learning required huge computational resources to be any good and therefore I decided to do CSE@TUM to be able to harness supercomputers in favour of my research.
-I've been working with some algorithms developed there and it made it look like a good idea.
-The prof' Cremers lab further increased attractivity.
-So did the DAAD scholarship, which meant I could focus on my studies.
+My observation was that deep learning, although superior to older approaches, required huge computational resources and datasets to be any good and therefore I decided to study computational science to be able to harness supercomputers.
+TUM seemed to be a great choice due to the computer vision teams operating there.
 
@@
 I learned how difficult it can be to introduce an innovative approach in a corporate setting.
-This, together with my earlier industrial experiences conviced me that I do want to pursue a PhD.
-I love solving scientific problems that do not have “the best” solution or reference specification and I vastly prefer it to chiseling icons in a UI.
+This, together with my earlier industrial experience, convinced me that I do want to pursue a PhD.
+I love solving scientific problems that do not have ``the best'' solution or reference specification and I vastly prefer it to chiseling icons in an user interface.
 
-Currently I'm working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof.
+Currently, I am working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof.
 Cremers.
-Our recent project is the introspective capacity of neural networks.
-Introspective capacity is understood as the ability of a classifier to assess uncertainty of its result given the data [ref].
-It is an important problem, because in mobile robotics or medical settings a wrong classification might lead to the lose of life.
-While high classification accuracy is certainly desired, it is even more important to know how certain classification results are.
-Our intermediate results show that neural networks, augmented by additional layers and a novel cost function, can be jointly trained for classification and uncertainty estimation.
-The topic might expand into my master thesis, which I am intending to write during the summer semester.
+In our recent project we investigate the introspective capacity of neural networks.
+It is understood as the ability of a classifier to assess uncertainty of its prediction given the data [ref].
+It is a paramount problem in mobile robotics or medicine where a wrong classification might lead to loss of life.
+While high classification accuracy is desired, it is impossible to assess whether a given prediction is accurate in the test setting, where no labels are available.
+It is, therefore, vital to assess uncertainty of predictions.
+Our intermediate results show that neural networks, augmented with additional layers and a novel cost function, can be jointly trained for classification and uncertainty estimation.
+The topic might expand into my master thesis.
 
-In my PhD research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks, topics that I do not have much experience in.
-The first problem fascinates me in that it seems to be similar to how humans learn.
-It is far easier to specify a reward function or to designate an expert as in the inverse version of the problem than to provide supervised training data.
-This means that a bulk of reinforcement learning can be done in the unsupervised fashion, which is, I believe, THE way of bringing the cutting edge AI technology to the masses.
-RNNs, on the other hand, are well suited for processing sequential information.
-As such, this model might contribute to decreasing the bulk of computation required e.g.
-in object classification in videos.
-One way of doing that would be to introduce beliefs about what might be in the image in the form of priors computed from the preceeding elements of the sequence.
-Another interesting use case would be optical flow computation from a sequence of images.
-The only end-to-end neural network for optical flow computation, while efficient and accurate, requires that the consecutive images of a video sequence are input in pairs.
-I would like to investigate an RNN designs capale of computing optical flow while presented with a single image at a time.
+In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks.
+The former fascinates me since it is similar to how humans learn.
+Specifying a reward function  or, as in the inverse problem, designating an expert is far easier than providing supervised training data.
+Recurrent neural networks, on the other hand, are well suited to sequential information processing.
+If beliefs about content could be introduced in a form of priors inferred from previous elements of the sequence, it might be possible to increase object classification accuracy in videos and to shrink the network size.
+It would reduce the amount of computation needed, possibly leading to an energy efficient system for portable devices.
+Another interesting problem is an end-to-end RNN for optical flow computation.
+The only end-to-end neural network, while efficient and accurate, does not use temporal information.
+It computes optical flow for any two possibly unrelated images [ref].
+I would like to investigate RNN designs capable of computing optical flow while presented with a single image at a time.
+It might lead to better accuracy with fewer computations, thus better energy efficiency.
 
-Passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely well motivated to take part in the Traction Europe workshop.
-I am confident that I will be able to contribute significantly to and benefit immensely from the workshop.
+Passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely well motivated to carry out research as a PhD student at the University of Oxford.
+I am confident that I will be able to contribute significantly to and benefit immensely from my stay there.
 Thank you for considering my application.
 \end{document}
+
+% 1. What's Needed.
+% 
+% General area of research
+% Relevant skills and experience
+% Your reasons for applying to Oxford
+% Why you consider yourself a strong applicant.
+% 
+% 2. Introspection.
+% 
+% Intelectual Development.
+% 
+% a. Matrix and IRobot hooked me up on AI and Robotics
+% b. Found B.Sc. in Robotics at WUT and went for it
+% c. Disapointed due to mechanical engineering oriented programme
+% d. Course in CV and mobile robotics -> that’s where AI is, should do CS
+% e. Internship in Samsung in CV: ML is really cool, need more CS skills
+% f. Thesis in Object Recognition
+% g. 1 semester of CS undegrad courses
+% h. CV/ML needs HPC for large scale -> CSE with focus on CV/ML
+% i. Want to do CV/ML for robotics as PhD -> RL
+% 
+% Role Models: Andrew Ng and Elon Musk.
+% 
+% Both extremely creative, a scientist and an applied engineer, both with huge vision.
+One is changing the world of AI/ML and education, the other is chaining the world by reaching for technologies far out of time.
+Bold, don’t fear to lose, have a dream and want to change the world.
+They don’t care about problems being too hard or too big.
+They just make stuff happen.
+% 
+% 3. Outline.
+% 
+% 1. Genesis
+% 1. I always liked physics and maths
+% 2. As a photographer I was interested in vision and perception
+% 3. I was fascinated by robotics and AI, although I had no idea what that even was
+% 
+% 2. What Happened
+% 1. Went to study Robotics @ WUT
+% 2.
+Turned out that most of Robotics is mechanical and electrical engineering with a nudge of control theory.
+I got more excited about general software development and got into it UNTIL my first class in computer vision where I learned about machine learning.
+% 3. Decided I want to do this and went to the Samsung R&D Lab to work objcect classification and was doing bachelor thesis on the same topic but in 3D with kinect
+% 4. Turnened out that TUM is quite strong in the area and I decided to go for a Masters hers, DAAD helped a lot
+% 5.
+I'm working with the Prof.
+Cremers' group on deep learning for computer vision.
+I'm really into it and I want to get a PhD.
+I love solving difficult research problems that have no right or wrong solution.
+I hate being tied up by the specs and working on that tiny little widget to make the UI better.
+% 
+% 3. Outlook
+% 1. Introspective capacity is vital for robotics. I'm looking into it in the context of deep learning and I'd like to make it work, but it's not sth I want to do a PhD on; Masters might be good
+% 2.
+Reinforcement Learning fascinates me in that it is, together with the inverse problem, more or less how we learn.
+I'd like to study it further for I believe it is THE way of bringing ML to the masses.
+It's way easier to specify a reward  function or to find an expert than to gather supervised training datasets.
+I'm interested into human learning processes as well and I wonder if we can port them to the neural network settings
+% 3.
+Recurrent neural networks seem suitable to a plethora of various tasks.
+E.g.
+object classification, and even more object detection, seem to be doing a lot of redundant work when dealing with e.g.
+video sequences.
+I'd like to work on itroducing knowledge of what might be in the sequence as priors based on what we've already seen there.
+Another thing is to use RNNs to compute optical flow from a sequence, but using one image at a time.
+Also RNNs for optical flow from DVS would be cool.

### 2015-11-08 ML/hw4/hw.tex
ml homework

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
% \pagestyle{empty}
\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 4} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section{Problem 1}

Let $|H|$ denote number of heads and $|T|$ number of tails. We are interested in sequences like: 'H', 'TH', 'TTTH' etc. Since these are coin tosses, probability of events corresponding to these sequences are $\frac{1}{2}$, $\frac{1}{4}$, $\frac{1}{8}$ and so on. Therefore:

\begin{equation}
 E[|H|] = \sum_{i=1}^{\infty} \frac{1}{2^i} = 1
\end{equation}
\begin{equation}
 E[|T|\ = 0 + \frac{1}{4} \cdot 1 + \frac{1}{8} \cdot 2 + \frac{1}{16} \cdot 3 + ... = \frac{1}{2} \sum_{i=1}^{\infty} \frac{1}{2^i} \cdot i = \frac{1}{2} \cdot 2 = 1
\end{equation}

\section{Problem 2}

Let $U$ number of urns, $M$ number of balls in each urn, $N$ number of balls draw with replacement, $n_B$ number of back balls draw, $B$ the event of drawing a single black ball. We know that the probability of drawing a single black ball from the urn $u$ is $P(B|u) = \frac{u}{M}$ and the probability of drawing $n_B$ black urns from the same urn is $P(n_B|u) = \binom{N}{n_B} P(B|u)^{n_b}(1-P(B|u))^{N-n_b}$. Due to the Bayes rule:
\begin{equation}
 P(u|n_B) = \frac{P(n_B|u)P(u)}{P(n_b)}
\end{equation}
where $P(u) = \frac{1}{U}$ and 
\begin{equation}
 P(n_b) = \sum_{u=1}^U P(n_B|u)P(u) = \frac{1}{U} \sum_{u=1}^U \binom{N}{n_B} \left(\frac{u}{M}\right)^{n_b}\left(1-\frac{u}{M}\right)^{N-n_b}
\end{equation}

Since we don't know which urn we have, we must rely on the data for the purpose of computing the probability of the next drawn ball being black. It is the MLE estimate of $\theta = P(B|u) = \frac{n_B}{M} = 0.3$

\section{Problem 3}

\begin{equation}
 P(X=x|\theta) = \theta^x (1-\theta)^{1-x}
\end{equation}
\begin{equation}
 \log(P(X=x|\theta)) = \log\left(\theta^x (1-\theta)^{1-x}\right) = x \log{\theta} + (1-x)\log(1-\theta)
\end{equation}

\begin{equation}
 \theta_{MLE} \Leftrightarrow 0 = \frac{\partial}{\partial \theta} \log(P(X=x|\theta)) = \frac{x}{\theta} - \frac{1-x}{1-\theta}
\end{equation}
\begin{equation}
 \theta_{MLE} = x
\end{equation}

\section{Problem 4}

We have that the prior is given by $Beta(\mu, a, b)$ with $E[Beta(\mu, a, b)] = \frac{a}{a+b}$, the posterior is given by $Beta(\mu, m + a, N - m + b)$ with $E[Beta(\mu, m + a, N - m + b)] = \frac{m+a}{N+a+b}$ and $\mu_{MLE} = \frac{m}{N}$.

\begin{equation}
 \begin{align}
    &E[Beta(\mu, m + a, N - m + b)] = \\
    &= \frac{m+a}{N+a+b} \\
    &= \frac{a+b}{N+a+b} \frac{m+a}{a+b} \\
    &= \frac{a+b}{N+a+b} \frac{a}{a+b} + \frac{N}{N+a+b} \frac{m}{N} \\
    &= \lambda \frac{a}{a+b} + (1-\lambda) \frac{m}{N} \\
    &= \lambda E[Beta(\mu, a, b)] + (1 - \lambda) \mu_{MLE}
 \end{align}
\end{equation}

\section{Problem 5}

For n i.i.d. samples $x_i \sim  Poi(\lambda) = \frac{e^{-\lambda} \lambda^x}{x!}$ we have
\begin{equation}
  P(x_1, \ldots, x_n | \lambda) = \prod_{i=1}^{n} P(x_i | \lambda) = \frac{e^{-n\lambda} \lambda^{\sum_{i=1}^{n} x_i}}{\sum_{i=1}^{n} x!}
\end{equation}
\begin{equation}
  log\left(P(x_1, \ldots, x_n | \lambda)\right) \propto -n\lambda + \log(\lambda) \sum_{i=1}^{n} x_i
\end{equation}
\begin{equation}
  \frac{\partial}{\partial \lambda} log\left(P(x_1, \ldots, x_n | \lambda)\right) \propto -n + \frac{1}{\lambda} \sum_{i=1}^{n} x_i = 0
\end{equation}
\begin{equation}
 \lambda_{MLE} = \frac{1}{n} \sum_{i=1}^{n} x_i
\end{equation}

It is unbiased, since due to the linearity of the expected value operator we have:

\begin{equation}
 E[\lambda_{MLE}] = E[\frac{1}{n} \sum_{i=1}^{n} x_i] = \frac{1}{n} \sum_{i=1}^{n} E[x_i] = \frac{1}{n} \sum_{i=1}^{n} \lambda = \lambda
\end{equation}

Gamma distribution is given by 
\begin{equation}
    Gamma(\alpha, \beta) = \frac{\beta^\alpha}{\Gamma(\alpha)} \lambda^{\alpha-1} e^{-\beta \lambda} \propto \lambda^{\alpha-1} e^{-\beta \lambda}                            
\end{equation}

and the posterior:

\begin{equation}
\begin{align}
  &P(\lambda | \mathrm{D}, \alpha, \beta)\\ 
  &\propto P(\mathrm{D} | \lambda) P(\lambda | \alpha, \beta)\\
  &= e^{-n\lambda} \lambda^{\sum_{i=1}^{n} x_i}\lambda^{\alpha-1} e^{-\beta \lambda} \\
  &= \lambda^{\alpha-1 + \sum_{i=1}^{n} x_i} e^{-(n + \beta) \lambda} 
\end{align}
\end{equation}

therefore

\begin{equation}
  \frac{\partial}{\partial \lambda} log\left(P(\lambda | \mathrm{D}, \alpha, \beta)\right) \propto -n + \beta + \frac{1}{\lambda} \left( \alpha - 1 + \sum_{i=1}^{n} x_i \right) = 0
\end{equation}

finally

\begin{equation}
 \lambda_{MAP} = \frac{1}{n + \beta} \left( \alpha - 1 + \sum_{i=1}^{n} x_i \right)
\end{equation}

\end{document}

### 2015-11-10 phd/sop/sop.tex
remove matrix intro and small corrections

@@
 
-I like challenges.
-After reading an article about genetic engineering I have chosen Biology as my major in the secondary school.
-With the biggest challenge I faced being memorization of some facts, I decided to change my major to Maths \& Physics I always loved.
-I would have chosen Informatics if only I had any idea what that meant.
-Coming from a background where CS is associated with technicians repairing “broken” PCs, my idea of CS was very far from the romantic vision of young startup founders set out to change the world.
-Luckily, I have been exposed to Matrix and IRobot, both filled with autonomous intelligent machines ruling the world.
-Fascinated, I decided to find out if it is possible and how make it happen.
-Not that I wanted to destroy the world, no, but the sheer concept of creating a machine that could improve itself, or the technological singularity, name I didn't know at the time, was enticing.
-
-My journey began at the best tech university in the country, where I enrolled for the BSc in Robotics.
+My journey with computer science began at the best tech university in the country, where I enrolled in the BSc in Robotics.
 I learned a lot about physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, have I realized how a truly intelligent entity could be controlled.
-This, paired with my passion for vision and perception I've developed as a photojournalist, steered me towards my first computer vision course, where I learned about image processing and machine learning.
-I immediately knew that this was the area I wanted to focus on.
-It was not only challenging, since there weren't any general purpose systems good enough for deployment in the real world, but it was also well aligned with my original passion for intelligent machines.
+Paired with my passion for vision and perception, which I have developed as a photojournalist, it steered me towards my first computer vision course.
+I immediately knew that this intersection of image processing and machine learning was the area I wanted to focus on.
+Without any systems good enough for real world deployment, it was challanging as well as aligned with my original passion for intelligent machines.
 
@@
 The final step was to optimize it and port to the Android platform.
-In my bachelor thesis I investigated how spatial information affect classification performance by implementing a similar system for Kinect-gathered point cloud classification.
-I continued my work at Samsung, where I worked on deep learning techniques for object classification and duplicate detection.
+In my bachelor thesis I investigated how spatial information affects classification performance by implementing a similar system for Kinect-gathered point cloud classification.
+Afterwards, I continued my work at Samsung, where I turned to deep learning techniques for object classification and image duplicate detection.
 
-My observation was that deep learning, although superior to older approaches, required huge computational resources and datasets to be any good and therefore I decided to study computational science to be able to harness supercomputers.
-TUM seemed to be a great choice due to the computer vision teams operating there.
-
-This summer I did an internship at Bloomberg in London, where I've been working on fraud detection in financial transactions.
-The problem can be cast as unsupervised anomaly detection with further verification in a supervised settings.
-I learned how difficult it can be to introduce an innovative approach in a corporate setting.
-This, together with my earlier industrial experience, convinced me that I do want to pursue a PhD.
-I love solving scientific problems that do not have ``the best'' solution or reference specification and I vastly prefer it to chiseling icons in an user interface.
-
+My observation was that deep learning, although often superior to older approaches, required huge computational resources and datasets.
+To mitigate this problem I went to study computational science at the Technical University in Munich.
+With cutting-edge computer vision research groups it seemed a great choice.
 Currently, I am working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof.
@@
 It is a paramount problem in mobile robotics or medicine where a wrong classification might lead to loss of life.
-While high classification accuracy is desired, it is impossible to assess whether a given prediction is accurate in the test setting, where no labels are available.
+While high classification accuracy is desired, it is impossible to assess whether a given prediction is accurate in a test setting, where no labels are available.
 It is, therefore, vital to assess uncertainty of predictions.
@@
 
+This summer I did an internship at Bloomberg in London, where I worked on fraud detection in financial transactions.
+The problem can be cast as unsupervised anomaly detection with further verification in a supervised settings.
+I learned how difficult it can be to introduce an innovative approach in a corporate setting.
+This, together with my earlier industrial experience, convinced me that I do want to pursue a PhD.
+I love solving scientific problems that do not have ``the best'' solution or a reference specification by going into the deepest details.
+
 In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks.
 The former fascinates me since it is similar to how humans learn.
-Specifying a reward function  or, as in the inverse problem, designating an expert is far easier than providing supervised training data.
+Specifying a reward function  or designating an expert in the inverse problem  is far easier than providing supervised training data.
 Recurrent neural networks, on the other hand, are well suited to sequential information processing.
@@
 I would like to investigate RNN designs capable of computing optical flow while presented with a single image at a time.
-It might lead to better accuracy with fewer computations, thus better energy efficiency.
+It might lead to a better accuracy with fewer computations, thus better energy efficiency.
 
 Passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely well motivated to carry out research as a PhD student at the University of Oxford.
-I am confident that I will be able to contribute significantly to and benefit immensely from my stay there.
+I am confident that I will be able to contribute significantly to and benefit immensely from my stay in Oxford.
 Thank you for considering my application.

### 2015-11-14 ML/hw5/hw.tex
hw5 ml

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
% \pagestyle{empty}
\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 5} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

There are four names in the hat: Martin, Alexi, Igor, and Valery. The name that is randomly drawn wins the lottery. We define three different events:
\begin{description}
 \item{$X_1$} Martin or Alexi wins.
 \item {$X_2$} Alexi or Igor wins.
 \item {$X_3$} Igor or Martin wins.
\end{description}

We have $P(X_1) = P(X_2) = P(X_3) = 0.5$ and $P(X_1, X_2) = P(X_1, X_3) = P(X_2, X_3) = 0.25$, so the events are pairwise independent. But P$(X_1, X_2, X_3) = 0$, not $0.125$, as it would if the events were all mutual independent.

\section*{Problem 2}

We use the following identities:\\
\begin{equation}
Var[X] = E[X^2] + E[X]^2\label{eq:eq1}
\end{equation}
\begin{equation}
E[X+Y] = E[X]+E[Y]\label{eq:eq2}
\end{equation}
Thus we have\\
\begin{align*}
\begin{split}
Var[X+Y] &\stackrel{\eqref{eq:eq1}}{=} E\left[(X+Y)^2\right]-E[X+Y]^2\\
&\stackrel{\eqref{eq:eq2}}{=}E\left[X^2+2XY+Y^2\right] -\left(E[X]+E[Y]\right)^2\\
&\stackrel{\eqref{eq:eq2}}{=}E\left[X^2\right] + 2E[XY] + E\left[Y^2\right] -E[X]^2-2E[X]E[Y]-E[Y]^2\\
&=Var[X]+Var[Y] + 2\left(E[XY] - E[X]E[Y]\right) = Var[X]+Var[Y]+2Cov(X,Y).
\end{split}
\end{align*}

\section*{Problem 3}

\begin{equation}
Var[X + Y] = E[(X+Y)^2]-(E[X] + E[Y])^2 = E[X^2] - E[X]^2 + E[Y]^2 - E[Y]^2 + 2E[XY] - 2E[X]E[Y]
\end{equation}

And hence,

\begin{equation}
Var[X + Y ] = Var[X] + Var[Y ] + 2Cov[X,Y].
\end{equation}

\section*{Problem 4}

$Z = \begin{pmatrix} X \\ Y \end{pmatrix} \sim \mathrm{N}(\mu_Z, \Sigma_Z)$, with $\mu_Z = \begin{pmatrix} \mu_X \\ \mu_Y \end{pmatrix}$ and $\Sigma_Z = \begin{pmatrix} \sigma_X^2 & Cov(X, Y) \\ Cov(X, Y) & \sigma_Y^2 \end{pmatrix}$. Since $\rho(X, Y) = 0$ also $Cov(X, Y) = 0$ and therefore $\Sigma_Z = \begin{pmatrix} \sigma_X^2 & 0 \\ 0 & \sigma_Y^2 \end{pmatrix}$ is diagonal. We can compute the conditional probabilities $p(X|Y) \sim \mathrm{N}(\mu_{X|Y}, \Sigma_{X|Y})$ and $p(Y|X) \sim \mathrm{N}(\mu_{Y|X}, \Sigma_{Y|X})$ using the formula (4.69) from Murphy p. 111. We get:

\begin{equation}
\begin{align}
 \mu_{X|Y} = \mu_X + Cov(X, Y) \frac{1}{\sigma_y^2}(Y - \mu_Y) = \mu_X \\
 \Sigma_{X|Y} = \sigma_X^2 - Cov(X, Y) \frac{1}{\sigma_Y^2} Cov(Y, X) = \sigma_X^2
\end{align}
\end{equation}

thus $P(X|Y) = P(X)$, proves that $X$ and $Y$ are independent.

\begin{equation}
\end{equation}

\end{document}

### 2015-11-18 phd/sop/sop.tex
add references

@@
 \pagestyle{empty}
+
+\setlength{\parindent}{0em}
+\setlength{\parskip}{1em}
 
@@
 
-My journey with computer science began at the best tech university in the country, where I enrolled in the BSc in Robotics.
-I learned a lot about physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, have I realized how a truly intelligent entity could be controlled.
+%The Statement of Purpose should describe succinctly your reasons for applying to the proposed %program at Stanford, your preparation for this field of study, research interests, future career %plans, and other aspects of your background and interests which may aid the admissions committee %in evaluating your aptitude and motivation for graduate study.
+The Statement of Purpose should not %exceed two pages.
+Artificial Intelligence has always fascinated me.
+Long before understanding what computer science stands for have I wondered why AI was not achieved yet.
+To explore it further, I enrolled in BSc in Robotics at the best technical university in the country.
+Soon after it became apparent that pursuing a PhD is one of the few ways that could get me closer to the answer.
+
+Robotics is far more than AI - or so I would learn during the first few years of undergrad.
+I studied maths, physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, have I realized how a truly intelligent entity could be controlled.
 Paired with my passion for vision and perception, which I have developed as a photojournalist, it steered me towards my first computer vision course.
 I immediately knew that this intersection of image processing and machine learning was the area I wanted to focus on.
-Without any systems good enough for real world deployment, it was challanging as well as aligned with my original passion for intelligent machines.
+Without any systems good enough for real world deployment, it was challenging as well as aligned with my original passion for intelligent machines.
 
@@
 With cutting-edge computer vision research groups it seemed a great choice.
+The numerical linear algebra and scientific computing, comprising the majority of the content, help enormously with optimization and implementation of algorithms.
 Currently, I am working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof.
@@
 In our recent project we investigate the introspective capacity of neural networks.
-It is understood as the ability of a classifier to assess uncertainty of its prediction given the data [ref].
+It is understood as the ability of a classifier to assess uncertainty of its prediction given the data \cite{introspective}.
 It is a paramount problem in mobile robotics or medicine where a wrong classification might lead to loss of life.
@@
 This, together with my earlier industrial experience, convinced me that I do want to pursue a PhD.
-I love solving scientific problems that do not have ``the best'' solution or a reference specification by going into the deepest details.
+I love solving scientific problems, which do not have ``the best'' solution or a reference specification, by going into the deepest details.
 
@@
 If beliefs about content could be introduced in a form of priors inferred from previous elements of the sequence, it might be possible to increase object classification accuracy in videos and to shrink the network size.
-It would reduce the amount of computation needed, possibly leading to an energy efficient system for portable devices.
 Another interesting problem is an end-to-end RNN for optical flow computation.
 The only end-to-end neural network, while efficient and accurate, does not use temporal information.
-It computes optical flow for any two possibly unrelated images [ref].
+It computes optical flow for any two possibly unrelated images \cite{flownet}.
 I would like to investigate RNN designs capable of computing optical flow while presented with a single image at a time.
-It might lead to a better accuracy with fewer computations, thus better energy efficiency.
+Solution of both problems might thus lead to smaller networks and reduced amounts of computation, possibly leading to higher energy efficiency.
 
@@
 Thank you for considering my application.
+
+\bibliographystyle{ieeetr}
+\bibliography{bibliography}
 \end{document}

### 2015-11-18 phd/sop/sop.tex
alfredos grammar and style corrections

@@
 The Statement of Purpose should not %exceed two pages.
+
 Artificial Intelligence has always fascinated me.
-Long before understanding what computer science stands for have I wondered why AI was not achieved yet.
-To explore it further, I enrolled in BSc in Robotics at the best technical university in the country.
-Soon after it became apparent that pursuing a PhD is one of the few ways that could get me closer to the answer.
+Long before understanding what computer science stands for I wondered why AI has not been achieved yet.
+To explore it further, I enrolled in BSc in Robotics at the best technical university in my country.
+Soon after, it became apparent that pursuing a PhD was one of the few ways that could get me closer to the answer.
 
 Robotics is far more than AI - or so I would learn during the first few years of undergrad.
-I studied maths, physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, have I realized how a truly intelligent entity could be controlled.
+I studied maths, physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, did I realize how a truly intelligent entity could be controlled.
 Paired with my passion for vision and perception, which I have developed as a photojournalist, it steered me towards my first computer vision course.
 I immediately knew that this intersection of image processing and machine learning was the area I wanted to focus on.
-Without any systems good enough for real world deployment, it was challenging as well as aligned with my original passion for intelligent machines.
+Without any systems good enough for real world deployment, $it was challenging as well as aligned with my original passion for intelligent machines.
+$
 
-To learn from the best, I went for an internship to the Computer Vision Lab at Samsung R\&D in Warsaw.
+To learn from the best, I joined Computer Vision Lab at Samsung R\&D in Warsaw for an internship.
 My project was to design and implement a Bag of Words-based pipeline for object classification.
@@
 I focused my research mainly on the visual codebook generation within Bag of Words.
-The final step was to optimize it and port to the Android platform.
+The final step was to optimize it and then port it to the Android platform.
 In my bachelor thesis I investigated how spatial information affects classification performance by implementing a similar system for Kinect-gathered point cloud classification.
@@
 My observation was that deep learning, although often superior to older approaches, required huge computational resources and datasets.
-To mitigate this problem I went to study computational science at the Technical University in Munich.
+To mitigate this problem I went to study computational science at the Technical University of Munich.
 With cutting-edge computer vision research groups it seemed a great choice.
-The numerical linear algebra and scientific computing, comprising the majority of the content, help enormously with optimization and implementation of algorithms.
+Numerical linear algebra and scientific computing, comprising the majority of the coursework, help enormously with optimization and implementation of algorithms.
 Currently, I am working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof.
@@
 It is understood as the ability of a classifier to assess uncertainty of its prediction given the data \cite{introspective}.
-It is a paramount problem in mobile robotics or medicine where a wrong classification might lead to loss of life.
+It is a paramount problem in mobile robotics and medicine where wrong classification might lead to loss of life.
 While high classification accuracy is desired, it is impossible to assess whether a given prediction is accurate in a test setting, where no labels are available.
-It is, therefore, vital to assess uncertainty of predictions.
+It is, therefore, vital to assess the uncertainty of predictions.
 Our intermediate results show that neural networks, augmented with additional layers and a novel cost function, can be jointly trained for classification and uncertainty estimation.
@@
 I would like to investigate RNN designs capable of computing optical flow while presented with a single image at a time.
-Solution of both problems might thus lead to smaller networks and reduced amounts of computation, possibly leading to higher energy efficiency.
+Solving both problems might thus lead to smaller networks and reduced amounts of computation, possibly increasing energy efficiency of such systems.
 
-Passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely well motivated to carry out research as a PhD student at the University of Oxford.
+My passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely motivated to carry out research as a PhD student at the University of Oxford.
 I am confident that I will be able to contribute significantly to and benefit immensely from my stay in Oxford.

### 2015-11-21 ML/hw6/hw.tex
add ML hw6 prob1

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
% \pagestyle{empty}
\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 6} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

Let $\hat{\Phi} = \begin{pmatrix} \Phi \\ \sqrt{\lambda}I_{p \times M} \end{pmatrix} \in \mathrm{R}^{(N+p) \times M}$ be the augmented design matrix and $\hat{\mathbf{z}} = \begin{pmatrix} \mathbf{z} \\ \mathbf{0}_p \end{pmatrix} \in \mathrm{R}^{N+p}$ the augmented target vector. Maximum likelihood estimation of parameters $ \mathbf{w}$ for linear regression is equivalent to minimizing energy given by $E(\mathbf{w}) = (\mathbf{z} - \Phi^T\mathbf{w})^T(\mathbf{z} - \Phi^T\mathbf{w})$. The minimizer $\mathbf{w}_{MLE} = (\Phi^T\Phi)^{-1}\Phi^T\mathbf{z}$ is agnostic of the form and contents of $\mathbf{w}$, $\mathbf{z}$ or $\Phi$. Therefore, we can write the minimizer of the energy resulting from the augmented design matrix and target vector as

\begin{equation}
\begin{align}
 \hat{\mathbf{w}}_{MLE} &= (\hat{\Phi}^T\hat{\Phi})^{-1}\hat{\Phi}^T\hat{\mathbf{z}} \\
	       &= \left( \begin{pmatrix} \Phi^T & \sqrt{\lambda}I_{p \times M}^T \end{pmatrix} \begin{pmatrix} \Phi \\ \sqrt{\lambda}I_{p \times M} \end{pmatrix} \right)^{-1} \begin{pmatrix} \Phi^T & \sqrt{\lambda}I_{p \times M}^T \end{pmatrix} \begin{pmatrix} \mathbf{z} \\ \mathbf{0}_p \end{pmatrix} \\
	       &= \left( \lambda I_{M \times M} + \Phi^T\Phi \right)^{-1} \Phi^T \mathbf{z} \\
	       &= \mathbf{w}_{MLE}^{ridge}
\end{align}
\end{equation}

\end{document}

### 2015-11-21 phd/sop/sop.tex
make SoP more Oxford-specific; this version was submited

@@
 
-Artificial Intelligence has always fascinated me.
+Artificial Intelligence has always tantalized me.
 Long before understanding what computer science stands for I wondered why AI has not been achieved yet.
 To explore it further, I enrolled in BSc in Robotics at the best technical university in my country.
-Soon after, it became apparent that pursuing a PhD was one of the few ways that could get me closer to the answer.
-
-Robotics is far more than AI - or so I would learn during the first few years of undergrad.
-I studied maths, physics and mechanics, material science and manufacturing technologies; not until my first programming class, however, did I realize how a truly intelligent entity could be controlled.
+I studied maths, physics and mechanics, material science and manufacturing technologies; programming and software-oriented courses were my favourite, however.
 Paired with my passion for vision and perception, which I have developed as a photojournalist, it steered me towards my first computer vision course.
 I immediately knew that this intersection of image processing and machine learning was the area I wanted to focus on.
-Without any systems good enough for real world deployment, $it was challenging as well as aligned with my original passion for intelligent machines.
-$
+It became apparent that pursuing a PhD was one of the few ways that could get me closer to solving the AI challange.
 
@@
 My project was to design and implement a Bag of Words-based pipeline for object classification.
-Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of automatically learned low level features in convolutional neural networks (CNNs).
-I focused my research mainly on the visual codebook generation within Bag of Words.
+Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of automatically learned low level features in convolutional neural networks.
+I focused my research primarily on the visual codebook generation within Bag of Words.
 The final step was to optimize it and then port it to the Android platform.
@@
 This summer I did an internship at Bloomberg in London, where I worked on fraud detection in financial transactions.
-The problem can be cast as unsupervised anomaly detection with further verification in a supervised settings.
-I learned how difficult it can be to introduce an innovative approach in a corporate setting.
+The problem can be cast as unsupervised anomaly detection with further verification in a supervised setting.
+I learned how difficult it can be to introduce an innovative approach in a corporate environment.
 This, together with my earlier industrial experience, convinced me that I do want to pursue a PhD.
@@
 In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks.
-The former fascinates me since it is similar to how humans learn.
+The former fascinates me, for it is similar to how humans learn.
 Specifying a reward function  or designating an expert in the inverse problem is far easier than providing supervised training data.
@@
 Solving both problems might thus lead to smaller networks and reduced amounts of computation, possibly increasing energy efficiency of such systems.
+
+I believe that AIMS is a perfect programme for me due to the following reasons.
+Firstly, it is multidisciplinary with robotics as its main motivation.
+I feel it makes it more practical, since it is about solving real world problems.
+Secondly, while I have a number of research ideas, I am not confident planning a 3 year research project just yet.
+The first year of courses and interaction with the faculty is a great means of solving this issue.
+Finally, the Mobile Robotics Group research is extremely close to my interests, e.g.
+introspective capacity of classifiers.
 

### 2015-11-21 cv3d_seminar/meanfield.m
notes on the rt3dcv seminar paper + paper

@@ -1,0 +1,51 @@
+colormap(copper);
+
+% an extraordinarily realistic image
+a = -ones(50,50);
+a(11:40,11:40) = ones(30,30);
+imagesc(a);
+axis image;
+fprintf('original image\n');
+pause
+
+% corrupt with white noise
+[n, m] = size(a);
+sigma = 1;
+b = a + randn(n, m)*sigma;
+imagesc(b)
+axis image;
+fprintf('corrupted image\n');
+pause
+imagesc(b>0)
+axis image;
+fprintf('naive pixel classification\n');
+pause
+  
+% constants for mean field iteration: shift operators and
+% the parameters field strength, temperature, and learning rate 
+% (it is very insensitive to parameters)
+shu = spdiags(ones(m,1),1,m,m);
+shd = spdiags(ones(m,1),-1,m,m);
+shl = spdiags(ones(n,1),1,n,n);
+shr = spdiags(ones(n,1),-1,n,n);
+field = 10;
+temp = 5;
+rate = .5;
+
+% mean field iteration
+c = zeros(n,m);
+imagesc(c)
+axis image;
+fprintf('iteration 0\n');
+pause
+for i = 1:35
+  c = (1-rate)* c + rate * tanh((b + field * (shl*c + shr*c + c*shu + c*shd)) / temp);
+  imagesc(c)
+  axis image;
+  fprintf('iteration %d\n', i);
+  pause
+end
+
+imagesc(c > 0);
+axis image;
+fprintf('final pixel classification\n');

### 2015-11-28 ML/hw7/hw.tex
ml hw7 problems 5-9

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{epstopdf}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}

\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 7} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 5}

Let $\mathbf{w}^0 = \mathbf{0}$ be an initial set of weights and $b^0 = 0$ an initial bias. Let $z_k^i \in \{-1, 1\}$ denote the classification result of the $i^{th}$ training sample $\mathbf{x}^i$, $i = 1, \ldots, N$ using weights $\mathbf{w}^k$ and bias $b^k$. The training algorithm iterates as long as there is at least one misclassification and computes:
\begin{equation}
\begin{align}
  \mathbf{w}^{k+1} &= \mathbf{w}^k + \sum_{t^i \neq z_k^i} t^i \mathbf{x}^i \\
  b^{k+1} &= b^k + \sum_{t^i \neq z_k^i} t^i
\end{align}
\label{eq:train}
\end{equation}

\section*{Problem 6}

\begin{figure}[!ht]
 \includegraphics[width=\textwidth]{perceptron6}
 \caption{Decision boundary after 0-3 iterations.}
 \label{fig:prob6}
\end{figure}

Figure \ref{fig:prob6} shows the decision boundary after 0-3 training iterations. After the third iteration it does not change anymore; The norm of the weights $||\mathbf{w}^k||$ grows to infinity, however.

\section*{Problem 7}

Let $z_k^i \in \{-1, 1\}$ denote the classification result of the $i^{th}$ training sample $\mathbf{x}^i$, $i = 1, \ldots, N$. Since $\mathbf{x}^i$ are linearly separable, it holds that $ \mathbf{x}^i > 0 \implies t^i \widetilde{\mathbf{w}}^T \mathbf{x}^i > 0$. If we define $\gamma = \min\limits_{\mathbf{x}^i} t^i \widetilde{\mathbf{w}}^T \mathbf{x}^i$ then:

\begin{equation}
 \widetilde{\mathbf{w}}^T \mathbf{w}^k = \widetilde{\mathbf{w}}^T \mathbf{w}^{k-1} + \sum_{t^i \neq z_{k-1}^i} t^i \widetilde{\mathbf{w}}^T \mathbf{x}^i \geq \widetilde{\mathbf{w}}^T \mathbf{w}^{k-1} + \gamma \geq \widetilde{\mathbf{w}}^T \mathbf{w}^{0} + k \gamma = k \gamma
\end{equation}

\section*{Problem 8}

From equation \ref{eq:train} and since $||\mathbf{x}^i|| < R$ and the number of misclassified samples at each iteration is at most $N$ we have

\begin{equation}
 \begin{align}
  ||\mathbf{w}^k||^2 &\leq ||\mathbf{w}^{k-1}||^2 + \sum_{t^i \neq z_{k-1}^i} ||\mathbf{x}^i||^2 \\
		     &\leq ||\mathbf{w}^{k-1}||^2 + \sum_{t^i \neq z_{k-1}^i} R^2 \\
		     &\leq ||\mathbf{w}^{k-1}||^2 + N R^2 \\
		     &\leq k N R^2 \\
 \end{align}
\end{equation}

\section*{Problem 9}
  From problems 7 and 8 we know that $||\widetilde{\mathbf{w}}^T \mathbf{w}^k|| \geq k \gamma$ and $||\mathbf{w}^k||^2 < k R^2$. Therefore:

\begin{equation}
 \frac{k^2 \gamma^2}{k R^2} = \frac{k \gamma^2}{R^2} \leq \frac{\left( \widetilde{\mathbf{w}}^T \mathbf{w}^k \right)^2}{||\mathbf{w}^k||^2}
\end{equation}

Since $\left( \widetilde{\mathbf{w}}^T \mathbf{w}^k \right)^2 = \sum_i \widetilde{\mathbf{w}}_i^2 (\mathbf{w}^k)_i^2 = \sum_i \widetilde{\mathbf{w}}_i^2 \sum_i (\mathbf{w}^k)_i^2 = ||\widetilde{\mathbf{w}}||^2 ||\mathbf{w}^k||^2$ and

\begin{equation}
 \frac{k \gamma^2}{R^2} \leq \||\widetilde{\mathbf{w}}||^2
\end{equation}

\begin{equation}
 k \leq \frac{||\widetilde{R^2 \mathbf{w}}||^2}{\gamma^2}
\end{equation}

\end{document}

### 2015-11-28 ML/hw7/perceptron.m

@@ -1,0 +1,67 @@
+function perceptron()
+close all
+% x = [
+%     0 0;
+%     -0.1 0.1;
+%     -0.3 -0.2;
+%     0.2 0.1;
+%     0.2 -0.1;
+%     -1.1 -1;
+%     -1.3 -1.2;
+%     -1 -1;
+%     1 1;
+%     0.9 1.2;
+%     1.1 1;
+%     ];
+
+% y = [1 1 1 1 1 -1 -1 -1 -1 -1 -1]';
+
+x = [-0.7 0.8; -0.9 0.6; -0.3 -0.2; -0.6 0.7; 0.6 -0.8; 0.2 -0.5; 0.3 0.2];
+y = [1 1 1 1 -1 -1 -1]';
+
+ 
+as = x(y == 1, :);
+bs = x(y == -1, :);
+
+
+xx = [min(x(:))-1 max(x(:))+1];
+
+N = 4;
+w = [0 0]';
+b = 0;
+
+figure(1)
+for k = 1:N
+    subplot(2, 2, k)
+    title(sprintf('Iteration %d', k-1));
+   
+    z = out(x, w, b);
+    diff = (z ~= y);
+    
+    errors = sum(diff);
+    slope = -w(1) / w(2);
+    fprintf('Iter = %d, Errors = %d, Slope = %f\n', k-1, errors, slope);
+    
+    hold on
+    plot(as(:, 1), as(:, 2), 'mo', 'MarkerSize', 10, 'LineWidth', 4);
+    plot(bs(:, 1), bs(:, 2), 'rx', 'MarkerSize', 10, 'LineWidth', 4);
+    yy = slope * xx - b / w(2);
+    plot(xx, yy, 'b', 'LineWidth', 2);
+    axis([xx xx]);
+    pause
+    
+    w = w + sum(repmat(y(diff), [1 2]) .* x(diff, :))';
+    b = b + sum(y(diff));
+end
+
+
+end
+
+
+function a = activ(x, w, b)
+    a = x * w + b;
+end
+
+function o = out(x, w, b)
+    o = -1 + 2 * (activ(x, w, b) >= 0);
+end

### 2015-12-01 cv3d_seminar/template.tex
seminar paper with template

% \documentclass[journal, a4paper]{IEEEtran}
\documentclass{llncs}
\usepackage{llncsdoc}

% \usepackage{graphicx}   
% \usepackage{url}
% \usepackage{amsmath}

\begin{document}

	\title{SemanticPaint}
	\author{Adam Kosiorek
	\thanks{Advisor: M.Eng.~Keisuke Tateno, Chair for Computer Aided Medical Procedures \& Augmented Reality, TUM, WS 2015/16.}}
	\markboth{Master Seminar: Recent Trends in 3D Computer Vision}{}
	\institute{}
	\maketitle

\begin{abstract}
	The short abstract (50-80 words) is intended to give the reader an overview of the work.
\end{abstract}

\section{Introduction}

  Capturing your own enviornment has never been easier. SemanticPaint can register your surroundings which, after undergoing a low-level 3D reconstruction, can be semanticly segmented in an interactive way. Not only it works in real time but also requires no pretraining. Adding new object categories on the fly is facilitated by online model updates. The user is provided with instantenous feedback and can re-label any object to correct errors. SemanticPaint makes capturing customized enviornment models with object classes particular to the user's interest easy and efficient.

  The pipeline starts with capturing the enviornment as a stream of noisy RGBD images and combining them into a updated 3D model in an online fashion. The user can choose which objects to label and can do so by ``touching'' a small part of an object or encircling one with his hand and uttering the label. It is recognized by a standard speech recognition system. The label and the information about the affected data points are further passed to a Streaming Random Forest classifier which constantly learns and labels all visible voxels. To further improve classification results a spatially dense labeling is produced by an efficient mean-field inference algorithm. One of the biggest strength of SemanticPaint is the efficiency of each part of the pipeline, which translates to real time performance. Algorithms used in the pipeline were adapted to work on volumentric data in the TSDF format directly in order to avoding the costly conversions to mesh or point-cloud formats. To allow this, the Voxel Oriented Patch features --- a new type of a descriminative feature describing the voxel space --- has been designed. 
  The contributions can be summarized as follows: 3D semantic modeling system, Streaming Random Forest, efficient Mean-Field inference, Voxel Oriented Patch features.

  Numerous applications are possible: (1) building large scale datasets of 3D objects or whole scenes for use in large-scale computer vision systems (2) using the dense semantic labeling of 3D enviornments in robot navigation or to aid people with impaird sight and (3) map enviornments for use in augmanted reality scenerios or games.

  The rest of the paper is organised as follows: Section 2. describes the related work, section 3 details internal data handling, section 4 describes the efficient mean-field inference algorithm, section 5 outlint the Streaming Random Forest classifiers.....

\section{Related Work}

  \cite{VCRF} uses a Voxel-based CRF for simultaneous segmentation and reconstruction. Each voxel contains information about visibility and occlusion as well as group membership. The first two are used to to improve reconstruction by mitigating depth-map noise. The visibility values are constrained by that each ray from the camera can hit only one visible vortex. The group membership information encodes priors given by bounding boxes of detected objects. A graph-cut algorithm is used for inference, which is performed globally. No computational performance was reported.

%   Outline:
%   1) what has generally been done and use cases

    Capturing the geometry of the surrounding world has been a long standing problem. We have managed to reconstruct digital heritage and construct world-scale 3D models with remarkable quality using offline processing from multiple images. Since low-cost RGBD sensors and powerful GPUs have become available, whch enabled online 3D scanning, augmented reality or using 3D enviornment models for navigation purposes.

%   2) scene understanding

      There are a number of approaches to scene understanding. Some of them work on 2D RGB images, other reconstruct geometry from multiple RGB images but still use only 2D data for classification purposes and back-project the results into the 3D model. Yet other approaches focus on geometric data exclusively. Since RGBD sensors are available there has been a growing interest in working with RGBD data, point clouds or voxel representations. There has been some work on segmenting 3D scenes and meshes, detect objects in small scenes or replace them with synthetic models.

      There has been some work on semantic segmentation of 3D meshes; These methods consider mostly noise-free meshes, do not work in real time and use geometric features only. Some work has been done on matching scan data with databases of synthetic 3D models. It would allow to replace noisy point clouds with detailed models to improve reconstruction quality and decrease memory requirements by exploiting repetetiveness. Usually a model is split into parts individually matched against an offline-built database. It is too slow for real-time usage.

      There exists an online SLAM system that recognizes objects and perform online model updatres, but it supports only a single object class and relatively small scenes. What's appealing is that it does semantic recognition and reconstruction.

      There are no approaches that can handle an outdoor setting. Most of the work focused on image classification. No system works in real-time nor does online model updates. They also do not use ful 3d information, or do so in a global optimization setting which is slow.

% Areas:
%   1. Geometry Acquisition and 3D Reconstruction Techniques:
%     1) digitized cultural heritage with remarkable quality [Levoy et al. 2000]
%   
%     2) world-scale, Internet-accessible, 3D maps reconstructed using street-side [Pollefeys et al. 2008], aerial [Hirschmuller 2008] and online photo collections [Snavely et al.2006; Shan et al. 2013]
% 
%     3) Methods for real-time dense reconstructions, even over large physical scales, with only a single commodity depth or RGB camera have been demonstrated [Rusinkiewicz et al. 2002; Newcombe et al. 2011; Izadi et al. 2011; Newcombe et al. 2011; Chen et al. 2013; Nießner et al. 2013; Pradeep et al. 2013]. Applications such as: live 3D scanning, physically-plausible augmented reality, autonomous robot or vehicle guidance, and 3D fabrication.
%       
%   2. Scene Understanding and Semantic Modeling - interpreting the content of captured 3D models:
%   
%     1) 2D: automatically partition RGB image into semantically meaningful regions [Shotton et al. 2006]
%     
%     2) geometric reasoning to extract 3D structure from single RGB images [Gupta et al. 2010] or exlicit object detection [Yao et al. 2012]).
%     
%     3) RGB + Depth [Silberman and Fergus 2011; Silberman et al. 2012; Couprie et al. 2013; Ren et al. 2012; K ̈ahler and Reid 2013]
%     
%     4) 3D point clouds [Brostow et al.2008; Koppula et al. 2011; Anand et al. 2013; St ̈uckler et al. 2013]
%     
%     5) meshes [Valentin et al. 2013]
%     
%     6) voxel representations [Kim et al. 2013; Karpathy et al. 2013; Salas-Moreno et al. 2013; H ̈ane et al.2013]. 
%       
%     7) semantic segmentation of meshes [Chen et al. 2009]
%     
%     9) dense segmentation of reconstructed scenes [Lin et al. 2013; Sengupta et al. 2013; Ladick`y et al. 2012; Valentin et al. 2013], even in an online manner [Herbstet al. 2014]
%     
%     10) localize objects in small scenes [Abdelrahman et al.2013; Bonde et al. 2013; Karpathy et al. 2013; Lin et al. 2013]
%     
%     11) replace objects with synthetic models [Salas-Moreno et al. 2013; Kim et al. 2012; Shao et al. 2012; Nan et al. 2012; Wang et al. 2014].
%       
%     12) capturing large and compelling datasets which have moved from traditional 2D object images to RGB-D and full 3D scenes [Xiao et al. 2010; Xiao et al. 2013; Geiger et al. 2012].
%     
%     
%   3. Computer Graphics
%   
%     1) automatically segmenting 3D meshes into semantic parts [Chen et al. 2009; Kalogerakis et al. 2010; Shapira et al. 2010; Kim et al. 2013], including incremental depth camera-based methods [Shen et al. 2012]. Most of these methods consider only connected noise-free meshes, and geometric properties, ignoring the appearance. Furthermore, these techniques operate only on single objects, and do not operate in real-time.
%     
%     2) matching scan data to synthetic 3D model databases [Kim et al. 2012; Nan et al. 2012; Shao et al. 2012], with the aim to replace  noisy point clouds with detailed CAD models. These approaches are compelling in that they increase final reconstruction fidelity and exploit repetition of objects to minimize the memory footprint. These systems first perform automatic or interactive segmentation of the scene into constituent parts which are then individually matched to the model database. However, these techniques require a model database to be built and learned offline, and the test-time matching techniques can take seconds to minutes to perform.
%     
%     5) online SLAM system that can recognize objects and update the model live [Salas-Moreno et al. 2013] the model database is still captured and generated offline. Only a single object class (chair) is recognized and it is unclear how the system can support larger surfaces such as floors, walls and ceilings. However, this system demonstrates the power of semantic recognition alongside the reconstruction process, improving relocalization, memory efficiency, and loop closure.
%     
%     6) This type of semantic information has also been explored in the context of bundle adjustment [Fioraio and Di Stefano 2013], and extended to sparse map representations [Ramos et al. 2008; Castle et al. 2007].
%     
%   4. Outdoor Scene Labeling:
%   
%     1) classification of images [Brostow et al. 2008; Posner et al. 2009; Ladick`y et al. 2012]. 
%     
%     2) Dense semantic 3D reconstruction with labeling performed on the images and projectred to the final model, which limits the use of full 3D geometry in their inference [Sengupta et al. 2013].
%     
%     3) Joint volumetric dense reconstruction and semantic segmentation using computationally complex global optimization [H ̈ane et al. 2013].
%     
%     4) Decomposition of outdoor scenes into semantic part; employs 3D model matching techniques similar to [Kim et al. 2012; Nan et al. 2012; Shao et al. 2012] to create reconstructions from LiDAR data [Lin et al. 2013]. None of these systems operate in a real-time or in an online manner.    
%     
%   5. Indoor Scene Recognition:
%   
%     1) using RGB-D sensors [Silberman and Fergus 2011; Silberman et al. 2012; Couprie et al. 2013; Ren et al. 2012; K ̈ahler and Reid 2013]. Classification or recognition is performed in image-space, along with 3D priors to aid segmentation. Again these systems fail to exploit full 3D geometry and are the counterpart of image-based segmentation but for RGB-D frames. 
%     
%     2) [Valentin et al. 2013] exploits 3D meshes and geometric and appearance features for improved inference in outdoor and indoor scenes.
%     
%     3) [Kim et al. 2013] use a voxel-based conditional random field (CRF) for segmentation and occupancy-grid based reconstruction.
%     
%     However, these techniques are not efficient enough to be used in an online system, and operate only on coarse reconstructions.
%     
%     
% Our approach differs from these systems in several compelling
% ways. Firstly our system runs entirely online and interactively, in-
% cluding data capture, feature computation, labeling, segmentation,
% learning and filtering. Second, our pipeline leads to robust and dense
% object labels directly on the acquired 3D model. Finally, in our
% system, the user is ‘in the loop’ during the labeling and training
% process, allowing the system to evolve to new object classes in an
% online fashion, and allowing the user to label a minimal amount and
% correct any mistakes interactively. This allows the user to rapidly
% build up models personalized to their spaces and goals

\section{Pipeline}
\section{Mean-Field Inference}
\section{Streaming Random Forest}
\section{Voxel-Oriented Patch Features}
\section{Qualitative Results}
\section{Quantitative Results}
\section{Discussion}
\section{Concolusions}

% 	\begin{table}[!hbt]
% 		% Center the table
% 		\begin{center}
% 		% Title of the table
% 		\caption{Simulation Parameters}
% 		\label{tab:simParameters}
% 		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
% 		\begin{tabular}{|c|c|}
% 			% To create a horizontal line, type \hline
% 			\hline
% 			% To end a column type &
% 			% For a linebreak type \\
% 			Information message length & $k=16000$ bit \\
% 			\hline
% 			Radio segment size & $b=160$ bit \\
% 			\hline
% 			Rate of component codes & $R_{cc}=1/3$\\
% 			\hline
% 			Polynomial of component encoders & $[1 , 33/37 , 25/37]_8$\\
% 			\hline
% 		\end{tabular}% 		\end{center}
% 	\end{table}
% 
% 	\begin{figure}[!hbt]
% 		% Center the figure.
% 		\begin{center}
% 		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
% 		\includegraphics[width=\columnwidth]{plot_tf}
% 		% Create a subtitle for the figure.
% 		\caption{Simulation results on the AWGN channel. Average throughput $k/n$ vs $E_s/N_0$.}
% 		Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
% 		\label{fig:tf_plot}
% 		\end{center}
% 	\end{figure}

\bibliographystyle{splncs}
\begin{thebibliography}{1}

	%Each item starts with a \bibitem{reference} command and the details thereafter.
	\bibitem{HOP96} % Transaction paper
	J.~Hagenauer, E.~Offer, and L.~Papke. Iterative decoding of binary block
	and convolutional codes. {\em IEEE Trans. Inform. Theory},
	vol.~42, no.~2, pp.~429–-445, Mar. 1996.
	\bibitem{MJH06} % Conference paper
	T.~Mayer, H.~Jenkac, and J.~Hagenauer. Turbo base-station cooperation for intercell interference cancellation. {\em IEEE Int. Conf. Commun. (ICC)}, Istanbul, Turkey, pp.~356--361, June 2006.

	\bibitem{Proakis} % Book
	J.~G.~Proakis. {\em Digital Communications}. McGraw-Hill Book Co.,
	New York, USA, 3rd edition, 1995.

	\bibitem{talk} % Web document
	F.~R.~Kschischang. Giving a talk: Guidelines for the Preparation and Presentation of Technical Seminars.
	\url{http://www.comm.toronto.edu/frank/guide/guide.pdf}.

	\bibitem{VCRF}
	KIM , B.-S., KOHLI , P., AND SAVARESE , S. 2013. 3D scene understanding by voxel-CRF. In Proc. ICCV.

\end{thebibliography}

% Your document ends here!
\end{document}

### 2015-12-02 phd/sop/stanford.tex
stanford sop

@@
 
+\setlength{\parindent}{0em}
+\setlength{\parskip}{1em}
+
 \begin{document}
@@
 {\Large Statement of Purpose} \\[.3in]
-{\large Robert S. French}
+{\large Adam Kosiorek}
 \end{center}
 
-\vspace*{.5in}
+% \vspace*{.5in}
 
-My primary interests in computer science are computer architecture and
-compiler technology.  I have participated in several projects during
-my undergraduate career at MIT that have provided me with experience
-in these areas, and I have recently specialized in parallel processing
-systems.
+%The Statement of Purpose should describe succinctly your reasons for applying to the proposed %program at Stanford, your preparation for this field of study, research interests, future career %plans, and other aspects of your background and interests which may aid the admissions committee %in evaluating your aptitude and motivation for graduate study.
+The Statement of Purpose should not %exceed two pages.
 
-My interest in computer architecture was first kindled during the
-design contest for the MIT course ``Computation Structures.''  The
-contest consists of redesigning an 8-bit microcoded processor to
-improve its performance.  Our two-person team constructed a 16-bit
-design for the processor and achieved at speedup of over 750 times.
-As part of the design, I implemented a microcode assembler, a
-microcode simulator, various levels of logic simulators, and over 80
-pages of microcode.  The contest taught me a great deal about
-low-level system design and the tradeoffs involved in designing fast
-systems.
+Artificial Intelligence has always tantalized me.
+Long before understanding what computer science stands for I wondered why AI has not been achieved yet.
+To explore it further, I enrolled in BSc in Robotics at the best technical university in my country.
+I studied maths, physics and mechanics, material science and manufacturing technologies; programming and software-oriented courses were my favourite, however.
+Paired with my passion for vision and perception, which I have developed as a photojournalist, it steered me towards my first computer vision course.
+I immediately knew that this intersection of image processing and machine learning was the area I wanted to focus on.
+It became apparent that pursuing a PhD was one of the few ways that could get me closer to solving the AI challange.
 
-My next architecture project was for the course ``Digital Systems
-Laboratory,'' in which my partner and I implemented a hardware
-simulator for a highly-parallel, fine-grained optical computer.  The
-system was based on recent research by Alan Huang at Bell Laboratories
-in ``computational origami,'' a simple method of simulating an
-arbitrarily large number of processors using a small set of processing
-elements with long delay lines.  As part of the project I designed a
-compiler and a placement and routing system for allocating processors.
-I presented a paper on the algorithm at the MIT-ACM Undergraduate
-Computer Science Conference in April, 1989, and was invited to present
-my findings at Bell Labs.
+To learn from the best, I joined Computer Vision Lab at Samsung R\&D in Warsaw for an internship.
+My project was to design and implement a Bag of Words-based pipeline for object classification.
+Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of automatically learned low level features in convolutional neural networks.
+I focused my research primarily on the visual codebook generation within Bag of Words.
+The final step was to optimize it and then port it to the Android platform.
+In my bachelor thesis I investigated how spatial information affects classification performance by implementing a similar system for Kinect-gathered point cloud classification.
+Afterwards, I continued my work at Samsung, where I turned to deep learning techniques for object classification and image duplicate detection.
 
-I designed and implemented part of {\em Zephyr\/}, a distributed
-multicast message system, as an employee of MIT's Project Athena.
-Project Athena is an eight year experiment in the use of computers in
-education, and consists of approximately 1,000 workstations connected
-by a campus-wide network.  This setup provides a unique environment in
-which to develop scalable distributed software.  I investigated
-various communications methods, and was the primary architect of the
-protocol and the authentication system used by {\em Zephyr\/}.  Our
-development team published a paper on {\em Zephyr\/} in the Winter
-1988 Usenix Conference Proceedings.
+My observation was that deep learning, although often superior to older approaches, required huge computational resources and datasets.
+To mitigate this problem I went to study computational science at the Technical University of Munich.
+With cutting-edge computer vision research groups it seemed a great choice.
+Numerical linear algebra and scientific computing, comprising the majority of the coursework, help enormously with optimization and implementation of algorithms.
+Currently, I am working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof.
+Cremers.
+In our recent project we investigate the introspective capacity of neural networks.
+It is understood as the ability of a classifier to assess uncertainty of its prediction given the data \cite{introspective}.
+It is a paramount problem in mobile robotics and medicine where wrong classification might lead to loss of life.
+While high classification accuracy is desired, it is impossible to assess whether a given prediction is accurate in a test setting, where no labels are available.
+It is, therefore, vital to assess the uncertainty of predictions.
+Our intermediate results show that neural networks, augmented with additional layers and a novel cost function, can be jointly trained for classification and uncertainty estimation.
+The topic might expand into my master thesis.
 
-My current research is for my undergraduate thesis, in which I am
-investigating the use of fractals to measure communication locality in
-distributed systems.  I hope to extend this research and develop new
-algorithms for efficient processor allocation.
+This summer I did an internship at Bloomberg in London, where I worked on fraud detection in financial transactions.
+The problem can be cast as unsupervised anomaly detection with further verification in a supervised setting.
+I learned how difficult it can be to introduce an innovative approach in a corporate environment.
+This, together with my earlier industrial experience, convinced me that I do want to pursue a PhD.
+I love solving scientific problems, which do not have ``the best'' solution or a reference specification, by going into the deepest details.
 
-I plan to continue my study of architecture and compiler technology in
-graduate school, and I feel that Stanford is the ideal place to study
-these fields.  Stanford offers an excellent selection of courses and
-supports a vigorous research program in my areas of interest.  I enjoy
-developing theories and building hardware systems on which to test
-them, and hope to be involved in applied research at Stanford.
+In my doctoral research project I would like to focus on recurrent neural networks for computer vision.
+RNNs are well suited to sequential information processing and, while often used for NLP tasks, they have been somehwat underutilized in computer vision.
+If beliefs about content could be introduced in a form of priors inferred from previous elements of a sequence, it might be possible to increase object classification accuracy in videos and to shrink the network size.
+Another interesting problem is an end-to-end RNN for optical flow computation.
+The only end-to-end neural network, while efficient and accurate, does not use temporal information.
+It computes optical flow for any two possibly unrelated images \cite{flownet}.
+I would like to investigate RNN designs capable of computing optical flow while presented with a single image at a time.
+Solving both problems might thus lead to smaller networks and reduced amounts of computation, possibly increasing energy efficiency of such systems.
 
-After receiving a Ph.D., I plan to pursue an academic career in
-computer science research.
+I believe that the CS PhD programme at Stanford is perfect for me due to the following reasons.
+Firstly, the courses taught by world renown scientists e.g.
+Prof.
+Fei-Fei's ``High-Level Vision Behaviors, Neurons and Computational Models'' provide an extraordinary introduction to research in deep learning I would like to conduct.
+Secondly, research of the Stanford Visian Lab is multimodal in that it successfuly merges natural language processing with computer vision \cite{imageannot}.
+I find it very relevant, for it might lead to the improvement of living standard for people with impaired sight.
+Finally, Stanford is \emph{the} most exciting area for any computer scientist to live in.
+The diversity and the amount of talent there bolster personal development like no other place in the world.
 
+My passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely motivated to pursue a PhD programme at the Stanford University.
+I am confident that I will be able to contribute significantly to and benefit immensely from my stay at Stanford.
+Thank you for considering my application.
+
+\bibliographystyle{ieeetr}
+\bibliography{bibliography}
 \end{document}

### 2015-12-02 ML/hw6/hw.tex
ml homework and some random previously unsubmitted files for paralle_programming and HiWi

@@
 
-\begin{equation}
+%\begin{equation}
 \begin{align}
@@
 \end{align}
+%\end{equation}
+
+\section*{Problem 2}
+
+We want to prove the following identity:
+
+\[ (M + vv^T)^{-1} = M^{-1} - \frac{(M^{-1} v) (v^T M^{-1})  }{1 + v^T M^{-1} v} \]
+
+To prove the identity we will use the fact that for any matrix M:
+
+\[
+M M^{-1} = I
+\]
+
+where I is the identity matrix. So we multiply both sides by the inverse of the left-hand side
+
+\[
+M + vv^T
+\]
+
+which results in
+
+\begin{equation}
+I = I + M^{-1}vv^T - \frac{M^{-1} v v^T + M^{-1} v v^T M^{-1} v v^T}{1 + v^T M^{-1} v}
 \end{equation}
 
+$ v^T M^{-1} v $ is a scalar, so:
+
+\[
+M^{-1} v v^T + M^{-1} v v^T M^{-1} v v^T = ( 1 + v^T M^{-1} v ) M^{-1} v v^T
+\]
+
+(5) then simplifies to:
+
+\[
+I = I + M^{-1}vv^T - M^{-1}vv^T = I
+\]
+
+so the identity is verified.
+
+Using the identity above, we need to prove the following inequality:
+
+\begin{equation}
+\sigma^2_{N+1} (x) \leq \sigma^2_{N} (x)
+\end{equation}
+
+We use the definition of $\sigma^2_{N}$ and the definition of $S_N$ (here $ \Phi_N $ is the design matrix with the first N observations).
+
+\[
+\sigma^2_{N} (x) = \frac{1}{\beta} + \phi(x)^T S_N \phi(x)
+\]
+
+\[
+S^{-1}_{N} = \alpha I + \beta (\Phi^T_{N}) \Phi_N
+\]
+
+We now use the following identity:
+
+\[
+(\Phi^T_{N}) \Phi_N = \sum_{k=1}^N \phi(x_k) \phi(x_k)^T
+\]
+
+so we can write
+
+\[
+(\phi^T_{N+1}) \phi_{N+1} =(\phi^T_{N}) \phi_N + \phi(x_{N+1}) \phi(x_{N+1})^T
+\]
+
+and hence
+
+\[
+S^{-1}_{N+1} = \alpha I + \beta (\Phi^T_{N+1}) \Phi_{N+1} = \alpha I + \beta ((\Phi^T_{N}) \Phi_N + \phi(x_{N+1}) \phi(x_{N+1})^T) = S^{-1}_{N} + \phi(x_{N+1}) \phi(x_{N+1})^T
+\]
+
+Substituting this into $ \sigma^2_{N+1} (x) $ and using the above matrix identity we get:
+
+\[
+\sigma^2_{N+1} (x) = \frac{1}{\beta} + \phi(x)^T S_{N+1} \phi(x)
+\]
+
+\[
+ = \frac{1}{\beta} + \phi(x)^T ({S^{-1}_{N} + \phi(x_{N+1}) \phi(x_{N+1})^T })^{-1} \phi(x)
+\]
+
+\[
+ = \frac{1}{\beta} + \phi(x)^T (S_{N} - \frac{S_{N} \phi(x_{N+1}) \phi(x_{N+1})^T S_{N}} {1 + \phi(x_{N+1})^T S_{N} \phi(x_{N+1})}) \phi(x) \]
+
+\begin{equation}
+ = \frac{1}{\beta} + \phi(x)^T S_{N} \phi(x) - \frac{ \phi(x)^T S_{N} \phi(x_{N+1}) \phi(x_{N+1})^T S_{N} \phi(x)} {1 + \phi(x_{N+1})^T S_{N} \phi(x_{N+1})}
+\end{equation}
+
+The nominator of the last fraction is greather than zero. Now note that the $S^{-1}_N$ is positive semi-definite since $\beta$ is greater than zero, $\alpha$ is greater than zero, and:
+
+\[
+x^T S^{-1}_{N} x = x^T (\alpha I + \beta \Phi^{T}_{N} \Phi_N  ) x = x^T \alpha I x + x^T \beta \Phi^{T}_{N} \Phi_N x = \alpha x^T I x + \beta (\Phi_N x)^T \Phi_N x
+\]
+
+A matrix A has an eigenvalue if and only if $A^{-1}$ has eigenvalue $\frac{1}{\lambda}$, as
+
+\[
+Av = \lambda v \Leftrightarrow A^{-1} A v = \lambda A^{-1} v \Leftrightarrow \frac{1}{\lambda} v = A^{-1} v 
+\]
+
+With that and the fact that the eigenvalues of positive semi-definite matrices are nonnegative we have shown
+that $S_N$ also is positive semi-definite. Therefore
+
+\[
+\phi(x_{N+1})^T S_{N} \phi(x_{N+1}) \geq 0
+\]
+
+and so the denominator of the last fraction in eq. (7) is positive. We thus have proven the inequality.
+
+\section*{Problem 3}
+Since we spent no money, the model boils down to $y = 10 + \mathcal{N}(0,4)$. The Gaussian is centered on zero and, therefore, we can say that the probability to get more than 10 likes is $0.5$.
+
+\section*{Problem 4}
+
+Spending 1 EUR on advertisements our model gives us 
+\begin{equation}
+y = 15 + \mathcal{N}(0,4) \hspace{0.5cm} \Rightarrow \hspace{0.5cm} \operatorname{E}[y] = 15
+\end{equation}
+with \hspace{0.5cm} $\operatorname{E}[\mathcal{N}(0,4)] = 0$. Hence, we expect 15 likes.
+
 \end{document}

### 2015-12-02 ML/hw7/hw.tex
ml homework and some random previously unsubmitted files for paralle_programming and HiWi

@@
 \vspace*{.5in}
+
+\section*{Problem 1}
+If the data set is linearly separable, any decision boundary separating the two classes will have the property
+\begin{equation}
+w^T\phi(x^n) = \begin{cases} \geq 0 & \mbox{if } z^n = 1, \\ < 0 & otherwise. \\ \end{cases}
+\label{eq:}
+\end{equation}
+
+Moreover, the log-likelihood 
+\begin{align}
+L(w) &= \log p(\{z^i\}|w,\{\phi(x^i)\})\\
+&= \sum^{N}_{n = 1}\left[z^n\log \sigma(w^T\phi(x^n)) + (1-z^n)(\log(1-\sigma(w^T\phi(x^n))))\right]
+\end{align}
+
+will be maximized when $y^n = \sigma(w^T\phi(x^n)) = z^n \hspace{0.3cm}\forall n$.
+This will be the case when the sigmoid function is saturated, which occurs when its argument, $w^T\phi$, goes to $\pm \inf$, i.e., when the magnitude of $w$ goes to infinity.
+
+\section*{Problem 2}
+The basis function 
+\begin{equation}
+\phi(x_1,x_2) = x_1x_2
+\end{equation}
+
+enables us to linearly seperate the crosses from the circles. The hyperplane would go through the origin of the coordinate system.
 

### 2015-12-02 HiWi/matlab/multiply_tests.m

@@ -1,16 +1,15 @@
 function multiply_tests()
     clc, clear all
     
-    sizeA1 = 4;
+    sizeA1 = 3;
     sizeA2 = 3;
-    sizeB1 = 5;
+    sizeB2 = 4;
     
     A = createA(sizeA1, sizeA2)
-    B = createB(sizeB1, sizeA1)
-    C = createC(sizeB1, sizeA2)
+    B = createB(sizeA2, sizeB2)
+    C = createC(sizeA1, sizeB2)
     
-    BtimesA = B*A
-    CplusBtimesA = C + BtimesA
+    AtimesB = A * B
     
     
 end

### 2015-12-02 parallel_programming/nested_reduction.cpp

@@ -1,0 +1,32 @@
+#include <omp.h>
+#include <iostream>
+
+using namespace std;
+
+int main(int argc, char** argv) {
+
+
+	int a[10][10];
+	int b[10]  = {0};
+
+	for(int i = 0; i < 10; ++i) {
+		for(int j = 0; j < 10; ++j) {
+			a[i][j] = i * 10 + j;
+		}
+	}
+
+
+#pragma omp parallel for
+	for(int i = 0; i < 10; ++i) {
+
+#pragma omp for reduction(+: b[i]) nowait
+		for(int j = 0; j < 10; ++j) {
+			b[i] += a[i][j];
+		}
+	}
+	
+	for(int i = 0; i < 10; ++i) {
+		cout << b[i] << ", ";
+        }
+        cout << endl;
+}

### 2015-12-02 cv3d_seminar/paper.tex
seminar presentation template

--- 2015-12-01 cv3d_seminar/template.tex
@@
 
-  \cite{VCRF} uses a Voxel-based CRF for simultaneous segmentation and reconstruction.
-Each voxel contains information about visibility and occlusion as well as group membership.
-The first two are used to to improve reconstruction by mitigating depth-map noise.
-The visibility values are constrained by that each ray from the camera can hit only one visible vortex.
-The group membership information encodes priors given by bounding boxes of detected objects.
-A graph-cut algorithm is used for inference, which is performed globally.
-No computational performance was reported.
-
 %   Outline:
@@
 They also do not use ful 3d information, or do so in a global optimization setting which is slow.
+
+  \cite{Monofusion} enables 3D reconstruction of small since using a single off-the-shelf RGB camera.
+It uses a sparse tracking method to first estimate the camera's pose and then select key frames and relative to them secondary frames from which 3D stereo reconstruction is perfomed.
+Results achieved are similar to KinectFusion with the only limitation being low precision of reconstruction of textureless surfaces.
+
+  \cite{ChangeDetection} is a first step towards online simultaneous registration and segmentation.
+Using RGBD images, the framework constructs a model of the enviornment and updates it each time a new frame comes in.
+When a significant change in the model is detected, the resulting model is split into a static and a dynamic part, where the latter is assumed to have moved in space.
+The movement is detected by comparing the expected and observed intensity values at each voxel.
+The system is online, but is far from real time with 0.7 to 2s processing time per frame.
+
+  \cite{mesh}
+
+  \cite{VCRF} uses a Voxel-based CRF for simultaneous segmentation and reconstruction.
+Each voxel contains information about visibility and occlusion as well as group membership.
+The first two are used to to improve reconstruction by mitigating depth-map noise.
+The visibility values are constrained by that each ray from the camera can hit only one visible vortex.
+The group membership information encodes priors given by bounding boxes of detected objects.
+A graph-cut algorithm is used for inference, which is performed globally.
+No computational performance was reported.
 
@@
 
+	\bibitem{Monofusion}
+	PRADEEP , V., RHEMANN , C., IZADI , S., ZACH , C., BLEYER , M., AND BATHICHE , S. 2013. Monofusion: Real-time 3D reconstruction of small scenes with a single web camera. In Proc. ISMAR.
+
+	\bibitem{ChangeDetection}
+	HERBST , E., HENRY , P., AND FOX , D. 2014. Toward online 3-d object segmentation and mapping. In IEEE International Conference on Robotics and Automation (ICRA).
+
+	\bibitem{mesh}
+	VALENTIN , J. P., SENGUPTA , S., WARRELL , J., SHAHROKNI , A., AND TORR , P. H. 2013. Mesh based semantic modelling for indoor and outdoor scenes. In Proc. CVPR.
+
 \end{thebibliography}

### 2015-12-02 cv3d_seminar/presentation.tex
seminar presentation template

\documentclass[mathserif]{beamer}
\input{macro}

\mode<presentation>
{

  	\usetheme{Pittsburgh}

  	%\usetheme[numbers,
	 %pageofpages=of,% String used between the current page and the total page count.
         % bullet=circle,% Use circles instead of squares for bullets.
          %titleline=true,% Show a line below the frame title.
          %]{Singapore}

%  \setbeamertemplate{footline}[frame number]
%  \usefonttheme[onlysmall]{structurebold}
%  \setbeamercovered{dynamic}
% \setbeamercovered{transparent}

\usebackgroundtemplate{\includegraphics[height=\paperheight]{template}}

\setbeamertemplate{frametitle}{
	\vskip30pt
 	\usebeamercolor[blue!60!green]{frametitle}
  	\centering
  	\insertframetitle
	\vskip5pt
	\color{blue!60!green}{\hrule}
	}
}

\setbeamercolor{math text}{fg=orange!80!black}
\newcommand{\hl}[1]{\textcolor{red!80!black}{\underline{#1}}}

\title[short title]{SemanticPaint}
\author[short presentator]{Adam Kosiorek}
\institute[CAMP]{Computer Aided Medical Procedures \\
Technische Universit\"at M\"unchen}
\date[]{10.12.2015}

\begin{document}

%---------------------------------------------------------------------------------------

\begin{frame}
\titlepage
\end{frame}

%---------------------------------------------------------------------------------------

\begin{frame}{Outline}
\tableofcontents
\end{frame}

%---------------------------------------------------------------------------------------
\section{Introduction}
\begin{frame}
\frametitle{Introduction}

\end{frame}

%---------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Two columns}
\begin{columns}[t]
\begin{column}{0.45\linewidth}
Left
\end{column}
\begin{column}{0.45\linewidth}
Right
\end{column}
\end{columns}
\end{frame}

%---------------------------------------------------------------------------------------
\section{State of the art}

\begin{frame}
\frametitle{Another section}
\end{frame}

%---------------------------------------------------------------------------------------
\begin{frame}
\frametitle{How to include an image}
\begin{center}
\includegraphics[width=0.3\linewidth]{template}
\end{center}
\end{frame}

%---------------------------------------------------------------------------------------
\end{document}

### 2015-12-02 cv3d_seminar/paper.tex
related works finished

@@
 
-  Capturing your own enviornment has never been easier.
-SemanticPaint can register your surroundings which, after undergoing a low-level 3D reconstruction, can be semanticly segmented in an interactive way.
+  Capturing your own environment has never been easier.
+SemanticPaint can register your surroundings which, after undergoing a low-level 3D reconstruction, can be semantically segmented in an interactive way.
 Not only it works in real time but also requires no pretraining.
 Adding new object categories on the fly is facilitated by online model updates.
-The user is provided with instantenous feedback and can re-label any object to correct errors.
-SemanticPaint makes capturing customized enviornment models with object classes particular to the user's interest easy and efficient.
+The user is provided with instantaneous feedback and can re-label any object to correct errors.
+SemanticPaint makes capturing customized environment models with object classes particular to the user's interest easy and efficient.
 
-  The pipeline starts with capturing the enviornment as a stream of noisy RGBD images and combining them into a updated 3D model in an online fashion.
+  The pipeline starts with capturing the environment as a stream of noisy RGBD images and combining them into a updated 3D model in an online fashion.
 The user can choose which objects to label and can do so by ``touching'' a small part of an object or encircling one with his hand and uttering the label.
@@
 One of the biggest strength of SemanticPaint is the efficiency of each part of the pipeline, which translates to real time performance.
-Algorithms used in the pipeline were adapted to work on volumentric data in the TSDF format directly in order to avoding the costly conversions to mesh or point-cloud formats.
-To allow this, the Voxel Oriented Patch features --- a new type of a descriminative feature describing the voxel space --- has been designed.
+Algorithms used in the pipeline were adapted to work on volumetric data in the TSDF format directly in order to avoiding the costly conversions to mesh or point-cloud formats.
+To allow this, the Voxel Oriented Patch features --- a new type of a discriminative feature describing the voxel space --- has been designed.
   The contributions can be summarized as follows: 3D semantic modeling system, Streaming Random Forest, efficient Mean-Field inference, Voxel Oriented Patch features.
 
-  Numerous applications are possible: (1) building large scale datasets of 3D objects or whole scenes for use in large-scale computer vision systems (2) using the dense semantic labeling of 3D enviornments in robot navigation or to aid people with impaird sight and (3) map enviornments for use in augmanted reality scenerios or games.
+  Numerous applications are possible: (1) building large scale datasets of 3D objects or whole scenes for use in large-scale computer vision systems (2) using the dense semantic labeling of 3D environments in robot navigation or to aid people with impaired sight and (3) map environments for use in augmented reality scenarios or games.
 
-  The rest of the paper is organised as follows: Section 2.
-describes the related work, section 3 details internal data handling, section 4 describes the efficient mean-field inference algorithm, section 5 outlint the Streaming Random Forest classifiers.....
+  The rest of the paper is organized as follows: Section 2.
+describes the related work, section 3 details internal data handling, section 4 describes the efficient mean-field inference algorithm, section 5 outline the Streaming Random Forest classifiers.....
 
@@
 
-%   Outline:
-%   1) what has generally been done and use cases
+  \textbf{Acquisition and Reconstruction.} Capturing the geometry of the surrounding world has been a long standing problem \cite{old}.
+An offline processing of multiple images allowed to reconstruct digital heritage and construct world-scale 3D models with remarkable quality \cite{heritage}.
+The inception of low-cost RGBD sensors and powerful GPUs enabled online 3D scanning, augmented reality or using 3D environment models for navigation purposes \cite{fusion}.
+\cite{Monofusion} enables 3D reconstruction of small scenes using a single off-the-shelf RGB camera.
+It uses a sparse tracking method to first estimate the camera's pose and then select key frames and relative to them secondary frames, from which 3D stereo reconstruction is performed.
+The achieved results are similar to KinectFusion with the only limitation being that the precision of texture-less surfaces' reconstruction is somewhat lacking.
 
-    Capturing the geometry of the surrounding world has been a long standing problem.
-We have managed to reconstruct digital heritage and construct world-scale 3D models with remarkable quality using offline processing from multiple images.
-Since low-cost RGBD sensors and powerful GPUs have become available, whch enabled online 3D scanning, augmented reality or using 3D enviornment models for navigation purposes.
-
-%   2) scene understanding
-
-      There are a number of approaches to scene understanding.
-Some of them work on 2D RGB images, other reconstruct geometry from multiple RGB images but still use only 2D data for classification purposes and back-project the results into the 3D model.
-Yet other approaches focus on geometric data exclusively.
-Since RGBD sensors are available there has been a growing interest in working with RGBD data, point clouds or voxel representations.
-There has been some work on segmenting 3D scenes and meshes, detect objects in small scenes or replace them with synthetic models.
-
-      There has been some work on semantic segmentation of 3D meshes; These methods consider mostly noise-free meshes, do not work in real time and use geometric features only.
-Some work has been done on matching scan data with databases of synthetic 3D models.
-It would allow to replace noisy point clouds with detailed models to improve reconstruction quality and decrease memory requirements by exploiting repetetiveness.
-Usually a model is split into parts individually matched against an offline-built database.
-It is too slow for real-time usage.
-
-      There exists an online SLAM system that recognizes objects and perform online model updatres, but it supports only a single object class and relatively small scenes.
-What's appealing is that it does semantic recognition and reconstruction.
-
-      There are no approaches that can handle an outdoor setting.
-Most of the work focused on image classification.
-No system works in real-time nor does online model updates.
-They also do not use ful 3d information, or do so in a global optimization setting which is slow.
-
-  \cite{Monofusion} enables 3D reconstruction of small since using a single off-the-shelf RGB camera.
-It uses a sparse tracking method to first estimate the camera's pose and then select key frames and relative to them secondary frames from which 3D stereo reconstruction is perfomed.
-Results achieved are similar to KinectFusion with the only limitation being low precision of reconstruction of textureless surfaces.
-
-  \cite{ChangeDetection} is a first step towards online simultaneous registration and segmentation.
-Using RGBD images, the framework constructs a model of the enviornment and updates it each time a new frame comes in.
-When a significant change in the model is detected, the resulting model is split into a static and a dynamic part, where the latter is assumed to have moved in space.
-The movement is detected by comparing the expected and observed intensity values at each voxel.
-The system is online, but is far from real time with 0.7 to 2s processing time per frame.
-
-  \cite{mesh}
-
-  \cite{VCRF} uses a Voxel-based CRF for simultaneous segmentation and reconstruction.
+  \noindent
+  \textbf{Scene Understanding.} Object recognition, detection and segmentation has been done with 2D RGB images, RGBD data, point clouds and volumetric representations.
+\cite{VCRF} uses a Voxel-based CRF for simultaneous reconstruction and segmentation.
 Each voxel contains information about visibility and occlusion as well as group membership.
@@
 The group membership information encodes priors given by bounding boxes of detected objects.
-A graph-cut algorithm is used for inference, which is performed globally.
+A graph-cut algorithm is used for global inference.
+The whole scene has to be registered beforehand, since any change in the CRF's structure would require restarting the inference algorithm.
 No computational performance was reported.
+\cite{ChangeDetection} is a first step towards online simultaneous registration and segmentation.
+Using RGBD images, the framework constructs a model of the environment and updates it each time a new frame comes in.
+When a significant change in the model is detected, the resulting model is split into a static and a dynamic part, where the latter is assumed to have moved in space.
+The movement is detected by comparing the expected and observed intensity values at each voxel.
+The system is online, but is far from real time with 0.7 to 2s processing time per frame.
+\cite{mesh} shows that converting to another data representation can help to improve segmentation accuracy and inference speed.
+TSDF model is used to reconstruct a 3D scene from a sequence of RGBD images.
+The volumetric representation is triangulated and a 3D mesh is recovered.
+Next, visual features are computed directly on images and projected into the 3D model, while geometric features are computed on the mesh directly.
+Segmentation is done via a CRF.
+The approach is tested on the augmented KITTI and NYU datasets for indoor and outdoor scene segmentation where it delivers state-of-the-art results.
 
@@
 
-\section{Pipeline}
-\section{Mean-Field Inference}
-\section{Streaming Random Forest}
-\section{Voxel-Oriented Patch Features}
-\section{Qualitative Results}
-\section{Quantitative Results}
-\section{Discussion}
-\section{Concolusions}
+% \section{Pipeline}
+% \section{Mean-Field Inference}
+% \section{Streaming Random Forest}
+% \section{Voxel-Oriented Patch Features}
+% \section{Qualitative Results}
+% \section{Quantitative Results}
+% \section{Discussion}
+% \section{Conclusions}
 
@@
 
-	%Each item starts with a \bibitem{reference} command and the details thereafter.
-	\bibitem{HOP96} % Transaction paper
-	J.~Hagenauer, E.~Offer, and L.~Papke. Iterative decoding of binary block
-	and convolutional codes. {\em IEEE Trans. Inform. Theory},
-	vol.~42, no.~2, pp.~429–-445, Mar. 1996.
-	\bibitem{MJH06} % Conference paper
-	T.~Mayer, H.~Jenkac, and J.~Hagenauer. Turbo base-station cooperation for intercell interference cancellation. {\em IEEE Int. Conf. Commun. (ICC)}, Istanbul, Turkey, pp.~356--361, June 2006.
-
-	\bibitem{Proakis} % Book
-	J.~G.~Proakis. {\em Digital Communications}. McGraw-Hill Book Co.,
-	New York, USA, 3rd edition, 1995.
-
-	\bibitem{talk} % Web document
-	F.~R.~Kschischang. Giving a talk: Guidelines for the Preparation and Presentation of Technical Seminars.
-	\url{http://www.comm.toronto.edu/frank/guide/guide.pdf}.
+  \bibitem{old}
+  ROBERTS , L. G. 1963. Machine perception of three-dimensional solids. Ph.D. thesis, Massachusetts Institute of Technology.
 
@@
 
+  \bibitem{heritage}
+  LEVOY , M., PULLI , K., CURLESS , B., RUSINKIEWICZ , S., KOLLER , D., EREIRA , L., GINZTON , M., ANDERSON , S., DAVIS , J., G NSBERG , J., ET AL .
+2000.
+The digital Michelangelo project: 3D scanning of large statues.
+In Proc.
+SIGGRAPH.
+ACM.
+
+  \bibitem{fusion}
+  NEWCOMBE , R.
+A., IZADI , S., HILLIGES , O., MOLYNEAUX , D., KIM , D., DAVISON , A.
+J., KOHLI , P., SHOTTON , J., HODGES , S., AND FITZGIB-BON , A.
+2011.
+KinectFusion: Real-time dense surface mapping and tracking.
+In Proc.
+ISMAR.
+
 \end{thebibliography}

### 2015-12-03 cv3d_seminar/paper.tex
streaming random forests finished

@@
 
-% \usepackage{graphicx}   
+\usepackage{graphicx}   
 % \usepackage{url}
@@
 
-  The pipeline starts with capturing the environment as a stream of noisy RGBD images and combining them into a updated 3D model in an online fashion.
+  The pipeline starts with capturing the environment as a stream of noisy RGBD images and combining them into a  3D model updated in an online fashion.
 The user can choose which objects to label and can do so by ``touching'' a small part of an object or encircling one with his hand and uttering the label.
@@
 
+\section{Space Representation}
+
+  The space is represented by a 3D volume constructed from the sequence of RGBD images using Truncated Signed Distance Function (TSDF) \cite{TSDF}.
+Large-scale scene handling is facilitated by the hashed volumetric representation of \cite{hashed_voxel} with a 6mm$^{3}$ voxel.
+Each voxel contains: TSDF distance, colour information in the LAB space, a weight for averaging distance values, a user annotation class, a predictied class, an outcome of the mean-field inference from before and after classification and the probability of each class as inferred by the mean-field algorithm, for each class.
+The total number of bits per voxel is $96 + \textrm{[number of classes]} \cdot 16$.
+
+\section{Dynamic Conditional Random Field}
+\section{Efficient Mean-Field Inference}
+\section{Streaming Random Forests}
+
+  A random forest is an ensamble of classification or regression trees, whose outputs are combined to produce a result with lower variance \cite{murphy}.
+Each tree is typically trained only on a subset of the training data $S$ comprised of pairs $(i, l)$ where $i$ is a sample and $l$ its label.
+Let $f(i; \theta) \in \{L, R\}$ denote the binary split function with learned parameters $\theta$, which specify a feature this node uses.
+The tree's output amounts to the probability distribution $P(c_i = l | i)$ stored at each node.
+Trees learn in a greedy way: for each node a split function is chosen from a distribution $\Theta$ of candidate split functions to maximize the information gain.
+
+  \begin{equation} \label{eq:infogain}
+   G(S, S^L, S^R) = H(S) - \sum_{d \in \{L, R\}} \frac{|S^d|}{|S|}H(S^d)
+  \end{equation}
+
+  Where $H(S) = - \sum_{(l, i) \in S} p(c_i = l) \log{p(c_i = l)}$ is Shannon Entropy.
+Training set is then split into subsets $S^d(\theta)$, $d \in \{L, R\}$ used to train child nodes.
+This procedure requires that all training data is available.
+
+  Online Random Forest \cite{online_forest} allow runtime model updates and they differ in that for a new node $n$ they generate a distribution of split function $\Theta_n$ and maintain a set of label statistics for each of them, updated when a new sample comes to the node.
+Label statistics are sufficient to compute information gain and thus the node $n$ is further split if it has seen a predefined number of samples or if the minimal value of the information gain is low enough.
+Storing and udpdating statistics in leaf nodes is computationally expensive which limits runtime performanc
+
+\begin{figure}[!ht]
+  \center
+  \includegraphics[width=0.75\textwidth]{figures/forest}
+  \caption{Splitting reservoirs in Streaming Random Forest.}
+  \label{fig:forest}
+\end{figure}
+
+  Streaming Random Forest begins by storing a reservoir $R_n$ of at most $K$ samples stored in a list $T_n$ at the root node.
+It represents an unbiased sample of all training data seen so far.
+Initially, the first $K$ samples are stored.
+Then, a sample in the reservoir is exchanged with an incoming one with probability $p = \frac{K}{m_n}$, where $m_n$ is the number of samples seen so far at node $n$.
+Splitting the node requires computing the objective function for each $\theta \in \Theta_n$ over $R_n$.
+$H(R_n)$ is computed from the normalized class label histogram $P(l|R_n)$.
+Let $|R_n|$ be the amount of samples stored in the reservoir $n$ and let $R_n^d$ be the splitting induced by $\theta$.
+Finally, the split function is chosen as $f_n = \arg \max_{\theta \in \Theta_n} G(S, S^L, S^R | \theta)$  and set the number of samples seen by the child node to $|m_n^d| = |R_n^d| \max {\left(1, \frac{m_n}{K}\right)}$.
+Since the total number of samples stored is bounded by $K$ this approach uses less memory.
+Constructing child reservoirs from parent reservoirs (cf.
+fig \ref{fig:forest}) lessen the computational load.
+It also outperform Online Random Forests in classification accuracy.
+
+\section{Voxel Oriented Patches}
+
+\section{Putting it All Together}
+
 % \section{Pipeline}
@@
 
+  \bibitem{TSDF}
+  C URLESS , B.
+AND L EVOY , M.
+1996.
+A volumetric method for building complex models from range images.
+In Proceedings of the 23rd annual conference on Computer graphics and interactive techniques.
+ACM, 303–312.
+
+  \bibitem{hashed_voxel}
+  N IESSNER , M., Z OLLH OFER , M., I ZADI , S., AND S TAMMINGER , M. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
+
+  \bibitem{online_forest}
+  S AFFARI , A., L EISTNER , C., S ANTNER , J., G ODEC , M., AND B ISCHOF , H. 2009. On-line random forests. In IEEE ICCV Workshop.
+
+  \bibitem{reservoir}
+  [Vitter 1985]
+
+  \bibitem{murphy}
+  murphy's ML
+
 \end{thebibliography}

### 2015-12-03 cv3d_seminar/paper.tex
crf

@@
 % \usepackage{url}
-% \usepackage{amsmath}
+\usepackage{amsmath}
 
@@
 
-\section{Dynamic Conditional Random Field}
-\section{Efficient Mean-Field Inference}
+\section{Voxel Oriented Patches}
+
+\begin{figure}
+ \center
+ \includegraphics[width=0.75\textwidth]{figures/vop}
+ \caption{Voxel Oriented Patch. Colours shown in RGB for illustration purposes.}
+ \label{fig:vop}
+\end{figure}
+
+  Even the most powerful classifier will be useless if inappropriate features are supplied.
+Since speed is crucial, this paper contributes Voxel Oriented Patch (VOP) features, which can be computed efficiently from the TSDF volume representation (cf.
+figure \ref{fig:vop}.
+Let $V_i$ be a VOP and $\mathbf{n}^i$ the normal of voxel i.
+We create an image patch of dimensions $r \times r$  centered on voxel $i$ from the plane $(\mathbf{p} - \mathbf{p}_i) \cdot \mathbf(n)_i = 0$.
+The patch containes colour values stored in the TSDF on the plane in CIELab to mitigate illumination effects.
+Additionaly, $V_i$ stores distance to the nearest dominant horizontal surface.
+$r = 13$ with a resolution of $10\frac{mm}{pixel}$ is used.
+Rotation invariance is achieved exactly as in \cite{sift}.
+The patch is rotated according to the dominant gradient direction.
+It is computed by creating a histogram of gradient directions with 36 bins, where the directions are weighted by gradient magnitudes and a gaussian window centered at the voxel.
+Furthermore, a quadratic polynomial is fit to the three biggest values of the histogram, whose maximum is the final directiion.
+
 \section{Streaming Random Forests}
 
-  A random forest is an ensamble of classification or regression trees, whose outputs are combined to produce a result with lower variance \cite{murphy}.
+  A random forest is an ensamble of classification or regression trees, whose outputs are combined to produce a result with lower variance \cite{online_forest}.
 Each tree is typically trained only on a subset of the training data $S$ comprised of pairs $(i, l)$ where $i$ is a sample and $l$ its label.
@@
 
-\section{Voxel Oriented Patches}
+\section{Dynamic Conditional Random Field}
+Conditional Random Field (CRF) [Lafferty et al.2001].
+
+ Conditional Random Fields are often used for segmentation \cite{related_work}, where availability of all data is assumed.
+Here, the underlying model changes constantly (new frames, labels or different SRF prediction), which requires a novel inference alogorithm.
+The class of each voxel is modeled by a random variable $x_i$.
+The label distribution over the volume $\mathcal{V}$ factorizes into likelihood terms $\psi_i(x_i)$ and prior terms $\psi_{ij}(x_i, x_j)$.
+Let $\mathcal{E}_i$ be a neighbourhood of $v_i$, $\mathbf{D}_t $ volumetric data at time $t$ and $\mathbf{x}$ a vector of all $x_i$.
+The posterior distribution is given by
+
+\begin{equation} \label{eq:posterior}
+ P(\mathbf{x}|\mathbf{D}) = \prod_{i \in \mathcal{V}} \left( \psi_i(x_i) \prod_{j \in \mathcal{E}_i} \psi_{ij}(x_i, x_j) \right) 
+\end{equation}
+
+Neighbourhood of 6 cm is used. Negative log likelihood of eq. \ref{eq:posterior} defines the energy
+
+\begin{equation} \label{eq:energy}
+ E_t(\mathbf{x}) = \sum_{i \in \mathcal{V}} \left( \phi_i(x_i) + \sum_{j \in \mathcal{E}_i} \phi_{ij} (x_i, x_j) \right) + K
+\end{equation}
+
+Unary and pairwise potentials $\phi_i(x_i)$ and $\phi_{ij}(x_i, x_j)$ encode the cost of assigning label to $v_i$ and a cost of assuming different labels by voxels in a neighbourhood, respectively.
+They are constatly changing due to the chainging predictions of the random forest and user interaction.
+
+Initially, all voxels are encouraged to take a background label by setting $\phi_i(l) = 0$ for background class and grater than zero otherwise.
+Later, these are replaced by the random forest predictions for each voxel and as a result of user interactions.
+
+When the user touches an object, a set of touched voxels $\mathcal(H)_S$ is registered.
+Their unary potentials are set to $\phi_i(l) = \infty$ if their labels are different then the specified one, and zero if the same, thus imposing the labeling.
+If user labels a region more than once, the old labeling are simply overwritten.
+
+To process the encircling action, the labeling is projected into the current frame.
+Next, a GMM is fittend with the foreground class taken as a convex hull of the encirvled region and background as the rest of the image.
+For every pixel in a bounding box around the user annotation the unary potential are updated as
+
+\begin{equation}
+ \phi_i(l) =
+  \begin{cases}
+    \log P_E(fg|\mathbf{a}_i)       & \quad \text{if } l = fg\\
+    \log (1 - P_E(fg|\mathbf{a}_i))  & \quad \text{if } l = bg\\
+  \end{cases}
+\end{equation}
+
+with $P_E(fg|\mathbf{a}_i)$ the probability of assuming the foreground label, which is a normalized likelihood $P_(\mathbf{a}_i|fg)$ from the GMM.
+
+\begin{equation}
+  P_E(fg|\mathbf{a}_i) = \frac{P_(\mathbf{a}_i|fg) }{P_(\mathbf{a}_i|fg) + P_(\mathbf{a}_i)|bg}
+\end{equation}
+
+Smoothness is achieved by the standard Potts model by assigning a disconuity cost $\phi_{ij}(x_i, x_j) = \lambda_{ij}$ if voxels have different labels and zero otherwise, with $\lambda_{ij}$ a function of difference in position, intensity and normal directions.
+
+\section{Efficient Mean-Field Inference}
 
@@
 
-  \bibitem{murphy}
-  murphy's ML
+  \bibitem{sift}
+  L OWE , D. G. 1999. Object recognition from local scale-invariant features. In Proc. ICCV.
 

### 2015-12-04 cv3d_seminar/paper.tex
meanfield inference

@@
-% \documentclass[journal, a4paper]{IEEEtran}
 \documentclass{llncs}
@@
 \usepackage{graphicx}   
-% \usepackage{url}
 \usepackage{amsmath}
@@
 
+  The GPU implementation can handle up to 17 million voxel classifications per second, while between 3 and 10 million voxels are visible at any one time \cite{hashed_voxel}.
+To improve performance, voxels are batched and only 1 batch is evaluated per frame.
+
 \section{Dynamic Conditional Random Field}
-Conditional Random Field (CRF) [Lafferty et al.2001].
 
- Conditional Random Fields are often used for segmentation \cite{related_work}, where availability of all data is assumed.
-Here, the underlying model changes constantly (new frames, labels or different SRF prediction), which requires a novel inference alogorithm.
+ Conditional Random Fields are often used for segmentation \cite{related_work}, but they usually assume availability of all data.
+Here, a pairwise CRF \cite{crf} with a time dependent underlying model is used.
+It requires a novel inference alogorithm that can handle updates of the geometry and user specified labes.
 The class of each voxel is modeled by a random variable $x_i$.
@@
 
-Unary and pairwise potentials $\phi_i(x_i)$ and $\phi_{ij}(x_i, x_j)$ encode the cost of assigning label to $v_i$ and a cost of assuming different labels by voxels in a neighbourhood, respectively.
-They are constatly changing due to the chainging predictions of the random forest and user interaction.
+  Unary and pairwise potentials $\phi_i(x_i)$ and $\phi_{ij}(x_i, x_j)$ encode the cost of assigning a label to $v_i$ and a cost of assuming different labels by voxels in a neighbourhood, respectively.
+These costs are constatly changing due to the chainging predictions of the random forest and user interaction.
 
   Initially, all voxels are encouraged to take a background label by setting $\phi_i(l) = 0$ for background class and grater than zero otherwise.
-Later, these are replaced by the random forest predictions for each voxel and as a result of user interactions.
-
-When the user touches an object, a set of touched voxels $\mathcal(H)_S$ is registered.
-Their unary potentials are set to $\phi_i(l) = \infty$ if their labels are different then the specified one, and zero if the same, thus imposing the labeling.
-If user labels a region more than once, the old labeling are simply overwritten.
+When the user touches an object, a set of touched voxels $\mathcal{H}_S$ is registered and their unary potentials set to $\phi_i(l) = \infty$ if their labels are different then the specified one, and zero if they are the same, thus imposing the labeling.
+If user labels a region more than once, the old labelings are simply overwritten.
 
   To process the encircling action, the labeling is projected into the current frame.
-Next, a GMM is fittend with the foreground class taken as a convex hull of the encirvled region and background as the rest of the image.
+Next, a GMM is fitted with the foreground class taken as a convex hull of the encirvled region and background as the rest of the image.
 For every pixel in a bounding box around the user annotation the unary potential are updated as
@@
 
-Smoothness is achieved by the standard Potts model by assigning a disconuity cost $\phi_{ij}(x_i, x_j) = \lambda_{ij}$ if voxels have different labels and zero otherwise, with $\lambda_{ij}$ a function of difference in position, intensity and normal directions.
+  For all voxels that haven't been explicitly labaled by the user unitary potentials are updated with the random forest predictions as $\phi_i(l) = -\log P_F(x_i = l | \mathbf{D})$.
+Finally, smoothness is achieved by the standard Potts model by assigning a disconuity cost $\phi_{ij}(x_i, x_j) = \lambda_{ij}$ if voxels have different labels and zero otherwise, with $\lambda_{ij}$ a function of difference in position, intensity and normal directions.
 
@@
 
+The mean-field inference algorithms follows \cite{inference}.
+It is adapted to the GPU programming model and to the constantly changing energy landscape.
+First, the original prbability distribution $P(\mathbf{x})$ is approximated by $Q(\mathbf{x})$ under KL-divergence $KL(Q||P)$.
+It is chosen such that the marginal of each random variable is independent, that is $Q(\mathbf{x}) = \prod_i Q_i(\mathbf{x)_i})$.
+It is iteratively refined with updates derived from the fixed point of KL-divergence \cite{kl} as
+
+\begin{equation}
+ Q_i^t(l) = \frac{1}{Z_i}e^{M_i(l)} \text{, } t = 1, \ldots, T
+\end{equation}
+\begin{equation}
+ M_i(l) = \phi_i(l) + \sum_{l' \in \mathcal{L}} \sum_{j \in \mathcal{N}(i)} Q_j^{t-1}(l')\phi_{ij}(l, l')
+\end{equation}
+\begin{equation}
+ Z_i = \sum_{i \in \mathcal{L}} e^{-M_i(l)}
+\end{equation}
+\begin{equation}
+ x_i^\star = \arg \max_{l \in \mathcal{L}} Q_i^T(l)
+\end{equation}
+
+It is guaranteed to converge under some conditions \cite{inference}.
+However, these guarantees do not hold here due to changing energy landscape.
+Moreover, the energy distribution is assumed to change only gradualy instead of initializing enerrgies at a given frame from a uniform distribution, it is initialized with the energies of the previous frame.
+It works well in practice, speeds up the process and allows visually pleasing label propagation effect.
+
+SRF classification results would impact the final segmentation after several frames due to their effect on unitary potential.
+To speed up the process, however, the initial energy value at a new frame is taken as a weighted sum of the value from the previous frame and the SRF prediction, that is
+
+\begin{equation}
+ \widetilde{Q}_i^t(x_i) = \gamma Q_i^{t-1}(x_i) + (1 - \gamma) P_F^{t-1}(x_i = l | \mathbf{D}) \text{, } \gamma \in [0, 1]
+\end{equation}
+
 \section{Putting it All Together}
+\section{Results}
+\section{Future Work}
+\section{Conclusions}
 
@@
   \bibitem{TSDF}
-  C URLESS , B.
-AND L EVOY , M.
+  CURLESS , B.
+AND LEVOY , M.
 1996.
@@
   \bibitem{hashed_voxel}
-  N IESSNER , M., Z OLLH OFER , M., I ZADI , S., AND S TAMMINGER , M. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
+  NIESSNER , M., Z OLLH OFER , M., IZADI , S., AND STAMMINGER , M. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
 
   \bibitem{online_forest}
-  S AFFARI , A., L EISTNER , C., S ANTNER , J., G ODEC , M., AND B ISCHOF , H. 2009. On-line random forests. In IEEE ICCV Workshop.
+  SAFFARI , A., LEISTNER , C., SANTNER , J., G ODEC , M., AND BISCHOF , H. 2009. On-line random forests. In IEEE ICCV Workshop.
 
   \bibitem{reservoir}
-  [Vitter 1985]
+  VITTER , J. S. 1985. Random sampling with a reservoir. ACM TOMS 11, 1.
 
   \bibitem{sift}
-  L OWE , D. G. 1999. Object recognition from local scale-invariant features. In Proc. ICCV.
+  LOWE , D. G. 1999. Object recognition from local scale-invariant features. In Proc. ICCV.
+
+  \bibitem{crf}
+  LAFFERTY , J., M C C ALLUM , A., AND PEREIRA , F. C. 2001. Conditional random fields: Probabilistic models for segmenting and labeling sequence data.
+
+  \bibitem{inference}
+  [Kr ̈ahenb ̈uhl and Koltun 2011]
+
+  \bibitem{kl}
+  [Koller and Friedman 2009]
 
 \end{thebibliography}
-
-% Your document ends here!
 \end{document}

### 2015-12-04 cv3d_seminar/paper.tex
finished contentwise, needs to be shortened and polished

@@
 \section{Putting it All Together}
+
+    The user comes into a room and touches a table uttering a label at the same time.
+Then, he touches a chair giving another label.
+He switches into test mode and goes into another room.
+All tables and chair are marked by the system accordingly.
+It works by feeding Voxel Oriented Patches associated with each pixel to the Streaming Random Forest, which learn to predict voxel categories in an online fashion.
+User interactions are facilitated by a Conditial Random Field that models the label distribution.
+An efficient mean-field algorithm, which amortizes classification of the whole scene among several frames, is used on top of it to smooth predicted labels.
+
 \section{Results}
+  Each part of the pipeline is evaluated separately.
+Several video sequences were recorded to evaluate segmentation and VOP features.
+A set of keyframe covering the whole scene from each sequence was hand-labeled to create groudtruth data.
+They were further projected into the 3D volume, resulting in 4176, 12346, 7583, 8916 frames for \emph{LivingRoom}, \emph{Kitchen}, \emph{Bedroom}, \emph{Desk} sequences respectively, of wich around a third was used for testing.
+Table \ref{fig:segm_results} summarises segmantation results.
+User interaction gives very accurate labeling, but it is pertinent to single object instances only.
+SRF prediction works quite well and is improved by the inference algorithm.
+Table \ref{fig:vop_results} compares VOPs against other features.
+The numbers reflect the percentage of correctly classified pixels with SRF using different features.
+VOP clearly outperforms all other features.
+$\Delta$ RGB mean stands for the mean difference of two randomly selected patches.
+
+\begin{table}[!ht]	\label{fig:segm_results}
+ \center
+ \caption{Segmentation Results.}
+ \begin{tabular}{c|c|c|c|c|c}
+  \textbf{Component} & \textbf{LivingRoom} & \textbf{Bedroom} & \textbf{Kitchen} & \textbf{Desk} & \textbf{Average} \\ \hline
+  User Interaction & 99.35\% & 97.61\% & 96.09\% & 97.73\% & 97.7\% \\ 
+  Forest Prediction & 94.57\% & 88.31\% & 82.58\% & 90.29\% & 88.94\% \\
+  Final Inference & 96.26\% & 95.19\% & 90.69\% & 95.55\% & 94.42\%
+ \end{tabular}
+\end{table}
+
+\begin{table}[!ht]	\label{fig:vop_results}
+ \center
+ \caption{Comparison of VOP against other features.}
+ \begin{tabular}{c|c|c|c|c|c}
+  \textbf{Feature} & \textbf{LivingRoom} & \textbf{Bedroom} & \textbf{Kitchen} & \textbf{Desk} & \textbf{Average} \\ \hline
+  VOP & \textbf{94.57\%} & \textbf{88.31\%} & 82.58\% & \textbf{90.29\%} & \textbf{88.94\%} \\
+  $\Delta$ RGB mean & 80\% & 71.84\% & 76.29\% & 73.42\% & 75.39\% \\
+  Depth probe& 77.54\% & 61.79\% & \textbf{84.9\%} & 68.9\% & 73.06\% \\
+  Color Probe& 56.39\% & 65.68\% & 60.77\% & 60.74\% & 60.9\% \\
+  SURF & 43.74\% & 67.12\% & 57\% & 58.13\% & 56.5\% \\
+  SPIN & 58.77\% & 43.22\% & 48.41\% & 36.1\% & 46.63\% \\
+ \end{tabular}
+\end{table}
+
+  SRF was compared against Online Random Forests (ORF) \cite{online_forest} and Hoeffding trees (HT) \cite{htree}.
+In order to simulate non-IID data, a dataset of 300 objects of 51 categories \cite{dataset} was used.
+One revolution of each object was recorded with an RGBD camera from three different viewpoints.
+Online setting was created by sequentially adding new categories.
+Two thirds of each viewpoint of each object is used for training.
+Results are summarized by figure \ref{fig:srf_results}.
+SRF clearly outperforms its competitors.
+
+\begin{figure}[!ht]	 \label{fig:srf_results}
+ \center
+ \includegraphics[width=0.75\textwidth]{figures/srf_results}
+ \caption{Comparison of Streaming Random Forest against Online Random Forest and Hoeffding Trees.}
+\end{figure}
+
+\section{Discussion}
+
+  This paper contributes an interactive system that enables users to create customized models of 3D enviornments.
+It can be used to gather groundtruth for large-scale visual recognition systems, robot navigation and to aid partially sighted people.
+There are some limitations, however.
+Use of the system is now restricted to simple, indoor scenes.
+Computational load increases with the size of the scene while prediction accuracy drops with the number of object categories.
+There are also some failure cases as shwon in figure \ref{fig:failures}.
+
+\begin{figure}[!ht]	 \label{fig:failures}
+ \center
+ \includegraphics[width=0.75\textwidth]{figures/failures}
+ \caption{Failure cases.}
+\end{figure}
+
+Combing colour and depth data requires careful sensor calibration.
+Lack thereof results in misclassification at object boundaries.
+Also strong changes of lighting conditions impair classification ability at certain viewpoints.
+What is more, no global context is used for classification, which could be improved if some prior knowledge were introduced.
+There are two disjoint modes: training and testing, which limits possible use cases.
+
+\section{Summary}
+
 \section{Future Work}
-\section{Conclusions}
 
-% \section{Pipeline}
-% \section{Mean-Field Inference}
-% \section{Streaming Random Forest}
-% \section{Voxel-Oriented Patch Features}
-% \section{Qualitative Results}
-% \section{Quantitative Results}
-% \section{Discussion}
-% \section{Conclusions}
-
-% 	\begin{table}[!hbt]
-% 		% Center the table
-% 		\begin{center}
-% 		% Title of the table
-% 		\caption{Simulation Parameters}
-% 		\label{tab:simParameters}
-% 		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
-% 		\begin{tabular}{|c|c|}
-% 			% To create a horizontal line, type \hline
-% 			\hline
-% 			% To end a column type &
-% 			% For a linebreak type \\
-% 			Information message length & $k=16000$ bit \\
-% 			\hline
-% 			Radio segment size & $b=160$ bit \\
-% 			\hline
-% 			Rate of component codes & $R_{cc}=1/3$\\
-% 			\hline
-% 			Polynomial of component encoders & $[1 , 33/37 , 25/37]_8$\\
-% 			\hline
-% 		\end{tabular}% 		\end{center}
-% 	\end{table}
-% 
-% 	\begin{figure}[!hbt]
-% 		% Center the figure.
-% 		\begin{center}
-% 		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
-% 		\includegraphics[width=\columnwidth]{plot_tf}
-% 		% Create a subtitle for the figure.
-% 		\caption{Simulation results on the AWGN channel. Average throughput $k/n$ vs $E_s/N_0$.}
-% 		Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
-% 		\label{fig:tf_plot}
-% 		\end{center}
-% 	\end{figure}
+  - introducing prior
+  - improving user interaction system (verbal priors)
+  - improving scalability
+  - using pure geometrical features
+  - moving to outdoor settings
+  - better scaling
 

### 2015-12-05 ML/hw8/hw.tex
hw8 ml

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{epstopdf}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}

\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 8} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 5}

\begin{figure}[!ht]
 \center
 \includegraphics[width=\textwidth]{prob5}
 \label{fig:primal}
\end{figure}

As can be clearly seen in the above plot, $x^\star$ = 2 is the optimal solution.

\section*{Problem 6}

\begin{equation}
 \begin{align}
  L(x, \alpha) &= f_0(x) + \alpha f_1(x) \\
    &= x^2 + 1 + \alpha (x-2)(x-4) \\
    &= (1 + \alpha) x^2 + 6 \alpha x + (1 + 8 \alpha)
 \end{align}
\end{equation}

\begin{figure}[!ht] \label{fig:prob6}
 \center
 \includegraphics[width=\textwidth]{prob6}
 \caption{Lagrangian $L(x, \alpha)$ for different values of $\alpha$}.
\end{figure}

The value of the objective function is given by the lagrangian $L(x, \alpha)$ evaluated at $\alpha = 0$, portrayed by the red curve in Fig. \ref{fig:prob6}. Lagrangian is smaller than the objective function for $x \in (2, 4)$ and greater than the objective function for $x \notin [2, 4]$. Values at $x = 2$ and $x = 4$ are unaffected by $\alpha$. The upper bound for $\min_x L(x, \alpha) = L(2, \alpha)$ for all $\alpha \geq 0$ is $5$.

\section*{Problem 7}

\begin{equation}
 g(\alpha) = \min_x L(x, \alpha) \implies \frac{\partial}{\partial x} L(x, \alpha) = 0 \implies x^\star = \frac{3 \alpha}{1 + \alpha}
\end{equation}

\begin{equation}
 g(\alpha) = \frac{- \alpha^2 + 9 \alpha + 1}{1 + \alpha}
\end{equation}

The dual problem is plotted in Fig. \ref{fig:dual} and it is given by

\begin{equation}
 \begin{align}
  \max_\alpha g(\alpha) \\
  \text{subject to } \alpha \geq 0
 \end{align}
\end{equation}

\begin{figure}[!ht] \label{fig:dual}
 \center
 \includegraphics[width=\textwidth]{prob7}
 \caption{Lagrange dual function $g(\alpha)$.}
\end{figure}

\section*{Problem 8}

\begin{equation}
 \alpha^\star = \arg \max_\alpha g(\alpha) 
\end{equation}
\begin{equation}
 0 = \frac{\partial}{\partial \alpha} g(\alpha) = \frac{- \alpha^2 - 2 \alpha +8}{(1 + \alpha)^2} \text{ and } \alpha \geq 0
\end{equation}

The dual optimal solution $\alpha^\star = 2$ and $g(\alpha^\star) = 5$.

\section*{Problem 9}

\begin{equation}
 x^\star = \frac{3 \alpha^\star}{1 + \alpha^\star} = \frac{6}{1 + 2} = 2
\end{equation}

The optimal solution is given by $f_0(x^\star) = 5$, which is equal to the dual optimal value.

\section*{Problem 10}

The constraint $f_1$ is active. We can see it also on the plot of the primal problem in Fig. \ref{fig:primal}, since the feasible region is bounded by the zero-crossings of this constraint and the objective function attains its minimum at one of them.

\end{document}

### 2015-12-05 ML/hw8/ml8.m

@@ -1,0 +1,59 @@
+
+close all, clear all
+
+f0 = @(x) x.^2 + 1;
+f1 = @(x) (x-2) .* (x-4);
+
+
+
+xx = linspace(0, 5, 100);
+
+%% problem 5
+xx0 = linspace(2, 4, 100);
+
+figure(1)
+hold on
+plot(xx, f0(xx), 'r-', 'LineWidth', 2);
+plot(xx, f1(xx), 'b-', 'LineWidth', 2);
+plot(xx0, f1(xx0), 'g-', 'LineWidth', 2);
+plot([2 2], [-10 70], 'g-');
+plot([4 4], [-10 70], 'g-');
+title('function and its constraint');
+xlabel('x');
+ylabel('f(x)');
+legend('f0', 'f1', 'feasible region');
+ 
+
+%% problem 6
+
+as = [0 0.5 1 3 4 5 8];
+colors = 'rgbkcmy';
+L = @(x, a) f0(x) + a .* f1(x);
+
+figure(2)
+hold on
+
+
+for i = 1:numel(as)
+    c = colors(i);
+    a = as(i);
+    plot(xx, L(xx, a), c, 'LineWidth', 2);
+end
+title('Lagrangian for different alphas');
+xlabel('x')
+ylabel('L(x, a)');
+legend('a = 0', 'a = 0.5', 'a = 1', 'a = 3', 'a = 4', 'a = 5', 'a = 8');
+
+
+%% problem 7
+g = @(a) (-a.^2 + 9 * a +1) ./ (1 + a);
+aa = linspace(0, 8, 100);
+
+figure(3)
+plot(aa, g(aa), 'b-', 'LineWidth', 2);
+title('Lagrange dual function');
+xlabel('a');
+ylabel('g(a)');
+
+
+

### 2015-12-05 cv3d_seminar/paper.tex
rework up to Putting it together section

@@
   Capturing your own environment has never been easier.
-SemanticPaint can register your surroundings which, after undergoing a low-level 3D reconstruction, can be semantically segmented in an interactive way.
+SemanticPaint can register your surroundings which, after 3D reconstruction, can be semantically segmented in an interactive way.
 Not only it works in real time but also requires no pretraining.
@@
   The pipeline starts with capturing the environment as a stream of noisy RGBD images and combining them into a  3D model updated in an online fashion.
-The user can choose which objects to label and can do so by ``touching'' a small part of an object or encircling one with his hand and uttering the label.
-It is recognized by a standard speech recognition system.
-The label and the information about the affected data points are further passed to a Streaming Random Forest classifier which constantly learns and labels all visible voxels.
+The user can choose which objects to label and can do so by ``touching'' a small part of an object or encircling it with his hand and uttering the label.
+The label and the affected data points are further passed to a Streaming Random Forest (SRF) classifier which constantly learns and labels all visible voxels.
 To further improve classification results a spatially dense labeling is produced by an efficient mean-field inference algorithm.
-One of the biggest strength of SemanticPaint is the efficiency of each part of the pipeline, which translates to real time performance.
-Algorithms used in the pipeline were adapted to work on volumetric data in the TSDF format directly in order to avoiding the costly conversions to mesh or point-cloud formats.
+One of the biggest strengths of SemanticPaint is the efficiency of each part of the pipeline, which translates to real time performance.
+Algorithms used in the pipeline were adapted to work on volumetric data in the TSDF format directly in order to avoiding costly conversions to mesh or point-cloud formats.
 To allow this, the Voxel Oriented Patch features --- a new type of a discriminative feature describing the voxel space --- has been designed.
-  The contributions can be summarized as follows: 3D semantic modeling system, Streaming Random Forest, efficient Mean-Field inference, Voxel Oriented Patch features.
-
-  Numerous applications are possible: (1) building large scale datasets of 3D objects or whole scenes for use in large-scale computer vision systems (2) using the dense semantic labeling of 3D environments in robot navigation or to aid people with impaired sight and (3) map environments for use in augmented reality scenarios or games.
-
-  The rest of the paper is organized as follows: Section 2.
-describes the related work, section 3 details internal data handling, section 4 describes the efficient mean-field inference algorithm, section 5 outline the Streaming Random Forest classifiers.....
 
@@
 
-  \textbf{Acquisition and Reconstruction.} Capturing the geometry of the surrounding world has been a long standing problem \cite{old}.
+  \textbf{Acquisition and Reconstruction } Capturing the geometry of the surrounding world has been a long standing problem \cite{old}.
 An offline processing of multiple images allowed to reconstruct digital heritage and construct world-scale 3D models with remarkable quality \cite{heritage}.
-The inception of low-cost RGBD sensors and powerful GPUs enabled online 3D scanning, augmented reality or using 3D environment models for navigation purposes \cite{fusion}.
+The inception of low-cost RGBD sensors and powerful GPUs enabled online 3D scanning, augmented reality and using 3D environment models for navigation purposes \cite{fusion}.
 \cite{Monofusion} enables 3D reconstruction of small scenes using a single off-the-shelf RGB camera.
@@
   \noindent
-  \textbf{Scene Understanding.} Object recognition, detection and segmentation has been done with 2D RGB images, RGBD data, point clouds and volumetric representations.
+  \textbf{Scene Understanding } Object recognition, detection and segmentation has been done with 2D RGB images, RGBD data, point clouds and volumetric representations.
 \cite{VCRF} uses a Voxel-based CRF for simultaneous reconstruction and segmentation.
@@
 The first two are used to to improve reconstruction by mitigating depth-map noise.
-The visibility values are constrained by that each ray from the camera can hit only one visible vortex.
+The visibility values are constrained by that each ray from the camera can hit only one visible vertex.
 The group membership information encodes priors given by bounding boxes of detected objects.
@@
 
-  Even the most powerful classifier will be useless if inappropriate features are supplied.
-Since speed is crucial, this paper contributes Voxel Oriented Patch (VOP) features, which can be computed efficiently from the TSDF volume representation (cf.
-figure \ref{fig:vop}.
+  This work relies heavily on classification and since even the most powerful classifier is useless if inappropriate features are supplied this paper contributes Voxel Oriented Patch (VOP) features.
+They are efficiently computed from the TSDF volume representation (cf.
+figure \ref{fig:vop}).
 Let $V_i$ be a VOP and $\mathbf{n}^i$ the normal of voxel i.
-We create an image patch of dimensions $r \times r$  centered on voxel $i$ from the plane $(\mathbf{p} - \mathbf{p}_i) \cdot \mathbf(n)_i = 0$.
+An image patch of dimensions $r \times r$  is centered on voxel $i$ from the plane $(\mathbf{p} - \mathbf{p}_i) \cdot \mathbf(n)_i = 0$.
 The patch containes colour values stored in the TSDF on the plane in CIELab to mitigate illumination effects.
@@
 Let $f(i; \theta) \in \{L, R\}$ denote the binary split function with learned parameters $\theta$, which specify a feature this node uses.
-The tree's output amounts to the probability distribution $P(c_i = l | i)$ stored at each node.
+The tree's output amounts to the probability distribution $P_F(x_i = l | \mathbf{D})$ stored at each node.
 Trees learn in a greedy way: for each node a split function is chosen from a distribution $\Theta$ of candidate split functions to maximize the information gain.
@@
 
-  Streaming Random Forest begins by storing a reservoir $R_n$ of at most $K$ samples stored in a list $T_n$ at the root node.
+  Streaming Random Forest begins by creating a reservoir $R_n$ of at most $K$ samples stored in a list $T_n$ at the root node.
 It represents an unbiased sample of all training data seen so far.
@@
 Constructing child reservoirs from parent reservoirs (cf.
-fig \ref{fig:forest}) lessen the computational load.
-It also outperform Online Random Forests in classification accuracy.
+fig \ref{fig:forest}) lessens the computational load.
 
   The GPU implementation can handle up to 17 million voxel classifications per second, while between 3 and 10 million voxels are visible at any one time \cite{hashed_voxel}.
-To improve performance, voxels are batched and only 1 batch is evaluated per frame.
+To improve performance, voxels are batched and only a single batch is evaluated per frame.
 
@@
 When the user touches an object, a set of touched voxels $\mathcal{H}_S$ is registered and their unary potentials set to $\phi_i(l) = \infty$ if their labels are different then the specified one, and zero if they are the same, thus imposing the labeling.
-If user labels a region more than once, the old labelings are simply overwritten.
+If user labels a region more than once, the old labels are simply overwritten.
 
   To process the encircling action, the labeling is projected into the current frame.
-Next, a GMM is fitted with the foreground class taken as a convex hull of the encirvled region and background as the rest of the image.
-For every pixel in a bounding box around the user annotation the unary potential are updated as
+Next, a GMM is fit with the foreground class taken as a convex hull of the encirvled region and background as the rest of the image.
+For every pixel in a bounding box around the user annotation the unary potentials are updated as
 
@@
     \begin{cases}
-      \log P_E(fg|\mathbf{a}_i)       & \quad \text{if } l = fg\\
-      \log (1 - P_E(fg|\mathbf{a}_i))  & \quad \text{if } l = bg\\
+      \log P_E(fg|\mathbf{a}_i)       & \quad \text{if } l = \text{fg}\\
+      \log (1 - P_E(fg|\mathbf{a}_i))  & \quad \text{if } l = \text{bg}\\
     \end{cases}
@@
 
-  with $P_E(fg|\mathbf{a}_i)$ the probability of assuming the foreground label, which is a normalized likelihood $P_(\mathbf{a}_i|fg)$ from the GMM.
+  with $P_E($fg$|\mathbf{a}_i)$ the probability of assuming the foreground label, which stems from the normalized likelihood $P(\mathbf{a}_i|$fg$)$ from the GMM.
 
   \begin{equation}
-    P_E(fg|\mathbf{a}_i) = \frac{P_(\mathbf{a}_i|fg) }{P_(\mathbf{a}_i|fg) + P_(\mathbf{a}_i)|bg}
+    P_E(\text{fg}|\mathbf{a}_i) = \frac{P(\mathbf{a}_i|\text{fg}) }{P(\mathbf{a}_i|\text{fg}) + P(\mathbf{a}_i|\text{bg})}
   \end{equation}
 
-  For all voxels that haven't been explicitly labaled by the user unitary potentials are updated with the random forest predictions as $\phi_i(l) = -\log P_F(x_i = l | \mathbf{D})$.
+  For all voxels that haven't been explicitly labaled by the user unitary potentials are updated with the predicted values as $\phi_i(l) = -\log P_F(x_i = l | \mathbf{D})$.
 Finally, smoothness is achieved by the standard Potts model by assigning a disconuity cost $\phi_{ij}(x_i, x_j) = \lambda_{ij}$ if voxels have different labels and zero otherwise, with $\lambda_{ij}$ a function of difference in position, intensity and normal directions.
@@
 
-The mean-field inference algorithms follows \cite{inference}.
-It is adapted to the GPU programming model and to the constantly changing energy landscape.
+The mean-field inference algorithm is adapted from \cite{inference} to handle the constantly changing energy landscape and implemented on GPU.
 First, the original prbability distribution $P(\mathbf{x})$ is approximated by $Q(\mathbf{x})$ under KL-divergence $KL(Q||P)$.
-It is chosen such that the marginal of each random variable is independent, that is $Q(\mathbf{x}) = \prod_i Q_i(\mathbf{x)_i})$.
-It is iteratively refined with updates derived from the fixed point of KL-divergence \cite{kl} as
+$Q(\mathbf{x})$ is chosen such that the marginal of each random variable is independent, that is $Q(\mathbf{x}) = \prod_i Q_i(\mathbf{x)_i})$.
+Iterative updates yield:
 
@@
 \end{equation}
-\begin{equation}
- Z_i = \sum_{i \in \mathcal{L}} e^{-M_i(l)}
-\end{equation}
-\begin{equation}
- x_i^\star = \arg \max_{l \in \mathcal{L}} Q_i^T(l)
-\end{equation}
 
-It is guaranteed to converge under some conditions \cite{inference}.
-However, these guarantees do not hold here due to changing energy landscape.
-Moreover, the energy distribution is assumed to change only gradualy instead of initializing enerrgies at a given frame from a uniform distribution, it is initialized with the energies of the previous frame.
-It works well in practice, speeds up the process and allows visually pleasing label propagation effect.
-
-SRF classification results would impact the final segmentation after several frames due to their effect on unitary potential.
-To speed up the process, however, the initial energy value at a new frame is taken as a weighted sum of the value from the previous frame and the SRF prediction, that is
+with $Z_i$ a normalizing factor and final class chosen as the minizer of $Q_i^T$.
+Energy distribution is assumed to change only gradualy.
+Moreover, SRF classification results would impact the final segmentation after several frames due to their effect on unitary potentials.
+To speed up the process the initial energy value at a new frame is set to a weighted sum of the previous frame's state and the SRF prediction.
+It works well in practice and allows visually pleasing label propagation effect.
 
@@
 There are also some failure cases as shwon in figure \ref{fig:failures}.
+
+   Numerous applications are possible: (1) building large scale datasets of 3D objects or whole scenes for use in large-scale computer vision systems (2) using the dense semantic labeling of 3D environments in robot navigation or to aid people with impaired sight and (3) map environments for use in augmented reality scenarios or games.
 

### 2015-12-05 ML/hw8/hw.tex
hw8 with input from Martin

@@
 
+\section*{Problem 1}
+
+The k-nearest neighbours algorithm in the feature space by introducing the feature map is
+
+\begin{equation}
+|\phi(x) - \phi(x^{(s_i)})|_2.
+\label{eq:}
+\end{equation}
+
+In order that it only depends on the scalar product in feature space $K(\phi(x),\phi(x^{(s_i)})) = \phi(x)^T\phi(y)$, we square this. Using that operation, the k-nearest neighbours don't change.
+
+\begin{align}
+|\phi(x) - \phi(x^{(s_i)})|_2^2 =& (\phi(x) - \phi(x^{(s_i)}))^T(\phi(x) - \phi(x^{(s_i)})) \\
+=& \phi(x)^T\phi(x) -2\phi(x)^T\phi(x^{(s_i)}) + \phi(x^{(s_i)})^T\phi(x^{(s_i)}).
+\label{eq:}
+\end{align}
+
+When searching for the k training samples that minimize this function, the first term stays constant. Therefore, we can drop the term and must find k training samples $x^{(s_i)}$ that minimize
+
+\begin{equation}
+ \phi(x^{(s_i)})^T\phi(x^{(s_i)})-2\phi(x)^T\phi(x^{(s_i)}) = K(x^{(s_i)}, x^{(s_i)}) - 2K(x, x^{(s_i)}).
+\end{equation}
+
+\section*{Problem 2}
+
+The definition for a convex function is
+
+\begin{equation}
+f(\alpha x + (1-\alpha)y) \leq \alpha f(x) +(1-\alpha)f(y).
+\end{equation}
+
+Plugging this into $h(x) = f(x)+g(x)$, gives us
+
+\begin{align}
+h(\alpha x + (1-\alpha)y) =& f(\alpha x + (1-\alpha)y) + g(\alpha x + (1-\alpha)y)\\
+\leq& \alpha f(x) +(1-\alpha)f(y) +\alpha g(x) +(1-\alpha)g(y)\\
+=& \alpha h(x) +(1-\alpha)h(y).
+\end{align}
+
+The proof of the scaled function being convex is
+
+\begin{align}
+u(\alpha x + (1-\alpha)y) =& cf(\alpha x + (1-\alpha)y)\\
+\leq& c\left(\alpha f(x) +(1-\alpha)f(y)\right)\\
+=& \alpha u(x) +(1-\alpha)u(y)
+\end{align}
+
+\section*{Problem 3}
+
+By definition of convexity for every $\lambda$ we have for arbitrary x,y and $0\leq\alpha\leq1$
+
+\begin{equation}
+f_\lambda((1-\alpha)x + \alpha y) \leq (1-\alpha)f_\lambda(x) + \alpha f_\lambda(y).
+\end{equation}
+
+By definition of maximum we have
+
+\begin{equation}
+(1-\alpha)\max\limits_{\lambda}f_\lambda(x) + \alpha \max\limits_{\lambda}f_\lambda(y) \geq (1-\alpha)f_{\lambda '}(x) + \alpha f_{\lambda '}(y)
+\end{equation}
+
+for any $\lambda '$, which is the maximizer of $f_\lambda((1-\alpha)x + \alpha y)$ w.r.t. $\lambda$. We obtain
+
+\begin{align}
+(1-\alpha)f_{\lambda '}(x) + \alpha f_{\lambda '}(y) \geq& f_{\lambda '}(1-\alpha)x + \alpha y)\\
+=& \max\limits_{\lambda}f_\lambda((1-\alpha)(x + \alpha y).
+\end{align}
+
+\section*{Problem 4}
+The Lagrangian is given by
+
+\begin{equation}
+L(x,\alpha) = f_0(x) + \sum_{i}{\alpha_i f_i(x)}.
+\end{equation}
+
+This can also be interpreted as a family of functions $L_x(\alpha) = L(x,\alpha)$, that means we interpret $x$ as fixed. Thus $L_x(\alpha)$ has the form
+
+\begin{equation}
+L_x(\alpha) = C_{0,x} + \sum_{i}{C_{i,x} \alpha_i},
+\end{equation}
+
+where the $Cs$ are constants. By applying the definition of a convex function we see immediately that $L_x(\alpha)$ is convex for any $Cs$.
+The Lagrange dual function is given by the pointwise minimum of this family of functions,
+
+\begin{equation}
+g(\alpha) = \min\limits_{x}L_x(\alpha) = -\max\limits_{x}(-L_x(\alpha)).
+\end{equation}
+
+The negative Lagragian $-L_x(\alpha)$ is also convex, because the negation only changes the sign of the
+constants $C_{0,x}$ and $C_{i,x}$. Thus, as shown in the previous exercise,$-g(\alpha)$ is convex and $g(\alpha)$ is concave.
+
 \section*{Problem 5}
 
-\begin{figure}[!ht]
- \center
- \includegraphics[width=\textwidth]{prob5}
- \label{fig:primal}
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{ex5.jpg}
+\caption{The objective $f_0(x)$ and its constraint $f_1(x)$.}
+\label{optimizationprob}
 \end{figure}
 
-As can be clearly seen in the above plot, $x^\star$ = 2 is the optimal solution.
+The feasible region is marked green which is from 2 to 4 and the solution of the optimization problem is 2, also called the minimizer. In figure \ref{optimizationprob} the minimizer is marked orange.
 
@@
     &= x^2 + 1 + \alpha (x-2)(x-4) \\
-    &= (1 + \alpha) x^2 + 6 \alpha x + (1 + 8 \alpha)
+    &= (1 + \alpha) x^2 - 6 \alpha x + (1 + 8 \alpha)
  \end{align}
@@
 
-\begin{figure}[!ht] \label{fig:prob6}
+\begin{figure}[!ht] 
  \center
@@
  \caption{Lagrangian $L(x, \alpha)$ for different values of $\alpha$}.
+ \label{figprob6}
 \end{figure}
@@
 The value of the objective function is given by the lagrangian $L(x, \alpha)$ evaluated at $\alpha = 0$, portrayed by the red curve in Fig.
-\ref{fig:prob6}.
+\ref{figprob6}.
 Lagrangian is smaller than the objective function for $x \in (2, 4)$ and greater than the objective function for $x \notin [2, 4]$.
@@
 
-\begin{figure}[!ht] \label{fig:dual}
+\begin{figure}[!ht] 
  \center
@@
  \caption{Lagrange dual function $g(\alpha)$.}
+ \label{fig:dual}
 \end{figure}
@@
 We can see it also on the plot of the primal problem in Fig.
-\ref{fig:primal}, since the feasible region is bounded by the zero-crossings of this constraint and the objective function attains its minimum at one of them.
+\ref{optimizationprob}, since the feasible region is bounded by the zero-crossings of this constraint and the objective function attains its minimum at one of them.
 

### 2015-12-05 cv3d_seminar/paper.tex
rework all

@@
 
- Conditional Random Fields are often used for segmentation \cite{related_work}, but they usually assume availability of all data.
+ Conditional Random Fields are often used for segmentation \cite{VCRF}, but they usually assume availability of all data.
 Here, a pairwise CRF \cite{crf} with a time dependent underlying model is used.
@@
 
-\section{Putting it All Together}
-
-    The user comes into a room and touches a table uttering a label at the same time.
+% \section{Putting it All Together}
+% 
+%   The user comes into a room and touches a table uttering a label at the same time.
 Then, he touches a chair giving another label.
@@
 A set of keyframe covering the whole scene from each sequence was hand-labeled to create groudtruth data.
-They were further projected into the 3D volume, resulting in 4176, 12346, 7583, 8916 frames for \emph{LivingRoom}, \emph{Kitchen}, \emph{Bedroom}, \emph{Desk} sequences respectively, of wich around a third was used for testing.
+They were further projected into the 3D volume, resulting in 4176, 12346, 7583, 8916 frames for \emph{LivingRoom}, \emph{Kitchen}, \emph{Bedroom} and \emph{Desk} sequences respectively, of wich around a third was used for testing.
 Table \ref{fig:segm_results} summarises segmantation results.
-User interaction gives very accurate labeling, but it is pertinent to single object instances only.
-SRF prediction works quite well and is improved by the inference algorithm.
+User interaction gives very accurate labeling, but it affects a single object instance only.
+SRF prediction works well and is improved by further inference.
 Table \ref{fig:vop_results} compares VOPs against other features.
@@
 
-\begin{table}[!ht]	\label{fig:segm_results}
+\begin{table}[!ht]	
  \center
@@
  \end{tabular}
+ \label{fig:segm_results}
 \end{table}
 
-\begin{table}[!ht]	\label{fig:vop_results}
+\begin{table}[!ht]	
  \center
@@
  \end{tabular}
+ \label{fig:vop_results}
 \end{table}
@@
 
-\begin{figure}[!ht]	 \label{fig:srf_results}
+\begin{figure}[!ht]
  \center
@@
  \caption{Comparison of Streaming Random Forest against Online Random Forest and Hoeffding Trees.}
+ \label{fig:srf_results}
 \end{figure}
@@
 There are some limitations, however.
-Use of the system is now restricted to simple, indoor scenes.
+Use of the system is now restricted to simple indoor scenes.
 Computational load increases with the size of the scene while prediction accuracy drops with the number of object categories.
-There are also some failure cases as shwon in figure \ref{fig:failures}.
 
-   Numerous applications are possible: (1) building large scale datasets of 3D objects or whole scenes for use in large-scale computer vision systems (2) using the dense semantic labeling of 3D environments in robot navigation or to aid people with impaired sight and (3) map environments for use in augmented reality scenarios or games.
-
-\begin{figure}[!ht]	 \label{fig:failures}
+\begin{figure}[!ht]	 
  \center
@@
  \caption{Failure cases.}
+ \label{fig:failures}
 \end{figure}
 
+  The system has also some limitations.
+Combing colour and depth data requires careful sensor calibration.
 Combing colour and depth data requires careful sensor calibration.
 Lack thereof results in misclassification at object boundaries.
-Also strong changes of lighting conditions impair classification ability at certain viewpoints.
-What is more, no global context is used for classification, which could be improved if some prior knowledge were introduced.
-There are two disjoint modes: training and testing, which limits possible use cases.
-
-\section{Summary}
+Additionaly, strong changes of lighting conditions impair classification ability at certain viewpoints.
+These failures are shwon in figure \ref{fig:failures}.
+Moreover, computational performance drops with the increasing size of the scene and classification accuracy decreases with the number of classes.
+What is more, no global context is used for classification.
 
@@
 
-  - introducing prior
-  - improving user interaction system (verbal priors)
-  - improving scalability
-  - using pure geometrical features
-  - moving to outdoor settings
-  - better scaling
+  There is big room for improrvement.
+Introducting priors associated with the type of enviornment or with an object class (\emph{e.g.} vertical walls) could greatly improve classification accuracy.
+Failure cases from Fig.
+\ref{fig:failures} could be avoided if pure geometric features were used.
+Some research is required on scaling the system to bigger scenes and moving to outdoor enviornments.
 
@@
   \bibitem{hashed_voxel}
-  NIESSNER , M., Z OLLH OFER , M., IZADI , S., AND STAMMINGER , M. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
+  NIESSNER , M., ZOLLHOFER , M., IZADI , S., AND STAMMINGER , M. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
 
   \bibitem{online_forest}
-  SAFFARI , A., LEISTNER , C., SANTNER , J., G ODEC , M., AND BISCHOF , H. 2009. On-line random forests. In IEEE ICCV Workshop.
+  SAFFARI , A., LEISTNER , C., SANTNER , J., GODEC , M., AND BISCHOF , H. 2009. On-line random forests. In IEEE ICCV Workshop.
 
@@
   \bibitem{crf}
-  LAFFERTY , J., M C C ALLUM , A., AND PEREIRA , F. C. 2001. Conditional random fields: Probabilistic models for segmenting and labeling sequence data.
+  LAFFERTY , J., McCALLUM , A., AND PEREIRA , F. C. 2001. Conditional random fields: Probabilistic models for segmenting and labeling sequence data.
 
   \bibitem{inference}
-  [Kr ̈ahenb ̈uhl and Koltun 2011]
+  KRAHENBÜHL, P. AND KOLTUN , V. 2011. Efficient inference in fully connected CRFs with Gaussian edge potentials. In NIPS.
 
   \bibitem{kl}
-  [Koller and Friedman 2009]
+  KOLLER , D. AND FRIEDMAN , N. 2009. Probabilistic Graphical Models: Principles and Techniques. MIT Press
+
+  \bibitem{htree}
+  DOMINGOS , P. AND HULTEN , G. 2000. Mining high-speed data streams. In Proc. SIGKDD.
+
+  \bibitem{dataset}
+  LAI , K., BO , L., REN , X., , AND FOX , D. 2011. A large-scale hierarchical multi-view rgb-d object dataset. In Proc. ICRA.
 

### 2015-12-05 cv3d_seminar/paper.tex
spellcheck

@@
 Large-scale scene handling is facilitated by the hashed volumetric representation of \cite{hashed_voxel} with a 6mm$^{3}$ voxel.
-Each voxel contains: TSDF distance, colour information in the LAB space, a weight for averaging distance values, a user annotation class, a predictied class, an outcome of the mean-field inference from before and after classification and the probability of each class as inferred by the mean-field algorithm, for each class.
+Each voxel contains: TSDF distance, colour information in the LAB space, a weight for averaging distance values, a user annotation class, a predicted class, an outcome of the mean-field inference from before and after classification and the probability of each class as inferred by the mean-field algorithm, for each class.
 The total number of bits per voxel is $96 + \textrm{[number of classes]} \cdot 16$.
@@
 An image patch of dimensions $r \times r$  is centered on voxel $i$ from the plane $(\mathbf{p} - \mathbf{p}_i) \cdot \mathbf(n)_i = 0$.
-The patch containes colour values stored in the TSDF on the plane in CIELab to mitigate illumination effects.
-Additionaly, $V_i$ stores distance to the nearest dominant horizontal surface.
+The patch contains colour values stored in the TSDF on the plane in CIELab to mitigate illumination effects.
+Additionally, $V_i$ stores distance to the nearest dominant horizontal surface.
 $r = 13$ with a resolution of $10\frac{mm}{pixel}$ is used.
@@
 It is computed by creating a histogram of gradient directions with 36 bins, where the directions are weighted by gradient magnitudes and a gaussian window centered at the voxel.
-Furthermore, a quadratic polynomial is fit to the three biggest values of the histogram, whose maximum is the final directiion.
+Furthermore, a quadratic polynomial is fit to the three biggest values of the histogram, whose maximum is the final direction.
 
@@
 
-  A random forest is an ensamble of classification or regression trees, whose outputs are combined to produce a result with lower variance \cite{online_forest}.
+  A random forest is an ensemble of classification or regression trees, whose outputs are combined to produce a result with lower variance \cite{online_forest}.
 Each tree is typically trained only on a subset of the training data $S$ comprised of pairs $(i, l)$ where $i$ is a sample and $l$ its label.
@@
 Label statistics are sufficient to compute information gain and thus the node $n$ is further split if it has seen a predefined number of samples or if the minimal value of the information gain is low enough.
-Storing and udpdating statistics in leaf nodes is computationally expensive which limits runtime performanc
+Storing and updating statistics in leaf nodes is computationally expensive which limits runtime performance
 
@@
 Here, a pairwise CRF \cite{crf} with a time dependent underlying model is used.
-It requires a novel inference alogorithm that can handle updates of the geometry and user specified labes.
+It requires a novel inference algorithm that can handle updates of the geometry and user specified labels.
 The class of each voxel is modeled by a random variable $x_i$.
@@
   Unary and pairwise potentials $\phi_i(x_i)$ and $\phi_{ij}(x_i, x_j)$ encode the cost of assigning a label to $v_i$ and a cost of assuming different labels by voxels in a neighbourhood, respectively.
-These costs are constatly changing due to the chainging predictions of the random forest and user interaction.
+These costs are constantly changing due to the changing predictions of the random forest and user interaction.
 
@@
   To process the encircling action, the labeling is projected into the current frame.
-Next, a GMM is fit with the foreground class taken as a convex hull of the encirvled region and background as the rest of the image.
+Next, a GMM is fit with the foreground class taken as a convex hull of the encircled region and background as the rest of the image.
 For every pixel in a bounding box around the user annotation the unary potentials are updated as
@@
 
-  For all voxels that haven't been explicitly labaled by the user unitary potentials are updated with the predicted values as $\phi_i(l) = -\log P_F(x_i = l | \mathbf{D})$.
-Finally, smoothness is achieved by the standard Potts model by assigning a disconuity cost $\phi_{ij}(x_i, x_j) = \lambda_{ij}$ if voxels have different labels and zero otherwise, with $\lambda_{ij}$ a function of difference in position, intensity and normal directions.
+  For all voxels that haven't been explicitly labeled by the user unitary potentials are updated with the predicted values as $\phi_i(l) = -\log P_F(x_i = l | \mathbf{D})$.
+Finally, smoothness is achieved by the standard Potts model by assigning a discontinuity cost $\phi_{ij}(x_i, x_j) = \lambda_{ij}$ if voxels have different labels and zero otherwise, with $\lambda_{ij}$ a function of difference in position, intensity and normal directions.
 
@@
 The mean-field inference algorithm is adapted from \cite{inference} to handle the constantly changing energy landscape and implemented on GPU.
-First, the original prbability distribution $P(\mathbf{x})$ is approximated by $Q(\mathbf{x})$ under KL-divergence $KL(Q||P)$.
+First, the original probability distribution $P(\mathbf{x})$ is approximated by $Q(\mathbf{x})$ under KL-divergence $KL(Q||P)$.
 $Q(\mathbf{x})$ is chosen such that the marginal of each random variable is independent, that is $Q(\mathbf{x}) = \prod_i Q_i(\mathbf{x)_i})$.
@@
 
-with $Z_i$ a normalizing factor and final class chosen as the minizer of $Q_i^T$.
-Energy distribution is assumed to change only gradualy.
+with $Z_i$ a normalizing factor and final class chosen as the minimizer of $Q_i^T$.
+Energy distribution is assumed to change only gradually.
 Moreover, SRF classification results would impact the final segmentation after several frames due to their effect on unitary potentials.
@@
 Several video sequences were recorded to evaluate segmentation and VOP features.
-A set of keyframe covering the whole scene from each sequence was hand-labeled to create groudtruth data.
-They were further projected into the 3D volume, resulting in 4176, 12346, 7583, 8916 frames for \emph{LivingRoom}, \emph{Kitchen}, \emph{Bedroom} and \emph{Desk} sequences respectively, of wich around a third was used for testing.
-Table \ref{fig:segm_results} summarises segmantation results.
+A set of key frames covering the whole scene from each sequence was hand-labeled to create groundtruth data.
+They were further projected into the 3D volume, resulting in 4176, 12346, 7583, 8916 frames for \emph{LivingRoom}, \emph{Kitchen}, \emph{Bedroom} and \emph{Desk} sequences respectively, of which around a third was used for testing.
+Table \ref{fig:segm_results} summarizes segmentation results.
 User interaction gives very accurate labeling, but it affects a single object instance only.
@@
 
-  This paper contributes an interactive system that enables users to create customized models of 3D enviornments.
+  This paper contributes an interactive system that enables users to create customized models of 3D environments.
 It can be used to gather groundtruth for large-scale visual recognition systems, robot navigation and to aid partially sighted people.
@@
 Lack thereof results in misclassification at object boundaries.
-Additionaly, strong changes of lighting conditions impair classification ability at certain viewpoints.
-These failures are shwon in figure \ref{fig:failures}.
+Additionally, strong changes of lighting conditions impair classification ability at certain viewpoints.
+These failures are shown in figure \ref{fig:failures}.
 Moreover, computational performance drops with the increasing size of the scene and classification accuracy decreases with the number of classes.
@@
 
-  There is big room for improrvement.
-Introducting priors associated with the type of enviornment or with an object class (\emph{e.g.} vertical walls) could greatly improve classification accuracy.
+  There is big room for improvement.
+Introducing priors associated with the type of environment or with an object class (\emph{e.g.} vertical walls) could greatly improve classification accuracy.
 Failure cases from Fig.
 \ref{fig:failures} could be avoided if pure geometric features were used.
-Some research is required on scaling the system to bigger scenes and moving to outdoor enviornments.
+Some research is required on scaling the system to bigger scenes and moving to outdoor environments.
 

### 2015-12-05 cv3d_seminar/paper.tex
add references and abstract

@@
 \begin{abstract}
-	The short abstract (50-80 words) is intended to give the reader an overview of the work.
+  We describe SemanticPaint \cite{SemanticPaint} --- a system for capturing and  interactive segmentation of 3D enviornments.
+It allows adding new object categories at runtime, updates the model with and infers labels of newly seen data in an online fashion.
 \end{abstract}
@@
   Capturing your own environment has never been easier.
-SemanticPaint can register your surroundings which, after 3D reconstruction, can be semantically segmented in an interactive way.
+SemanticPaint \cite{SemanticPaint} register your surroundings which, after 3D reconstruction, can be semantically segmented in an interactive way.
 Not only it works in real time but also requires no pretraining.
@@
   \bibitem{htree}
-  DOMINGOS , P. AND HULTEN , G. 2000. Mining high-speed data streams. In Proc. SIGKDD.
+  DOMINGOS, P. AND HULTEN , G. 2000. Mining high-speed data streams. In Proc. SIGKDD.
 
@@
 
+  \bibitem{SemanticPaint}
+  VALENTIN, J., VINEET, V., CHENG, M.-M., KIM, D., SHOTTON, J., KOHLI, P., NIESSNER, M., CRIMINISI, A., IZADI, S., TORR, P.
+2015.
+SemanticPaint: Interactive 3D Labeling and Learning at your Fingertips.
+SIGGRAPH.
+
 \end{thebibliography}

### 2015-12-06 cv3d_seminar/presentation.tex
rework up to Putting it together section
references and outline for presentation

@@
 
+% 1. introduction
+%   - lots of research in registration in visual understanding since the beginning of CV
+%   - none approach so far achieved robust real time segmentation
+%   - this paper introduces an end-to-end pipeline for interactive registration and segmentation of 3D enviornments
+%   
+% 2. State of the Art
+%   - 
+%   
+% 3. Pipeline
+%   - regristration and reconstruction
+%   - classification
+%   - crf and mean-field inference for smoothing
+%   - VOPs
+% 
+%  4. Results
+%   - dataset for segmentation and VOPs
+%   - segmentation
+%   - VOPs
+%   - dataset for SRF
+%   - SRF
+%   
+% 5. Discussion
+%   - applications
+%   - failures
+%   
+% 6. Summary
+%   - what it is about
+%   
+% 7. Future Work
+%   - scaling
+%   - pure geometry
+%   - (verbal) priors
+
 %---------------------------------------------------------------------------------------
@@
 %---------------------------------------------------------------------------------------
+\section{State of the Art}
+\begin{frame}
+\frametitle{Introduction}
+
+\end{frame}
+
+%---------------------------------------------------------------------------------------
+\section{Pipeline}
+\begin{frame}
+\frametitle{Introduction}
+
+\end{frame}
+
+%---------------------------------------------------------------------------------------
+\section{Results}
+\begin{frame}
+\frametitle{Segmentation}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Features}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Streaming Random Forest}
+
+\end{frame}
+
+%---------------------------------------------------------------------------------------
+\section{Discussion and Outlook}
+\begin{frame}
+\frametitle{Summary}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Failures}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Future Work}
+
+\end{frame}
+
+%---------------------------------------------------------------------------------------
 
@@
 %---------------------------------------------------------------------------------------
-\section{State of the art}
+\section{State of the Art}
 
@@
 %---------------------------------------------------------------------------------------
+\begin{frame}
+\frametitle{References}
+\fontsize{6pt}{7.2}\selectfont
+\begin{itemize}
+  \item Roberts, L. G. 1963. Machine perception of three-dimensional solids. Ph.D. thesis, Massachusetts Institute of Technology.
+
+  \item Kim, B.-S. et. al. 2013. 3D scene understanding by voxel-CRF. In Proc. ICCV.
+
+  \item Pradeep, V. et. al. 2013. Monofusion: Real-time 3D reconstruction of small scenes with a single web camera. In Proc. ISMAR.
+
+  \item Herbst, E. et.al. 2014. Toward online 3-d object segmentation and mapping. In IEEE International Conference on Robotics and Automation (ICRA).
+
+  \item Valentin, J. P. et. al. 2013. Mesh based semantic modelling for indoor and outdoor scenes. In Proc. CVPR.
+
+  \item Levoy , M. et. al. 2000. The digital Michelangelo project: 3D scanning of large statues. In Proc. SIGGRAPH. ACM.
+
+  \item Newcombe , R. A. et. al. 2011. KinectFusion: Real-time dense surface mapping and tracking. In Proc. ISMAR.
+
+  \item Curless , B.
+et.
+al.
+1996.
+A volumetric method for building complex models from range images.
+In Proceedings of the 23rd annual conference on Computer graphics and interactive techniques.
+ACM, 303–312.
+
+  \item Niessner , M. et. al. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
+\end{itemize}
+\end{frame}
+
+%---------------------------------------------------------------------------------------
+\begin{frame}
+\frametitle{References cont'd}
+\fontsize{6pt}{7.2}\selectfont
+\begin{itemize}
+  \item Saffari , A. et. al. 2009. On-line random forests. In IEEE ICCV Workshop.
+
+  \item Vitter , J. S. 1985. Random sampling with a reservoir. ACM TOMS 11, 1.
+
+  \item Lower , D. G. 1999. Object recognition from local scale-invariant features. In Proc. ICCV.
+
+  \item Lafferty , J. et. al. 2001. Conditional random fields: Probabilistic models for segmenting and labeling sequence data.
+
+  \item Ktahenbül, P. et. al. 2011. Efficient inference in fully connected CRFs with Gaussian edge potentials. In NIPS.
+
+  \item Koller , D. et.al  , N. 2009. Probabilistic Graphical Models: Principles and Techniques. MIT Press
+
+  \item Domingos, P. et. al. 2000. Mining high-speed data streams. In Proc. SIGKDD.
+
+  \item Lai, K. et. al. 2011. A large-scale hierarchical multi-view rgb-d object dataset. In Proc. ICRA.
+
+  \item Valentin, J. et. al. 2015. SemanticPaint: Interactive 3D Labeling and Learning at your Fingertips. SIGGRAPH.
+\end{itemize}
+\end{frame}
+
+%---------------------------------------------------------------------------------------
 \end{document}

### 2015-12-06 cv3d_seminar/paper.tex
results

@@
   $\Delta$ RGB mean & 80\% & 71.84\% & 76.29\% & 73.42\% & 75.39\% \\
-  Depth probe& 77.54\% & 61.79\% & \textbf{84.9\%} & 68.9\% & 73.06\% \\
+  Depth Probe& 77.54\% & 61.79\% & \textbf{84.9\%} & 68.9\% & 73.06\% \\
   Color Probe& 56.39\% & 65.68\% & 60.77\% & 60.74\% & 60.9\% \\

### 2015-12-06 cv3d_seminar/presentation.tex
results

@@
-\documentclass[mathserif]{beamer}
+\documentclass[mathserif, 8pt]{beamer}
 \input{macro}
+
+% \usepackage[font=footnotesize,labelfont=bf]{caption}
+% \setbeamerfont{caption}{size=\scriptsize}
 
@@
 
+\begin{table}[!ht]
+  \fontsize{10pt}{7.2}\selectfont
+ \center
+ \caption{Segmentation Results}
+\noindent\makebox[\linewidth]{
+  \begin{tabular}{c|c|c|c|c|c}
+    \textbf{Component} & \textbf{LivingRoom} & \textbf{Bedroom} & \textbf{Kitchen} & \textbf{Desk} & \textbf{Average} \\ \hline
+    User Interaction & 99.35\% & 97.61\% & 96.09\% & 97.73\% & 97.7\% \\ 
+    Forest Prediction & 94.57\% & 88.31\% & 82.58\% & 90.29\% & 88.94\% \\
+    Final Inference & 96.26\% & 95.19\% & 90.69\% & 95.55\% & 94.42\%
+  \end{tabular}
+ }
+ \label{fig:segm_results}
+\end{table}
+
 \end{frame}
@@
 
+\begin{table}[!ht]
+ \fontsize{10pt}{7.2}\selectfont
+ \center
+ \caption{Feature Comparison}
+ \noindent\makebox[\linewidth]{
+  \begin{tabular}{c|c|c|c|c|c}
+  \textbf{Feature} & \textbf{LivingRoom} & \textbf{Bedroom} & \textbf{Kitchen} & \textbf{Desk} & \textbf{Average} \\ \hline
+  VOP & \textbf{94.57\%} & \textbf{88.31\%} & 82.58\% & \textbf{90.29\%} & \textbf{88.94\%} \\
+  $\Delta$ RGB mean & 80\% & 71.84\% & 76.29\% & 73.42\% & 75.39\% \\
+  Depth Probe& 77.54\% & 61.79\% & \textbf{84.9\%} & 68.9\% & 73.06\% \\
+  Color Probe& 56.39\% & 65.68\% & 60.77\% & 60.74\% & 60.9\% \\
+  SURF & 43.74\% & 67.12\% & 57\% & 58.13\% & 56.5\% \\
+  SPIN & 58.77\% & 43.22\% & 48.41\% & 36.1\% & 46.63\% \\
+ \end{tabular}
+ }
+ \label{fig:vop_results}
+\end{table}
+
 \end{frame}
@@
 \frametitle{Streaming Random Forest}
+
+\begin{columns}
+ \begin{column}{0.5\textwidth}
+  \begin{figure}[!ht]
+  \center
+  \includegraphics[width=\textwidth]{figures/srf_ap}
+  \vspace{-0.25cm}
+  \caption{Average Precision}
+  \label{fig:srf_ap}
+  \end{figure}
+  \vspace{-0.5cm}
+  \begin{figure}[!ht]
+  \center
+  \includegraphics[width=\textwidth]{figures/srf_iu}
+  \vspace{-0.25cm}
+  \caption{Intersection/Union}
+  \label{fig:srf_iu}
+  \end{figure}
+ \end{column}
+
+ \begin{column}{0.5\textwidth}
+  Data:\\
+  300 objects\\
+  51 classes\\
+  full revolution\\
+  3 points of view \\
+  \vspace{0.5cm}
+  SRF - Streaming Random Forest\\
+  ORF - Online Random Forest\\
+  HT - Hoeffding Tree \\
+ \end{column}
+\end{columns}
 
@@
 \end{columns}
-\end{frame}
-
-%---------------------------------------------------------------------------------------
-\section{State of the Art}
-
-\begin{frame}
-\frametitle{Another section}
 \end{frame}

### 2015-12-06 cv3d_seminar/presentation.tex
add section seprator slide
all but SRF and Related

@@
-\documentclass[mathserif, 8pt]{beamer}
+\documentclass[mathserif, 10pt]{beamer}
 \input{macro}
@@
 }
-
+\setbeamertemplate{footline}[frame number]
 \setbeamercolor{math text}{fg=orange!80!black}
 \newcommand{\hl}[1]{\textcolor{red!80!black}{\underline{#1}}}
+
+\AtBeginSection[]{
+  \begin{frame}
+  \vfill
+  \centering
+  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
+    \usebeamerfont{title}\insertsectionhead\par%
+  \end{beamercolorbox}
+  \vfill
+  \end{frame}
+}
 
@@
 \begin{frame}
-\frametitle{Introduction}
+\frametitle{Acquisition and Reconstruction}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Scene Understanding}
 
@@
 \begin{frame}
-\frametitle{Introduction}
+\frametitle{Voxel Oriented Patch features}
+
+\begin{columns}
+ \begin{column}{0.5\textwidth}
+  \begin{figure}
+  \center
+  \includegraphics[width=\textwidth]{figures/vop}
+  \caption{Colours shown in RGB for illustration purposes.}
+  \label{fig:vop}
+\end{figure}
+ \end{column}
+ \begin{column}{0.5\textwidth}
+    $(\mathbf{p} - \mathbf{p}_i) \cdot \mathbf(n)_i = 0$ \\
+    $r \times r$, $r = 13px$ with $10\frac{mm}{pixel}$\\
+    CIELab\\
+    Rotated to dominant gradient direction
+
+ \end{column}
+
+\end{columns}
+\end{frame}
+
+\begin{frame}
+\frametitle{Streaming Random Forest}
+\end{frame}
+
+\begin{frame}
+\frametitle{SRF - Reservoir Splitting}
+
+\begin{figure}[!ht]
+  \center
+  \includegraphics[width=\textwidth]{figures/forest}
+%   \caption{Splitting reservoirs in Streaming Random Forest.}
+  \label{fig:forest}
+\end{figure}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Dynamic Conditional Random Field}
+
+  \begin{equation} \label{eq:posterior}
+  P(\mathbf{x}|\mathbf{D}) = \prod_{i \in \mathcal{V}} \left( \psi_i(x_i) \prod_{j \in \mathcal{E}_i} \psi_{ij}(x_i, x_j) \right) 
+  \end{equation}
+
+  \begin{equation} \label{eq:energy}
+  E_t(\mathbf{x}) = \sum_{i \in \mathcal{V}} \left( \phi_i(x_i) + \sum_{j \in \mathcal{E}_i} \phi_{ij} (x_i, x_j) \right) + K
+  \end{equation}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{CRF - User Interactions}
+
+  Touching:
+  \begin{equation}
+  \phi_i(l) =
+    \begin{cases}
+      0      & \quad \text{if } l = l_S\\
+      \infty  & \quad \text{otherwise}\\
+    \end{cases}
+  \end{equation}
+
+  Encircling:
+  \begin{equation}
+  \phi_i(l) =
+    \begin{cases}
+      \log P_E(fg|\mathbf{a}_i)       & \quad \text{if } l = \text{fg}\\
+      \log (1 - P_E(fg|\mathbf{a}_i))  & \quad \text{if } l = \text{bg}\\
+    \end{cases}
+  \end{equation}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{CRF - Predictions and Smoothnes}
+
+  Predictions:
+  \begin{equation}
+  \phi_i(l) = -\log P_F(x_i = l | \mathbf{D})
+  \end{equation}
+
+  Smoothnes:
+  \begin{equation}
+      \phi_{ij}(x_i, x_j)  = \theta_p e^{-||\mathbf{p}_i - \mathbf{p}_j||}  + \theta_a e^{-||\mathbf{a}_i - \mathbf{a}_j||} + \theta_n e^{-||\mathbf{n}_i - \mathbf{n}_j||}
+  \end{equation}
+\end{frame}
+
+\begin{frame}
+\frametitle{Mean-Field Inference}
+$P(\mathbf{x})$ approximated by $Q(\mathbf{x})$ under $KL(Q||P)$:
+\begin{equation}
+ Q_i^t(l) = \frac{1}{Z_i}e^{M_i(l)} \text{, } t = 1, \ldots, T
+\end{equation}
+\begin{equation}
+ M_i(l) = \phi_i(l) + \sum_{l' \in \mathcal{L}} \sum_{j \in \mathcal{N}(i)} Q_j^{t-1}(l')\phi_{ij}(l, l')
+\end{equation}
+
+Frame at time $t$ initialized with:
+\begin{equation}
+ \widetilde{Q}_i^t(x_i) = \gamma Q_i^{t-1}(x_i) + (1 - \gamma) P_F^{t-1}(x_i = l | \mathbf{D}) \text{, } \gamma \in [0, 1]
+\end{equation}
 
@@
 \end{columns}
-
 \end{frame}
@@
 \frametitle{Summary}
+\begin{itemize}
+ \item customized models of 3D enviornments
+ \item fully interactive
+ \item online and real time
+ \item no pretraining
+\end{itemize}
 
@@
 \frametitle{Failures}
+\begin{columns}
+ \begin{column}{0.5\textwidth}
+  \begin{figure}[!ht]	 
+  \center
+  \includegraphics[width=\textwidth]{figures/failures}
+  \caption{Failure cases.}
+  \label{fig:failures}
+  \end{figure}
+ \end{column}
+
+ \begin{column}{0.5\textwidth}
+  \begin{itemize}
+   \item bleeding
+   \item illumination change
+   \item viewpoint change
+  \end{itemize}
+ \end{column}
+\end{columns}
 
@@
 \frametitle{Future Work}
+\begin{itemize}
+ \item class priors for different enviornments
+ \item priors for class properties (vertical walls)
+ \item discriminative geometrical features
+ \item outdoor enviornments
+ \item better scalability
+\end{itemize}
 
-\end{frame}
-
-%---------------------------------------------------------------------------------------
-
-\begin{frame}
-\frametitle{Two columns}
-\begin{columns}[t]
-\begin{column}{0.45\linewidth}
-Left
-\end{column}
-\begin{column}{0.45\linewidth}
-Right
-\end{column}
-\end{columns}
-\end{frame}
-
-%---------------------------------------------------------------------------------------
-\begin{frame}
-\frametitle{How to include an image}
-\begin{center}
-\includegraphics[width=0.3\linewidth]{template}
-\end{center}
 \end{frame}

### 2015-12-07 cv3d_seminar/presentation.tex
presentation + notes

@@
 \author[short presentator]{Adam Kosiorek}
-\institute[CAMP]{Computer Aided Medical Procedures \\
-Technische Universit\"at M\"unchen}
+\institute[CAMP]{Advisor: M.Eng.~Keisuke Tateno}
 \date[]{10.12.2015}
@@
 
+\begin{table}
+\noindent\makebox[\linewidth]{
+ \begin{tabular}{c|c|c}
+  Who & What & How \\ \hline
+  Levoy et. al. 2000 & world-scale 3D models & offline, from online images\\ \hline
+  Newcombe et. al. 2011  & online 3D scanning & KinectFusion \\ \hline
+  Pradeep et.al. 2013 & 3D reconstructiona & sparse tracking and  \\
+    & with 1 RGB camera  & stereo reconstruction \\
+    && on par with KinectFusion
+ \end{tabular}
+}
+\end{table}
+
 \end{frame}
@@
 \frametitle{Scene Understanding}
+
+\begin{table}
+\noindent\makebox[\linewidth]{
+ \begin{tabular}{c|c|c}
+  Who & What & How \\ \hline
+  Kim et. al. 2013 & reconstruction & Voxel-based CRF \\ 
+    & segmentation   & with visibility contraints\\ \hline
+
+  Herbst et. al. 2014  & registration & online \\
+     & segmentation & change detection\\ \hline
+
+  Valentin et. al. 2013  & inference on & RGB and geom. features  \\
+    & mesh from TSDF  & CRF segmentation \\
+ \end{tabular}
+}
+\end{table}
 
@@
 \begin{frame}
+\frametitle{Random Forest}
+
+\begin{columns}
+ \begin{column}{0.5\textwidth}
+  \begin{figure}
+  \center
+  \includegraphics[width=\textwidth]{figures/rf}
+  \caption{Single tree}
+\end{figure}
+ \end{column}
+ \begin{column}{0.5\textwidth}
+    bagged trees\\
+    bootstraped data\\
+
+    voting for final result\\
+    greedy training\\
+    off-line, all data at once\\
+    $(i, l) \in \mathcal{S}$ - (voxel, label) pairs\\
+    $f(i, \theta)$ - split functions\\
+    $\Theta$ - distribution of split functions\\
+    $P_F(x_i = l | \mathbf{D})$\\
+ \end{column}
+
+\end{columns}
+\end{frame}
+
+\begin{frame}
 \frametitle{Streaming Random Forest}
+
+  Information Gain:
+  \begin{equation} \label{eq:infogain}
+    G(S, S^L, S^R) = H(S) - \sum_{d \in \{L, R\}} \frac{|S^d|}{|S|}H(S^d)
+  \end{equation}
+  Shannon Entropy:
+  \begin{equation}
+   H(S) = - \sum_{(l, i) \in S} p(c_i = l) \log{p(c_i )}
+  \end{equation}
 \end{frame}

### 2015-12-08 cv3d_seminar/presentation.tex
some updates

--- 2015-12-06 cv3d_seminar/presentation.tex
@@
 \section{Introduction}
-\begin{frame}
-\frametitle{Introduction}
-
-\end{frame}
 
@@
 \section{State of the Art}
+
 \begin{frame}
-\frametitle{Acquisition and Reconstruction}
+\frametitle{Scene Understanding}
+
+\begin{table}
+\noindent\makebox[\linewidth]{
+ \begin{tabular}{c|c|c}
+  Who & What & How \\ \hline
+
+  Valentin et. al. 2013  & inference on & RGB and geom. features  \\
+  & mesh from TSDF  & CRF segmentation \\ \hline
+
+  Kim et. al. 2013 & reconstruction & Voxel-based CRF \\ 
+    & segmentation   & with visibility contraints\\ \hline
+
+  Herbst et. al. 2014  & reconstruction & online model updates \\
+     & segmentation & change detection\\ \hline
+
+ \end{tabular}
+}
+\end{table}
 
@@
 \begin{frame}
-\frametitle{Scene Understanding}
+\frametitle{Model-based SLAM}
+
+\begin{table}
+\noindent\makebox[\linewidth]{
+ \begin{tabular}{c|c|c}
+  Who & What & How \\ \hline
+  Newcombe et. al. 2011  & online 3D SLAM & model-based tracking \\
+  && global TSDF volume \\ \hline
+
+  Salas-Moreno et al. 2013 & object-level & offline object database \\
+  & SLAM & pose-object graph\\ \hline
+
+  Pradeep et.al. 2013 & 3D reconstruction & sparse tracking and  \\
+    & with 1 RGB camera  & stereo reconstruction \\
+    && on par with KinectFusion
+ \end{tabular}
+}
+\end{table}
 
@@
 \section{Pipeline}
+
+\begin{frame}
+\frametitle{Pipeline Overview}
+\begin{figure}
+ \includegraphics[width=\textwidth]{figures/pipeline}
+\end{figure}
+
+\end{frame}
+
 \begin{frame}
@@
 \begin{frame}
+\frametitle{Random Forest}
+
+\begin{columns}
+ \begin{column}{0.5\textwidth}
+  \begin{figure}
+  \center
+  \includegraphics[width=\textwidth]{figures/rf}
+  \caption{Single tree}
+\end{figure}
+ \end{column}
+ \begin{column}{0.5\textwidth}
+    bagged trees\\
+    greedy training\\
+    bootstraped data\\    
+    off-line, all data at once\\
+    voting for final result\\
+
+    $(i, l) \in \mathcal{S}$ - (voxel, label) pairs\\
+    $f(i, \theta)$ - split functions\\
+    $\Theta$ - distribution of split functions\\
+    $P_F(x_i = l | \mathbf{D})$ - class conditional probability\\
+ \end{column}
+
+\end{columns}
+\end{frame}
+
+\begin{frame}
 \frametitle{Streaming Random Forest}
+  \begin{itemize}
+   \item Node n: Reservoir $R_n$ with a list of samples $T_n$, $|T_n| \leq K$
+   \item First $K$ samples added
+   \item Current samples swapped with new ones with decreasing probability
+   \item Split node if: $|R_n| > N$
+  \end{itemize}
+
+  Information Gain:
+  \vspace{-0.25cm}
+  \begin{equation} \label{eq:infogain}
+    G(R_n, R_n^L, R_n^R) = H(R_n) - \sum_{d \in \{L, R\}} \frac{|R_n^d|}{|R_n|}H(R_n^d)
+  \end{equation}
+  Shannon Entropy:
+  \vspace{-0.25cm}
+  \begin{equation}
+   H(R_n) = - \sum_{(l, i) \in T_n} p(c_i = l) \log{p(c_i = l)}
+  \end{equation}
+
+  \vspace{-0.25cm}
+  $H(R_n)$ computed from a node's class distribution
 \end{frame}
@@
 \frametitle{SRF - Reservoir Splitting}
-
+$m_n$ - number of samples seen at node $n$\\
+$P(l | T_n)$ - normalized class distribution of $R_n$
+\vspace{-0.25cm}
 \begin{figure}[!ht]
@@
 
+Joint class probability distribution for the volume $\mathcal{V}$:
   \begin{equation} \label{eq:posterior}
@@
 
+\vspace{-0.5cm}
+Labeling Energy at time $t$:
   \begin{equation} \label{eq:energy}
@@
   \end{equation}
+
+\vspace{-0.5cm}  
+where:\\
+$\phi_i(x_i)$ - cost of assigning a label \\
+$\phi_{ij}(x_i, x_j)$ - cost of assuming different labels\\
+$\mathcal{E}_i$ - neighbourhood of voxel $i$
 
@@
 \frametitle{CRF - User Interactions}
-
+% \vspace{-1cm}
   Touching:
+  \begin{columns}
+   \begin{column}{0.25\textwidth}
+    \begin{figure}
+    \includegraphics[width=\textwidth]{figures/touch}
+    \end{figure}
+    \end{column}
+    \begin{column}{0.75\textwidth}
       \begin{equation}
@@
 	  \begin{cases}
-      0      & \quad \text{if } l = l_S\\
+	    0      & \quad \text{if } l = l_T\\
 	    \infty  & \quad \text{otherwise}\\
@@
       \end{equation}
+      \center
+      $T$ --- touched pixels
+   \end{column}
 
+  \end{columns}
+
+  \vspace{0.5cm}
   Encircling:
+  \vspace{-0.5cm}
+    \begin{columns}
+   \begin{column}{0.25\textwidth}
+    \begin{figure}
+    \includegraphics[width=\textwidth]{figures/circle}
+    \end{figure}
+   \end{column}
+    \begin{column}{0.75\textwidth}
       \begin{equation}
@@
   \end{equation}
+
+  \center
+  $P_E$ from GMM\\
+  fg --- inside\\
+  bg --- outside
+   \end{column}
+
+  \end{columns}
 
@@
   \end{equation}
+  $P_F$ --- Streaming Random Forest prediction
 
+  \vspace{0.5cm}
   Smoothnes:
@@
   \end{equation}
+
+  $\theta_p$, $\theta_a$, $\theta_n$ --- paramters \\
+  $\mathbf{p}_i$ --- position\\
+  $\mathbf{a}_i$ --- appearance\\
+  $\mathbf{n}_i$ --- normal vector
+
 \end{frame}
@@
 \begin{equation}
- M_i(l) = \phi_i(l) + \sum_{l' \in \mathcal{L}} \sum_{j \in \mathcal{N}(i)} Q_j^{t-1}(l')\phi_{ij}(l, l')
+ M_i(l) = \phi_i(l) + \sum_{l' \in \mathcal{L}} \sum_{j \in \mathcal{E}_i} Q_j^{t-1}(l')\phi_{ij}(l, l')
 \end{equation}
@@
 \section{Results}
+\begin{frame}
+\frametitle{Segmentation}
+  \begin{figure}
+   \includegraphics[width=\textwidth]{figures/results}
+  \end{figure}
+
+\end{frame}
+
 \begin{frame}
@@
 
+\end{frame}
+
+\begin{frame}
+\frametitle{Features}
+\vspace{-0.25cm}
+ \begin{figure}[!ht]
+  \center
+  \includegraphics[width=\textwidth]{figures/results_vop}
+%   \caption{Average Precision}
+%   \label{fig:srf_ap}
+  \end{figure}
 \end{frame}
@@
 
-  \item Levoy , M. et. al. 2000. The digital Michelangelo project: 3D scanning of large statues. In Proc. SIGGRAPH. ACM.
+  \item Salas-Moreno, R. F. et. al. 2013. SLAM++: Simultaneous localisation and mapping at the level of objects. In Proc. CVPR.
 

### 2015-12-09 cv3d_seminar/presentation.tex
finished notes, add Q&A slide with SemanticPaint diagram
add inline notes

@@
 \input{macro}
-
+\usepackage{pdfpcnotes}
+\usepackage{multimedia}
 % \usepackage[font=footnotesize,labelfont=bf]{caption}
@@
 \titlepage
+\pnote{name}
+\pnote{framework for interactive capturing of 3D enviornments}
 \end{frame}
@@
 \tableofcontents
+\pnote{Let's start by seeing the framework in action}
 \end{frame}
@@
      & segmentation & change detection\\ \hline
-
  \end{tabular}
@@
 \end{table}
+
+\pnote{research on: RFB, D, point cloud, mesh, volume}
+\pnote{most offline, not real time; CNN for class & detetect, CRF and variation for segm}
+\pnote{Valentin}
+\pnote{depth into volumetric TSDF, triangulated, mesh}
+\pnote{rgb on img, geo on mesh}
+\pnote{KITTI, NYU; mesh speed and accuracy}
+\pnote{}
+\pnote{Kim}
+\pnote{voxel <- visibility and occupancy, vis constrained}
+\pnote{model and reduce noise in depth maps}
+\pnote{offline, global inference}
+\pnote{}
+\pnote{Herbst}
+\pnote{detect change after model updates}
+\pnote{static unchanged, dynamic changed}
+\pnote{online but not real time, 2s per frame}
 
@@
 
+\pnote{SLAM important in robotics, SemanticPaint relies heavily}
+\pnote{major contribution by Newcombe, KinectFusion; 3D reconstruction from a single depth sensor}
+\pnote{model-based tracking to match frame against model}
+\pnote{global volume in TSDF}
+\pnote{}
+\pnote{extends KinectFusion -> object classification}
+\pnote{match objects from the scene against a database}
+\pnote{obj-pose graph, sparse, compact repr}
+\pnote{sparse, requires offline database}
+\pnote{}
+\pnote{3D from single RGB camera}
+\pnote{tracking, key and secondary frames, stereo reconsutrction}
+\pnote{small scenes, good performance for textured}
 \end{frame}
@@
 
+\pnote{depth + rgb = TSDF}
+\pnote{user interactions = touched obj into CRF => priors}
+\pnote{also into RF => predict labels for other objects}
+\pnote{feedback loop: CRF gives labels from SRF, RF is trained on CRF}
 \end{frame}
@@
     Rotated to dominant gradient direction
+ \end{column}
+\end{columns}
 
- \end{column}
-
-\end{columns}
 \end{frame}
@@
   $H(R_n)$ computed from a node's class distribution
+
+  \pnote{SRF inherently online; some terminology}
+  \pnote{the behind the reservoir: samples over huge time window -> unbiased sample of all data}
+  \pnote{objective: Information Gain from normalized class distribution at each node}
 \end{frame}
@@
 
+\pnote{create node + distribution of split functions}
+\pnote{split node -> sweep through R for every split function and compute class distributions}
+\pnote{no explicit R computations are needed, just the splits}
+\pnote{pass the statistics to child nodes}
 \end{frame}
@@
 
+\pnote{CRFS subclass of PGMs; classification of sequences; global contex}
+\pnote{in CV: segmentation; here: smoot out segm and user inter}
+\pnote{}
+\pnote{psi - likelihood and prior}
+\pnote{distr -> NLL -> energy -> minimize}
+\pnote{phi -> cost of label, smoothnes}
+\pnote{eps -> neigbourhood, 6cm}
 \end{frame}
@@
 \frametitle{CRF - User Interactions}
-% \vspace{-1cm}
+
   Touching:
@@
    \end{column}
-
   \end{columns}
 
+\pnote{Note -> changing energy landscape <- user interactions}
+\pnote{touch and say -> set potentials; 0 if correct or inf otherwise}
+\pnote{}
+\pnote{circle more complication -> convex hull of selection as fg, others bg}
+\pnote{fit GMM and infer in a boundix box around selection}
+\pnote{to update potentials}
 \end{frame}
@@
 
+\pnote{Next come forest predictions -> update pot of otherwise unaffected voxels -> NLL of current class}
+\pnote{smothness -> priors based on differences}
 \end{frame}
@@
 
+\pnote{Then comes mean-field inference}
+\pnote{approximate original P(x) under KL with Q(x)}
+\pnote{iterative algo to refine}
+\pnote{gradual change of energy landscape}
+\pnote{single update in each frame}
+\pnote{to speed up RF's impact -> initialization}
 \end{frame}
@@
 
+\pnote{Let's look at some segmentation results}
+\pnote{We have... for 2 different scenes}
+\pnote{inference improves results quite a lot}
+\pnote{we can relabel and correct mistakes}
 \end{frame}
@@
 
+\pnote{if we look at percentage}
+\pnote{iteraction very well}
+\pnote{RF ok}
+\pnote{mean-field smoothing improves}
 \end{frame}
@@
   \end{figure}
+
+\pnote{Next, let's look at features}
+\pnote{compared indirectly by comparing how RF behaves while using different features}
+\pnote{discriminative power}
+\pnote{here: 3 best results; VOP most consistent}
+
 \end{frame}
@@
 
+\pnote{Again, if we look at percentage, VOP the best}
+\pnote{Kitchen is worse -> harsh illumination}
+\pnote{makes RGB feature inconclusive}
 \end{frame}
@@
 \end{columns}
+
+\pnote{SRF against two other tree-based online classifiers ORF and HT}
+\pnote{simulate online setting: add new classes and PoVs sequentially}
+\pnote{SRF outperforms comfortably, more effiiicient}
+\pnote{Sad fact: accurayc drops with classes}
+\pnote{handling coplicated scenes tricky}
 \end{frame}
@@
 
+\pnote{To sum up: delivers pipeline that lets capture personalized 3D envs as you need}
+\pnote{fully interactive, req no pretrain and works online}
+\pnote{possible apps: robot navi, sighted people, datasets for large-scale CV}
 \end{frame}
@@
 
+\pnote{Unf, some failures}
+\pnote{Bleeding, or misclassif at object edges <- errors of depth and colour sensor alignm}
+\pnote{big illumination and viewpoint changes -> errors}
 \end{frame}
@@
 \begin{itemize}
+  \item discriminative geometrical features
+  \item priors for class properties (vertical walls)
  \item class priors for different enviornments
- \item priors for class properties (vertical walls)
- \item discriminative geometrical features
  \item outdoor enviornments
@@
 
+\pnote{Some future work to be done}
+\pnote{Geo features -> prevent bleeding}
+\pnote{class priors (vertical walls) and for diff envs -> better accuracy}
+\pnote{outdoor envs -> passive sensors}
+\pnote{bigger scenes and more classes -> scalable algos}
 \end{frame}
@@
 
+\begin{frame}
+\frametitle{Q\&A?}
+\begin{figure}
+ \includegraphics[width=0.8\textwidth]{figures/SemanticPaint}
+\end{figure}
+
+\end{frame}
+
 %---------------------------------------------------------------------------------------

### 2015-12-10 ML/hw8/hw.tex
final version hw8 ML

@@
 
-The k-nearest neighbours algorithm in the feature space by introducing the feature map is
+The distance to a training sample in feature space is given by
 
@@
 
-In order that it only depends on the scalar product in feature space $K(\phi(x),\phi(x^{(s_i)})) = \phi(x)^T\phi(y)$, we square this. Using that operation, the k-nearest neighbours don't change.
+In order that it only depends on the scalar product in feature space $K(x, x^{(s_i)}) = \phi(x)^T\phi(y)$, we square this. Using that operation, the k-nearest neighbours don't change.
 
@@
 =& \max\limits_{\lambda}f_\lambda((1-\alpha)(x + \alpha y).
-\end{align}
-
-\section*{Problem 4}
-The Lagrangian is given by
-
-\begin{equation}
-L(x,\alpha) = f_0(x) + \sum_{i}{\alpha_i f_i(x)}.
-\end{equation}
-
-This can also be interpreted as a family of functions $L_x(\alpha) = L(x,\alpha)$, that means we interpret $x$ as fixed. Thus $L_x(\alpha)$ has the form
-
-\begin{equation}
-L_x(\alpha) = C_{0,x} + \sum_{i}{C_{i,x} \alpha_i},
-\end{equation}
-
-where the $Cs$ are constants. By applying the definition of a convex function we see immediately that $L_x(\alpha)$ is convex for any $Cs$.
-The Lagrange dual function is given by the pointwise minimum of this family of functions,
+\end{align}s
 
@@
 The negative Lagragian $-L_x(\alpha)$ is also convex, because the negation only changes the sign of the
-constants $C_{0,x}$ and $C_{i,x}$. Thus, as shown in the previous exercise,$-g(\alpha)$ is convex and $g(\alpha)$ is concave.
+constants $C_{0,x}$ and $C_{i,x}$. Thus, as shown in the previous exercise,$-g(\alpha)$ is convex and $g(\alpha)$ is concave
 

### 2015-12-12 cv3d_seminar/paper.tex
all but SRF and Related
final paper version; shortened and reworked a little

@@
 \usepackage{amsmath}
+\usepackage{subcaption}
 
@@
 
-	\title{SemanticPaint}
+	\title{An Overview of SemanticPaint}
 	\author{Adam Kosiorek
-	\thanks{Advisor: M.Eng.~Keisuke Tateno, Chair for Computer Aided Medical Procedures \& Augmented Reality, TUM, WS 2015/16.}}
+	\thanks{Advisor: M.Eng.~Keisuke Tateno, Chair for Computer Aided Medical Procedures~\&~Augmented Reality, TUM, WS 2015/16.}}
 	\markboth{Master Seminar: Recent Trends in 3D Computer Vision}{}
@@
 \begin{abstract}
-  We describe SemanticPaint \cite{SemanticPaint} --- a system for capturing and  interactive segmentation of 3D enviornments.
+  We describe SemanticPaint \cite{SemanticPaint} --- a system for capturing and  interactive segmentation of 3D environments.
 It allows adding new object categories at runtime, updates the model with and infers labels of newly seen data in an online fashion.
@@
 
-  The pipeline starts with capturing the environment as a stream of noisy RGBD images and combining them into a  3D model updated in an online fashion.
-The user can choose which objects to label and can do so by ``touching'' a small part of an object or encircling it with his hand and uttering the label.
-The label and the affected data points are further passed to a Streaming Random Forest (SRF) classifier which constantly learns and labels all visible voxels.
-To further improve classification results a spatially dense labeling is produced by an efficient mean-field inference algorithm.
-One of the biggest strengths of SemanticPaint is the efficiency of each part of the pipeline, which translates to real time performance.
-Algorithms used in the pipeline were adapted to work on volumetric data in the TSDF format directly in order to avoiding costly conversions to mesh or point-cloud formats.
-To allow this, the Voxel Oriented Patch features --- a new type of a discriminative feature describing the voxel space --- has been designed.
-
 \section{Related Work}
 
-  \textbf{Acquisition and Reconstruction } Capturing the geometry of the surrounding world has been a long standing problem \cite{old}.
-An offline processing of multiple images allowed to reconstruct digital heritage and construct world-scale 3D models with remarkable quality \cite{heritage}.
-The inception of low-cost RGBD sensors and powerful GPUs enabled online 3D scanning, augmented reality and using 3D environment models for navigation purposes \cite{fusion}.
-\cite{Monofusion} enables 3D reconstruction of small scenes using a single off-the-shelf RGB camera.
-It uses a sparse tracking method to first estimate the camera's pose and then select key frames and relative to them secondary frames, from which 3D stereo reconstruction is performed.
-The achieved results are similar to KinectFusion with the only limitation being that the precision of texture-less surfaces' reconstruction is somewhat lacking.
+  \textbf{Scene Understanding } Object recognition, detection and segmentation has been done with colour and RGBD images, point clouds, meshes and volumetric representations.
+\cite{mesh}  uses a TSDF model to reconstruct a 3D scene from a sequence of RGBD images.
+The volumetric representation is triangulated and a 3D mesh is recovered.
+Next, visual features are computed on images and projected into the 3D model, while geometric features are computed on the mesh directly.
+Segmentation is done via a CRF.
+This approach achieves state-of-the-art results on KITTI and NYU datasets for indoor and outdoor scene segmentation.
+\cite{VCRF} uses a Voxel-based CRF for offline simultaneous reconstruction and segmentation.
+Each voxel contains information about visibility and occlusion as well as group membership.
+The first two are used to improve reconstruction by mitigating depth-map noise.
+The visibility values are constrained by that each ray from the camera can hit only one visible vertex.
+The group membership information encodes priors given by bounding boxes of detected objects.
+\cite{ChangeDetection} is a first step towards online registration and segmentation.
+Using RGBD images, the framework constructs a model of the environment and updates it with each new frame.
+When a significant change in the model is detected, it is split into a static and a dynamic part, where the latter is assumed to have moved in space.
+The movement is detected by comparing the expected and observed intensity values at each voxel.
+The system is online, but is far from real time with 0.7 to 2s processing time per frame.
 
   \noindent
-  \textbf{Scene Understanding } Object recognition, detection and segmentation has been done with 2D RGB images, RGBD data, point clouds and volumetric representations.
-\cite{VCRF} uses a Voxel-based CRF for simultaneous reconstruction and segmentation.
-Each voxel contains information about visibility and occlusion as well as group membership.
-The first two are used to to improve reconstruction by mitigating depth-map noise.
-The visibility values are constrained by that each ray from the camera can hit only one visible vertex.
-The group membership information encodes priors given by bounding boxes of detected objects.
-A graph-cut algorithm is used for global inference.
-The whole scene has to be registered beforehand, since any change in the CRF's structure would require restarting the inference algorithm.
-No computational performance was reported.
-\cite{ChangeDetection} is a first step towards online simultaneous registration and segmentation.
-Using RGBD images, the framework constructs a model of the environment and updates it each time a new frame comes in.
-When a significant change in the model is detected, the resulting model is split into a static and a dynamic part, where the latter is assumed to have moved in space.
-The movement is detected by comparing the expected and observed intensity values at each voxel.
-The system is online, but is far from real time with 0.7 to 2s processing time per frame.
-\cite{mesh} shows that converting to another data representation can help to improve segmentation accuracy and inference speed.
-TSDF model is used to reconstruct a 3D scene from a sequence of RGBD images.
-The volumetric representation is triangulated and a 3D mesh is recovered.
-Next, visual features are computed directly on images and projected into the 3D model, while geometric features are computed on the mesh directly.
-Segmentation is done via a CRF.
-The approach is tested on the augmented KITTI and NYU datasets for indoor and outdoor scene segmentation where it delivers state-of-the-art results.
+  \textbf{Model-based SLAM } KinectFusion \cite{fusion} has brought online 3D reconstruction from a single depth sensor.
+It uses model-based tracking to match an incoming frame against a global volume in the TSDF \cite{TSDF} format.
+SLAM++ extends KinectFusion by object classification capabilities.
+It classifies objects by matching object in the scene against a database of objects in real time.
+It then builds an object-pose graph, which is a very sparse and compact representation of the world.
+It works online, but requires a previously prepared object database.
+\cite{monofusion} enables 3D reconstruction of small scenes using a single RGB camera.
+It uses a sparse tracking method to first estimate the camera's pose and then select key frames and relative to them secondary frames, from which 3D stereo reconstruction is performed.
+The achieved results are similar to KinectFusion with the only limitation being the poor precision at texture-less surfaces.
 
@@
 
-\section{Space Representation}
+\section{Pipeline Overview}
 
-  The space is represented by a 3D volume constructed from the sequence of RGBD images using Truncated Signed Distance Function (TSDF) \cite{TSDF}.
-Large-scale scene handling is facilitated by the hashed volumetric representation of \cite{hashed_voxel} with a 6mm$^{3}$ voxel.
-Each voxel contains: TSDF distance, colour information in the LAB space, a weight for averaging distance values, a user annotation class, a predicted class, an outcome of the mean-field inference from before and after classification and the probability of each class as inferred by the mean-field algorithm, for each class.
-The total number of bits per voxel is $96 + \textrm{[number of classes]} \cdot 16$.
+  \begin{figure}[!ht]
+    \center
+    \includegraphics[width=0.75\textwidth]{figures/pipeline}
+    \caption{The SemanticPaint Pipeline.}
+    \label{fig:pipeline}
+  \end{figure}
+
+    The pipeline (cf.
+Fig.
+\ref{fig:pipeline})  starts with capturing the environment as a stream of noisy RGBD images and combining them into a  3D model updated in an online fashion, similarly to KinectFusion \cite{fusion}.
+The user can choose which objects to label and does so by ``touching'' a small part of an object or encircling it with his hand and uttering the label.
+The label and the affected data points are used to update a Conditional Random Field (CRF) segmentation model.
+They are also passed to a Streaming Random Forest (SRF), an online version of Random Forest classifier, which learns to predict labels of previously unseen elements of the scene.
+Finally, when the user activates the `test` mode, the SRFs predictions are used to update the CRF.
+One of the biggest strengths of SemanticPaint is the efficiency of each part of the pipeline, which translates to real time performance.
+Algorithms used in the pipeline were adapted to work on volumetric data in the TSDF \cite{TSDF} format directly.
 
@@
 
-  This work relies heavily on classification and since even the most powerful classifier is useless if inappropriate features are supplied this paper contributes Voxel Oriented Patch (VOP) features.
-They are efficiently computed from the TSDF volume representation (cf.
-figure \ref{fig:vop}).
-Let $V_i$ be a VOP and $\mathbf{n}^i$ the normal of voxel i.
-An image patch of dimensions $r \times r$  is centered on voxel $i$ from the plane $(\mathbf{p} - \mathbf{p}_i) \cdot \mathbf(n)_i = 0$.
+  Voxel Oriented Patch (VOP, cf.
+Fig.
+\ref{fig:vop}) features are used to guide classification efforts.
+They are efficiently computed from the TSDF volume representation.
+Let $V_i$ be a VOP and $\mathbf{n}_i$ the normal of voxel i.
+An image patch of dimensions $r \times r$  is centered on voxel $i$ and taken from the plane $(\mathbf{p} - \mathbf{p}_i) \cdot \mathbf{n}_i = 0$.
 The patch contains colour values stored in the TSDF on the plane in CIELab to mitigate illumination effects.
@@
 Rotation invariance is achieved exactly as in \cite{sift}.
-The patch is rotated according to the dominant gradient direction.
-It is computed by creating a histogram of gradient directions with 36 bins, where the directions are weighted by gradient magnitudes and a gaussian window centered at the voxel.
-Furthermore, a quadratic polynomial is fit to the three biggest values of the histogram, whose maximum is the final direction.
 
@@
 
-  Online Random Forest \cite{online_forest} allow runtime model updates and they differ in that for a new node $n$ they generate a distribution of split function $\Theta_n$ and maintain a set of label statistics for each of them, updated when a new sample comes to the node.
-Label statistics are sufficient to compute information gain and thus the node $n$ is further split if it has seen a predefined number of samples or if the minimal value of the information gain is low enough.
-Storing and updating statistics in leaf nodes is computationally expensive which limits runtime performance
-
 \begin{figure}[!ht]
@@
 
-  Streaming Random Forest begins by creating a reservoir $R_n$ of at most $K$ samples stored in a list $T_n$ at the root node.
-It represents an unbiased sample of all training data seen so far.
+  Streaming Random Forest is an inherently online variant of a Random Forest classifier.
+It begins by creating a reservoir $R_n$ of at most $K$ samples stored in a list $T_n$ at the root node, which represents an unbiased sample of all training data seen so far.
 Initially, the first $K$ samples are stored.
@@
 Splitting the node requires computing the objective function for each $\theta \in \Theta_n$ over $R_n$.
-$H(R_n)$ is computed from the normalized class label histogram $P(l|R_n)$.
+$H(R_n)$ is computed from the normalized class histogram $P(l|R_n)$.
 Let $|R_n|$ be the amount of samples stored in the reservoir $n$ and let $R_n^d$ be the splitting induced by $\theta$.
 Finally, the split function is chosen as $f_n = \arg \max_{\theta \in \Theta_n} G(S, S^L, S^R | \theta)$  and set the number of samples seen by the child node to $|m_n^d| = |R_n^d| \max {\left(1, \frac{m_n}{K}\right)}$.
-Since the total number of samples stored is bounded by $K$ this approach uses less memory.
+Since the total number of samples stored is bounded by $K$ this approach uses only limited memory.
 Constructing child reservoirs from parent reservoirs (cf.
-fig \ref{fig:forest}) lessens the computational load.
-
-  The GPU implementation can handle up to 17 million voxel classifications per second, while between 3 and 10 million voxels are visible at any one time \cite{hashed_voxel}.
-To improve performance, voxels are batched and only a single batch is evaluated per frame.
+Fig \ref{fig:forest}) lessens computational load.
 
@@
 
-  Neighbourhood of 6 cm is used. Negative log likelihood of eq. \ref{eq:posterior} defines the energy
+  By taking negative log likelihood of eq. \ref{eq:posterior} we arrive at the energy
 
@@
 
-  Unary and pairwise potentials $\phi_i(x_i)$ and $\phi_{ij}(x_i, x_j)$ encode the cost of assigning a label to $v_i$ and a cost of assuming different labels by voxels in a neighbourhood, respectively.
-These costs are constantly changing due to the changing predictions of the random forest and user interaction.
+  Unary potential $\phi_i(x_i)$ encodes the cost of assigning a label to $v_i$, while $\phi_{ij}(x_i, x_j)$ is a prior cost of assuming different labels.
 
-  Initially, all voxels are encouraged to take a background label by setting $\phi_i(l) = 0$ for background class and grater than zero otherwise.
-When the user touches an object, a set of touched voxels $\mathcal{H}_S$ is registered and their unary potentials set to $\phi_i(l) = \infty$ if their labels are different then the specified one, and zero if they are the same, thus imposing the labeling.
+  Initially, all voxels are encouraged to take the background label by penalizing all other ones.
+When the user touches an object, a set of touched voxels $\mathcal{H}_S$ is registered and their unary potentials set to $\phi_i(l) = \infty$ if their labels are different then the specified one, thus imposing the labeling.
 If user labels a region more than once, the old labels are simply overwritten.
-
-  To process the encircling action, the labeling is projected into the current frame.
-Next, a GMM is fit with the foreground class taken as a convex hull of the encircled region and background as the rest of the image.
+To process an encircling action, a GMM is fit with the foreground class taken as a convex hull of the encircled region and the background class as the rest of the image.
 For every pixel in a bounding box around the user annotation the unary potentials are updated as
@@
 
-  with $P_E($fg$|\mathbf{a}_i)$ the probability of assuming the foreground label, which stems from the normalized likelihood $P(\mathbf{a}_i|$fg$)$ from the GMM.
-
-  \begin{equation}
-    P_E(\text{fg}|\mathbf{a}_i) = \frac{P(\mathbf{a}_i|\text{fg}) }{P(\mathbf{a}_i|\text{fg}) + P(\mathbf{a}_i|\text{bg})}
-  \end{equation}
-
+  with $P_E($fg$|\mathbf{a}_i)$ the probability of assuming the foreground label.
 For all voxels that haven't been explicitly labeled by the user unitary potentials are updated with the predicted values as $\phi_i(l) = -\log P_F(x_i = l | \mathbf{D})$.
@@
 Two thirds of each viewpoint of each object is used for training.
-Results are summarized by figure \ref{fig:srf_results}.
+Results are summarized by Fig.
+\ref{fig:srf_results}.
 SRF clearly outperforms its competitors.
 
-\begin{figure}[!ht]
- \center
- \includegraphics[width=0.75\textwidth]{figures/srf_results}
- \caption{Comparison of Streaming Random Forest against Online Random Forest and Hoeffding Trees.}
+  \begin{figure}
+    \centering
+    \begin{subfigure}{.5\textwidth}
+      \centering
+      \includegraphics[width=\linewidth]{figures/srf_ap}
+      \label{fig:srf_ap}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+      \centering
+      \includegraphics[width=\linewidth]{figures/srf_iu}
+      \label{fig:srf_iu}
+    \end{subfigure}
+    \caption{Comparison of Streaming Random Forest against Online Random Forest and Hoeffding Trees}
     \label{fig:srf_results}
@@
 It can be used to gather groundtruth for large-scale visual recognition systems, robot navigation and to aid partially sighted people.
-There are some limitations, however.
-Use of the system is now restricted to simple indoor scenes.
-Computational load increases with the size of the scene while prediction accuracy drops with the number of object categories.
 
@@
 
-  The system has also some limitations.
+  The system has some limitations, however.
 Combing colour and depth data requires careful sensor calibration.
@@
 Additionally, strong changes of lighting conditions impair classification ability at certain viewpoints.
-These failures are shown in figure \ref{fig:failures}.
+These failures are shown in Fig.
+\ref{fig:failures}.
 Moreover, computational performance drops with the increasing size of the scene and classification accuracy decreases with the number of classes.
@@
 
-  There is big room for improvement.
+  There is some room for improvement.
 Introducing priors associated with the type of environment or with an object class (\emph{e.g.} vertical walls) could greatly improve classification accuracy.
@@
 
-  \bibitem{old}
-  ROBERTS , L. G. 1963. Machine perception of three-dimensional solids. Ph.D. thesis, Massachusetts Institute of Technology.
+  \bibitem{VCRF}
+  Kim, B.-S., Kohli, P., and Savarese, S. 2013. 3d Scene Understanding By Voxel-Crf. In Proc. ICCV.
 
-  \bibitem{VCRF}
-  KIM , B.-S., KOHLI , P., AND SAVARESE , S. 2013. 3D scene understanding by voxel-CRF. In Proc. ICCV.
-
-  \bibitem{Monofusion}
-  PRADEEP , V., RHEMANN , C., IZADI , S., ZACH , C., BLEYER , M., AND BATHICHE , S. 2013. Monofusion: Real-time 3D reconstruction of small scenes with a single web camera. In Proc. ISMAR.
+  \bibitem{monofusion}
+  Pradeep, V., Rhemann, C., Izadi, S., Zach, C., Bleyer, M., and Bathiche, S. 2013. Monofusion: Real-Time 3d Reconstruction Of Small Scenes With A Single Web Camera. In Proc. ISMAR.
 
   \bibitem{ChangeDetection}
-  HERBST , E., HENRY , P., AND FOX , D. 2014. Toward online 3-d object segmentation and mapping. In IEEE International Conference on Robotics and Automation (ICRA).
+  Herbst, E., Henry, P., and Fox, D. 2014. Toward Online 3-D Object Segmentation and Mapping. In IEEE International Conference On Robotics and Automation (ICRA).
 
   \bibitem{mesh}
-  VALENTIN , J. P., SENGUPTA , S., WARRELL , J., SHAHROKNI , A., AND TORR , P. H. 2013. Mesh based semantic modelling for indoor and outdoor scenes. In Proc. CVPR.
-
-  \bibitem{heritage}
-  LEVOY , M., PULLI , K., CURLESS , B., RUSINKIEWICZ , S., KOLLER , D., EREIRA , L., GINZTON , M., ANDERSON , S., DAVIS , J., G NSBERG , J., ET AL .
-2000.
-The digital Michelangelo project: 3D scanning of large statues.
-In Proc.
-SIGGRAPH.
-ACM.
+  Valentin, J. P., Sengupta, S., Warrell, J., Shahrokni, A., and Torr, P. H. 2013. Mesh Based Semantic Modelling For Indoor and Outdoor Scenes. In Proc. CVPR.
 
   \bibitem{fusion}
-  NEWCOMBE , R.
-A., IZADI , S., HILLIGES , O., MOLYNEAUX , D., KIM , D., DAVISON , A.
-J., KOHLI , P., SHOTTON , J., HODGES , S., AND FITZGIB-BON , A.
+  Newcombe, R.
+A., Izadi, S., Hilliges, O., Molyneaux, D., Kim, D., Davison, A.
+J., Kohli, P., Shotton, J., Hodges, S., and Fitzgib-Bon, A.
 2011.
-KinectFusion: Real-time dense surface mapping and tracking.
+Kinectfusion: Real-Time Dense Surface Mapping and Tracking.
 In Proc.
@@
   \bibitem{TSDF}
-  CURLESS , B.
-AND LEVOY , M.
+  Curless, B.
+and Levoy, M.
 1996.
-A volumetric method for building complex models from range images.
-In Proceedings of the 23rd annual conference on Computer graphics and interactive techniques.
+A Volumetric Method For Building Complex Models From Range Images.
+In Proceedings Of The 23rd Annual Conference On Computer Graphics and Interactive Techniques.
 ACM, 303–312.
 
-  \bibitem{hashed_voxel}
-  NIESSNER , M., ZOLLHOFER , M., IZADI , S., AND STAMMINGER , M. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
-
   \bibitem{online_forest}
-  SAFFARI , A., LEISTNER , C., SANTNER , J., GODEC , M., AND BISCHOF , H. 2009. On-line random forests. In IEEE ICCV Workshop.
+  Saffari, A., Leistner, C., Santner, J., Godec, M., and Bischof, H. 2009. On-Line Random Forests. In IEEE ICCV Workshop.
 
   \bibitem{reservoir}
-  VITTER , J. S. 1985. Random sampling with a reservoir. ACM TOMS 11, 1.
+  Vitter, J. S. 1985. Random Sampling With A Reservoir. ACM Toms 11, 1.
 
   \bibitem{sift}
-  LOWE , D. G. 1999. Object recognition from local scale-invariant features. In Proc. ICCV.
+  Lowe, D. G. 1999. Object Recognition From Local Scale-Invariant Features. In Proc. ICCV.
 
   \bibitem{crf}
-  LAFFERTY , J., McCALLUM , A., AND PEREIRA , F. C. 2001. Conditional random fields: Probabilistic models for segmenting and labeling sequence data.
+  Lafferty, J., Mccallum, A., and Pereira, F. C. 2001. Conditional Random Fields: Probabilistic Models For Segmenting and Labeling Sequence Data.
 
   \bibitem{inference}
-  KRAHENBÜHL, P. AND KOLTUN , V. 2011. Efficient inference in fully connected CRFs with Gaussian edge potentials. In NIPS.
-
-  \bibitem{kl}
-  KOLLER , D. AND FRIEDMAN , N. 2009. Probabilistic Graphical Models: Principles and Techniques. MIT Press
+  Krahenbühl, P. and Koltun, V. 2011. Efficient Inference In Fully Connected Crfs With Gaussian Edge Potentials. In NIPS.
 
   \bibitem{htree}
-  DOMINGOS, P. AND HULTEN , G. 2000. Mining high-speed data streams. In Proc. SIGKDD.
+  Domingos, P. and Hulten, G. 2000. Mining High-Speed Data Streams. In Proc. SIGKDD.
 
   \bibitem{dataset}
-  LAI , K., BO , L., REN , X., , AND FOX , D. 2011. A large-scale hierarchical multi-view rgb-d object dataset. In Proc. ICRA.
+  Lai, K., Bo, L., Ren, X.,, and Fox, D. 2011. A Large-Scale Hierarchical Multi-View Rgb-D Object Dataset. In Proc. ICRA.
 
   \bibitem{SemanticPaint}
-  VALENTIN, J., VINEET, V., CHENG, M.-M., KIM, D., SHOTTON, J., KOHLI, P., NIESSNER, M., CRIMINISI, A., IZADI, S., TORR, P.
+  Valentin, J., Vineet, V., Cheng, M.-M., Kim, D., Shotton, J., Kohli, P., Niessner, M., Criminisi, A., Izadi, S., Torr, P.
 2015.
-SemanticPaint: Interactive 3D Labeling and Learning at your Fingertips.
+SemanticPaint: Interactive 3d Labeling and Learning At Your Fingertips.
 SIGGRAPH.

### 2015-12-13 ML/hw9/hw.tex
hw9

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{epstopdf}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}

\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework } \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

Let us assume without any loss of generality that $y > x$. Since $t \in ]0, 1[$ we see that $tx + (1-t)y = z \in ]x, y[$. Therefore, the definition of a strictly convex function does not consider the function's graph at points $x$ and $y$. A graph of strictly convex function must lie strictly below a straight line from $f(x)$ to $f(y)$ and $f$ cannot be linear \emph{e.g.} $f(x) = ax + b$ . It is different from a convex function, whose graph can be a straight line.

\section*{Problem 2}

If a function is convex, it can be a straight line $f(x) = ax + b$. Since $a$ can be equal to $0$, the function $f(x) = b$ is a convex function for which $\forall_{x \in \mathcal{R}} f(x) = b$. Therefore an infinite number of points $x^*$ satisfies $x^* = \arg \min_x f(x)$.

\section*{Problem 3}

If a function $f$ is strictly convex, the graph of $f$ must lie below a straight line between points $(x, f(x))$ and $(y, f(y))$ for any $x \neq y$. We know that $\nabla f(x^*) = 0$. Let's take a point $x_0$ such that $y = c$ for some constant $c$ is a horizontal line that crosses $f(x)$ at the point $x_0$. We know that at least a part of the graph of $f$ must lie below the line $y = c$. We can move this line to a point just above $f(x^*)$ such that $f(x^*)$ is the only point below the line. Therefore, there is only a single point that satifies $f(x^*) = min_x f(x)$ for a strictly convex function f.

\section*{Problem 4}

\begin{figure}[!ht]
  \includegraphics[width=\linewidth]{convex}
  \caption{An example of a strictly convex function $y = x^2$ and a non-strictly convex function $y = x$.}
  \label{fig:convex}
\end{figure}

As in Fig. \ref{fig:convex}.

\section*{Problem 5}

We cannot apply this feature to data directly. It is infinite dimensional and would require infinite memory to store and time to compute.

\section*{Problem 6}

\begin{equation}
\begin{align}
 K(x, y) &= \sum_{i=1}^\infty \phi_{\infty, i}(x)\phi_{\infty, i}(y) \\
	 &= \sum_{i=1}^\infty \frac{1}{i!} e^\frac{-(x^2 + y^2)}{2 \sigma^2} \left( \frac{xy}{\sigma^2} \right)^i \\
	 &= e^\frac{-(x^2 + y^2)}{2 \sigma^2} \sum_{i=1}^\infty \frac{1}{i!} \left( \frac{xy}{\sigma^2} \right)^i \\
	 &= e^\frac{-(x^2 + y^2)}{2 \sigma^2} \sum_{i=1}^\infty \frac{1}{i!} \left( z \right)^i
 \end{align}
\end{equation}

where $z = \frac{xy}{\sigma^2}$. From the Taylor expansion of $e^z$ around $0$ we have

\begin{equation}
  e^z = 1 + z + \frac{1}{2!} z^2 + \ldots = 1 + \sum_{i=1}^\infty \frac{1}{i!} z^i
\end{equation}

and therefore $\sum_{i=1}^\infty \frac{1}{i!} z^i = e^z - 1$. Finally

\begin{equation}
\begin{align}
 K(x, y) &= e^\frac{-(x^2 + y^2)}{2 \sigma^2} \left( e^\frac{xy}{\sigma^2} - 1 \right) \\
         &= e^\frac{-(x - y)^2}{2 \sigma^2} - e^\frac{-(x^2 + y^2)}{2 \sigma^2}
\end{align}
\end{equation}

which is easy to compute. We should be concerned with overfitting, since an infinite dimensional feature space translates to very high complexity of the model. It can fit training data perfectly, but generalize badly. It is at least partially mitigated by SVM, because it looks for a large margin decision boundary, which improves generalization.

\section*{Problem 7}

For every $x, y$ it holds that $x^2 + y^2 \geq (x - y)^2$ which implies that $e^\frac{-(x - y)^2}{2 \sigma^2} \geq e^\frac{-(x^2 + y^2)}{2 \sigma^2}$ and $K(x, y) \geq 0$. A linear decision boundary is given by $f(x) = w^T \phi_\infty(x)$ where the parameters $w$ are given by a linear combination of training examples $w = \sum_{i=1}^N \alpha_i y^{(i)} \phi_\infty(x^{(i)})$, where $y^{(i)} \in \{0, 1\}$ is label of the $i^{th}$ data sample. We have

\begin{equation}
 \begin{align}
  f(x^{(j)}) &= \sum_{i=1}^N \alpha_i y^{(i)} \phi_\infty(x^{(i)})^T \phi_\infty(x^{(j)}) \\
       &= \sum_{i=1}^N  \alpha_i y^{(i)} K(x^{(i)}, x^{(j)}) \\
       &= \sum_{i=1}^N  \alpha_i y^{(i)} \left( e^\frac{-\left(x^{(i)} - x^{(j)}\right)^2}{2 \sigma^2} - e^\frac{-\left(\left(x^{(i)}\right)^2 + \left(x^{(j)}\right)^2\right)}{2 \sigma^2} \right)
 \end{align}
\end{equation}

Let $\alpha_i = 1$ for every $i$ and set $\sigma = 10^{-6} \min_{i, j} \left(x^{(i)} - x^{(j)}\right)^2$. Now, the pair of closest points in the feature space is at least a million units of $\sigma$ apart from each other and we can write that

\begin{equation}
 \begin{align}
    f(x^{(j)}) &= y^{(j)} + \epsilon
 \end{align}
\end{equation}

It means that each training sample exclusively defines labels of all points in its neighbourhood.

\end{document}

### 2015-12-13 ML/hw9/ml8.m
ml hw9

@@ -1,59 +1,0 @@
-
-close all, clear all
-
-f0 = @(x) x.^2 + 1;
-f1 = @(x) (x-2) .* (x-4);
-
-
-
-xx = linspace(0, 5, 100);
-
-%% problem 5
-xx0 = linspace(2, 4, 100);
-
-figure(1)
-hold on
-plot(xx, f0(xx), 'r-', 'LineWidth', 2);
-plot(xx, f1(xx), 'b-', 'LineWidth', 2);
-plot(xx0, f1(xx0), 'g-', 'LineWidth', 2);
-plot([2 2], [-10 70], 'g-');
-plot([4 4], [-10 70], 'g-');
-title('function and its constraint');
-xlabel('x');
-ylabel('f(x)');
-legend('f0', 'f1', 'feasible region');
- 
-
-%% problem 6
-
-as = [0 0.5 1 3 4 5 8];
-colors = 'rgbkcmy';
-L = @(x, a) f0(x) + a .* f1(x);
-
-figure(2)
-hold on
-
-
-for i = 1:numel(as)
-    c = colors(i);
-    a = as(i);
-    plot(xx, L(xx, a), c, 'LineWidth', 2);
-end
-title('Lagrangian for different alphas');
-xlabel('x')
-ylabel('L(x, a)');
-legend('a = 0', 'a = 0.5', 'a = 1', 'a = 3', 'a = 4', 'a = 5', 'a = 8');
-
-
-%% problem 7
-g = @(a) (-a.^2 + 9 * a +1) ./ (1 + a);
-aa = linspace(0, 8, 100);
-
-figure(3)
-plot(aa, g(aa), 'b-', 'LineWidth', 2);
-title('Lagrange dual function');
-xlabel('a');
-ylabel('g(a)');
-
-
-

### 2016-01-09 phd/sop/posner.tex
research proposal for MRG

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
\pagestyle{empty}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

\begin{center}
{\large PhD Research Proposal}\\[.1in]
{\large Unsupervised Tuning of Neural Networks for Object Classification in Sequential Data with Dynamically Chosen Object Classes}\\[.1in]
{\large Adam Kosiorek}
\end{center}

\section{Introduction}

Suppose that an autonomous agent moves through a large highly differentiated space. It can enter private and public buildings, parks, factories, crowded or desolate places. In order to interact with its environment, it has to recognize surrounding objects. High accuracy classification of variable environments with hundreds of object classes requires computationally expensive approaches, huge labeled datasets and is not good enough for many practical applications yet \cite{imagenet}.

One approach of solving this problem relies on the fact that filters used in Convolutional Neural Networks (CNNs) respond strongly to visually similar inputs. It suggests using several simpler CNN-based classifiers specialized for subsets of possible object classes. Some of these classifiers, however, could have huge overlaps in types of objects they are trained to classify. Therefore, I suggest another approach. I believe that this problem can be solved by a (possibly recurrent) CNN classifier, whose filters are tuned online in an unsupervised way and whose final fully-connected layers are constantly calibrated on a small labeled training set. Additionally, if the set of classes this classifier predicts could be altered at run-time, it might be possible to achieve better accuracy at a lower computational cost. Finally, it results in a classifier that is constantly adapting to its changing environment.

\section{Project Outline}

There are three essential phases in this project: (1) Developing a recurrent CNN (RCNN) for object classification in sequential data, (2) unsupervised tuning of the network with simultaneous supervised calibration and (3) dynamically changing prediction classes. I will discuss each of the phases in detail.

\subsection{RCNN for Object Classification in Sequential Data}

CNNs are well suited to object classification, while RNNs with \emph{e.g.} LSTMs can model short and long term dependencies in the data \cite{rcnn}. The first stage of the project is to develop an RCNN for object classification in RGB videos, possibly augmented with depth data or working on pure geometrical data \emph{e.g.} point clouds. Training problems might arise due to high correlation between subsequent video frames. They can be mitigated by holding a reservoir of seen samples \cite{reservoir} or by an experience replay mechanism \cite{dqn}. The resulting system should resemble \cite{rcnn} and it should take around 3 months to complete.

\subsection{Unsupervised Tuning with Supervised Calibration}

Unsupervised learning is superior to supervised learning in that it needs no labels and can act on raw data, which is typically available in unlimited quantities. It has been shown that unsupervised training of neural networks provides good initialization for supervised finetuning \cite{autoencoders} as well as delivers good features that can be used for classification by feeding them to a classifier directly \cite{adversarial}. The next step is to use a combination of supervised and unsupervised approach for lifelong learning and adaptation to the changing environment.

Firstly, the CNN part of the system (and possibly the recurrent part as well) is tuned in an unsupervised way as an autoencoder or in an adversarial setting \cite{adversarial}. This should adjust the convolutional filters so that they are highly responsive to recently seen object classes. Secondly, due to changing response pattern of the CNN, the last fully-connected layers of the network, which act as a classifier, can decalibrate. This could be addressed by simultaneous fine-tuning of the whole network on a (small) training set.

In order for this approach to work, appropriate RCNN architectures need to be found and some adjustments to learning algorithms might be needed. It would be beneficial to merge unsupervised and supervised tuning into a unified end-to-end approach. For now, it is not clear how those two tasks can be done at the same time and yet separately. Some challenges with computational complexity of this approach might arise. Since it operates on a real working system (mobile robot), the adjustment mechanism should work in or close to real-time. I expect this phase to take 12 to 15 months.

\subsection{Dynamically Changing Prediction Classes}

As an agent moves through the environment, a type of a scene it is in might change and with it categories of objects that are likely to be present there. A classifier of fixed architecture might be very complex and it might waste processing power if it is to care about every possible object class regardless of the scene type. The purpose of this stage of the project would be to find out how the classes a classifier can predict could be changed at runtime.

One possible solution is the following. Let A the main classifier be fitted to the most prevalent classes in the training set and let B another classifier that is able to recognize whole groups of object classes. At any given point in time A can recognize only a subset of all classes, while B maintains a ranking of groups of classes according to their probability given the data seen so far. When the ranking changes significantly, \emph{i.e.} a group becomes very probable that hasn't been probable before, the final fully connected layers of A could be refitted to predict a union of the most probable groups of classes as given by B. It can be seen as an attention mechanism that pays higher attention to most probable classes given the characteristic of the environment, which slightly resembles \cite{memory}.

Challenges of this phase might arise from computational complexity, since refitting might require considerable resources and from finding a well-performing algorithm for scoring groups of classes or detecting changes of a scene type. The final goal would be to produce an end-to-end differentiable architecture, that can learn the classes it should predict from the data by gradient-based methods. I anticipate it to take around 12 months, with the remaining research time dedicated to unifying and improving the whole architecture, experiments and casting parts of the approach as a Reinforcement Learning problem.

\bibliographystyle{ieeetr}
\bibliography{bibliography}
\end{document}

### 2016-01-09 phd/sop/toronto.tex
research proposal for MRG

--- 2015-12-02 phd/sop/stanford.tex
@@
 I immediately knew that this intersection of image processing and machine learning was the area I wanted to focus on.
-It became apparent that pursuing a PhD was one of the few ways that could get me closer to solving the AI challange.
+It became apparent that pursuing a PhD was one of the few ways that could get me closer to solving the AI challenge.
 
@@
 
-I believe that the CS PhD programme at Stanford is perfect for me due to the following reasons.
+I believe that the CS PhD programme at the University of Toronto is perfect for me due to the following reasons.
+Firstly, the faculty has an extensive research experience in neural networks for computer vision, natural language and multi-modal settings.
+
 Firstly, the courses taught by world renown scientists e.g.
 Prof.
-Fei-Fei's ``High-Level Vision Behaviors, Neurons and Computational Models'' provide an extraordinary introduction to research in deep learning I would like to conduct.
-Secondly, research of the Stanford Visian Lab is multimodal in that it successfuly merges natural language processing with computer vision \cite{imageannot}.
+Raquel Urtasun and Prof.
+Richard Zemel ``Probabilistic Learning and Reasoning'' provide an extraordinary introduction to research in deep learning I would like to conduct.
+Secondly, research of the Machine Learning group at the U of T is multimodal in that it successfuly merges natural language processing with computer vision \cite{show}.
 I find it very relevant, for it might lead to the improvement of living standard for people with impaired sight.
-Finally, Stanford is \emph{the} most exciting area for any computer scientist to live in.
-The diversity and the amount of talent there bolster personal development like no other place in the world.
 

### 2016-01-10 phd/sop/mrg.tex
mrg proposal draft

--- 2015-11-21 phd/sop/sop.tex
@@
 \begin{center}
-{\Large Statement of Purpose} \\[.3in]
+{\Large Research Proposal} \\[.1in]
 {\large Adam Kosiorek}
@@
 
-\vspace*{.5in}
-
-%The Statement of Purpose should describe succinctly your reasons for applying to the proposed %program at Stanford, your preparation for this field of study, research interests, future career %plans, and other aspects of your background and interests which may aid the admissions committee %in evaluating your aptitude and motivation for graduate study.
-The Statement of Purpose should not %exceed two pages.
+\vspace*{.1in}
 
@@
 To explore it further, I enrolled in BSc in Robotics at the best technical university in my country.
-I studied maths, physics and mechanics, material science and manufacturing technologies; programming and software-oriented courses were my favourite, however.
-Paired with my passion for vision and perception, which I have developed as a photojournalist, it steered me towards my first computer vision course.
-I immediately knew that this intersection of image processing and machine learning was the area I wanted to focus on.
-It became apparent that pursuing a PhD was one of the few ways that could get me closer to solving the AI challange.
+Programming and software-oriented courses were my favourite and paired with my passion for vision and perception, which I have developed as a photojournalist, they steered me towards my first computer vision course.
+At that moment it became apparent that pursuing a PhD was one of the few ways that could get me closer to solving the AI challange.
 
-To learn from the best, I joined Computer Vision Lab at Samsung R\&D in Warsaw for an internship.
-My project was to design and implement a Bag of Words-based pipeline for object classification.
-Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of automatically learned low level features in convolutional neural networks.
-I focused my research primarily on the visual codebook generation within Bag of Words.
-The final step was to optimize it and then port it to the Android platform.
+To learn from the best, I joined Computer Vision Lab at Samsung R\&D, where I implemented a Bag of Words-based pipeline for object classification.
+Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of learned low level features in convolutional neural networks.
+Before jumping into deep learning for object classification and image duplicate detection, I ported the project to Android and researched codebook generation within BoW.
 In my bachelor thesis I investigated how spatial information affects classification performance by implementing a similar system for Kinect-gathered point cloud classification.
-Afterwards, I continued my work at Samsung, where I turned to deep learning techniques for object classification and image duplicate detection.
 
-My observation was that deep learning, although often superior to older approaches, required huge computational resources and datasets.
+My observation was that deep learning, although often superior to other approaches, requires huge computational resources and datasets.
 To mitigate this problem I went to study computational science at the Technical University of Munich.
-With cutting-edge computer vision research groups it seemed a great choice.
 Numerical linear algebra and scientific computing, comprising the majority of the coursework, help enormously with optimization and implementation of algorithms.
-Currently, I am working with Caner Hazirbas and Rudolph Triebel of the Computer Vision group headed by Prof.
-Cremers.
-In our recent project we investigate the introspective capacity of neural networks.
+Currently, I am working with Rudolph Triebel of the Computer Vision group on the introspective capacity of neural networks.
 It is understood as the ability of a classifier to assess uncertainty of its prediction given the data \cite{introspective}.
 It is a paramount problem in mobile robotics and medicine where wrong classification might lead to loss of life.
-While high classification accuracy is desired, it is impossible to assess whether a given prediction is accurate in a test setting, where no labels are available.
-It is, therefore, vital to assess the uncertainty of predictions.
 Our intermediate results show that neural networks, augmented with additional layers and a novel cost function, can be jointly trained for classification and uncertainty estimation.
@@
 
-This summer I did an internship at Bloomberg in London, where I worked on fraud detection in financial transactions.
+Last summer I did an internship at Bloomberg in London, where I worked on fraud detection in financial transactions.
 The problem can be cast as unsupervised anomaly detection with further verification in a supervised setting.
 I learned how difficult it can be to introduce an innovative approach in a corporate environment.
-This, together with my earlier industrial experience, convinced me that I do want to pursue a PhD.
-I love solving scientific problems, which do not have ``the best'' solution or a reference specification, by going into the deepest details.
 
 In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks.
-The former fascinates me, for it is similar to how humans learn.
-Specifying a reward function  or designating an expert in the inverse problem is far easier than providing supervised training data.
-Recurrent neural networks, on the other hand, are well suited to sequential information processing.
-If beliefs about content could be introduced in a form of priors inferred from previous elements of the sequence, it might be possible to increase object classification accuracy in videos and to shrink the network size.
-Another interesting problem is an end-to-end RNN for optical flow computation.
-The only end-to-end neural network, while efficient and accurate, does not use temporal information.
-It computes optical flow for any two possibly unrelated images \cite{flownet}.
-I would like to investigate RNN designs capable of computing optical flow while presented with a single image at a time.
-Solving both problems might thus lead to smaller networks and reduced amounts of computation, possibly increasing energy efficiency of such systems.
-
-I believe that AIMS is a perfect programme for me due to the following reasons.
-Firstly, it is multidisciplinary with robotics as its main motivation.
-I feel it makes it more practical, since it is about solving real world problems.
-Secondly, while I have a number of research ideas, I am not confident planning a 3 year research project just yet.
-The first year of courses and interaction with the faculty is a great means of solving this issue.
-Finally, the Mobile Robotics Group research is extremely close to my interests, e.g.
-introspective capacity of classifiers.
 
@@
 \end{document}
-
-% 1. What's Needed.
-% 
-% General area of research
-% Relevant skills and experience
-% Your reasons for applying to Oxford
-% Why you consider yourself a strong applicant.
-% 
-% 2. Introspection.
-% 
-% Intelectual Development.
-% 
-% a. Matrix and IRobot hooked me up on AI and Robotics
-% b. Found B.Sc. in Robotics at WUT and went for it
-% c. Disapointed due to mechanical engineering oriented programme
-% d. Course in CV and mobile robotics -> that’s where AI is, should do CS
-% e. Internship in Samsung in CV: ML is really cool, need more CS skills
-% f. Thesis in Object Recognition
-% g. 1 semester of CS undegrad courses
-% h. CV/ML needs HPC for large scale -> CSE with focus on CV/ML
-% i. Want to do CV/ML for robotics as PhD -> RL
-% 
-% Role Models: Andrew Ng and Elon Musk.
-% 
-% Both extremely creative, a scientist and an applied engineer, both with huge vision.
-One is changing the world of AI/ML and education, the other is chaining the world by reaching for technologies far out of time.
-Bold, don’t fear to lose, have a dream and want to change the world.
-They don’t care about problems being too hard or too big.
-They just make stuff happen.
-% 
-% 3. Outline.
-% 
-% 1. Genesis
-% 1. I always liked physics and maths
-% 2. As a photographer I was interested in vision and perception
-% 3. I was fascinated by robotics and AI, although I had no idea what that even was
-% 
-% 2. What Happened
-% 1. Went to study Robotics @ WUT
-% 2.
-Turned out that most of Robotics is mechanical and electrical engineering with a nudge of control theory.
-I got more excited about general software development and got into it UNTIL my first class in computer vision where I learned about machine learning.
-% 3. Decided I want to do this and went to the Samsung R&D Lab to work objcect classification and was doing bachelor thesis on the same topic but in 3D with kinect
-% 4. Turnened out that TUM is quite strong in the area and I decided to go for a Masters hers, DAAD helped a lot
-% 5.
-I'm working with the Prof.
-Cremers' group on deep learning for computer vision.
-I'm really into it and I want to get a PhD.
-I love solving difficult research problems that have no right or wrong solution.
-I hate being tied up by the specs and working on that tiny little widget to make the UI better.
-% 
-% 3. Outlook
-% 1. Introspective capacity is vital for robotics. I'm looking into it in the context of deep learning and I'd like to make it work, but it's not sth I want to do a PhD on; Masters might be good
-% 2.
-Reinforcement Learning fascinates me in that it is, together with the inverse problem, more or less how we learn.
-I'd like to study it further for I believe it is THE way of bringing ML to the masses.
-It's way easier to specify a reward  function or to find an expert than to gather supervised training datasets.
-I'm interested into human learning processes as well and I wonder if we can port them to the neural network settings
-% 3.
-Recurrent neural networks seem suitable to a plethora of various tasks.
-E.g.
-object classification, and even more object detection, seem to be doing a lot of redundant work when dealing with e.g.
-video sequences.
-I'd like to work on itroducing knowledge of what might be in the sequence as priors based on what we've already seen there.
-Another thing is to use RNNs to compute optical flow from a sequence, but using one image at a time.
-Also RNNs for optical flow from DVS would be cool.

### 2016-01-10 phd/sop/mrg.tex
research proposal for mrg

@@
 To learn from the best, I joined Computer Vision Lab at Samsung R\&D, where I implemented a Bag of Words-based pipeline for object classification.
-Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of learned low level features in convolutional neural networks.
+Consequently, I learned about the state-of-the-art in keypoint detection and feature description, which later helped me to appreciate the importance of learned low level features in Convolutional Neural Networks (CNNs).
 Before jumping into deep learning for object classification and image duplicate detection, I ported the project to Android and researched codebook generation within BoW.
@@
 
-In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks.
+In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks, preferably within the Mobile Robotics Group.
+Following is one of the research problems that attracts me (I am also open to other problems).
+Suppose that a mobile robot moves through highly differentiated space with multiple scene types.
+High accuracy classification of hundreds of object classes requires computationally expensive approaches and huge labeled datasets \cite{imagenet}.
+Since some objects can be more likely to appear in a particular type of a scene, it might be possible to develop a recurrent CNN that adapts to the changing environment.
+One way of accomplishing this is to have the filters of the CNN tuned online in an unsupervised way, in either autoencoder \cite{autoencoders} or adversarial \cite{adversarial} setting.
+It would make the CNN more responsive to the recently seen object classes.
+It might lead to the decalibration of the final fully-connected layers of the network, which might be mitigated by simultaneous fine tuning on a (small) training set.
+Additionally, if the set of classes predicted by the classifier could be altered at run-time, it might be possible to achieve higher accuracy at lower computational cost.
+This could be implemented by having A --- a recurrent CNN fitted to the most prevalent object classes in the training set and B --- a second classifier able to predict whole groups of classes.
+At any given point in time A would predict a subset of all classes, while B would maintain a probability-weighted ranking of groups of classes.
+When a significant change in the ranking occurs, A could be refitted to the union of the most probable groups of classes.
+
+I expect this project to take around 3 years.
+The first stage, lasting about 3 months, would be to develop a recurrent CNN for object classification in sequential data similar to \cite{rcnn}.
+The next stage would be focused on unsupervised tuning with supervised calibration of the recurrent CNN.
+The design of appropriate architectures, adjusting learning algorithms and merging unsupervised tuning with supervised calibration in a unified approach might take 12 to 15 months.
+Finally, implementing dynamically changing prediction classes requires solving a number of challenges.
+Refitting of the final fully-connected layers would have to be made fast enough to work in real-time, since the system should work on a mobile robot.
+An algorithm for scoring groups of classes and change of a scene type would have to be developed.
+The ultimate goal would be to cast it as a differentiable problem, so that object classes to predict could be learnt with gradient-based approach, similar to how memory is learnt in \cite{memory}.
+This phase might take around 15 months.
 

### 2016-01-10 ML/hw11/hw.tex
nn homework

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{epstopdf}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}

\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 11} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

Consider a neural network with an input layer, a hidden layer and an output layer. Assume a linear activation function in the output layer and let the output layer take form of a single neuron. The output of the network can be expressed as

\begin{equation}
 y(x) = V^T a(x) + c = V^T f(z) = V^T f(W^T x + b)
\end{equation}

where $V$ weight matrix from the hidden layer to the output, $a$ the activation of the hidden layer, $f$ the hidden layer's nonlinearity, which acts on its input element wise, $W$ the weight matrix from the input to the hidden layer, $b$ and $c$ are biases and $x$ the input vector.

Sigmoid is a scaled and translated version of tanh, since

\begin{equation}
\begin{align}
 2 \sigma(2x - 1) = \frac{2}{1+e^{-2x}} - 1 &= \frac{2e^{2x}}{e^{2x}+1} - 1 = \frac{e^{2x}-1}{e^{2x}+1} &= tanh(x)
 \end{align}
\end{equation}

therefore, if we take $W = 2 \widetilde{W}$ and $b = 2 \widetilde{b}$ it is obvious that

\begin{equation}
 2 \sigma(W^T x + b) - 1 = tanh(\widetilde{W}^T x + \widetilde{b}) 
\end{equation}

Weights $V$ and the bias $c$ of the final layer can easily account for scaling and translation of the resulting hidden's layer activation.

\section*{Problem 2}

\begin{equation}
 \begin{align}
  \frac{d}{dx} \sigma(x) = \frac{e^{-x}}{(1 + e^{-x})^2} = \frac{1 + e^{-x}}{(1 + e^{-x})^2} - \frac{1}{(1 + e^{-x})^2} = \sigma(x) - \sigma^2(x) = \sigma(x)(1 - \sigma(x))
 \end{align}
\end{equation}

\begin{equation}
 \begin{align}
  \frac{d}{dx} tanh(x) &= \frac{2 e^{2x} \left( (e^{2x} + 1) - (e^{2x} - 1) \right)}{(e^{2x} + 1)^2} \\
  &= \frac{\left( (e^{2x} + 1) + (e^{2x} - 1) \right) \left( (e^{2x} + 1) - (e^{2x} - 1) \right)}{(e^{2x} + 1)^2} \\
  &=  \frac{(e^{2x} + 1)^2 - (e^{2x} - 1)^2}{(e^{2x} + 1)^2} \\
  &= 1 - \frac{(e^{2x} - 1)^2}{(e^{2x} + 1)^2} = 1 - tanh^2(x)
 \end{align}
\end{equation}

\end{document}

### 2016-01-10 phd/sop/mrg.tex
after Martin's remarks

@@
 Programming and software-oriented courses were my favourite and paired with my passion for vision and perception, which I have developed as a photojournalist, they steered me towards my first computer vision course.
-At that moment it became apparent that pursuing a PhD was one of the few ways that could get me closer to solving the AI challange.
+At that moment it became apparent that pursuing a DPhil degree was one of the few ways that could get me closer to solving the AI challange.
 
@@
 In my bachelor thesis I investigated how spatial information affects classification performance by implementing a similar system for Kinect-gathered point cloud classification.
+Last summer I did an internship at Bloomberg in London, where I worked on fraud detection in financial transactions.
+The problem can be cast as unsupervised anomaly detection with further verification in a supervised setting.
 
-My observation was that deep learning, although often superior to other approaches, requires huge computational resources and datasets.
+Deep learning, although often superior to other approaches, requires huge computational resources and datasets.
 To mitigate this problem I went to study computational science at the Technical University of Munich.
@@
 
-Last summer I did an internship at Bloomberg in London, where I worked on fraud detection in financial transactions.
-The problem can be cast as unsupervised anomaly detection with further verification in a supervised setting.
-I learned how difficult it can be to introduce an innovative approach in a corporate environment.
-
-In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks, preferably within the Mobile Robotics Group.
-Following is one of the research problems that attracts me (I am also open to other problems).
+In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks within the Mobile Robotics Group.
+Following is one of the research problems that attracts me.
 Suppose that a mobile robot moves through highly differentiated space with multiple scene types.
@@
 
-My passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely motivated to carry out research as a PhD student at the University of Oxford.
+My passion for solving technical problems coupled with my demonstrated skills show that I am ready and extremely motivated to carry out research as a DPhil student at the University of Oxford.
 I am confident that I will be able to contribute significantly to and benefit immensely from my stay in Oxford.

### 2016-01-13 phd/sop/mrg.tex
mrg final version

@@
 
-In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks within the Mobile Robotics Group.
+In my doctoral research project I would like to focus on deep neural networks for reinforcement learning and recurrent neural networks within the Mobile Robotics Group (MRG).
 Following is one of the research problems that attracts me.
@@
 
-I expect this project to take around 3 years.
+This project fits well with MRG's research on Experience-Based Classification \cite{experience} as it solves the problem of when to retrain a classifier.
+I expect it to take around 3 years.
 The first stage, lasting about 3 months, would be to develop a recurrent CNN for object classification in sequential data similar to \cite{rcnn}.

### 2016-01-31 ML/hw12/hw.tex
ML hw12 - EM and GMM

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{graphicx}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}

\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learning Homework 12} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

Let $x_n, \mu_k \in \mathcal{R}^D$ and $\Sigma_k = \Sigma = \sigma^2I \in \mathcal{R}^{D \times D}$. Since $\Sigma$ is fixed we do not reestimate it. The probability of observing a single samples $x_n$ simplifies to 
\begin{equation}
 p(x_n|\mu_k, \Sigma) = \frac{1}{(2 \pi )^\frac{D}{2} \sigma} e^{-\frac{1}{2 \sigma^2} ||x_n - \mu_k||^2}
\end{equation}

It follows that the responsibilities $r_{nk}$ are given by a softmax function, that is

\begin{equation}
 r_{nk} = \frac{ \pi_k e^{-\frac{1}{2 \sigma^2} ||x_n - \mu_k||^2} } { \sum_j \pi_j e^{-\frac{1}{2 \sigma^2} ||x_n - \mu_j||^2}} = 
 \begin{cases}
  1 & \text{if } k = \arg \min_j ||x_n - \mu_j|| \\                                                                                                                                  
  0 & \text{otherwise}                                                                                                                                 \end{cases}
\end{equation}

which results in hard assignments of samples to clusters. According to (9.40) in Bishop, the expected value of the complete-data log likelihood is given by

\begin{equation}
 \begin{align}
  L = \sum_{n=1}^{N} \sum_{k=1}^{K} r_{nk} \left( \log{\pi_k} + \log{\mathcal{N}(x_n|\mu_k, \Sigma_k)} \right)
 \end{align}
\end{equation}

In the limit, the Gaussian distribution simplifies to

\begin{equation}
 \lim_{\sigma \rightarrow 0} \mathcal{N}(x_n|\mu_k, \Sigma_k) = \lim_{\sigma \rightarrow 0} -\log{\sigma} - \frac{1}{2 \sigma^2} ||x_n - \mu_k||^2 + \text{const.} \propto -||x_n - \mu_k||^2
\end{equation}

Finally, maximizing the the log-likelihood, given by the formula below, corresponds to the minimization of the K-Means error function.

\begin{equation}
 \begin{align}
  L = \sum_{n=1}^{N} \sum_{k=1}^{K} -r_{nk} ||x_n - \mu_k||^2
 \end{align}
\end{equation}

\section*{Problem 2}

\begin{equation}
 p(x) = \sum_k \pi_k \mathcal{N}(x|\mu_k, \Sigma_k)
\end{equation}

Suppose we have two random variables of the same dimensionality: $X \sim \mathcal{N}(x|\mu_x, \Sigma_x)$ and $Y \sim \mathcal{N}(y|\mu_y, \Sigma_y)$. Let $z = \begin{pmatrix} x \\ y \end{pmatrix} \sim \mathcal{N}(x|\mu_z, \Sigma_z)$ with $\mu_z = \begin{pmatrix} \mu_x \\ \mu_y \end{pmatrix}$ and $\Sigma_z = \begin{pmatrix} \Sigma_x & 0 \\ 0 & \Sigma_y \end{pmatrix}$ We can construct a matrix $A$ such that $Az = \pi_x x + \pi_y y$ It follows that $Az \sim \mathcal{N}(\mu_{xy}, \Sigma_{xy})$ with $\mu_{xy} = \pi_x \mu_x + \pi_y \mu_y$ and  $\Sigma_{xy} = \pi_x^2 \Sigma_x + \pi_y^2 \Sigma_y$. This scheme generalizes to an arbitrary number of components. Specifically, a Gaussian Mixture Model with $K$ components is characterized by the following values:

\begin{equation}
  \begin{align}	
    E[x] &= \sum_{k=1}^K \pi_k \mu_k \\
    Cov(x) &= \sum_{k=1}^K \pi_k^2 \Sigma_k
  \end{align}
\end{equation}

\end{document}
